---
title: "Calculate all Pin Bills for 2021"
author: "Alea Wilbur"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

# Get Pin and Tax Bill Data

There are 1,865,694 pins in cook county and  22,972,902 partial tax bills (i.e. the amount charged by each taxing agency to each pin) when using the tax_bill() command for 2021 data.

File 2 sums pin level bill totals, parcel level bill totals, and class and tax code sums. 



```{r setup}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval=FALSE)


library(tidyverse)
library(DBI)
library(data.table)
library(ggspatial)
library(gstat)
library(here)
library(httr)
library(jsonlite)
library(ptaxsim)
library(sf)
library(stars)
library(glue)

# Create the DB connection with the default name expected by PTAXSIM functions
ptaxsim_db_conn <- DBI::dbConnect(RSQLite::SQLite(), "./ptaxsim.db/ptaxsim-2021.0.4.db")



```


There are 4228 unique tax codes in Cook County in 2021.   

```{r cook-pins}

cook_agency_names <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT DISTINCT agency_num, agency_name
  FROM agency_info
  "
)

 


# has all tax codes and the taxing agency that taxes them. Tax code rates and agency rates. 
cook_tax_codes <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql("
  SELECT*
  FROM tax_code
  WHERE agency_num IN ({cook_agency_names$agency_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
)  
 


# There are 1,864,594 pins taxed by Cook County in 2021.
cook_pins <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql(
  "SELECT DISTINCT pin, class, tax_code_num
  FROM pin
  WHERE tax_code_num IN ({cook_tax_codes$tax_code_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
))
```


Uses tax_bill() to pull all bills for all pins in cook county:
 
```{r cook-taxbills}
# # Normal output from lookup_pin() command. 
# # Includes all types of exemptions

# exe_dt <- lookup_pin(2021, muni_pins$pin) %>%
#   setDT(key = c("year", "pin"))



taxbills_current <- tax_bill(2021,  
                  cook_pins$pin, 
                #  pin_dt = exe_dt, # default option, change for "no exemption" simulation
                  simplify = FALSE)

taxbills_current %>% write_csv("./Output/1_Get_All_Pins-CookPinTaxbills_2021_Actual.csv")

head(taxbills_current)

```
## Pulling incorporated area taxbills only

The code below is what we used originally to grab only pins that were taxed by municipalities. 

Going forward, I am pulling all pins for Cook, summing totals based on property classes and tax codes, and then I will filter out the tax codes being taxed by the taxing agency of interest. (First focus was municipalities and next is probably schools.)



```{r muni-only-pins, eval=FALSE}

options(digits=4, scipen = 999)

library(sf)
library(jsonlite)
library(httr)

# link to the API output as a JSON file
#muni_shp <- read_sf("https://gis.cookcountyil.gov/traditional/rest/services/politicalBoundary/MapServer/2/query?outFields=*&where=1%3D1&f=geojson")

#muni_shp <- read_json("muni_shp.json")
nicknames <- readxl::read_excel("./Necessary_Files/muni_shortnames.xlsx")


# Pull all agency names that exist, then use agency numbers associated with MUNI types to pull only the `muni_agency_names` object.
# 
# There are 1,878 taxing agencies. When grouped by minor_type, there are 134 muni agencies, 639 tif agencies, 30 townships, etc.
# 
# 
# Using the agency numbers for each municipality, I pull all tax codes that have an agency_num included in the muni_agency_names object. By narrowing the agencies down to just Municipality types, this prevents duplicate tax_codes from being pulled.

# grabs all unique muni names & numbs
# don't forget Cicero
muni_agency_names <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT DISTINCT agency_num, agency_name, minor_type
  FROM agency_info
  WHERE minor_type = 'MUNI'
  OR agency_num = '020060000'  

  "
)

muni_agency_names <- muni_agency_names %>% 
    mutate(first6 = str_sub(agency_num,1,6),
         first5 = str_sub(agency_num,1,5)) %>% 
  select(-minor_type)



#Makes a list of ALL taxing agencies, including TIFs, SSAs, etc.

# all agency names, numbers, and types
# includes TIF and non-TIF agencies
all_taxing_agencies <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT agency_num, agency_name, major_type, minor_type
  FROM agency_info
  "
) %>% mutate(first6 = str_sub(agency_num,1,6),
         first5 = str_sub(agency_num,1,5))


muni_agency_nums<- all_taxing_agencies %>% 
  filter(minor_type %in% c("MUNI") | 
           agency_num == "020060000") %>%
   select(agency_num)



# There are 3774 tax codes within Cook County that are taxed by Municipalities.  

# list of all taxcodes in municipalities. 
# This does NOT include unincorporated tax codes!!
muni_tax_codes <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql("
  SELECT*
  FROM tax_code
  WHERE agency_num IN ({muni_agency_names$agency_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
)
```


```{r muni-taxbills, eval=FALSE}
# Create list of all pins in Cook County
muni_pins <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql(
  "SELECT DISTINCT pin, class, tax_code_num
  FROM pin
  WHERE tax_code_num IN ({muni_tax_codes$tax_code_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
))



# Normal output from lookup_pin() command. Includes all types of exemptions
exe_dt <- lookup_pin(2021, muni_pins$pin) %>%
  setDT(key = c("year", "pin"))


taxbills_current <- tax_bill(2021, 
                  muni_pins$pin, 
                  pin_dt = exe_dt, # default option
                  simplify = FALSE)

taxbills_current %>% write_csv("./Output/1_Get_All_Pins-MuniPinTaxbills_2021_Actual.csv")

head(taxbills_current)

```

