---
title: "Calculate all Pin Bills for 2021"
author: "Alea Wilbur"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    df_print: paged
    code_folding: hide
    code_download: true
---

# Get Pin and Tax Bill Data

```{r setup}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)


library(tidyverse)
library(DBI)
library(data.table)
library(ggspatial)
library(gstat)
library(here)
library(httr)
library(jsonlite)
library(ptaxsim)
library(sf)
library(stars)
library(glue)

# Create the DB connection with the default name expected by PTAXSIM functions
ptaxsim_db_conn <- DBI::dbConnect(RSQLite::SQLite(), "./ptaxsim.db/ptaxsim-2021.0.4.db")



```


There are 4228 unique tax codes in Cook County in 2021.   

```{r agency_info}


cook_agency_names <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT DISTINCT agency_num, agency_name
  FROM agency_info
  "
)

# has all tax codes and the taxing agency that taxes them. Tax code rates and agency rates. 
cook_tax_codes <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql("
  SELECT*
  FROM tax_code
  WHERE agency_num IN ({cook_agency_names$agency_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
)  
  
```



```{r cook-pins}
# There are 1,864,594 pins taxed by Cook County in 2021.
cook_pins <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql(
  "SELECT DISTINCT pin, class, tax_code_num
  FROM pin
  WHERE tax_code_num IN ({cook_tax_codes$tax_code_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
))
```


There are 1,865,694 pins in cook county and  22,972,902 partial tax bills (i.e. the amount charged by each taxing agency to each pin) when using the tax_bill() command for 2021 data.

File 2 sums pin level bill totals, parcel level bill totals, and class and tax code sums. 
 
```{r cook-taxbills}
# # Normal output from lookup_pin() command. 
# # Includes all types of exemptions

# exe_dt <- lookup_pin(2021, muni_pins$pin) %>%
#   setDT(key = c("year", "pin"))



taxbills_current <- tax_bill(2021,  
                  cook_pins$pin, 
                #  pin_dt = exe_dt, # default option, change for "no exemption" simulation
                  simplify = FALSE)

taxbills_current %>% write_csv("1_Get_All_Pins-CookPinTaxbills_2021_Actual.csv")

head(taxbills_current)

```
## Pulling incorporated area taxbills only

The code below is what we used originally to grab only pins that were taxed by municipalities. 

Going forward, I am pulling all pins for Cook, summing totals based on property classes and tax codes, and then I will filter out the tax codes being taxed by the taxing agency of interest. (First focus was municipalities and next is probably schools.)




```{r muni-only-pins, eval=FALSE}

options(digits=4, scipen = 999)

library(sf)
library(jsonlite)
library(httr)

# link to the API output as a JSON file
#muni_shp <- read_sf("https://gis.cookcountyil.gov/traditional/rest/services/politicalBoundary/MapServer/2/query?outFields=*&where=1%3D1&f=geojson")

#muni_shp <- read_json("muni_shp.json")
nicknames <- readxl::read_excel("muni_shortnames.xlsx")


# Pull all agency names that exist, then use agency numbers associated with MUNI types to pull only the `muni_agency_names` object.
# 
# There are 1,878 taxing agencies. When grouped by minor_type, there are 134 muni agencies, 639 tif agencies, 30 townships, etc.
# 
# 
# Using the agency numbers for each municipality, I pull all tax codes that have an agency_num included in the muni_agency_names object. By narrowing the agencies down to just Municipality types, this prevents duplicate tax_codes from being pulled.

# grabs all unique muni names & numbs
# don't forget Cicero
muni_agency_names <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT DISTINCT agency_num, agency_name, minor_type
  FROM agency_info
  WHERE minor_type = 'MUNI'
  OR agency_num = '020060000'  

  "
)

muni_agency_names <- muni_agency_names %>% 
    mutate(first6 = str_sub(agency_num,1,6),
         first5 = str_sub(agency_num,1,5)) %>% 
  select(-minor_type)



#Makes a list of ALL taxing agencies, including TIFs, SSAs, etc.

# all agency names, numbers, and types
# includes TIF and non-TIF agencies
all_taxing_agencies <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT agency_num, agency_name, major_type, minor_type
  FROM agency_info
  "
) %>% mutate(first6 = str_sub(agency_num,1,6),
         first5 = str_sub(agency_num,1,5))


muni_agency_nums<- all_taxing_agencies %>% 
  filter(minor_type %in% c("MUNI") | 
           agency_num == "020060000") %>%
   select(agency_num)



# There are 3774 tax codes within Cook County that are taxed by Municipalities.  

# list of all taxcodes in municipalities. 
# This does NOT include unincorporated tax codes!!
muni_tax_codes <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql("
  SELECT*
  FROM tax_code
  WHERE agency_num IN ({muni_agency_names$agency_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
)
```


```{r notused_code, eval=FALSE, include = FALSE}

# has EAV values, extensions by agency_num
agency_dt <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT *
  FROM agency
  WHERE year = 2021
  "
)



# Agency number and agency name for all TIFs
TIF_agencies <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT DISTINCT agency_num, agency_name, major_type, minor_type
  FROM agency_info
  WHERE minor_type = 'TIF'
  "
)

unique_tif_taxcodes <- DBI::dbGetQuery(
  ptaxsim_db_conn, 
  glue_sql("
  SELECT DISTINCT tax_code_num
  FROM tax_code
  WHERE agency_num IN ({TIF_agencies$agency_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
)


tif_distrib <- DBI::dbGetQuery(
  ptaxsim_db_conn, 
  glue_sql("
  SELECT *
  FROM tif_distribution
  WHERE tax_code_num IN ({muni_tax_codes$tax_code_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
) %>% mutate(tax_code_num = as.character(tax_code_num))



all_taxing_agencies <- all_taxing_agencies %>% 
  left_join(muni_agency_names, by = c("first5", "first6")) %>% 
  rename(muni_name =  agency_name.y,
        muni_num = agency_num.y,
        agency_name = agency_name.x,
        agency_num = agency_num.x)


# combine taxing agency names and agency type to data table that has eav and extension values
agency_data <- right_join(agency_dt, all_taxing_agencies) %>% 
  # get rid of unneeded columns to make table outputs smaller
  select(-c(cty_dupage_eav:cty_livingston_eav)) %>% # drop some of the unused variables
  arrange(agency_num)
```


```{r muni-taxbills, eval=FALSE}
# Create list of all pins in Cook County
muni_pins <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql(
  "SELECT DISTINCT pin, class, tax_code_num
  FROM pin
  WHERE tax_code_num IN ({muni_tax_codes$tax_code_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
))



# Normal output from lookup_pin() command. Includes all types of exemptions
exe_dt <- lookup_pin(2021, muni_pins$pin) %>%
  setDT(key = c("year", "pin"))


taxbills_current <- tax_bill(2021, 
                  muni_pins$pin, 
                  pin_dt = exe_dt, # default option
                  simplify = FALSE)

taxbills_current %>% write_csv("1_Get_All_Pins-MuniPinTaxbills_2021_Actual.csv")

head(taxbills_current)

```

