---
title: 'Hypothetical Taxbills: Remove all Exemptions'
author: "Alea Wilbur"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    code_folding: hide
    code_download: yes
  pdf_document: default
---


```{r setup, warning=FALSE, message=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)


library(tidyverse)
library(data.table)
library(gstat)
library(ptaxsim)


```


```{r}
taxcode_taxrates <- read_csv("2_taxcode_taxrates.csv")

pin_eav <- read_csv("3_Exemption_Details_output-all_cook_pin_exemptions_2021_actual.csv") %>%
    filter(class_code !=0) %>%

  mutate(exempt_EAV = exe_homeowner + exe_senior + exe_freeze + exe_longtime_homeowner + 
        exe_disabled + exe_vet_returning + exe_vet_dis_lt50 + exe_vet_dis_50_69 + exe_vet_dis_ge70 + exe_abate) %>%
  mutate(zero_bill = ifelse(eav <= exempt_EAV, 1, 0))

# table(pin_eav$zero_bill)

table(pin_eav$major_class_code, pin_eav$zero_bill)

pin_eav <- left_join(pin_eav, taxcode_taxrates, by = c("tax_code_num" = "tax_code"))

# head(pin_eav)
```


```{r}
tax_bill_change <- pin_eav %>% 
  mutate(taxbase_current = eav-exempt_EAV,
         taxbase_new = eav,
    bill_current = tax_rate_current * (eav-exempt_EAV),
    bill_noexemps =  taxrate_new * eav,
    bill_change = bill_noexemps - bill_current,
    bill_pct_change = bill_change/(bill_current)) %>%
  
  select(pin, av, eav, tax_code_num, class_code, bill_change, bill_pct_change, bill_current, bill_noexemps, clean_name, taxrate_change, taxbase_current, taxbase_new, taxrate_new, tax_rate_current, major_class_code, eav, exempt_EAV)
```


```{r}
muni_class_stats <- tax_bill_change %>% 
  group_by(clean_name, class_code) %>% 
  arrange(taxbase_new)%>%
  summarise(
    median_bill_current = median(bill_current, na.rm=TRUE),
    median_bill_noexemps = median(bill_noexemps, na.rm=TRUE),
    median_AV = median(av, na.rm=TRUE),
    median_eav = median(eav, na.rm=TRUE),
    pin_count = n(),
        group_eav = sum(taxbase_current, na.rm = TRUE),
    group_eav_new = sum(taxbase_new,na.rm=TRUE),
    group_av = sum(av, na.rm=TRUE) ) %>% 
  mutate(bill_change = median_bill_noexemps - median_bill_current) %>%
  mutate_at(vars(bill_change, median_bill_current:group_av), funs(round(., 0)))


write_csv(muni_class_stats, "4_Taxbill_Hypotheticals-taxbill_change_foreachPropClass_perMuni.csv")

muni_class_stats %>% select(Municipality = clean_name, class_code, bill_change, everything())



muni_majorclass_stats <- tax_bill_change %>% 
  filter(class_code !=0) %>%
    arrange(taxbase_new)%>%

  group_by(clean_name, major_class_code) %>% 
  summarise(
    median_bill_current = median(bill_current, na.rm=TRUE),
    median_bill_noexemps = median(bill_noexemps, na.rm=TRUE),
    median_AV = median(av, na.rm=TRUE),
    median_eav = median(eav, na.rm=TRUE),
    pin_count = n(),
    group_eav_current = sum(taxbase_current, na.rm = TRUE),
    group_eav_new = sum(taxbase_new,na.rm=TRUE),
    group_av = sum(av, na.rm=TRUE) ) %>% 
  mutate(bill_change =median_bill_noexemps - median_bill_current ) %>%
  mutate_at(vars(bill_change, median_bill_current:group_av), funs(round(., 0)))


muni_majorclass_stats %>% select(Municipality = clean_name, major_class = major_class_code, bill_change, everything())

write_csv(muni_majorclass_stats, "4_Taxbill_Hypotheticals-taxbill_change_eachMajorClass_perMuni.csv")



muni_rachelgroups_stats <- tax_bill_change %>% 
  ungroup() %>%
  mutate(grouped_classes = ifelse(major_class_code  == "2", "Class 2", NA),
        grouped_classes = ifelse(major_class_code == "3", "Class 3", grouped_classes),
        grouped_classes = ifelse(is.na(grouped_classes), "Other Classes", grouped_classes)
       ) %>%
  group_by(clean_name, grouped_classes) %>%
  arrange(taxbase_new) %>%
   summarise(
    median_bill_current = median(bill_current, na.rm=TRUE),
    median_bill_noexemps = median(bill_noexemps, na.rm=TRUE),
    median_AV = median(av, na.rm=TRUE),
    median_eav = median(eav, na.rm=TRUE),
    pin_count = n(),
    group_eav_current = sum(taxbase_current, na.rm = TRUE),
    group_eav_new = sum(taxbase_new,na.rm=TRUE),
    group_av = sum(av, na.rm=TRUE) ) %>% 
  mutate(bill_change =  median_bill_noexemps - median_bill_current) %>% 
  mutate_at(vars(bill_change, median_bill_current:group_av), funs(round(., 0)))

muni_rachelgroups_stats %>% select(Municipality = clean_name,  grouped_classes, bill_change, everything())

write_csv(muni_rachelgroups_stats, "4_Taxbill_Hypotheticals-taxbill_change_20230814.csv")
```


```{r, fig.show='hold', out.width="50%"}
tax_bill_change %>% ungroup() %>%
  filter(major_class_code == 2 & bill_current < 30000 & clean_name == "Dolton") %>% ggplot() +
  geom_histogram(aes(x= bill_noexemps, bins = 100))+
    geom_vline(aes(xintercept=median(bill_current, na.rm=TRUE), color = "red")) +
    theme_classic()+theme(legend.position = "none") + 

  scale_x_continuous(limits = c(0,20000))


tax_bill_change %>% ungroup() %>%
  filter(major_class_code == 2 & bill_noexemps < 30000 & clean_name == "Dolton") %>% ggplot() +
  geom_histogram(aes(x= bill_noexemps, bins = 100)) +
  theme_classic() +
  geom_vline(aes(xintercept=median(bill_noexemps, na.rm=TRUE), color = "red")) + theme(legend.position = "none") +
    scale_x_continuous(limits = c(0,20000))
```


```{r, fig.show='hold', out.width="50%"}
tax_bill_change %>% ungroup() %>%
  filter(major_class_code == 2 & bill_current < 30000 & clean_name == "Dolton") %>% ggplot() +
  geom_histogram(aes(x= bill_current, bins = 100))+
    geom_vline(aes(xintercept=median(bill_current, na.rm=TRUE), color = "red")) +
    theme_classic()+theme(legend.position = "none") + 

  scale_x_continuous(limits = c(0,20000))


tax_bill_change %>% ungroup() %>%
  filter(major_class_code == 2 & bill_noexemps < 30000 & clean_name == "Dolton") %>% ggplot() +
  geom_histogram(aes(x= bill_noexemps, bins = 100)) +
  theme_classic() +
  geom_vline(aes(xintercept=median(bill_noexemps, na.rm=TRUE), color = "red")) + theme(legend.position = "none") +
    scale_x_continuous(limits = c(0,20000))
```

**Class 3 Tax bills: **
```{r, fig.show='hold', out.width="50%"}
tax_bill_change %>% ungroup() %>%
  filter(major_class_code == 3 #& bill_current < 20000
         ) %>% ggplot() +
  geom_histogram(aes(x= bill_current, bins = 50))+
    geom_vline(aes(xintercept=median(bill_current, na.rm=TRUE), color = "red")) +
    theme_classic()+theme(legend.position = "none") + 

  scale_x_continuous(limits = c(0,100000))


tax_bill_change %>% ungroup() %>%
  filter(major_class_code == 3 #& bill_noexemps < 20000
         ) %>% ggplot() +
  geom_histogram(aes(x= bill_noexemps, bins = 50)) +
  theme_classic() +
  geom_vline(aes(xintercept=median(bill_noexemps, na.rm=TRUE), color = "red")) + theme(legend.position = "none") +
    scale_x_continuous(limits = c(0,100000))
```


Class 2 bills in Chicago would only change $28 for the median Class 2 property.


```{r}

muni_rachelgroups_stats2 <- tax_bill_change %>% ungroup() %>%
  mutate(grouped_classes = ifelse(major_class_code  == "2", "Class 2", NA),
        grouped_classes = ifelse(major_class_code %in% c("3", "9"), "Class 3", grouped_classes),
        grouped_classes = ifelse(is.na(grouped_classes), "Other Classes", grouped_classes)
       ) %>%
  group_by(clean_name, grouped_classes) %>%
    arrange(taxbase_new) %>%

   summarise(
    median_bill_current = median(bill_current, na.rm=TRUE),
    median_bill_noexemps = median(bill_noexemps, na.rm=TRUE),
    median_AV = median(av, na.rm=TRUE),
    median_eav = median(eav, na.rm=TRUE),
    pin_count = n(),
    group_eav_current = sum(taxbase_current, na.rm = TRUE),
    group_eav_new = sum(taxbase_new,na.rm=TRUE),
    group_av = sum(av, na.rm=TRUE) ) %>% 
  mutate(bill_change =  median_bill_noexemps - median_bill_current) %>% 
  mutate_at(vars(bill_change, median_bill_current:group_av), funs(round(., 0)))

muni_rachelgroups_stats2 %>% select(clean_name,  grouped_classes, bill_change, everything())
```


# Alternate Property Categories

Uses owner, renter, commercial and industrial.

```{r}
class_dict <- read_csv("class_dict_expanded.csv") %>% 
  mutate(class_code = as.character(class_code))

muni_Alea_cat_stats <- tax_bill_change %>% 
    filter(class_code !=0) %>%
  mutate(class_code = as.character(class_code)) %>%
  left_join(class_dict) %>%
    arrange(taxbase_new)%>%

  group_by(clean_name, Alea_cat) %>% 
  summarise(
    median_bill_current = median(bill_current, na.rm=TRUE),
    median_bill_noexemps = median(bill_noexemps, na.rm=TRUE),
    median_AV = median(av, na.rm=TRUE),
    median_eav = median(eav, na.rm=TRUE),
    pin_count = n(),
    group_eav_current = sum(taxbase_current, na.rm = TRUE),
    group_eav_new = sum(taxbase_new,na.rm=TRUE),
    group_av = sum(av, na.rm=TRUE) ) %>% 
  mutate(bill_change = median_bill_noexemps -  median_bill_current ) %>%
  mutate_at(vars(bill_change, median_bill_current:group_av), funs(round(., 0)))


muni_Alea_cat_stats %>% 
  filter(Alea_cat == "Owner Occupied") %>% 
  select(Municipality = clean_name, bill_change, everything()) %>% 
  arrange(bill_change)
```

**Owner occupied tax bills:**

```{r, fig.show='hold', out.width="50%"}
tax_bill_change %>% 
    filter(class_code !=0) %>%
  mutate(class_code = as.character(class_code)) %>%
  left_join(class_dict) %>%
  filter(Alea_cat == "Owner Occupied" & bill_current < 30000) %>% ggplot() +
  geom_histogram(aes(x= bill_current, bins = 100))+
    geom_vline(aes(xintercept=median(bill_current, na.rm=TRUE), color = "red")) +
    theme_classic()+theme(legend.position = "none") + 

  scale_x_continuous(limits = c(0,30000))


tax_bill_change %>% 
    filter(class_code !=0) %>%
  mutate(class_code = as.character(class_code)) %>%
  left_join(class_dict) %>%
  filter(Alea_cat == "Owner Occupied" & bill_noexemps < 30000) %>% ggplot() +
  geom_histogram(aes(x= bill_noexemps, bins = 100)) +
  theme_classic() +
  geom_vline(aes(xintercept=median(bill_noexemps, na.rm=TRUE), color = "red")) + theme(legend.position = "none") +
    scale_x_continuous(limits = c(0,30000))
```

# "Cost" of Exemptions

```{r}
library(DBI)
library(ptaxsim)
library(glue)
# Create the DB connection with the default name expected by PTAXSIM functions
ptaxsim_db_conn <- DBI::dbConnect(RSQLite::SQLite(), "./ptaxsim.db/ptaxsim-2021.0.4.db")



muni_agency_names <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT DISTINCT agency_num, agency_name, minor_type
  FROM agency_info
  WHERE minor_type = 'MUNI'
  OR agency_num = '020060000'  

  "
)

muni_tax_codes <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql("
  SELECT*
  FROM tax_code
  WHERE agency_num IN ({muni_agency_names$agency_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
)
nicknames <- readxl::read_excel("muni_shortnames.xlsx")


tax_bill_change %>% 
    filter(tax_code_num %in% muni_tax_codes$tax_code_num) %>%

    filter(class_code !=0) %>%
  mutate(class_code = as.character(class_code)) %>%
  left_join(nicknames) %>%
  filter(Triad == "South" & bill_change > 0) %>%
  arrange(bill_change) %>%
  summarize(cost = sum(bill_change, na.rm=TRUE))
```

Covering all changes in tax bills in the South triad would cost $211 million? CHECK 

```{r}
tax_bill_change %>% 
    filter(tax_code_num %in% muni_tax_codes$tax_code_num) %>%

  filter(bill_pct_change > 0.15 | bill_current == 0) %>%
  summarize(cost = sum(bill_change, na.rm=TRUE))


tax_bill_change %>% 
  filter(tax_code_num %in% muni_tax_codes$tax_code_num) %>%
  filter(major_class_code == 2) %>% 
  filter(bill_pct_change > 0.15) %>% 
  arrange(bill_current) %>%
  summarize(cost = sum(bill_change, na.rm=TRUE))

tax_bill_change %>% 
  filter(tax_code_num %in% muni_tax_codes$tax_code_num) %>%
  filter(major_class_code == 2) %>% 
  filter(bill_current== 0) %>% 
  arrange(bill_current) %>%
    summarize(cost = sum(bill_change, na.rm=TRUE))
```


```{r}
tax_bill_change %>% 
  filter(tax_code_num %in% muni_tax_codes$tax_code_num) %>%
  filter(major_class_code == 2) %>% 
  filter(bill_current != 0 & bill_pct_change > 0.15) %>% 
  arrange(bill_change)  %>%
    summarize(cost = sum(bill_change, na.rm=TRUE))

```


Check this -> Paying for all tax bills that experienced a 15% increase from the elimination of exemptions or had a \$0 taxbill would cost $825,963,411.