---
title: 'Hypothetical Taxbills: Remove all Exemptions'
author: "Alea Wilbur"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    code_folding: hide
    code_download: yes
  pdf_document: default
---


```{r setup, warning=FALSE, message=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)


library(tidyverse)
library(data.table)
library(gstat)
library(ptaxsim)


```

# Bring in data

The tax code taxrates were made in file `2_...rmd` and the exemptions file were created in file `3_.....rmd`

## Finding the "Median Property"

The median assessed value is used to determine the `median property` for various groups (e.g. the municipality, the 1-digit major property class, and 3-digit property class). 

The assessed value and original equalized assessed values come from the `pin` data table within PTAXSIM. This table also has every type of exemption that the property received and the amount of EAV that was exempt due to the exemption. 

```{r}
taxcode_taxrates <- read_csv("2_taxcode_taxrates.csv")

pin_eav <- read_csv("3_Exemption_Details_output-all_cook_pin_exemptions_2021_actual.csv") %>%
    filter(class_code !=0) %>%

  mutate(exempt_EAV = exe_homeowner + exe_senior + exe_freeze + exe_longtime_homeowner + 
        exe_disabled + exe_vet_returning + exe_vet_dis_lt50 + exe_vet_dis_50_69 + exe_vet_dis_ge70 + exe_abate) %>%
  mutate(zero_bill = ifelse(eav <= exempt_EAV, 1, 0))

# table(pin_eav$zero_bill)

table(pin_eav$major_class_code, pin_eav$zero_bill)

pin_eav <- left_join(pin_eav, taxcode_taxrates, by = c("tax_code_num" = "tax_code"))

# head(pin_eav)
```


```{r}
tax_bill_change <- pin_eav %>% 
  mutate(taxbase_current = eav-exempt_EAV,
         taxbase_new = eav,
    bill_current = tax_rate_current * (eav-exempt_EAV),
    bill_noexemps =  taxrate_new * eav,
    bill_change = bill_noexemps - bill_current,
    bill_pct_change = bill_change/(bill_current)) %>%
  
  select(pin, av, eav, tax_code_num, class_code, bill_change, bill_pct_change, bill_current, bill_noexemps, clean_name, taxrate_change, taxbase_current, taxbase_new, taxrate_new, tax_rate_current, major_class_code, eav, exempt_EAV)
```


```{r}
muni_class_stats <- tax_bill_change %>% 
  group_by(clean_name, class_code) %>% 
  arrange(taxbase_new)%>%
  summarise(
    median_bill_current = median(bill_current, na.rm=TRUE),
    median_bill_noexemps = median(bill_noexemps, na.rm=TRUE),
    median_AV = median(av, na.rm=TRUE),
    median_eav = median(eav, na.rm=TRUE),
    pin_count = n(),
        group_eav = sum(taxbase_current, na.rm = TRUE),
    group_eav_new = sum(taxbase_new,na.rm=TRUE),
    group_av = sum(av, na.rm=TRUE) ) %>% 
  mutate(bill_change = median_bill_noexemps - median_bill_current) %>%
  mutate_at(vars(bill_change, median_bill_current:group_av), funs(round(., 0)))


write_csv(muni_class_stats, "4_Taxbill_Hypotheticals-taxbill_change_foreachPropClass_perMuni.csv")

muni_class_stats %>% select(Municipality = clean_name, class_code, bill_change, everything())



muni_majorclass_stats <- tax_bill_change %>% 
  filter(class_code !=0) %>%
    arrange(taxbase_new)%>%

  group_by(clean_name, major_class_code) %>% 
  summarise(
    median_bill_current = median(bill_current, na.rm=TRUE),
    median_bill_noexemps = median(bill_noexemps, na.rm=TRUE),
    median_AV = median(av, na.rm=TRUE),
    median_eav = median(eav, na.rm=TRUE),
    pin_count = n(),
    group_eav_current = sum(taxbase_current, na.rm = TRUE),
    group_eav_new = sum(taxbase_new,na.rm=TRUE),
    group_av = sum(av, na.rm=TRUE) ) %>% 
  mutate(bill_change =median_bill_noexemps - median_bill_current ) %>%
  mutate_at(vars(bill_change, median_bill_current:group_av), funs(round(., 0)))


muni_majorclass_stats %>% select(Municipality = clean_name, major_class = major_class_code, bill_change, everything())

write_csv(muni_majorclass_stats, "4_Taxbill_Hypotheticals-taxbill_change_eachMajorClass_perMuni.csv")



muni_rachelgroups_stats <- tax_bill_change %>% 
  ungroup() %>%
  mutate(grouped_classes = ifelse(major_class_code  == "2", "Class 2", NA),
        grouped_classes = ifelse(major_class_code == "3", "Class 3", grouped_classes),
        grouped_classes = ifelse(is.na(grouped_classes), "Other Classes", grouped_classes)
       ) %>%
  group_by(clean_name, grouped_classes) %>%
  arrange(taxbase_new) %>%
   summarise(
    median_bill_current = median(bill_current, na.rm=TRUE),
    median_bill_noexemps = median(bill_noexemps, na.rm=TRUE),
    median_AV = median(av, na.rm=TRUE),
    median_eav = median(eav, na.rm=TRUE),
    pin_count = n(),
    group_eav_current = sum(taxbase_current, na.rm = TRUE),
    group_eav_new = sum(taxbase_new,na.rm=TRUE),
    group_av = sum(av, na.rm=TRUE) ) %>% 
  mutate(bill_change =  median_bill_noexemps - median_bill_current) %>% 
  mutate_at(vars(bill_change, median_bill_current:group_av), funs(round(., 0)))

muni_rachelgroups_stats %>% select(Municipality = clean_name,  grouped_classes, bill_change, everything())

write_csv(muni_rachelgroups_stats, "4_Taxbill_Hypotheticals-taxbill_change_20230814.csv")
```


```{r, fig.show='hold', out.width="50%"}
tax_bill_change %>% ungroup() %>%
  filter(major_class_code == 2 & bill_current < 30000 & clean_name == "Dolton") %>% ggplot() +
  geom_histogram(aes(x= bill_current, bins = 100))+
    geom_vline(aes(xintercept=median(bill_current, na.rm=TRUE), color = "red")) +
    theme_classic()+theme(legend.position = "none") + 

  scale_x_continuous(limits = c(0,20000))


tax_bill_change %>% ungroup() %>%
  filter(major_class_code == 2 & bill_noexemps < 30000 & clean_name == "Dolton") %>% ggplot() +
  geom_histogram(aes(x= bill_noexemps, bins = 100)) +
  theme_classic() +
  geom_vline(aes(xintercept=median(bill_noexemps, na.rm=TRUE), color = "red")) + theme(legend.position = "none") +
    scale_x_continuous(limits = c(0,20000))
```


```{r, fig.show='hold', out.width="50%"}
tax_bill_change %>% ungroup() %>%
  filter(major_class_code == 2 & bill_current < 30000 & clean_name == "Chicago") %>% ggplot() +
  geom_histogram(aes(x= bill_current, bins = 100))+
    geom_vline(aes(xintercept=median(bill_current, na.rm=TRUE), color = "red")) +
    theme_classic()+theme(legend.position = "none") + 

  scale_x_continuous(limits = c(0,20000))


tax_bill_change %>% ungroup() %>%
  filter(major_class_code == 2 & bill_noexemps < 30000 & clean_name == "Chicago") %>% ggplot() +
  geom_histogram(aes(x= bill_noexemps, bins = 100)) +
  theme_classic() +
  geom_vline(aes(xintercept=median(bill_noexemps, na.rm=TRUE), color = "red")) + theme(legend.position = "none") +
    scale_x_continuous(limits = c(0,20000))
```

**Class 3 Tax bills: **
```{r, fig.show='hold', out.width="50%"}
tax_bill_change %>% ungroup() %>%
  filter(major_class_code == 3 #& bill_current < 20000
         ) %>% ggplot() +
  geom_histogram(aes(x= bill_current, bins = 50))+
    geom_vline(aes(xintercept=median(bill_current, na.rm=TRUE), color = "red")) +
    theme_classic()+theme(legend.position = "none") + 

  scale_x_continuous(limits = c(0,100000))


tax_bill_change %>% ungroup() %>%
  filter(major_class_code == 3 #& bill_noexemps < 20000
         ) %>% ggplot() +
  geom_histogram(aes(x= bill_noexemps, bins = 50)) +
  theme_classic() +
  geom_vline(aes(xintercept=median(bill_noexemps, na.rm=TRUE), color = "red")) + theme(legend.position = "none") +
    scale_x_continuous(limits = c(0,100000))
```


Class 2 bills in Chicago would only change $28 for the median Class 2 property.


```{r}

muni_rachelgroups_stats2 <- tax_bill_change %>% ungroup() %>%
  mutate(grouped_classes = ifelse(major_class_code  == "2", "Class 2", NA),
        grouped_classes = ifelse(major_class_code %in% c("3", "9"), "Class 3", grouped_classes),
        grouped_classes = ifelse(is.na(grouped_classes), "Other Classes", grouped_classes)
       ) %>%
  group_by(clean_name, grouped_classes) %>%
    arrange(taxbase_new) %>%

   summarise(
    median_bill_current = median(bill_current, na.rm=TRUE),
    median_bill_noexemps = median(bill_noexemps, na.rm=TRUE),
    median_AV = median(av, na.rm=TRUE),
    median_eav = median(eav, na.rm=TRUE),
    pin_count = n(),
    group_eav_current = sum(taxbase_current, na.rm = TRUE),
    group_eav_new = sum(taxbase_new,na.rm=TRUE),
    group_av = sum(av, na.rm=TRUE) ) %>% 
  mutate(bill_change =  median_bill_noexemps - median_bill_current) %>% 
  mutate_at(vars(bill_change, median_bill_current:group_av), funs(round(., 0)))

muni_rachelgroups_stats2 %>% select(clean_name,  grouped_classes, bill_change, everything())
```


### Alternate Property Categories

Uses owner, renter, commercial and industrial.

```{r}
class_dict <- read_csv("class_dict_expanded.csv") %>% 
  mutate(class_code = as.character(class_code))

muni_Alea_cat_stats <- tax_bill_change %>% 
    filter(class_code !=0) %>%
  mutate(class_code = as.character(class_code)) %>%
  left_join(class_dict) %>%
    arrange(taxbase_new)%>%

  group_by(clean_name, Alea_cat) %>% 
  summarise(
    median_bill_current = median(bill_current, na.rm=TRUE),
    median_bill_noexemps = median(bill_noexemps, na.rm=TRUE),
    median_AV = median(av, na.rm=TRUE),
    median_eav = median(eav, na.rm=TRUE),
    pin_count = n(),
    group_eav_current = sum(taxbase_current, na.rm = TRUE),
    group_eav_new = sum(taxbase_new,na.rm=TRUE),
    group_av = sum(av, na.rm=TRUE) ) %>% 
  mutate(bill_change = median_bill_noexemps -  median_bill_current ) %>%
  mutate_at(vars(bill_change, median_bill_current:group_av), funs(round(., 0)))


muni_Alea_cat_stats %>% 
  filter(Alea_cat == "Owner Occupied") %>% 
  select(Municipality = clean_name, bill_change, everything()) %>% 
  arrange(bill_change)
```

**Owner occupied tax bills:**

```{r, fig.show='hold', out.width="50%"}
tax_bill_change %>% 
    filter(class_code !=0) %>%
  mutate(class_code = as.character(class_code)) %>%
  left_join(class_dict) %>%
  filter(Alea_cat == "Owner Occupied" & bill_current < 30000) %>% ggplot() +
  geom_histogram(aes(x= bill_current, bins = 100))+
    geom_vline(aes(xintercept=median(bill_current, na.rm=TRUE), color = "red")) +
    theme_classic()+theme(legend.position = "none") + 

  scale_x_continuous(limits = c(0,30000))


tax_bill_change %>% 
    filter(class_code !=0) %>%
  mutate(class_code = as.character(class_code)) %>%
  left_join(class_dict) %>%
  filter(Alea_cat == "Owner Occupied" & bill_noexemps < 30000) %>% ggplot() +
  geom_histogram(aes(x= bill_noexemps, bins = 100)) +
  theme_classic() +
  geom_vline(aes(xintercept=median(bill_noexemps, na.rm=TRUE), color = "red")) + theme(legend.position = "none") +
    scale_x_continuous(limits = c(0,30000))
```

# "Cost" of Exemptions

```{r}
library(DBI)
library(ptaxsim)
library(glue)
# Create the DB connection with the default name expected by PTAXSIM functions
ptaxsim_db_conn <- DBI::dbConnect(RSQLite::SQLite(), "./ptaxsim.db/ptaxsim-2021.0.4.db")



muni_agency_names <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT DISTINCT agency_num, agency_name, minor_type
  FROM agency_info
  WHERE minor_type = 'MUNI'
  OR agency_num = '020060000'  

  "
)

muni_tax_codes <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql("
  SELECT*
  FROM tax_code
  WHERE agency_num IN ({muni_agency_names$agency_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
)
nicknames <- readxl::read_excel("muni_shortnames.xlsx")


tax_bill_change %>% 
    filter(tax_code_num %in% muni_tax_codes$tax_code_num) %>%
    filter(class_code !=0) %>%
  mutate(class_code = as.character(class_code)) %>%
  left_join(nicknames) %>%
  filter(Triad == "South" & bill_change > 0) %>%
  arrange(bill_change) %>%
  summarize(cost = sum(bill_change, na.rm=TRUE))
```

Covering all changes in tax bills in the South triad would cost $211 million.  

```{r}
tax_bill_change %>% 
    filter(tax_code_num %in% muni_tax_codes$tax_code_num) %>%

  filter(bill_pct_change > 0.15 | bill_current == 0) %>%
  summarize(cost = sum(bill_change, na.rm=TRUE))


tax_bill_change %>% 
  filter(tax_code_num %in% muni_tax_codes$tax_code_num) %>%
  filter(major_class_code == 2) %>% 
  filter(bill_pct_change > 0.15) %>% 
  arrange(bill_current) %>%
  summarize(cost = sum(bill_change, na.rm=TRUE))

tax_bill_change %>% 
  filter(tax_code_num %in% muni_tax_codes$tax_code_num) %>%
#  filter(major_class_code == 2) %>% 
  filter(class_code >= 200 & class_code <=299)%>%
  filter(bill_current== 0) %>% 
  arrange(bill_current) %>%
    summarize(cost = sum(bill_change, na.rm=TRUE))
```


```{r}
tax_bill_change %>% 
  filter(tax_code_num %in% muni_tax_codes$tax_code_num) %>%
  filter(major_class_code == 2) %>% 
  filter(bill_current != 0 & bill_pct_change > 0.15) %>% 
  arrange(bill_change)  %>%
    summarize(cost = sum(bill_change, na.rm=TRUE))

```


Check this -> Paying for all tax bills that experienced a 15% increase from the elimination of exemptions or had a \$0 taxbill would cost $825,963,411.

```{r}
tax_bill_change %>% filter(clean_name == "Dolton") %>% # 8783
  filter(class_code >=200 & class_code <=299) %>%
  mutate(shittyvariable = median(bill_current, na.rm=TRUE),
         shittyvariable2 = median(bill_current)) %>% 
  select(shittyvariable, shittyvariable2, everything())
  
```

# Finding Similar Properties and their Taxbills

```{r}
library(glue)
ptaxsim_db_conn <- DBI::dbConnect(RSQLite::SQLite(), "./ptaxsim.db/ptaxsim-2021.0.4.db")

# has EAV values, extensions by agency_num
agency_dt <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT *
  FROM agency
  WHERE year = 2021
  "
)

# grabs all unique muni names & numbs
# don't forget Cicero
muni_agency_names <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT DISTINCT agency_num, agency_name, minor_type
  FROM agency_info
  WHERE minor_type = 'MUNI'
  OR agency_num = '020060000'  

  "
)


# list of all taxcodes in municipalities. 
# This does NOT include unincorporated tax codes!!
doltchi_tax_codes <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql("
  SELECT*
  FROM tax_code
  WHERE (agency_num = '030210000' OR agency_num = '030310000')
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
) %>% mutate(tax_code = as.character(tax_code_num))

```

```{r eval=FALSE}
taxbills_current <- read_csv("1_Get_All_Pins-CookPinTaxbills_2021_Actual.csv")
# 22,453,875 tax bills in 2021 in municipalities (incorporated areas). 
# 22,972,902 tax bills in all of Cook County in 2021 (incorporated and unincorporated)
# DOES INCLUDE unincorporated tax bills based on how we pulled the data in Step 1.



# 1,825,816 billed properties with 14-digit PINs  for incorporated areras
# 1,864,594 billed properties in Cook county (incorporated and unincorporated)
# 
pin14_bills_current <- taxbills_current %>%
  group_by(tax_code, class, pin) %>%
  
  mutate(total_bill = final_tax_to_dist + final_tax_to_tif) %>% # from each taxing agency
  
  summarize(total_billed = sum(total_bill, na.rm = TRUE), # total on someone's property tax bill
            av = first(av),
            eav = first(eav),
            pin_count_in_parcel = n(),
            final_tax_to_dist = sum(final_tax_to_dist, na.rm = TRUE),
            final_tax_to_tif = sum(final_tax_to_tif, na.rm = TRUE),
            tax_amt_exe = sum(tax_amt_exe, na.rm = TRUE), # revenue lost due to exemptions
            tax_amt_pre_exe = sum(tax_amt_pre_exe, na.rm = TRUE), # total rev before all exemptions
            tax_amt_post_exe = sum(tax_amt_post_exe, na.rm = TRUE), # total rev after all exemptions
            rpm_tif_to_cps = sum(rpm_tif_to_cps, na.rm = TRUE), # not used
            rpm_tif_to_rpm = sum(rpm_tif_to_rpm, na.rm=TRUE), # not used
            rpm_tif_to_dist = sum(rpm_tif_to_dist, na.rm=TRUE), # not used
            tif_share = mean(tif_share, na.rm=TRUE), # not used
  )  %>% 
  mutate(propclass_1dig = str_sub(class, 1, 1))

head(pin14_bills_current)


DoltonChicago <- pin14_bills_current %>% 
  filter(tax_code %in% doltchi_tax_codes$tax_code_num) %>%
  filter(class != "0")


DoltonChicago <- DoltonChicago %>% 
  mutate(tax_code = as.character(tax_code))%>%
  left_join(doltchi_tax_codes) %>% 
  group_by(agency_num, class) %>% arrange(av)
```


```{r}
DoltonChicago <- read_csv("Cholton_taxbills.csv")
nicknames <- readxl::read_excel("muni_shortnames.xlsx")

class_dict <- read_csv("class_dict.csv")


DoltonChicago %>% 
  filter(class == "203") %>% 
  group_by(agency_num, class) %>%
  summarize(medianbilll = median(total_billed),
            meanbill = mean(total_billed))


## Dolton ##
DoltonChicago %>% filter(agency_num == "030210000" & class == "203" & between(total_billed,3150,3152))
# av - 19000, eav = 57000
# had $670 "exempt" on their bill


DoltonChicago %>% filter(agency_num == "030210000" & class == "203" & between(total_billed,3150,3152)) %>%
  summarize(av = mean(av),
            eav = mean(eav),
            tax_amt_post_exe = mean(tax_amt_post_exe),
            tax_amt_exe = mean(tax_amt_exe),
            tax_amt_pre_exe = mean(tax_amt_pre_exe),
            pin_count = n())


## Chicago ## 

# Chicago only has 2 pins that are similar to Dolton's median pin (which had a lot of matches)
DoltonChicago %>% filter(agency_num == "030310000" & class == "203" & between(total_billed,3140,3160))

DoltonChicago %>% filter(agency_num == "030310000" & class == "203" & between(total_billed,3140,3160)) %>%
  summarize(av = mean(av),
            eav = mean(eav),
            tax_amt_post_exe = mean(tax_amt_post_exe),
            tax_amt_exe = mean(tax_amt_exe),
            tax_amt_pre_exe = mean(tax_amt_pre_exe),
            pin_count = n()
  )

# av = 
#exempt eav = 
```

```{r}
DoltonChicago %>% 
  filter(agency_num == "030210000" & class == "203" & between(av,9990,10000)) %>%
  summarize(
          tax_amt_post_exe = mean(tax_amt_post_exe),
      tax_amt_exe = mean(tax_amt_exe),
            tax_amt_pre_exe = mean(tax_amt_pre_exe),
            pin_count = n()
  )

# Chicago only has 2 pins that are similar to Dolton's median pin (which had a lot of matches)
DoltonChicago %>% 
  filter(agency_num == "030310000" & class == "203" & between(av,9990,10000)) %>%
    summarize(              
      tax_amt_post_exe = mean(tax_amt_post_exe),
      tax_amt_exe = mean(tax_amt_exe),
            tax_amt_pre_exe = mean(tax_amt_pre_exe),
            pin_count = n())
```

```{r}
DoltonChicago %>% 
  filter(agency_num == "030210000" & class == "203" & between(tax_amt_post_exe,3500,3600)) %>%
  summarize(
    tax_amt_post_exe = mean(tax_amt_post_exe),
    tax_amt_exe = mean(tax_amt_exe),
    tax_amt_pre_exe = mean(tax_amt_pre_exe),
    pin_count = n(),
    av = mean(av),
    eav = mean(eav)
  )

# Chicago only has 2 pins that are similar to Dolton's median pin (which had a lot of matches)
DoltonChicago %>% 
  filter(agency_num == "030310000" & class == "203" & between(tax_amt_post_exe,3500,3600)) %>%
  summarize(
    tax_amt_post_exe = mean(tax_amt_post_exe),
    tax_amt_exe = mean(tax_amt_exe),
    tax_amt_pre_exe = mean(tax_amt_pre_exe),
    pin_count = n(),
    av = mean(av),
    eav = mean(eav)
  )
```

