---
title: "Million Dollar Homes"
author: "Alea Wilbur"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    code_folding: hide
    code_download: yes
---



```{r setup, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)


library(tidyverse)
library(DBI)
library(data.table)
library(ggspatial)
library(gstat)
library(here)
library(httr)
library(jsonlite)
library(ptaxsim)
library(sf)
library(stars)
library(glue)
library(ggpattern)

# Create the DB connection with the default name expected by PTAXSIM functions
ptaxsim_db_conn <- DBI::dbConnect(RSQLite::SQLite(), "./ptaxsim.db/ptaxsim-2021.0.4.db")


options(digits=4, scipen = 999)


library(httr)
library(NatParksPalettes)

# link to the API output as a JSON file
muni_shp <- read_sf("https://gis.cookcountyil.gov/traditional/rest/services/politicalBoundary/MapServer/2/query?outFields=*&where=1%3D1&f=geojson")

cook_shp <- read_sf("https://gis.cookcountyil.gov/traditional/rest/services/plss/MapServer/1/query?outFields=*&where=1%3D1&f=geojson")



#muni_shp <- read_json("muni_shp.json")
nicknames <- readxl::read_excel("./Necessary_Files/muni_shortnames.xlsx")

class_dict <- read_csv("./Necessary_Files/class_dict_expanded.csv") %>% 
  mutate(class_code = as.character(class_code))



# `agency_dt` has all taxing agencies (but not TIFs) that existed each year and includes their total taxable base (cty_cook_eav), their levy, taxing rate, binary variables for if a municipality is home rule or not, as well as many other variables. tax_bill() uses this table for the taxable EAV that is used to  calculate the tax rates in the tax bills. For simulations, you must alter the taxable EAV or levy or other variables and then tell tax_bill() function to use the modified agency data table for simulated tax bills.
# 



# has EAV values, extensions by agency_num
agency_dt <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT *
  FROM agency
  WHERE year = 2021
  "
)


# cook_agency_names <- DBI::dbGetQuery(
#   ptaxsim_db_conn,
#   "SELECT DISTINCT agency_num, agency_name
#   FROM agency_info
#   "
# )
# 
#  
# 
# 
# # has all tax codes and the taxing agency that taxes them. Tax code rates and agency rates. 
# cook_tax_codes <- DBI::dbGetQuery(
#   ptaxsim_db_conn,
#   glue_sql("
#   SELECT*
#   FROM tax_code
#   WHERE agency_num IN ({cook_agency_names$agency_num*})
#   AND year = 2021
#   ",
#   .con = ptaxsim_db_conn
#   )
# )


muni_agency_names <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT DISTINCT agency_num, agency_name, minor_type
  FROM agency_info
  WHERE minor_type = 'MUNI'
  OR agency_num = '020060000'  

  "
)

muni_tax_codes <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql("
  SELECT*
  FROM tax_code
  WHERE agency_num IN ({muni_agency_names$agency_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
)  
  
tax_codes <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql("
  SELECT DISTINCT tax_code_num, tax_code_rate
  FROM tax_code
  WHERE year = 2021  
  ",
  .con = ptaxsim_db_conn
  )
)

## All tax codes. 
## tax codes within municipalities have additional info 
tc_muninames <- tax_codes %>% 
  left_join(muni_tax_codes) %>%
  left_join(muni_agency_names) %>% 
  select(-agency_rate) %>% 
  left_join(nicknames) %>% 
  select(-c(minor_type, short_name, `Column1`, `Most recent reassessed`, agency_number))


# Agency number and agency name for all TIFs
TIF_agencies <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT DISTINCT agency_num, agency_name, major_type, minor_type
  FROM agency_info
  WHERE minor_type = 'TIF'
  "
)

unique_tif_taxcodes <- DBI::dbGetQuery(
  ptaxsim_db_conn, 
  glue_sql("
  SELECT DISTINCT tax_code_num
  FROM tax_code
  WHERE agency_num IN ({TIF_agencies$agency_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
)


tif_distrib <- DBI::dbGetQuery(
  ptaxsim_db_conn, 
  glue_sql("
  SELECT *
  FROM tif_distribution
  WHERE tax_code_num IN ({muni_tax_codes$tax_code_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
) %>% mutate(tax_code_num = as.character(tax_code_num))



cross_county_lines <- c("030440000", "030585000", "030890000", "030320000", "031280000","030080000", "030560000", "031120000", "030280000", "030340000","030150000","030050000", "030180000","030500000","031210000")


cross_county_lines <- muni_agency_names %>% 
  filter(agency_num %in% cross_county_lines) %>% 
  left_join(nicknames, by = "agency_name")
```

# Changing General Homeowner Exemption - Scenarios


## Quartiles and Progressivity

> Download file 4C_joined_PINs_bills_and_exemptions.csv from the Box folder at this [link](https://uofi.box.com/s/cvp7pdjkx8kth7o0vg9itpra3z8vfkzq). Put file in your project folder and run code. Change file path if needed. 

```{r}
class_dict <- read_csv("./Necessary_Files/class_dict_singlefamcodes.csv") %>% 
  mutate(class_code = as.character(class_code)) # change variable type to character so the join works.

nicknames <- readxl::read_xlsx("./Necessary_Files/muni_shortnames.xlsx")

#pin_data2 <- read_csv("./Output/4C_joined_PINs_bills_and_exemptions.csv")

muni_taxrates <- read_csv("./Output/4C_muni_taxrates.csv")

#pin_data2 <- pin_data2 %>% left_join(class_dict)

muni_TC_fullyCook <- muni_tax_codes %>%
  filter(!agency_num %in% cross_county_lines)


joined_pins <- read_csv("./Output/4C_joined_PINs_bills_and_exemptions.csv") %>%
  mutate(tax_code_num = as.character(tax_code_num)) %>%  
  left_join(tc_muninames) %>% left_join(class_dict)

# all pins in munis fully within cook county that are some form of single-family, detached home
singfam_pins <- joined_pins %>% 
  filter(tax_code %in% muni_TC_fullyCook$tax_code_num) %>% # excludes county line crossing munis
  filter(Option2 == "Single-Family")
```

## Scenarios & tax rates

Similar to File 5_Exemption_Scenarios.rmd. 

- Calculate Class 2 Burden --> Calculate the amount of taxable EAV in the Municipality (for each scenario) and multiply it by the new composite tax rate (for each scenario).   


- Burden Share  = Taxable EAV within Property Class * Composite tax rate   

- Composite Tax Rate = (Municipal Levy / Taxable EAV )    


```{r}
## Bring in tax bills and exemption data for 2021 PINs ##
# 
# joined_pins <- read_csv("./Output/4C_joined_PINs_bills_and_exemptions.csv") %>%
#   mutate(tax_code_num = as.character(tax_code_num)) %>%  
#   left_join(tc_muninames) %>% left_join(class_dict)

MuniLevy <- joined_pins %>% 
  group_by(clean_name, agency_num) %>%
  
  summarize(MuniLevy = sum(final_tax_to_dist, na.rm = TRUE), # amount billed by munis with current exemptions in place
            current_nonTIF_EAV_post_exemps = sum(final_tax_to_dist/(tax_code_rate/100), na.rm = TRUE),
            current_TIF_increment_EAV = sum(final_tax_to_tif/(tax_code_rate/100), na.rm=TRUE),  
            current_Exempt_EAV = sum(tax_amt_exe/(tax_code_rate/100), na.rm=TRUE), 
            Total_EAV = sum((tax_amt_exe+final_tax_to_dist+final_tax_to_tif)/(tax_code_rate/100), na.rm = TRUE)) %>% 
  mutate(cur_muni_comp_rate = MuniLevy / current_nonTIF_EAV_post_exemps)

MuniLevy
```

```{r}
joined_pins <- joined_pins %>% 
  mutate(exe_neg10 = 0,
         exe_0 = ifelse(eav < 10000 & exe_homeowner!=0, eav, 
                             ifelse(eav>10000 & exe_homeowner!=0, 10000, 0 )),  #would be if there is no change in exemptions
         exe_plus10 = ifelse(eav < 20000 & exe_homeowner!=0, eav, 
                             ifelse(eav>20000 & exe_homeowner!=0, 20000, 0 )),
         exe_plus20 = ifelse(eav < 30000 & exe_homeowner!=0, eav, 
                             ifelse(eav>30000 & exe_homeowner!=0, 30000, 0 ) ),
         exe_plus30 = ifelse(eav < 40000 & exe_homeowner!=0, eav, 
                             ifelse(eav>40000 & exe_homeowner!=0, 40000, 0) ),
         exe_plus40 = ifelse(eav < 50000 & exe_homeowner!=0, eav, 
                             ifelse(eav>50000 & exe_homeowner!=0, 50000, 0) ),
         mil_home = ifelse(av*10 > 1000000, 1, 0))

scenario_calcs <- joined_pins %>%    
  group_by(clean_name) %>%

    summarize(MuniLevy = sum(final_tax_to_dist, na.rm = TRUE), # amount billed by munis with current exemptions in place
            current_nonTIF_EAV_post_exemps = sum(final_tax_to_dist/(tax_code_rate/100), na.rm = TRUE),
            current_TIF_increment_EAV = sum(final_tax_to_tif/(tax_code_rate/100), na.rm=TRUE),  
            current_Exempt_EAV = sum(tax_amt_exe/(tax_code_rate/100), na.rm=TRUE), 
            current_GHE = sum(exe_homeowner, na.rm=TRUE),
            Total_EAV = sum((tax_amt_exe+final_tax_to_dist+final_tax_to_tif)/(tax_code_rate/100), na.rm = TRUE),
            exe_neg10 = sum(exe_neg10),
            exe_0 = sum(exe_0), # no change, for comparison
            exe_plus10 = sum(exe_plus10),
            exe_plus20 = sum(exe_plus20),
            exe_plus30 = sum(exe_plus30),
            exe_plus40 = sum(exe_plus40),
            mil_home = sum(mil_home)) %>%

  # remove all GHE (up to 10,000 EAV added back to base per PIN), 
  # add exe_homeowner back to taxable base
  mutate(neg10_taxable_eav = Total_EAV - current_TIF_increment_EAV - current_Exempt_EAV + current_GHE,    # adds GHE exempt EAV back to taxable base and decreases tax rates
         plus10_taxable_eav = Total_EAV - current_TIF_increment_EAV - current_Exempt_EAV + current_GHE - exe_plus10, # will increase tax rates
         plus20_taxable_eav = Total_EAV - current_TIF_increment_EAV - current_Exempt_EAV + current_GHE - exe_plus20,
         plus30_taxable_eav = Total_EAV - current_TIF_increment_EAV - current_Exempt_EAV + current_GHE - exe_plus30,
         plus40_taxable_eav = Total_EAV - current_TIF_increment_EAV - current_Exempt_EAV + current_GHE - exe_plus40,
         scenario_noexemptions_taxable_eav = Total_EAV - current_TIF_increment_EAV) %>%
  
  mutate(tr_neg10 = MuniLevy / neg10_taxable_eav,
         tr_nochange = MuniLevy / current_nonTIF_EAV_post_exemps,
         tr_plus10 = MuniLevy / plus10_taxable_eav,
         tr_plus20 = MuniLevy / plus20_taxable_eav,
         tr_plus30 = MuniLevy / plus30_taxable_eav,
         tr_plus40 = MuniLevy / plus40_taxable_eav, 
         tax_rate_current = MuniLevy/current_nonTIF_EAV_post_exemps,
         taxrate_noexemps = MuniLevy /(Total_EAV - current_TIF_increment_EAV  ),
         taxrate_noTIFs = MuniLevy / (Total_EAV - current_Exempt_EAV),
         taxrate_noTIFs_orExemps = MuniLevy / Total_EAV) %>%
  select(clean_name, MuniLevy, tr_neg10:taxrate_noTIFs_orExemps, everything())

scenario_taxrates <- scenario_calcs %>%  select(clean_name, MuniLevy, tr_neg10:taxrate_noTIFs_orExemps) 

scenario_taxrates
```


# Ranked Properties and Muni Quartiles


```{r}
q = c(.10, .25, .5, .75, .9, .99)

## ranks properties that are considered single family homes in order of AV for each Muni
muni_quartiles <- joined_pins %>%
  filter(Option2 == "Single-Family") %>% 
  filter(tax_code %in% muni_TC_fullyCook$tax_code_num) %>%
  group_by(agency_name, clean_name) %>%
  arrange(av) %>%
  summarize(count_pins = n(), 
            min = min(av),
            quant10 = round(quantile(av, probs = q[1])), 
            quant25 = round(quantile(av, probs = q[2])), 
            quant50 = round(quantile(av, probs = q[3])),
            quant75 = round(quantile(av, probs = q[4])),
            quant90 = round(quantile(av, probs = q[5])),
            quant99 = round(quantile(av, probs = q[6])),
            max = max(av)
           ) %>% 
  arrange( desc( quant50))

muni_quartiles


## create rank variable for properties that fall within the quartiles +/- $500 range
munis_ranked <- joined_pins  %>%
  inner_join(muni_quartiles, by = c("agency_name", "clean_name")) %>% 
  mutate(rank = case_when(
    av > (quant10-500) & (av<quant10+500) ~ "q10",
    av > (quant25-500) & (av<quant25+500) ~ "q25",
    av > (quant50-500) & (av<quant50+500) ~ "q50",
    av > (quant75-500) & (av<quant75+500) ~ "q75",
    av > (quant90-500) & (av<quant90+500) ~ "q90",
    av > (quant99-500) & (av<quant99+500) ~ "q99")

    ) %>%
  select(clean_name, rank, av, pin, class, everything()) %>%
  left_join(nicknames)




munis_billchange <-  munis_ranked %>% 
  group_by(clean_name, rank) %>%
  left_join(scenario_taxrates) %>%
  arrange(av) %>%

    mutate(
   # ## Made negative tax bills!! ## #
         
         bill_neg10 = tr_neg10*(equalized_AV-all_exemptions+ exe_homeowner -exe_neg10),
         bill_current = cur_comp_TC_rate/100*(equalized_AV-all_exemptions),
         bill_plus10 =  tr_plus10*(equalized_AV-all_exemptions+ exe_homeowner -exe_plus10),
         bill_plus20 = tr_plus20*(equalized_AV-all_exemptions+ exe_homeowner -exe_plus20),
         bill_plus30 = tr_plus20*(equalized_AV-all_exemptions+ exe_homeowner -exe_plus30),
         bill_plus40 = tr_plus20*(equalized_AV-all_exemptions+ exe_homeowner -exe_plus40),
    
    # ## make bills $1 if they had been negative.    
         bill_neg10 = ifelse(bill_neg10 < 1, 1, bill_neg10),
         bill_current = ifelse(bill_current < 1, 1, bill_current),
         bill_plus10 = ifelse(bill_plus10 < 1, 1, bill_plus10),
         bill_plus20 = ifelse(bill_plus20 < 1, 1, bill_plus20),
         bill_plus30 = ifelse(bill_plus30 < 1, 1, bill_plus30),
         bill_plus40 = ifelse(bill_plus40 < 1, 1, bill_plus40)
         )%>%
  mutate(
         zerodol_bills_ghe0 = ifelse(bill_neg10 < 5, 1, 0),
         zerodol_bills_current = ifelse(bill_current < 5, 1, 0),
         zerodol_bills_ghe20 = ifelse(bill_plus10 < 5, 1, 0),
         zerodol_bills_ghe30 = ifelse(bill_plus20 < 5, 1, 0),
         zerodol_bills_ghe40 = ifelse(bill_plus30 < 5, 1, 0),
         zerodol_bills_ghe50 = ifelse(bill_plus40 < 5, 1, 0)
  ) %>%
  group_by(clean_name, rank, has_HO_exemp) %>% 
  summarize(median_AV = round(median(av)),
            median_EAV = round(median(eav)),
            mean_bill_neg10 = round(mean(bill_neg10, na.rm=TRUE)),
            mean_bill_cur = round(mean(bill_current, na.rm=TRUE)),
            mean_bill_plus10 = round(mean(bill_plus10, na.rm=TRUE)),
            mean_bill_plus20 = round(mean(bill_plus20, na.rm=TRUE)),
            mean_bill_plus30 = round(mean(bill_plus30, na.rm=TRUE)),
            mean_bill_plus40 = round(mean(bill_plus40, na.rm=TRUE)),
            
            # current perceived_savings = median(tax_amt_exe),
            tr_neg10 = round(mean(tr_neg10*100), digits = 2), 
            cur_comp_TC_rate = round(mean(cur_comp_TC_rate), digits = 2),
            tr_plus10 = round(mean(tr_plus10*100), digits = 2),
            tr_plus20 = round(mean(tr_plus20*100), digits = 2),
            tr_plus30 = round(mean(tr_plus30*100), digits = 2),
            tr_plus40 = round(mean(tr_plus40*100), digits = 2),
            pincount=n(),
            zerodol_bills_ghe0 = sum(zerodol_bills_ghe0),           
            zerodol_bills_current = sum(zerodol_bills_current),
            zerodol_bills_ghe20 = sum(zerodol_bills_ghe20),
            zerodol_bills_ghe30 = sum(zerodol_bills_ghe30),
            zerodol_bills_ghe40 = sum(zerodol_bills_ghe40),
            zerodol_bills_ghe50 = sum(zerodol_bills_ghe50),
            mill_homes = sum(mil_home)

  ) %>%
  arrange(has_HO_exemp, rank)


munis_billchange <- munis_billchange %>% left_join(muni_quartiles)

munis_billchange 
```





```{r milliondollar-homes}

munis_ranked %>% 
  mutate(mil_home = ifelse(av*10 > 1000000,1,0) ) %>%
 # filter(has_HO_exemp == 1) %>% 
  group_by(clean_name, has_HO_exemp) %>% 
  summarize(million_dollar_homes = sum(mil_home)) %>% 
  arrange(desc(million_dollar_homes))

munis_ranked %>% 
  mutate(mil_home = ifelse(av*10 > 1000000,1,0) ) %>%
  filter(has_HO_exemp == 1) %>% 
  group_by(clean_name) %>% 
  summarize(million_dollar_homes = sum(mil_home)) %>% 
  arrange(desc(million_dollar_homes))

munis_ranked %>% 
  mutate(mil_home = ifelse(av*10 > 1000000,1,0) ) %>%
  filter(has_HO_exemp == 1) %>% 
  summarize(million_dollar_homes = sum(mil_home))

munis_ranked %>% 
  mutate(mil_home = ifelse(av*10 > 1000000,1,0) ) %>%
  filter(has_HO_exemp == 1 & mil_home == 1) %>% 
  summarize(
    exempt_eav  = sum(exe_0),
    million_dollar_homes = sum(mil_home),
    bonus_rev = sum(cur_comp_TC_rate*exe_0))


```

Of these, 50,775 did not claim the GHE exemption. 17,107 million dollar homes did claim the GHE exemption in 2021.

Amount of money that could be "transferred" to areas impacted by high tax rates = 

Sum of (composite tax rate * exempt_EAV) = $1.3 billion