---
title: "Incentive Properties in Cook County, IL"
author: "AWM"
format: 
  html:
    df-print: paged
    code-fold: true
    code-download: true
    toc: true
    toc-location: left
    fig-cap-location: top
---

Incentive properties include any property within Cook County that has a property class between 600 and 899.

```{r setup, warning = FALSE, message = FALSE}
library(tidyverse)
library(ptaxsim)
library(DBI)
library(httr)
library(jsonlite)
library(glue)
library(sf)

knitr::opts_chunk$set(warning = FALSE, message = FALSE)



#ptaxsim_db_conn <- DBI::dbConnect(RSQLite::SQLite(), "./ptaxsim.db/ptaxsim-2021.0.4.db")

ptaxsim_db_conn <- DBI::dbConnect(RSQLite::SQLite(), "./ptaxsim.db/ptaxsim-2022.0.0.db")

# link to the API output as a JSON file
muni_shp <- read_sf("https://gis.cookcountyil.gov/traditional/rest/services/politicalBoundary/MapServer/2/query?outFields=*&where=1%3D1&f=geojson")

#muni_shp <- read_json("muni_shp.json")
nicknames <- readxl::read_excel("./Necessary_Files/muni_shortnames.xlsx")

class_dict <- read_csv("./Necessary_Files/class_dict_expanded.csv") %>% 
  mutate(class_code = as.character(class_code))

NBHs <- read_sf("https://datacatalog.cookcountyil.gov/resource/pcdw-pxtg.geojson")

chi_townships = c("WEST", "SOUTH", "HYDE PARK", "LAKE", "NORTH", "JEFFERSON", "LAKE VIEW", "ROGER PARK")

chi_townships2 <- c("Jefferson", "Hyde Park", "Lake", "Lake View", "North Chicago", "South Chicago", "West Chicago" )

class8township <- c("Bloom", "Bremen","Calumet","Rich", "Thornton")
```

# Incentive Properties In Cook County, IL

```{r}

source("helper_tc_muninames.R")

incentive_pins <- read_csv("./Output/7_output_incentive_classes.csv")

residential_pins <- read_csv("./Output/7_output_residential_pins.csv")

incentive_pins <- incentive_pins %>%  left_join(nicknames, by = "agency_name")


incentive_pins %>% 
 # left_join(nicknames, by = "agency_name") %>% 
  reframe(IncentiveProp_PINcount = n(),
          IncentiveProp_MuniAV = sum(av),
          .by=c(clean_name)) %>% 
  rename(
         Municipality = clean_name) %>%
    mutate(Municipality = ifelse(is.na(Municipality), "Unincorporated", Municipality)) %>%
  arrange(Municipality)

incentive_pins %>% 
#  left_join(nicknames, by = "agency_name") %>%
  filter(class > 599 & class < 900) %>% 
  group_by(clean_name, major_class_code) %>%
  summarize(pincount = n()) %>%
  pivot_wider(id_cols = clean_name, 
              names_from = "major_class_code", 
              values_from = "pincount") %>%
  select( Municipality = clean_name,
          "6A", "6B", "6C", "7A", "7B", "8") %>% 
  mutate(Municipality = ifelse(is.na(Municipality), "Unincorporated", Municipality)) %>%
  arrange(Municipality)
```

```{r }
#| column: page

incentive_pins %>% 
  #  left_join(nicknames,by = "agency_name") %>%

  filter(class > 599 & class < 900) %>% 
  group_by(clean_name, township_name, major_class_code) %>%
  summarize(pincount = n()) %>%
  pivot_wider(id_cols = c(clean_name, township_name), 
              names_from = "major_class_code", 
              values_from = "pincount")  %>% select( Municipality = clean_name, Township = township_name,
          "6A", "6B", "6C", "7A", "7B", "8") %>% 
  mutate(Municipality = ifelse(is.na(Municipality), "Unincorporated", Municipality)) %>%
  arrange(Municipality)

```

```{r nbhd-allclasses}
nbh_pincounts <- incentive_pins %>% 
  group_by(nbhd_code, class) %>%
  dplyr::summarize(pin_count = n(),
            av = sum(av)) %>%
  mutate(class_1dig = str_sub(class, 1, 1))

nbh_pincounts_majorclass <- incentive_pins %>% 
  group_by(major_class_code, major_class_type, nbhd_code) %>%
  dplyr::summarize(pin_count = n(),
            av = sum(av))

# table(nbh_pincounts_majorclass$major_class_code)


incentives_props <- nbh_pincounts_majorclass %>% 
  filter(major_class_code %in% c("6A", "6B", "6C", "7A", "7B", "8", "9")) %>%
  mutate(nbhd_code = as.character(nbhd_code))


fig_nbhd_pincount <- incentives_props %>% 
  group_by(nbhd_code) %>%
  summarize(pin_count = sum(pin_count)) %>%
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  ggplot(aes(fill = pin_count)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Number of PINs with Incentives in Assessor Neighborhoods", 
  caption = "Count of 14 digit PINs") +
  theme_void() + 
   theme(axis.ticks = element_blank(), axis.text = element_blank()) +
scale_fill_steps2(
  high = "darkblue", low = "orange",
  show.limits=TRUE,
  nice.breaks = TRUE,
na.value = NA,
  name = "Pin Count")


fig_nbhd_AV <- incentives_props %>% 
  group_by(nbhd_code) %>%
  summarize(av = sum(av)) %>%
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  filter(triad_name != "City") %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Assessed Value of Properties in 
       Assessor Neighborhoods", 
       caption = "Different classes of properties have different assessment levels. 
  Some change over time.") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    show.limits=TRUE,
    nice.breaks = FALSE,
   # limits = c(0, 90000000),
    na.value = NA,
    name = "AV",
    label = scales::dollar)



fig_nbhd_Chicago_AV <- incentives_props %>% 
  group_by(nbhd_code) %>%
  summarize(av = sum(av)) %>%
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  filter(triad_name == "City") %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Assessed Value of Properties in 
       Assessor Neighborhoods", 
       caption = "Different classes of properties have different assessment levels. 
  Some change over time.") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    show.limits=TRUE,
    nice.breaks = FALSE,
   # limits = c(0, 90000000),
    na.value = NA,
    name = "AV",
    label = scales::dollar)

```

```{r nhbd-class6and7}
nbh_pincounts_class <- incentive_pins %>% 
  group_by(nbhd_code, class) %>%
  dplyr::summarize(pin_count = n(),
            av = sum(av)) %>%
  mutate(class_1dig = str_sub(class, 1, 1))

nbh_pincounts_majorclass <- incentive_pins %>% 
  group_by(major_class_code, major_class_type, nbhd_code) %>%
  dplyr::summarize(pin_count = n(),
            av = sum(av))

# table(nbh_pincounts_majorclass$major_class_code)


incentives_props <- nbh_pincounts_majorclass %>% 
  filter(major_class_code %in% c("6A", "6B", "6C", "7A", "7B", "8")) %>%
  mutate(nbhd_code = as.character(nbhd_code))

fig_nbhd_C7AV <- incentives_props %>% 
  filter(major_class_code %in% c("7A", "7B", "7C") ) %>% 
  group_by(nbhd_code) %>%
  summarize(av = sum(av)) %>%
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  filter(triad_name != "City") %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Class 7 Properties - Assessed Value",
       subtitle = "AV of Buildings with Incentives in Assessor Neighborhoods",
       caption = "Over $90 million in Chicago") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "AV ($)",
    label = scales::dollar)

fig_nbhd_C7_Chicago_AV <- incentives_props %>% 
  filter(major_class_code %in% c("7A", "7B", "7C") ) %>% 
  group_by(nbhd_code) %>%
  summarize(av = sum(av)) %>%
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  filter(triad_name == "City") %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Class 7 Properties - Assessed Value",
       subtitle = "AV of Buildings with Incentives in Assessor Neighborhoods",
       caption = "Over $90 million in Chicago") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "AV ($)",
    label = scales::dollar)



fig_nbhd_C7_PC <- incentives_props %>% 
  filter(major_class_code %in% c("7A", "7B", "7C") ) %>%
  group_by(nbhd_code) %>%
  summarize(pincount = sum(pin_count)) %>% 
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
    filter(triad_name != "City") %>%

  ggplot(aes(fill = pincount)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Class 7 Properties",
       subtitle = "# of Buildings with Class 7 Incentives in Assessor Neighborhoods") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    show.limits=TRUE,
    na.value = NA,
    name = "Pin Count")

fig_nbhd_C7_Chicago_PC <- incentives_props %>% 
  filter(major_class_code %in% c("7A", "7B", "7C") ) %>%
  group_by(nbhd_code) %>%
  summarize(pincount = sum(pin_count)) %>% 
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
    filter(triad_name == "City") %>%

  ggplot(aes(fill = pincount)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Class 7 Properties in Chicago",
       subtitle = "# of Buildings with Class 7 Incentives in Assessor Neighborhoods") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    show.limits=TRUE,
    na.value = NA,
    name = "Pin Count")

fig_nbhd_C6_AV <- incentives_props %>% 
  filter(major_class_code %in% c("6A", "6B", "6C") ) %>%
  group_by(nbhd_code) %>%
  summarize(av = sum(av)) %>%
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Class 6 Properties",
       subtitle = "AV of Buildings with Incentives in Assessor Neighborhoods") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "AV ($)",
    label = scales::dollar)

fig_nbhd_C6_Chicago_AV <- incentives_props %>% 
  filter(major_class_code %in% c("6A", "6B", "6C") ) %>%
  group_by(nbhd_code) %>%
  summarize(av = sum(av)) %>%
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  filter(triad_name == "City") %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Class 6 Properties",
       subtitle = "AV of Buildings with Incentives in Assessor Neighborhoods") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "AV ($)",
    label = scales::dollar)
  
fig_nbhd_C6_PC <- incentives_props %>% 
  filter(major_class_code %in% c("6A", "6B", "6C") ) %>%
  group_by(nbhd_code) %>%
  summarize(pin_count= round(sum(pin_count)), digits = 0) %>%
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  ggplot(aes(fill = pin_count)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Class 6 Properties",
       subtitle = "# of Buildings with Incentives in Assessor Neighborhoods") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    #nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "Pin Count")

fig_nbhd_C6_Chicago_PC <- incentives_props %>% 
  filter(major_class_code %in% c("6A", "6B", "6C") ) %>%
  group_by(nbhd_code) %>%
  summarize(pin_count= round(sum(pin_count)), digits = 0) %>%
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  filter(triad_name == "City") %>%
  ggplot(aes(fill = pin_count)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Class 6 Properties",
       subtitle = "# of Buildings with Incentives in Assessor Neighborhoods") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    #nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "Pin Count")
```

## All Incentive Types Together

```{r muni-alltogether}
#| column: page
#| layout-ncol: 2
#| fig-cap: "All Incentive Property Classes: Municipality level and Assessor Neighborhood level"



# Add in names that merge with Muni shapefile names
muni_pincounts_majorclass <- incentive_pins %>% 
  left_join(nicknames) %>%
  group_by(clean_name, major_class_code, major_class_type, agency_name, shpfile_name) %>%
  dplyr::summarize(pin_count = n(),
            av = sum(av))

# table(muni_pincounts_majorclass$major_class_code) # Number of Munis with these major classes


fig_muni_AV <- muni_pincounts_majorclass %>% 
  filter(major_class_code %in% c("6A", "6B", "6C", "7A", "7B", "8")) %>%
  filter(clean_name != "Chicago") %>% 
  group_by(shpfile_name)%>%
  summarize(av = sum(av)) %>% 
  filter(!is.na(shpfile_name) ) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Properties with Any Type of Incentive Class",
       subtitle = "AV of Buildings with Incentives in Municipalities (Except Chicago).
       Chicago has around $500 million in AV from properties with incentives.") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
  #  nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
  limits = c(0, 90000000),
    name = "AV ($)",
    label = scales::dollar)

fig_muni_AV
fig_nbhd_AV
```

```{r muni-pincount}
#| column: page
#| layout-ncol: 2
fig_muni_pincount <- muni_pincounts_majorclass %>% 
  filter(major_class_code %in% c("6A", "6B", "6C", "7A", "7B", "8")) %>%
  filter(clean_name != "Chicago") %>% 
  group_by(shpfile_name)%>%
  summarize(pin_count = sum(pin_count)) %>% 
  filter(!is.na(shpfile_name)) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = pin_count)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Properties with Incentive Clases",
       subtitle = "# of Buildings with Incentives in Municipalities (Except Chicago)") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
  #  nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "# PINs")


fig_muni_pincount
fig_nbhd_pincount
```

## Class 6 Incentives

Class 6A: Industrial Incentive

Class 6B: Industrial Incentive - Designed to encourage industrial development throughout Cook County by offering a real estate tax incentive for the development of new industrial facilities, the rehabilitation of existing industrial structures, and the industrial reutilization of abandoned buildings.

Class 6C: Industrial Brownfield

```{r }
#| column: page
#| layout-ncol: 2
fig_muni_C6_AV <- muni_pincounts_majorclass %>% 
  filter(major_class_code %in% c("6A", "6B", "6C")) %>%
  filter(clean_name != "Chicago") %>% 
  group_by(shpfile_name)%>%
  summarize(av = sum(av)) %>% 
  filter(!is.na(shpfile_name)) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Properties with Class 6 Incentives",
       subtitle = "Assessed Value in Municipalities (Except Chicago)") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    nice.breaks = TRUE,
    show.limits=TRUE,
    na.value = NA,
    name = "AV ($)",
  label = scales::dollar)



fig_muni_C6_PC <- muni_pincounts_majorclass %>% 
  filter(major_class_code %in% c("6A", "6B", "6C")) %>%
 # filter(clean_name != "Chicago") %>% 
  group_by(shpfile_name)%>%
  summarize(pin_count = sum(pin_count)) %>% 
  filter(!is.na(shpfile_name)) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = pin_count)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Properties with Class 6 Incentives",
       subtitle = "# of Buildings with Incentives in Municipalities") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
  #  nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "# PINs")

fig_muni_C6_AV
fig_nbhd_C6_AV

fig_muni_C6_PC
fig_nbhd_C6_PC



```

## Class 7 Incentives

Class 7A(B): Intended to encourage commercial projects in areas determined to be "in need of commercial development.These projects have total development costs, exclusive of land, that DO NOT EXCEED(DO EXCEED) two million and would not be economically feasible without the incentive.

```{r }
#| column: page
#| layout-ncol: 2

fig_muni_C7_AV <- muni_pincounts_majorclass %>% 
  filter(major_class_code %in% c("7A", "7B", "7C")) %>%
  filter(clean_name != "Chicago") %>% 
  group_by(shpfile_name)%>%
  summarize(av = sum(av)) %>% 
  filter(!is.na(shpfile_name)) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Assessed Value of Properties with Class 7 Incentives",
       subtitle = "in Municipalities (Except Chicago)") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "AV ($)",
  label = scales::dollar)


fig_muni_C7_PC <- muni_pincounts_majorclass %>% 
  filter(major_class_code %in% c("7A", "7B", "7C")) %>%
  filter(clean_name != "Chicago") %>% 
  group_by(shpfile_name)%>%
  summarize(pin_count = sum(pin_count)) %>% 
  filter(!is.na(shpfile_name)) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = pin_count)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Properties with Class 7 Incentives",
       subtitle = "# of Buildings with Incentives in Municipalities (Except Chicago)") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
  #  nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "# PINs")

fig_muni_C7_AV
fig_nbhd_C7AV



fig_muni_C7_PC
fig_nbhd_C7_PC
```

## Class 8 Incentives

Class 8: Designed to encourage industrial and commercial development in areas of the county which are experiencing severe economic stagnation.

Class 8 properties exist in places besides the 5 townships:

Bloom, Bremen, Calumet, Rich, Thornton

```{r class8}
#| column: page
#| layout-ncol: 2

fig_nbhd_C8_PC <- incentives_props %>% 
  filter(major_class_code == "8") %>%
  group_by(nbhd_code)%>%
  summarize(pin_count = sum(pin_count)) %>% 
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  ggplot(aes(fill = pin_count)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Class 8 Properties",
       subtitle = "Number of PINs with Incentives in Assessor Neighborhoods", 
       caption = "Count of 14 digit PINs") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    #nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "Count")


fig_nbhd_C8 <- incentives_props %>% 
  filter(major_class_code == "8") %>%
  group_by(nbhd_code) %>%
  summarize(av = sum(av)) %>% 
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry),
         # color = "black",
          lwd = .1)  +
  labs(title = "Class 8 Properties",
       subtitle = "AV of Buildings with Incentives in Assessor Neighborhoods") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    nice.breaks = FALSE,
    show.limits=TRUE,
    limits = c(0, 13000000),
    na.value = NA,
    name = "AV ($)",
    label = scales::dollar)

fig_muni_C8_AV <- muni_pincounts_majorclass %>% 
  filter(major_class_code %in% c("8")) %>%
  filter(clean_name != "Chicago") %>% 
  group_by(shpfile_name)%>%
  summarize(av = sum(av)) %>% 
  filter(!is.na(shpfile_name)) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Properties with Class 8 Incentives",
       subtitle = "# of Buildings with Incentives in Municipalities (Except Chicago)") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "AV ($)",
  label = scales::dollar)



fig_muni_C8_PC <- muni_pincounts_majorclass %>% 
  filter(major_class_code %in% c("8")) %>%
 # filter(clean_name != "Chicago") %>% 
  group_by(shpfile_name)%>%
  summarize(pin_count = sum(pin_count)) %>% 
  filter(!is.na(shpfile_name)) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = pin_count)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Properties with Class 8 Incentives",
       subtitle = "# of Buildings with Incentives in Municipalities") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
  #  nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "# PINs")

fig_muni_C8_PC
fig_nbhd_C8_PC



```

> There are some Class 8 properties located in townships other than the Class 8 five chosen for the incentives.

```{r class8-2}
#| column: page
#| layout-ncol: 2
fig_muni_C8 <- muni_pincounts_majorclass %>% 
   filter(!is.na(shpfile_name)) %>% 

  filter(major_class_code %in% c("8")) %>%
  group_by(shpfile_name)%>%
  summarize(av = sum(av)) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry),color = "black", lwd = .05)+ 
  labs(title = "Class 8 Properties",
       subtitle = "AV of Buildings with Incentives in Assessor Neighborhoods") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    limits = c(0, 13000000),
    name = "AV ($)",
    label = scales::dollar)

fig_muni_C8
fig_nbhd_C8

```

# Chicago Specific Graphs

```{r}
#| column: page
#| layout-ncol: 2

# fig_nbhd_Chicago_AV All incentives

fig_nbhd_C6_Chicago_AV
fig_nbhd_C6_Chicago_PC
fig_nbhd_C7_Chicago_AV
fig_nbhd_C7_Chicago_PC
```

# Tables

```{r}
#| column: page

# By Muni and Major Property Class
incentive_pins %>% 
    mutate(tax_code_num = as.character(tax_code_num)) %>% 
  left_join(muni_tax_codes) %>%
    mutate(tax_code_num = as.character(tax_code_num)) %>% 
  left_join(muni_agency_names) %>%
  group_by(clean_name, major_class_code) %>%
  dplyr::summarize(pin_count = n(),
            av = sum(av))
```

# Areas with lots of exemptions and incentives

aka a reduced tax base

## Municpality Level

```{r}
pin_data <- read_csv("./Output/4C_joined_PINs_bills_and_exemptions.csv")
```

```{r}
muni_taxrates <- pin_data %>% 
  left_join(nicknames, by = "agency_name") %>%
  group_by(clean_name) %>%
  summarize(
    muni_av = sum(av, na.rm = TRUE),
    muni_eav = sum(eav, na.rm = TRUE),
    muni_equalized_AV = sum(equalized_AV, na.rm = TRUE),
    pins_in_muni = n(),
    muni_current_exemptions = sum(all_exemptions, na.rm = TRUE),
    muni_HO_exemps = sum(exe_homeowner, na.rm = TRUE),
    muni_comp_rate = mean(tax_code_rate, na.rm = TRUE), # Changed from first() to mean() on Nov 1
    final_tax_to_dist = sum(final_tax_to_dist, na.rm = TRUE), # used as LEVY amount!! 
    final_tax_to_tif = sum(final_tax_to_tif, na.rm = TRUE),
    tax_amt_exe = sum(tax_amt_exe, na.rm = TRUE), 
    tax_amt_pre_exe = sum(tax_amt_pre_exe, na.rm = TRUE), 
    tax_amt_post_exe = sum(tax_amt_post_exe, na.rm = TRUE),
    rpm_tif_to_cps = sum(rpm_tif_to_cps, na.rm = TRUE), # not used
    rpm_tif_to_rpm = sum(rpm_tif_to_rpm, na.rm=TRUE), # not used
    rpm_tif_to_dist = sum(rpm_tif_to_dist, na.rm=TRUE), # not used
    tif_share = mean(tif_share, na.rm=TRUE), # not used
  ) %>%
  
  rename(cur_comp_TC_rate = muni_comp_rate) %>%
  mutate(current_muni_taxable_eav = final_tax_to_dist/(cur_comp_TC_rate/100),
         new_muni_taxable_eav = final_tax_to_dist/(cur_comp_TC_rate/100) + muni_HO_exemps) %>%
  mutate(new_comp_TC_rate = (final_tax_to_dist / new_muni_taxable_eav)*100) %>%
  mutate(new_comp_TC_rate = ifelse(is.nan(new_comp_TC_rate), cur_comp_TC_rate, new_comp_TC_rate)) %>%
  select(clean_name, cur_comp_TC_rate, new_comp_TC_rate, current_muni_taxable_eav, new_muni_taxable_eav, everything())

muni_taxrates 
```

```{r}
#| column: page

prop_class_sums <- pin_data %>% 
    left_join(nicknames, by = "agency_name") %>%

  group_by(clean_name, major_class_code, major_class_type )  %>%
  
  summarize(
    av = sum(av, na.rm = TRUE),
    eav = sum(eav, na.rm = TRUE),
    equalized_AV = sum(equalized_AV, na.rm = TRUE),
    pins_in_class = n(),
    current_exemptions = sum(all_exemptions, na.rm = TRUE),
    HO_exemps = sum(exe_homeowner, na.rm = TRUE),
    tax_code_rate = mean(tax_code_rate, na.rm = TRUE), # Changed from first() to mean() on Nov 1
    final_tax_to_dist = sum(final_tax_to_dist, na.rm = TRUE), # used as LEVY amount!! 
    final_tax_to_tif = sum(final_tax_to_tif, na.rm = TRUE),
    tax_amt_exe = sum(tax_amt_exe, na.rm = TRUE), 
    tax_amt_pre_exe = sum(tax_amt_pre_exe, na.rm = TRUE), 
    tax_amt_post_exe = sum(tax_amt_post_exe, na.rm = TRUE),
    rpm_tif_to_cps = sum(rpm_tif_to_cps, na.rm = TRUE), # not used
    rpm_tif_to_rpm = sum(rpm_tif_to_rpm, na.rm=TRUE), # not used
    rpm_tif_to_dist = sum(rpm_tif_to_dist, na.rm=TRUE), # not used
    tif_share = mean(tif_share, na.rm=TRUE), # not used
  ) %>%
  
  mutate(total_bill_current = final_tax_to_dist + final_tax_to_tif) %>%
  rename(cur_comp_TC_rate = tax_code_rate) %>%
  mutate(current_taxable_eav = final_tax_to_dist/(cur_comp_TC_rate/100),
         new_taxable_eav = final_tax_to_dist/(cur_comp_TC_rate/100) + HO_exemps) %>%
  mutate(new_comp_TC_rate = (final_tax_to_dist / new_taxable_eav)*100) %>%
  mutate(new_comp_TC_rate = ifelse(is.nan(new_comp_TC_rate), cur_comp_TC_rate, new_comp_TC_rate)) %>%
  select(clean_name, major_class_code, HO_exemps, current_exemptions, pins_in_class, current_taxable_eav, new_taxable_eav,  everything())

# muni level by major_class summary
prop_class_sums 

prop_class_sums2 <- prop_class_sums %>%
  group_by(clean_name) %>%
  mutate(muni_PC = sum(pins_in_class, na.rm=TRUE),
         muni_taxable_eav = sum(current_taxable_eav, na.rm = TRUE),
         muni_equalized_av = sum(equalized_AV, na.rm=TRUE),
         muni_av = sum(av, na.rm=TRUE),
         pct_pins = pins_in_class / muni_PC,
         pct_taxable_eav = current_taxable_eav / muni_taxable_eav,
         pct_eq_eav = equalized_AV / muni_equalized_av,
         pct_av = av / muni_av,
         year = "2021"
  ) %>% 
  mutate_at(vars(pct_pins, pct_taxable_eav, pct_eq_eav, pct_av), funs(round(.,3)))

prop_class_sums2  

# write_csv(prop_class_sums2, "./Output/7_MC_muni_summaries_2021.csv") 

# write_csv(prop_class_sums, "./Output/7_TC_MC_summaries.csv")
```

```{r munilevel-exemptions, eval=FALSE, include=FALSE}
# prop_class_sums is at muni level
res_exempts_per_pin <- prop_class_sums %>% 
  filter(major_class_code == "2") %>% # Class 2 residential only
  mutate(avg_GHEexe_perPIN = HO_exemps/pins_in_class,
         avg_allexe_perPIN = current_exemptions / pins_in_class) %>% 
  select(agency_name, avg_GHEexe_perPIN, avg_allexe_perPIN)
exempts_per_pin


# Class2_data is also at Muni Level
Class2_data <- read_csv("./Output/5b_Class2_burdenshift.csv")

residential_data <- Class2_data %>% 
  mutate(exemptions_per_resPIN = C2_current_GHE/ C2_PC_permuni )  %>%
  select(clean_name, exemptions_per_resPIN, C2_EAV_pct)



fig_muni_AV <- Class2_data %>% 
  #filter(major_class_code %in% c("6A", "6B", "6C", "7A", "7B", "8")) %>%
 # filter(clean_name != "Chicago") %>% 
  group_by(shpfile_name)%>%
  summarize(av = sum(av)) %>% 
  filter(!is.na(shpfile_name) ) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Properties with Any Type of Incentive Class",
       subtitle = "AV of Buildings with Incentives in Municipalities (Except Chicago).
       Chicago has around $500 million in AV from properties with incentives.") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
  #  nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
  limits = c(0, 90000000),
    name = "AV ($)",
    label = scales::dollar)

```

## Neighborhood level join

```{r}

nbh_major_class <- read_csv("./Output/8_Current_Taxrates_per_nbh_by_majorclass.csv")
nbh_major_class


# neighborhoods with incentive properties
incentives_props <- nbh_major_class %>% 
  filter(major_class_code %in% c("6A", "6B", "6C", "7A", "7B", "8")) %>%
  mutate(nbhd_code = as.character(nbhd_code))
# 
# nbh_pincounts_class <- incentive_classes %>% 
#   group_by(nbhd_code, class) %>%
#   dplyr::summarize(pin_count = n(),
#             av = sum(av)) %>%
#   mutate(class_1dig = str_sub(class, 1, 1))
# 
# nbh_pincounts_majorclass <- incentive_classes %>% 
#   group_by(major_class_code, major_class_type, nbhd_code) %>%
#   dplyr::summarize(pin_count = n(),
#             av = sum(av))


```

```{r eval=FALSE}



# nbh_res_exemptions <- read_csv("./Output/8_nbh_sum_new_exe.csv") 
nbh_taxrates <- read_csv( "./Output/8_Current_Taxrates_per_nbh.csv")


nbh_data <- left_join(nbh_pincounts_majorclass, nbh_taxrates, by = "nbhd_code")


nbh_data <- nbh_data %>% 
  rename(all_PINs = pin_count.x,
         nbh_pins = pin_count.y)  %>%
  mutate(exempt_over_total_EAV = Exempt_EAV / Total_EAV,
         nbhd_code = as.character(nbhd_code),
         exempts_perPIN = Exempt_EAV / nbh_PINs,
         incentive_projects_pct = incentive_PINs / nbh_PINs,
         incentives_EAV = av * 3.0027 ,
         incentives_over_total_EAV = incentives_EAV / Total_EAV) %>%
  select(nbhd_code, exempt_over_total_EAV, nbh_PINs, major_class_PINs, exempts_perPIN, incentive_projects_pct, incentives_EAV, incentives_over_total_EAV, major_class_code, everything()) %>% arrange(nbhd_code)

nbh_data

```

```{r}
# nbh_res_exemptions <- read_csv("./Output/8_nbh_sum_new_exe.csv") 
nbh_taxrates <- read_csv( "./Output/8_Current_Taxrates_per_nbh.csv") %>% select(nbhd_code, pin_count, Total_EAV)


nbh_data <- left_join(nbh_major_class, nbh_taxrates, by = "nbhd_code")


nbh_data <- nbh_data %>% 
  rename(major_class_PINs = pin_count.x,
         nbh_PINs = pin_count.y)  %>%
  mutate(exempt_over_total_EAV = round(Exempt_EAV / Total_EAV.y, digits = 4),
         nbhd_code = as.character(nbhd_code),
         exempts_perPIN = round(Exempt_EAV / nbh_PINs, digits = 0),
         class_PIN_pct = major_class_PINs / nbh_PINs,
         #class_EAV = round(av * 3.0027, digits = 0),
         class_over_total_EAV = Total_EAV.x / Total_EAV.y) %>%
  select(nbhd_code, major_class_code, exempt_over_total_EAV, nbh_PINs, major_class_PINs, exempts_perPIN, class_PIN_pct, Total_EAV.x, class_over_total_EAV, everything()) %>% 
  arrange(nbhd_code)

nbh_data
```


```{r}
library(scales)
nbh_sum <- read_csv("./Output/nbh_summary.csv") %>%
    mutate(nbhd_code = as.character(nbhd_code))

nbh_MC_sum <- read_csv("./Output/nbh_MC_summary.csv") %>%
    mutate(nbhd_code = as.character(nbhd_code))

nbh_sum %>% 
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>% 
  ggplot(aes(fill = round(nbh_res_exe/nbh_PC))) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Average Exemptions/PIN in Neighborhoods", 
  caption = "Excludes Abatements") +
  theme_void() + 
   theme(axis.ticks = element_blank(), axis.text = element_blank()) +
scale_fill_steps2(
  high = "darkblue", low = "orange",
  midpoint = round(median(nbh_sum$nbh_res_exe/nbh_sum$nbh_PC, na.rm = TRUE)),
 nice.breaks = FALSE,
  show.limits=TRUE,
na.value = "gray",
  name = "Exempt EAV per C2 PIN",
  labels = scales::label_number(big.mark = ",")
) 
```

## Maps

```{r}
nbh_res <- nbh_data %>% 
  filter(major_class_code == 2) %>% 
  select(nbhd_code, major_class_code, 
         residential_PC = major_class_PINs, 
         res_EAV = Total_EAV.x, 
         res_exemptions = Exempt_EAV) %>%
  mutate(nbhd_code = as.character(nbhd_code),
         res_exe_per_res_pin = round(res_exemptions / residential_PC)
  )
  
  
#nbh_data2 <- nbh_data %>% left_join(nbh_res, by = "nbhd_code") %>% select(nbhd_code, all_PINs, residential_PC, incentive_PINs, everything()) %>% mutate(exemps_per_resPIN = res_exemptions / residential_PC) %>% arrange(nbhd_code)


fig4 <- nbh_res %>% 
  filter(major_class_code  == 2) %>%
  filter(residential_PC > 4) %>% 
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  ggplot(aes(fill = res_exe_per_res_pin)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Average exemptions per Residential Class 2 PIN") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "white",
    #nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "Exemptions per PIN")

fig4




fig2 <- nbh_data %>% 
  filter(major_class_code == "2") %>%
  #group_by(nbhd_code)%>%
  #summarize(pin_count = sum(pin_count)) %>% 
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  ggplot(aes(fill = exempts_perPIN)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Average exemption/PIN (divided by total PIN count of all property classes)") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "white",
    #nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "Exemptions per PIN")

fig2
```

```{r}
fig1 <- nbh_data %>% 
  filter(major_class_code == 2) %>%
  #filter(major_class_code %in% c("6A", "6B", "6C", "7A", "7B", "8")) %>%
  #group_by(nbhd_code)%>%
  #summarize(pin_count = sum(pin_count)) %>% 
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  ggplot(aes(fill = exempt_over_total_EAV)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Exempt EAV (from Residential Exemptons) / Total EAV",
      # caption = "Total EAV includes TIF and all EAV before exemptions are deducted"
       ) +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "forestgreen", low = "white",
    #nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "% Exempt")

fig1



fig3 <- nbh_data %>% 
  filter(major_class_code %in% c("6A", "6B", "6C", "7A", "7B", "8")) %>%
  group_by(nbhd_code)%>%
  summarize(incentives_over_total_EAV = sum(Total_EAV.x/Total_EAV.y) ) %>%
  #summarize(pin_count = sum(pin_count)) %>% 
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  ggplot(aes(fill = incentives_over_total_EAV)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "EAV of incentive properties / Total EAV") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "orange", low = "yellow",
    #nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "%")

fig3

```

# Incentive Properties and Appeals

- av_mailed - Initial assessed values when mailed to property owners, after Desk Review  
- av_certified - Assessed values after Assessor’s Office appeals  
- av_board - Assessed values after Board of Review appeals  
- av_clerk - Final assessed values used by the Clerk and Treasurer to calculate bills. Identical to av_board in 99.9% of cases   

```{r}
incentive_pins <- read_csv("./Output/7_output_incentive_classes.csv")

incentive_taxbills <- tax_bill(2008:2022, incentive_pins$pin, simplify = FALSE)
# pin data for all years if in SFHA


incent_pins_ptax <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql(
  "SELECT*
  FROM pin
  WHERE pin IN ({incentive_pins$pin*})
  ",
  .con = ptaxsim_db_conn
)) 
```

```{r}
#| label: fig-av-stages
#| fig-cap: "AV Stage Comparison"


incent_pins_ptax <- incent_pins_ptax %>% 
    mutate(major_class = str_sub(class, 1,1),
         major_class_fct = case_when(
           major_class == '0' ~ "0 & 1 - Vacant and Exempt",
           major_class == '1' ~ "0 & 1 - Vacant and Exempt",
           major_class == '2' ~ "2 - Residential",
           major_class == '3' ~ "3 & 5 - Commercial",
           major_class == '5' ~ "3 & 5 - Commercial",
           major_class == '6' ~ "6, 7, 8 - Incentive",
           major_class == '7' ~ "6, 7, 8 - Incentive",
           major_class == '8' ~ "6, 7, 8 - Incentive",
           TRUE ~ "Other"
         ))
  

incent_pins_ptax %>%  
    filter(tax_code_num < 70000) %>%

  group_by(year) %>%
  summarize(pins_per_year = n(),
            av_mailed = sum(av_mailed),
            av_certified = sum(av_certified),
            av_board = sum(av_board),
            av_clerk = sum(av_clerk),
          ) %>%
  ggplot() + 
  geom_line(aes(x = year, y = av_mailed)) +

    geom_line(aes(x = year, y = av_certified
               ), color = "green")+
    geom_line(aes(x = year, y = av_board
               ), color = "purple")+
  theme_minimal() + 
  scale_y_continuous(labels = scales::dollar) +
  labs(title = "Incentive Properties & Appeals: Excludes Chicago", subtitle = "Mailed AV and Clerk AV", caption = "Black is mailed, green is certified, purple is Board & Clerk AV",
       y = "Assessed Value")

incent_pins_ptax %>%  
  group_by(year) %>%
  summarize(pins_per_year = n(),
            av_mailed = sum(av_mailed),
            av_certified = sum(av_certified),
            av_board = sum(av_board),
            av_clerk = sum(av_clerk),
          ) %>%
  ggplot() + 
  geom_line(aes(x = year, y = av_mailed)) +
 # geom_line(aes(x = year, y = av_clerk
             #  ), color = "blue")+
    geom_line(aes(x = year, y = av_certified
               ), color = "green")+
    geom_line(aes(x = year, y = av_board
               ), color = "purple")+
  theme_minimal() + 
  scale_y_continuous(labels = scales::dollar) +
  labs(title = "Incentive Properties & Appeals", subtitle = "Mailed AV and Clerk AV", caption = "Black is mailed, green is certified, purple is Board & Clerk AV",
       y = "Assessed Value")



incent_pins_ptax %>%  
  filter(tax_code_num >= 70000) %>%
  group_by(year) %>%
  summarize(pins_per_year = n(),
            av_mailed = sum(av_mailed),
            av_certified = sum(av_certified),
            av_board = sum(av_board),
            av_clerk = sum(av_clerk),
          ) %>%
  ggplot() + 
  geom_line(aes(x = year, y = av_mailed)) +
    geom_line(aes(x = year, y = av_certified
               ), color = "green")+
    geom_line(aes(x = year, y = av_board
               ), color = "purple")+
  theme_minimal() + 
  scale_y_continuous(labels = scales::dollar) +
  labs(title = "Chicago only: Incentive Properties & Appeals", subtitle = "Mailed AV and Clerk AV", caption = "Black is mailed, green is certified, purple is Board & Clerk AV",
       y = "Assessed Value")
```

```{r}
incent_pins_ptax %>% 
  filter(av_mailed > av_clerk & 
                              year == 2021) %>%
  mutate(av_dif =  av_mailed - av_clerk) %>%
  arrange(-av_dif) %>%
  select(class, tax_code_num, av_dif, pin, everything())

incent_pins_ptax %>% 
  filter(av_mailed > av_clerk & 
                              year == 2022) %>%
  mutate(av_dif =  av_mailed - av_clerk) %>%
  arrange(-av_dif) %>%
  select(class, tax_code_num, av_dif, pin, everything())
```
```{r}
#| label: fig-muni-avappeals-ty21
#| column: page
#| layout-ncol: 2
#| fig-cap: "Tax Year 2021: Appeals & Incentive Properties"

source("helper_tc_muninames_2022.R")


# Add in names that merge with Muni shapefile names
muni_pincounts_majorclass <- incent_pins_ptax %>% 
  left_join(tc_muninames, by = "tax_code_num") %>%
  rename(year_ignore = year.y,
         year = year.x) %>%
  filter(av_mailed > av_clerk & 
                              year == 2021) %>%
    mutate(av_dif =  av_mailed - av_clerk) %>%
  group_by(clean_name, agency_name, shpfile_name) %>%
  dplyr::summarize(#pin_count = n(),
            av_dif = sum(av_dif))

# table(muni_pincounts_majorclass$major_class_code) # Number of Munis with these major classes


fig_muni_AV_w_chicago <- muni_pincounts_majorclass %>% 
  filter(!is.na(shpfile_name) ) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = av_dif) ) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Tax Year 2021") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "AV Decrease from Appeals",
    label = scales::dollar)

fig_muni_AV_nochicago <- muni_pincounts_majorclass %>% 

  filter(clean_name != "Chicago") %>% 
  filter(!is.na(shpfile_name) ) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = av_dif) ) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Tax Year 2021: Excludes Chicago") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
   nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
 # limits = c(0, 90000000),
    name = "AV Decrease from Appeals",
    label = scales::dollar)


fig_muni_AV_w_chicago
fig_muni_AV_nochicago
```

```{r}
#| label: fig-muni-avappeals-ty22
#| column: page
#| layout-ncol: 2
#| fig-cap: "Tax Year 2022 Appeals & Incentive Properties"

source("helper_tc_muninames_2022.R")


# Add in names that merge with Muni shapefile names
muni_pincounts_majorclass <- incent_pins_ptax %>% 
  left_join(tc_muninames) %>%

  filter(av_mailed > av_clerk & 
                              year == 2022) %>%
    mutate(av_dif =  av_mailed - av_clerk) %>%
  group_by(clean_name, agency_name, shpfile_name) %>%
  dplyr::summarize(#pin_count = n(),
            av_dif = sum(av_dif))


fig_muni_AV_w_chicago_2022 <- muni_pincounts_majorclass %>% 
  filter(!is.na(shpfile_name) ) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = av_dif) ) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Tax Year 2022") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "AV Decrease from Appeals",
    label = scales::dollar)

fig_muni_AV_nochicago_2022 <- muni_pincounts_majorclass %>% 
  filter(clean_name != "Chicago") %>% 
  filter(!is.na(shpfile_name) ) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = av_dif) ) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Tax Year 2022: Excludes Chicago") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
   nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
 # limits = c(0, 90000000),
    name = "AV Decrease from Appeals",
    label = scales::dollar)


fig_muni_AV_w_chicago_2022
fig_muni_AV_nochicago_2022
```
