---
title: "Incentive Properties in Cook County, IL"
author: "AWM"
output: 
  html_document:
    df_print: paged
    code_folding: hide
    code_download: true
---

```{r setup, warning = FALSE, message = FALSE}
library(tidyverse)
library(ptaxsim)
library(DBI)
library(httr)
library(jsonlite)
library(glue)
library(sf)

knitr::opts_chunk$set(warning = FALSE, message = FALSE)



ptaxsim_db_conn <- DBI::dbConnect(RSQLite::SQLite(), "./ptaxsim.db/ptaxsim-2021.0.4.db")

# link to the API output as a JSON file
muni_shp <- read_sf("https://gis.cookcountyil.gov/traditional/rest/services/politicalBoundary/MapServer/2/query?outFields=*&where=1%3D1&f=geojson")

#muni_shp <- read_json("muni_shp.json")
nicknames <- readxl::read_excel("./Necessary_Files/muni_shortnames.xlsx")

class_dict <- read_csv("./Necessary_Files/class_dict_expanded.csv") %>% 
  mutate(class_code = as.character(class_code))

NBHs <- read_sf("https://datacatalog.cookcountyil.gov/resource/pcdw-pxtg.geojson")



# has EAV values, extensions by agency_num
agency_dt <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT *
  FROM agency
  WHERE year = 2021
  "
)

cook_agency_names <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT DISTINCT agency_num, agency_name
  FROM agency_info
  "
)

 


# has all tax codes and the taxing agency that taxes them. Tax code rates and agency rates. 
cook_tax_codes <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql("
  SELECT*
  FROM tax_code
  WHERE agency_num IN ({cook_agency_names$agency_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
)

muni_agency_names <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT DISTINCT agency_num, agency_name, minor_type
  FROM agency_info
  WHERE minor_type = 'MUNI'
  OR agency_num = '020060000'  

  "
)

muni_tax_codes <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql("
  SELECT*
  FROM tax_code
  WHERE agency_num IN ({muni_agency_names$agency_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
)

muni_tax_codes <- muni_tax_codes %>% left_join(muni_agency_names)
```

# Maps!

<!--Assessor neighborhoods are 5 digits: first 2 for the township, last 3 are for the neighborhood within the township. These do not resemble the neighborhoods that many of us may think of but are instead of groups of properties with similar characteristics **(that are taxed in similar rates? double check)**-->

<!--Pull pins, neighborhood codes, property class, assessment triad, and other variables from Parcel Universe using the GET() command. Pulls data from online API, not PTAXSIM database. I am using this instead of PTAXSIM because neighborhood codes are not included in PTAXSIM tables: Tax codes are the smallest unit of measurement. **NOTE: Tax codes and neighborhoods do not appear to share borders?**-->


```{r eval=FALSE}
base_url <- "https://datacatalog.cookcountyil.gov/resource/tx2p-k2g9.json"

nbh_pins <- GET(
  base_url,
  query = list(
    tax_year = 2021,
 #   ward_num = 49,
   # property_city = "ROGERS PARK",
    `$select` = paste0(c("pin", "pin10", 
                         "class", 
                         "township_code", "township_name",
                         "nbhd_code", "census_puma_geoid",
                         #"lat","lon", 
                         "triad_name" 
                         ),
   collapse = ","),
   `$limit` = 500000000L
  )
)

nbh_pins2 <- fromJSON(rawToChar(nbh_pins$content))
head(nbh_pins2)
```




```{r eval=FALSE}
# There are 1,864,594 pins taxed by Cook County in 2021.
cook_pins <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql(
  "SELECT DISTINCT pin, class, tax_code_num
  FROM pin
  WHERE tax_code_num IN ({cook_tax_codes$tax_code_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
))

# finds all pins within a municipality
pin_data <- lookup_pin(2021, cook_pins$pin) %>%
  left_join(cook_pins, by = c("pin", "class")) %>% 
  mutate(class_1dig = str_sub(class, 1, 1))

# change variable type to character so the join works.
#class_dict$class_code <- as.character(class_dict$class_code)



# use the property class to make the major property types
# joins the class_dict file to the pin_data classes
#pin_data <- class_dict %>%
  #select(-c(assessment_level:reporting_group, class_desc:max_size)) %>%
#  right_join(pin_data, by = c("class_code" = "class"))



#parcel_exemps <- read_csv("./Output/3_Exemptions_Details_output-ParcelExemps.csv")

parcel_data <- left_join(pin_data, nbh_pins2, by = c("pin", "class"))

parcel_data <- parcel_data %>% 
  mutate(township = str_sub(nbhd_code,1,2),
         nbhd_3 = str_sub(nbhd_code, 3,5)) %>% 
  left_join(class_dict, by = c("class" = "class_code"))

parcel_data %>% head()

incentive_classes <- parcel_data %>% filter(class>599 & class < 900) %>%  mutate(pin10 = str_sub(pin,1, 10))  %>% left_join(muni_tax_codes)

write_csv(incentive_classes, "./Output/7_output_incentive_classes.csv")




```

```{r }
incentive_classes <- read_csv("./Output/7_output_incentive_classes.csv") %>% 
  filter(class>599 & class < 900)

  

incentives_perTC <- incentive_classes %>% 
  group_by(tax_code_num, major_class_code) %>% 
  summarize(av = sum(av), 
            pins_in_parcel = n())



table1 <- incentive_classes %>% 
  filter(class > 799 & class < 900) %>% 
  group_by(township_name, agency_name, major_class_code) %>%
  summarize(av = sum(av)) %>%
  pivot_wider(id_cols = c(township_name, agency_name), 
              names_from = "major_class_code", 
              values_from = "av") %>%
  rename(AV = `8`)

table2 <- incentive_classes %>% 
  filter(class > 799 & class < 900) %>% 
  group_by(township_name, agency_name, major_class_code) %>%
  summarize(pincount = n()) %>%
  pivot_wider(id_cols = c(township_name, agency_name), 
              names_from = "major_class_code", 
              values_from = "pincount") %>%
  rename(pincount = `8`)

left_join(table1, table2)

incentive_classes %>% 
  filter(class > 599 & class < 800) %>% 
  group_by(township_name, agency_name, major_class_code) %>%
  summarize(pincount = n()) %>%
  pivot_wider(id_cols = c(township_name, agency_name), 
              names_from = "major_class_code", 
              values_from = "pincount")

```

```{r nbhd-allclasses}
nbh_pincounts <- incentive_classes %>% 
  group_by(nbhd_code, class) %>%
  dplyr::summarize(pin_count = n(),
            av = sum(av)) %>%
  mutate(class_1dig = str_sub(class, 1, 1))

nbh_pincounts_majorclass <- incentive_classes %>% 
  group_by(major_class_code, major_class_type, nbhd_code) %>%
  dplyr::summarize(pin_count = n(),
            av = sum(av))

# table(nbh_pincounts_majorclass$major_class_code)


incentives_props <- nbh_pincounts_majorclass %>% 
  filter(major_class_code %in% c("6A", "6B", "6C", "7A", "7B", "8", "9")) %>%
  mutate(nbhd_code = as.character(nbhd_code))


fig_nbhd_pincount <- incentives_props %>% 
  group_by(nbhd_code) %>%
  summarize(pin_count = sum(pin_count)) %>%
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  ggplot(aes(fill = pin_count)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Number of PINs with Incentives in Assessor Neighborhoods", 
  caption = "Count of 14 digit PINs") +
  theme_void() + 
   theme(axis.ticks = element_blank(), axis.text = element_blank()) +
scale_fill_steps2(
  high = "darkblue", low = "orange",
  show.limits=TRUE,
  nice.breaks = TRUE,
na.value = NA,
  name = "Pin Count")


fig_nbhd_AV <- incentives_props %>% 
  group_by(nbhd_code) %>%
  summarize(av = sum(av)) %>%
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  filter(triad_name != "City") %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Assessed Value of Properties in 
       Assessor Neighborhoods", 
       caption = "Different classes of properties have different assessment levels. 
  Some change over time.") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    show.limits=TRUE,
    nice.breaks = FALSE,
   # limits = c(0, 90000000),
    na.value = NA,
    name = "AV",
    label = scales::dollar)



fig_nbhd_Chicago_AV <- incentives_props %>% 
  group_by(nbhd_code) %>%
  summarize(av = sum(av)) %>%
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  filter(triad_name == "City") %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Assessed Value of Properties in 
       Assessor Neighborhoods", 
       caption = "Different classes of properties have different assessment levels. 
  Some change over time.") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    show.limits=TRUE,
    nice.breaks = FALSE,
   # limits = c(0, 90000000),
    na.value = NA,
    name = "AV",
    label = scales::dollar)
fig_nbhd_Chicago_AV
```

```{r nhbd-class6and7}
nbh_pincounts_class <- incentive_classes %>% 
  group_by(nbhd_code, class) %>%
  dplyr::summarize(pin_count = n(),
            av = sum(av)) %>%
  mutate(class_1dig = str_sub(class, 1, 1))

nbh_pincounts_majorclass <- incentive_classes %>% 
  group_by(major_class_code, major_class_type, nbhd_code) %>%
  dplyr::summarize(pin_count = n(),
            av = sum(av))

# table(nbh_pincounts_majorclass$major_class_code)


incentives_props <- nbh_pincounts_majorclass %>% 
  filter(major_class_code %in% c("6A", "6B", "6C", "7A", "7B", "8")) %>%
  mutate(nbhd_code = as.character(nbhd_code))

fig_nbhd_C7AV <- incentives_props %>% 
  filter(major_class_code %in% c("7A", "7B", "7C") ) %>% 
  group_by(nbhd_code) %>%
  summarize(av = sum(av)) %>%
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  filter(triad_name != "City") %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Class 7 Properties - Assessed Value",
       subtitle = "AV of Buildings with Incentives in Assessor Neighborhoods",
       caption = "Over $90 million in Chicago") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "AV ($)",
    label = scales::dollar)

fig_nbhd_C7_Chicago_AV <- incentives_props %>% 
  filter(major_class_code %in% c("7A", "7B", "7C") ) %>% 
  group_by(nbhd_code) %>%
  summarize(av = sum(av)) %>%
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  filter(triad_name == "City") %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Class 7 Properties - Assessed Value",
       subtitle = "AV of Buildings with Incentives in Assessor Neighborhoods",
       caption = "Over $90 million in Chicago") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "AV ($)",
    label = scales::dollar)



fig_nbhd_C7_PC <- incentives_props %>% 
  filter(major_class_code %in% c("7A", "7B", "7C") ) %>%
  group_by(nbhd_code) %>%
  summarize(pincount = sum(pin_count)) %>% 
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
    filter(triad_name != "City") %>%

  ggplot(aes(fill = pincount)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Class 7 Properties",
       subtitle = "# of Buildings with Class 7 Incentives in Assessor Neighborhoods") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    show.limits=TRUE,
    na.value = NA,
    name = "Pin Count")

fig_nbhd_C7_Chicago_PC <- incentives_props %>% 
  filter(major_class_code %in% c("7A", "7B", "7C") ) %>%
  group_by(nbhd_code) %>%
  summarize(pincount = sum(pin_count)) %>% 
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
    filter(triad_name == "City") %>%

  ggplot(aes(fill = pincount)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Class 7 Properties in Chicago",
       subtitle = "# of Buildings with Class 7 Incentives in Assessor Neighborhoods") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    show.limits=TRUE,
    na.value = NA,
    name = "Pin Count")

fig_nbhd_C6_AV <- incentives_props %>% 
  filter(major_class_code %in% c("6A", "6B", "6C") ) %>%
  group_by(nbhd_code) %>%
  summarize(av = sum(av)) %>%
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Class 6 Properties",
       subtitle = "AV of Buildings with Incentives in Assessor Neighborhoods") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "AV ($)",
    label = scales::dollar)

fig_nbhd_C6_Chicago_AV <- incentives_props %>% 
  filter(major_class_code %in% c("6A", "6B", "6C") ) %>%
  group_by(nbhd_code) %>%
  summarize(av = sum(av)) %>%
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  filter(triad_name == "City") %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Class 6 Properties",
       subtitle = "AV of Buildings with Incentives in Assessor Neighborhoods") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "AV ($)",
    label = scales::dollar)
  
fig_nbhd_C6_PC <- incentives_props %>% 
  filter(major_class_code %in% c("6A", "6B", "6C") ) %>%
  group_by(nbhd_code) %>%
  summarize(pin_count= round(sum(pin_count)), digits = 0) %>%
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  ggplot(aes(fill = pin_count)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Class 6 Properties",
       subtitle = "# of Buildings with Incentives in Assessor Neighborhoods") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    #nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "Pin Count")

fig_nbhd_C6_Chicago_PC <- incentives_props %>% 
  filter(major_class_code %in% c("6A", "6B", "6C") ) %>%
  group_by(nbhd_code) %>%
  summarize(pin_count= round(sum(pin_count)), digits = 0) %>%
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  filter(triad_name == "City") %>%
  ggplot(aes(fill = pin_count)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Class 6 Properties",
       subtitle = "# of Buildings with Incentives in Assessor Neighborhoods") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    #nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "Pin Count")
```



## All Incentive Types Together

```{r muni-alltogether, out.width= "50%", fig.show='hold', fig.cap= "All Incentive Property Classes: Municipality level and Assessor Neighborhood level"}


# Add in names that merge with Muni shapefile names
muni_pincounts_majorclass <- incentive_classes %>% 
  left_join(nicknames) %>%
  group_by(clean_name, major_class_code, major_class_type, agency_name, shpfile_name) %>%
  dplyr::summarize(pin_count = n(),
            av = sum(av))

# table(muni_pincounts_majorclass$major_class_code) # Number of Munis with these major classes


fig_muni_AV <- muni_pincounts_majorclass %>% 
  filter(major_class_code %in% c("6A", "6B", "6C", "7A", "7B", "8")) %>%
  filter(clean_name != "Chicago") %>% 
  group_by(shpfile_name)%>%
  summarize(av = sum(av)) %>% 
  filter(!is.na(shpfile_name) ) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Properties with Any Type of Incentive Class",
       subtitle = "AV of Buildings with Incentives in Municipalities (Except Chicago).
       Chicago has around $500 million in AV from properties with incentives.") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
  #  nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
  limits = c(0, 90000000),
    name = "AV ($)",
    label = scales::dollar)

fig_muni_AV
fig_nbhd_AV
```


```{r muni-pincount, out.width= "50%", fig.show='hold'}
fig_muni_pincount <- muni_pincounts_majorclass %>% 
  filter(major_class_code %in% c("6A", "6B", "6C", "7A", "7B", "8")) %>%
  filter(clean_name != "Chicago") %>% 
  group_by(shpfile_name)%>%
  summarize(pin_count = round(sum(pin_count)) ) %>% 
  filter(!is.na(shpfile_name)) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = pin_count)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Properties with Incentive Clases",
       subtitle = "# of Buildings with Incentives in Municipalities (Except Chicago)") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
  #  nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "# PINs")


fig_muni_pincount
fig_nbhd_pincount
```

## Class 6 Incentives

Class 6A: Industrial Incentive

Class 6B: Industrial Incentive - Designed to encourage industrial development throughout Cook County by offering a real estate tax incentive for the development of new industrial facilities, the rehabilitation of existing industrial structures, and the industrial reutilization of abandoned buildings.

Class 6C: Industrial Brownfield

```{r out.width= "50%", fig.show='hold'}

fig_muni_C6_AV <- muni_pincounts_majorclass %>% 
  filter(major_class_code %in% c("6A", "6B", "6C")) %>%
  filter(clean_name != "Chicago") %>% 
  group_by(shpfile_name)%>%
  summarize(av = sum(av)) %>% 
  filter(!is.na(shpfile_name)) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Properties with Class 6 Incentives",
       subtitle = "Assessed Value in Municipalities (Except Chicago)") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
   # nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "AV ($)",
  label = scales::dollar)



fig_muni_C6_PC <- muni_pincounts_majorclass %>% 
  filter(major_class_code %in% c("6A", "6B", "6C")) %>%
 # filter(clean_name != "Chicago") %>% 
  group_by(shpfile_name)%>%
  summarize(pin_count = sum(pin_count)) %>% 
  filter(!is.na(shpfile_name)) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = pin_count)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Properties with Class 6 Incentives",
       subtitle = "# of Buildings with Incentives in Municipalities") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "# PINs")

fig_muni_C6_AV
fig_nbhd_C6_AV

fig_muni_C6_PC
fig_nbhd_C6_PC



```


## Class 7 Incentives

Class 7A(B): Intended to encourage commercial projects in areas determined to be "in need of commercial development.These projects have total development costs, exclusive of land, that DO NOT EXCEED(DO EXCEED) two million and would not be economically feasible without the incentive.



```{r out.width= "50%", fig.show='hold'}


fig_muni_C7_AV <- muni_pincounts_majorclass %>% 
  filter(major_class_code %in% c("7A", "7B", "7C")) %>%
  filter(clean_name != "Chicago") %>% 
  group_by(shpfile_name)%>%
  summarize(av = sum(av)) %>% 
  filter(!is.na(shpfile_name)) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Assessed Value of Properties with Class 7 Incentives",
       subtitle = "in Municipalities (Except Chicago)") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "AV ($)",
  label = scales::dollar)


fig_muni_C7_PC <- muni_pincounts_majorclass %>% 
  filter(major_class_code %in% c("7A", "7B", "7C")) %>%
  filter(clean_name != "Chicago") %>% 
  group_by(shpfile_name)%>%
  summarize(pin_count = sum(pin_count)) %>% 
  filter(!is.na(shpfile_name)) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = pin_count)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Properties with Class 7 Incentives",
       subtitle = "# of Buildings with Incentives in Municipalities (Except Chicago)") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
  #  nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "# PINs")

fig_muni_C7_AV
fig_nbhd_C7AV



fig_muni_C7_PC
fig_nbhd_C7_PC
```


## Class 8 Incentives


Class 8: Designed to encourage industrial and commercial development in areas of the county which are experiencing severe economic stagnation.


Class 8 properties exist in places besides the 5 townships:

Bloom, Bremen, Calumet, Rich, Thornton  


```{r class8, out.width= "50%", fig.show='hold'}
fig_nbhd_C8_PC <- incentives_props %>% 
  filter(major_class_code == "8") %>%
  group_by(nbhd_code)%>%
  summarize(pin_count = sum(pin_count)) %>% 
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  ggplot(aes(fill = pin_count)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Class 8 Properties",
       subtitle = "Number of PINs with Incentives in Assessor Neighborhoods", 
       caption = "Count of 14 digit PINs") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    #nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "Count")


fig_nbhd_C8 <- incentives_props %>% 
  filter(major_class_code == "8") %>%
  group_by(nbhd_code) %>%
  summarize(av = sum(av)) %>% 
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry),
         # color = "black",
          lwd = .1)  +
  labs(title = "Class 8 Properties",
       subtitle = "AV of Buildings with Incentives in Assessor Neighborhoods") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    nice.breaks = FALSE,
    show.limits=TRUE,
    limits = c(0, 13000000),
    na.value = NA,
    name = "AV ($)",
    label = scales::dollar)

fig_muni_C8_AV <- muni_pincounts_majorclass %>% 
  filter(major_class_code %in% c("8")) %>%
  filter(clean_name != "Chicago") %>% 
  group_by(shpfile_name)%>%
  summarize(av = sum(av)) %>% 
  filter(!is.na(shpfile_name)) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Properties with Class 8 Incentives",
       subtitle = "# of Buildings with Incentives in Municipalities (Except Chicago)") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "AV ($)",
  label = scales::dollar)



fig_muni_C8_PC <- muni_pincounts_majorclass %>% 
  filter(major_class_code %in% c("8")) %>%
 # filter(clean_name != "Chicago") %>% 
  group_by(shpfile_name)%>%
  summarize(pin_count = sum(pin_count)) %>% 
  filter(!is.na(shpfile_name)) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = pin_count)) + 
  geom_sf(aes(geometry = geometry),
          color = "black",
          lwd = .05)+ 
  labs(title = "Properties with Class 8 Incentives",
       subtitle = "# of Buildings with Incentives in Municipalities") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
  #  nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "# PINs")

fig_muni_C8_PC
fig_nbhd_C8_PC



```

> There are some Class 8 properties located in townships other than the Class 8 five chosen for the incentives.


```{r class8-2, out.width= "50%", fig.show='hold'}
fig_muni_C8 <- muni_pincounts_majorclass %>% 
   filter(!is.na(shpfile_name)) %>% 

  filter(major_class_code %in% c("8")) %>%
  group_by(shpfile_name)%>%
  summarize(av = sum(av)) %>% 
  full_join(muni_shp, by = c("shpfile_name" = "MUNICIPALITY")) %>%
  ggplot(aes(fill = av)) + 
  geom_sf(aes(geometry = geometry),color = "black", lwd = .05)+ 
  labs(title = "Class 8 Properties",
       subtitle = "AV of Buildings with Incentives in Assessor Neighborhoods") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "orange",
    nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    limits = c(0, 13000000),
    name = "AV ($)",
    label = scales::dollar)

fig_muni_C8
fig_nbhd_C8

```



```{r eval=FALSE}
class8Townships = read_csv("./Output/3_Exemptions_Details_output-ClassTaxcodeExemps.csv") %>% 
  mutate(class_code = as.character(class_code)) %>%
  filter(class_code >=800 & class_code <900) %>%
  left_join(class_dict) %>%
  mutate(tax_code_num = as.character(tax_code_num)) %>%
  left_join(township_tax_codes) %>%
  left_join(township_agency_names)%>% 
  distinct(agency_name) 

class8Townships
```

# Chicago Specific Graphs

```{r out.width= "50%", fig.show='hold'}

# fig_nbhd_Chicago_AV All incentives

fig_nbhd_C6_Chicago_AV
fig_nbhd_C6_Chicago_PC
fig_nbhd_C7_Chicago_AV
fig_nbhd_C7_Chicago_PC
```

# Tables

```{r}

incentive_classes %>% 
  group_by(agency_name, major_class_type) %>%
  summarize(av = sum(av),
           # pins = n()
           ) %>%
  pivot_wider(id_cols = agency_name, names_from = "major_class_type", values_from = "av")
  
incentive_classes %>% 
  group_by(agency_name, major_class_type) %>%
  summarize(av = sum(av),
            pins = n()) %>%
  pivot_wider(id_cols = agency_name, names_from = "major_class_type", values_from = "pins")

incentive_classes %>% 
  group_by(agency_name, major_class_type, major_class_code) %>%
  summarize(av = sum(av),
            pins = n()) %>%
  pivot_wider(id_cols = agency_name, names_from = "major_class_code", values_from = "av")

incentive_classes %>% 
  group_by(agency_name, major_class_type, major_class_code) %>%
  summarize(av = sum(av),
            pins = n()) %>%
  pivot_wider(id_cols = agency_name, names_from = "major_class_code", values_from = "pins")



# By Muni and Property Class
incentive_classes %>% 
    mutate(tax_code_num = as.character(tax_code_num)) %>% 
  left_join(muni_tax_codes) %>%
  mutate(tax_code_num = as.character(tax_code_num)) %>% 
  left_join(muni_agency_names) %>%
  group_by(agency_num, class) %>%
  dplyr::summarize(pin_count = n(),
            av = sum(av)) %>%
  mutate(class_1dig = str_sub(class, 1, 1))

# By Muni and Major Property Class
incentive_classes %>% 
    mutate(tax_code_num = as.character(tax_code_num)) %>% 
  left_join(muni_tax_codes) %>%
    mutate(tax_code_num = as.character(tax_code_num)) %>% 
  left_join(muni_agency_names) %>%
  group_by(agency_num, major_class_code) %>%
  dplyr::summarize(pin_count = n(),
            av = sum(av))
```

# Areas with lots of exemptions and incentives 

aka a reduced tax base

```{r}

pin_data <- read_csv("./Output/4C_joined_PINs_bills_and_exemptions.csv")


muni_taxrates <- pin_data %>% 
  group_by(agency_name)  %>%
  
  summarize(
    av = sum(av, na.rm = TRUE),
    eav = sum(eav, na.rm = TRUE),
    equalized_AV = sum(equalized_AV, na.rm = TRUE),
    pins_in_class = n(),
    current_exemptions = sum(all_exemptions, na.rm = TRUE),
    HO_exemps = sum(exe_homeowner, na.rm = TRUE),
    tax_code_rate = mean(tax_code_rate, na.rm = TRUE), # Changed from first() to mean() on Nov 1
    final_tax_to_dist = sum(final_tax_to_dist, na.rm = TRUE), # used as LEVY amount!! 
    final_tax_to_tif = sum(final_tax_to_tif, na.rm = TRUE),
    tax_amt_exe = sum(tax_amt_exe, na.rm = TRUE), 
    tax_amt_pre_exe = sum(tax_amt_pre_exe, na.rm = TRUE), 
    tax_amt_post_exe = sum(tax_amt_post_exe, na.rm = TRUE),
    rpm_tif_to_cps = sum(rpm_tif_to_cps, na.rm = TRUE), # not used
    rpm_tif_to_rpm = sum(rpm_tif_to_rpm, na.rm=TRUE), # not used
    rpm_tif_to_dist = sum(rpm_tif_to_dist, na.rm=TRUE), # not used
    tif_share = mean(tif_share, na.rm=TRUE), # not used
  ) %>%
  
  mutate(total_bill_current = final_tax_to_dist + final_tax_to_tif) %>%
  rename(cur_comp_TC_rate = tax_code_rate) %>%
  mutate(current_taxable_eav = final_tax_to_dist/(cur_comp_TC_rate/100),
         new_taxable_eav = final_tax_to_dist/(cur_comp_TC_rate/100) + HO_exemps) %>%
  mutate(new_comp_TC_rate = (final_tax_to_dist / new_taxable_eav)*100) %>%
  mutate(new_comp_TC_rate = ifelse(is.nan(new_comp_TC_rate), cur_comp_TC_rate, new_comp_TC_rate)) %>%
  select(agency_name, cur_comp_TC_rate, new_comp_TC_rate, current_taxable_eav, new_taxable_eav, everything())

muni_taxrates 
```

```{r}

prop_class_sums <- pin_data %>% 
  group_by(agency_name, major_class_code, major_class_type )  %>%
  
  summarize(
    av = sum(av, na.rm = TRUE),
    eav = sum(eav, na.rm = TRUE),
    equalized_AV = sum(equalized_AV, na.rm = TRUE),
    pins_in_class = n(),
    current_exemptions = sum(all_exemptions, na.rm = TRUE),
    HO_exemps = sum(exe_homeowner, na.rm = TRUE),
    tax_code_rate = mean(tax_code_rate, na.rm = TRUE), # Changed from first() to mean() on Nov 1
    final_tax_to_dist = sum(final_tax_to_dist, na.rm = TRUE), # used as LEVY amount!! 
    final_tax_to_tif = sum(final_tax_to_tif, na.rm = TRUE),
    tax_amt_exe = sum(tax_amt_exe, na.rm = TRUE), 
    tax_amt_pre_exe = sum(tax_amt_pre_exe, na.rm = TRUE), 
    tax_amt_post_exe = sum(tax_amt_post_exe, na.rm = TRUE),
    rpm_tif_to_cps = sum(rpm_tif_to_cps, na.rm = TRUE), # not used
    rpm_tif_to_rpm = sum(rpm_tif_to_rpm, na.rm=TRUE), # not used
    rpm_tif_to_dist = sum(rpm_tif_to_dist, na.rm=TRUE), # not used
    tif_share = mean(tif_share, na.rm=TRUE), # not used
  ) %>%
  
  mutate(total_bill_current = final_tax_to_dist + final_tax_to_tif) %>%
  rename(cur_comp_TC_rate = tax_code_rate) %>%
  mutate(current_taxable_eav = final_tax_to_dist/(cur_comp_TC_rate/100),
         new_taxable_eav = final_tax_to_dist/(cur_comp_TC_rate/100) + HO_exemps) %>%
  mutate(new_comp_TC_rate = (final_tax_to_dist / new_taxable_eav)*100) %>%
  mutate(new_comp_TC_rate = ifelse(is.nan(new_comp_TC_rate), cur_comp_TC_rate, new_comp_TC_rate)) %>%
  select(agency_name, major_class_code, HO_exemps, current_exemptions, pins_in_class, current_taxable_eav, new_taxable_eav,  everything())

# muni level by major_class summary
prop_class_sums 


```
## Neighborhood level join

```{r}

nbh_major_class <- read_csv("./Output/8_Current_Taxrates_per_nbh_by_majorclass.csv")
nbh_major_class


nbh_pincounts_class <- incentive_classes %>% 
  group_by(nbhd_code, class) %>%
  dplyr::summarize(pin_count = n(),
            av = sum(av)) %>%
  mutate(class_1dig = str_sub(class, 1, 1))

nbh_pincounts_majorclass <- incentive_classes %>% 
  group_by(major_class_code, major_class_type, nbhd_code) %>%
  dplyr::summarize(pin_count = n(),
            av = sum(av))

incentives_props <- nbh_pincounts_majorclass %>% 
  filter(major_class_code %in% c("6A", "6B", "6C", "7A", "7B", "8")) %>%
  mutate(nbhd_code = as.character(nbhd_code))




# nbh_res_exemptions <- read_csv("./Output/8_nbh_sum_new_exe.csv") 
nbh_taxrates <- read_csv( "./Output/8_Current_Taxrates_per_nbh.csv")


nbh_data <- left_join(nbh_pincounts_majorclass, nbh_taxrates, by = "nbhd_code")


nbh_data <- nbh_data %>% rename(all_PINs = pin_count.x,
                                incentive_PINs = pin_count.y)  %>%
  mutate(exempt_over_total_EAV = Exempt_EAV / Total_EAV,
         nbhd_code = as.character(nbhd_code),
         exempts_perPIN = Exempt_EAV / all_PINs,
         incentive_projects_pct = incentive_PINs / all_PINs,
         incentives_EAV = av * 3.0027 ,
         incentives_over_total_EAV = incentives_EAV / Total_EAV) %>%
  select(nbhd_code, exempt_over_total_EAV, all_PINs, incentive_PINs, exempts_perPIN, incentive_projects_pct, incentives_EAV, incentives_over_total_EAV, major_class_code, everything()) %>% arrange(nbhd_code)

nbh_data


```
```{r munilevel-exemptions, eval=FALSE, include=FALSE}
# prop_class_sums is at muni level
exempts_per_pin <- prop_class_sums %>% 
  filter(major_class_code == "2") %>% # Class 2 residential only
  mutate(avg_GHEexe_perPIN = HO_exemps/pins_in_class,
         avg_allexe_perPIN = current_exemptions / pins_in_class) %>% 
  select(agency_name, avg_GHEexe_perPIN, avg_allexe_perPIN)
exempts_per_pin


# Class2_data is also at Muni Level
Class2_data <- read_csv("./Output/5b_Class2_burdenshift.csv")

residential_data <- Class2_data %>% 
  mutate(exemptions_per_resPIN = C2_current_GHE/ C2_PC_permuni )  %>%
  select(clean_name, exemptions_per_resPIN, C2_EAV_pct)
```


```{r}
nbh_res <- nbh_major_class %>% 
  filter(major_class_code == 2) %>% 
  select(nbhd_code, major_class_code, 
         residential_PC = pin_count, 
         res_EAV = Total_EAV, 
         res_exemptions = Exempt_EAV) %>%
  mutate(nbhd_code = as.character(nbhd_code),
         res_exe_per_res_pin = round(res_exemptions / residential_PC)
  )
  
  
nbh_data2 <- nbh_data %>% left_join(nbh_res, by = "nbhd_code") %>% select(nbhd_code, all_PINs, residential_PC, incentive_PINs, everything()) %>% mutate(exemps_per_resPIN = res_exemptions / residential_PC) %>% arrange(nbhd_code)


fig4 <- nbh_data2 %>% 
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  ggplot(aes(fill = exemps_per_resPIN)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Average exemptions per PIN") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "white",
    #nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "Exemptions per PIN")

fig4
```

## Maps

```{r}
fig1 <- nbh_data %>% 
  #filter(major_class_code == "8") %>%
  #group_by(nbhd_code)%>%
  #summarize(pin_count = sum(pin_count)) %>% 
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  ggplot(aes(fill = exempt_over_total_EAV)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Exempt EAV / Total EAV",
      # caption = "Total EAV includes TIF and all EAV before exemptions are deducted"
       ) +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "forestgreen", low = "white",
    #nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "% Total EAV Exempt")

fig1



fig3 <- nbh_data %>% 
  #filter(major_class_code == "8") %>%
  #group_by(nbhd_code)%>%
  #summarize(pin_count = sum(pin_count)) %>% 
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  ggplot(aes(fill = incentives_over_total_EAV)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "EAV of incentive properties / Total EAV") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "orange", low = "yellow",
    #nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "%")

fig3


fig2 <- nbh_data %>% 
  #filter(major_class_code == "8") %>%
  #group_by(nbhd_code)%>%
  #summarize(pin_count = sum(pin_count)) %>% 
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>%
  ggplot(aes(fill = exempts_perPIN)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Average exemptions per PIN") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  scale_fill_steps2(
    high = "darkblue", low = "white",
    #nice.breaks = FALSE,
    show.limits=TRUE,
    na.value = NA,
    name = "Exemptions per PIN")

fig2
```

