---
title: "Assessor Neighborhoods"
author: "AWM"
format: 
  html:
    code-fold: true
---

<!--Assessor neighborhoods are 5 digits: first 2 for the township, last 3 are for the neighborhood within the township. These do not resemble the neighborhoods that many of us may think of but are instead of groups of properties with similar characteristics -->

<!--Pull pins, neighborhood codes, property class, assessment triad, and other variables from Parcel Universe using the GET() command. Pulls data from online API, not PTAXSIM database. I am using this instead of PTAXSIM because neighborhood codes are not included in PTAXSIM tables: Tax codes are the smallest unit of measurement. -->
# Using Assessor Neighborhoods

```{r setup, warning = FALSE, message = FALSE}
library(tidyverse)
library(ptaxsim)
library(DBI)
library(httr)
library(jsonlite)
library(glue)
library(sf)

knitr::opts_chunk$set(warning = FALSE, message = FALSE)


#NBHs<- read_csv("Assessor_-_Neighborhood_Boundaries.csv")

NBHs <- read_sf("https://datacatalog.cookcountyil.gov/resource/pcdw-pxtg.geojson")
```

## Aggregating pin data to new shapes

-   ::: {.callout-note appearance="minimal"}
    Assessor neighborhoods are 5 digits: first 2 for the township, last 3 are for the neighborhood within the township. These do not resemble the neighborhoods that many of us may think of but are instead of groups of properties with similar characteristics
    :::

Pull pins, neighborhood codes, property class, assessment triad, and other variables from Parcel Universe using the GET() command. Pulls data from online API, not PTAXSIM database. I am using this instead of PTAXSIM because neighborhood codes are not included in PTAXSIM tables: Tax codes are the smallest unit of measurement.


### Map

```{r}
nbh_sum <- read_csv("./Output/nbh_summary.csv") %>%
    mutate(nbhd_code = as.character(nbhd_code),
           township_code = str_sub(nbhd_code, 1,2))

nbh_MC_sum <- read_csv("./Output/nbh_MC_summary.csv") %>%
    mutate(nbhd_code = as.character(nbhd_code))

nbh_sum %>% 
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>% 
  ggplot(aes(fill = nbh_res_exe/nbh_PC)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Average Exemptions/PIN in Neighborhoods", 
  caption = "Excludes Abatements") +
  theme_void() + 
   theme(axis.ticks = element_blank(), axis.text = element_blank()) +
scale_fill_steps2(
  high = "darkblue", low = "orange",
  midpoint = median(nbh_sum$nbh_res_exe/nbh_sum$nbh_PC, na.rm = TRUE),
 nice.breaks = FALSE,
  show.limits=TRUE,
na.value = "gray",
  name = "Dollars",
  labels = scales::dollar
) 

nbh_sum %>% 
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>% 
  ggplot(aes(fill = nbh_exe_abate) ) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Exempt EAV from Abatements in Assessor Neighborhoods", 
  caption = "Abatements  comes from exe_abate in PTAXSIM data.") +
  theme_void() + 
   theme(axis.ticks = element_blank(), axis.text = element_blank()) +
scale_fill_steps2(
  high = "darkblue", low = "orange",
 # midpoint = mean(nbh_sum$nbh_exe_abate),
 nice.breaks = FALSE,
  show.limits=TRUE,
na.value = NA,
  name = "Dollars",
  labels = scales::dollar
)

nbh_sum %>% 

  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>% 
  ggplot(aes(fill = nbh_PC ) ) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "# PINs with Abatements in each Assessor Neighborhood", 
  caption = "Abatements comes from exe_abate in PTAXSIM data.") +
  theme_void() + 
   theme(axis.ticks = element_blank(), axis.text = element_blank()) +
scale_fill_steps2(
  high = "darkblue", low = "orange",
 nice.breaks = FALSE,
  show.limits=TRUE,
na.value = NA,
  name = "# PINs"
)

nbh_sum %>% 
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>% 
  ggplot(aes(fill = nbh_current_rate)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Composite Tax Rate in Assessor Neighborhoods") +
  theme_void() + 
   theme(axis.ticks = element_blank(), axis.text = element_blank()) +
scale_fill_steps2(
  high = "darkblue", low = "orange",
  midpoint = median(nbh_sum$nbh_current_rate, na.rm = TRUE),
 nice.breaks = FALSE,
  show.limits=TRUE,
na.value = "gray",
  name = "Tax Rate",
  labels = scales::percent
) 
```

```{r}
nbh_MC_sum %>%
    group_by(nbhd_code) %>%
    summarize(pct_taxable_EAV = nbh_nonTIF_EAV_post_exemps /Total_EAV
  ) %>%
 arrange(pct_taxable_EAV) %>%
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>% 
  ggplot(aes(fill = pct_taxable_EAV)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Percent of EAV that is taxable by taxing districts",
       caption = "Non-taxable EAV includes TIF EAV above the frozen EAV 
       and tax exempt EAV due to various exemptions") +
  theme_void() + 
   theme(axis.ticks = element_blank(), axis.text = element_blank()) +
scale_fill_steps2(
  high = "darkblue", low = "white",
# midpoint = median(pct_taxable_EAV, na.rm = TRUE),
 nice.breaks = FALSE,
  show.limits=TRUE,
na.value = "gray",
  name = "Percent EAV taxable",
  labels = scales::percent
) 
```


### Townships

```{r townships, eval=FALSE}
nbh_sum <- nbh_sum %>% 
  mutate(nbhd_code = as.character(nbhd_code),
    township_code = str_sub(nbhd_code, 1, 2))

# table(nbh_sum$township_code)


nbh_sum <- nbh_sum %>% 
  mutate(township_code2 =
           case_when(
             township_code == 70 ~ 11,  # Hyde Park
             township_code == 71 ~ 44, # Jefferson
             township_code == 72 ~ 38, #Lake
             township_code == 73 ~ 24, #Lake View
             township_code == 74 ~ 27, #North Chicago
             township_code == 75 ~ 5, # Rogers Park
             township_code == 76 ~ 29, # South Chicago
             township_code == 77 ~ 31, # West Chicago
           ))
#table(nbh_sum$township_code2)

nbh_sum <- nbh_sum %>% 
  mutate(township_code2 = ifelse(is.na(township_code2), township_code, township_code2))
         
#table(nbh_sum$township_code2)

# nbh_sum <- nbh_sum %>% 
#   mutate(township_code2 = 
#    case_when(
#     township_code == 11~70,  # Hyde Park
#     township_code == 44~71,  # Jefferson
#     township_code == 38~72,  #Lake
#     township_code == 24~73, #Lake View
#     township_code == 27~74, #North Chicago
#     township_code == 5~75, # Rogers Park
#     township_code == 29~76, # South Chicago
#     township_code == 31~77 # West Chicago
# ))

township_shp <- read_sf("https://gis.cookcountyil.gov/traditional/rest/services/politicalBoundary/MapServer/3/query?outFields=*&where=1%3D1&f=geojson") %>%
  mutate(ORIGOID = as.character(ORIGOID)) %>% left_join(nicknames, by = c("ORIGOID" = "ORIGOID"))



nbh_sum %>% 
  group_by(township_code2) %>%
  summarize(abatements = sum(nbh_exe_abate))%>%
  left_join(township_shp, 
            by = c("township_code2" = "ORIGOID")
            ) %>% 
  ggplot(aes(fill = abatements)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Composite Tax Rate in Assessor Neighborhoods") +
  theme_void() + 
   theme(axis.ticks = element_blank(), axis.text = element_blank()) +
scale_fill_steps2(
  high = "darkblue", low = "orange",
 # midpoint = median(nbh_sum$abatements, na.rm = TRUE),
 nice.breaks = FALSE,
  show.limits=TRUE,
 na.value = "gray",
  name = "Exempt EAV \nfrom Abatements",
  labels = scales::dollar
) 
```

