---
title: "Mapping Property Tax Burden Shift due to Exemptions"
subtitle: "Cook County School Districts"
author: "Alea Wilbur"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    df_print: paged
    code_folding: hide
    code_download: true
---

# Getting Data

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)


library(tidyverse)
library(DBI)
library(data.table)
library(ggspatial)
library(gstat)
library(here)
library(httr)
library(jsonlite)
library(ptaxsim)
library(sf)
library(stars)
library(glue)

# Create the DB connection with the default name expected by PTAXSIM functions
ptaxsim_db_conn <- DBI::dbConnect(RSQLite::SQLite(), "./ptaxsim.db/ptaxsim-2021.0.4.db")


options(digits=4, scipen = 999)

library(sf)
library(jsonlite)
library(httr)

# link to the API output as a JSON file
muni_shp <- read_sf("https://gis.cookcountyil.gov/traditional/rest/services/politicalBoundary/MapServer/2/query?outFields=*&where=1%3D1&f=geojson")

# nicknames <- readxl::read_excel("muni_shortnames.xlsx")


elem_schools_shp <- read_sf("https://gis.cookcountyil.gov/traditional/rest/services/clerkTaxDistricts/MapServer/2/query?outFields=*&where=1%3D1&f=geojson") %>%
  mutate(agency_num = str_pad(AGENCY, width = "9" ,side = "left", pad = "0"))


```



**Data Prep**
```{r}



base_url <- "https://datacatalog.cookcountyil.gov/resource/tx2p-k2g9.json"

# Grab all PINs with township variables

pins_2021 <- GET(
  base_url,
  query = list(
    year = 2021,
    `$select` = paste0(c("pin", "class", "township_code", "township_name",
                         "nbhd_code", "census_puma_geoid",
                         
                         "school_elementary_district_name",
                         "school_elementary_district_geoid"),
                        # "ward_num", 
                        # "ward_chicago_2021",
                       #  "chicago_community_area_name", "chicago_community_area_num",
                       #  "tax_cook_municipality_name", "tax_cook_municipality_number"
                     #  ),
   collapse = ","),
    `$limit` = 500000000L
  )
)

pins_2021 <- fromJSON(rawToChar(pins_2021$content))

```

Pull all agency names that exist, then use agency numbers associated with SCHOOL types to pull only the `school_agency_names` object.

There are 946 unique taxing agencies that existed in 2021.

```{r}
joined_pins <- read_csv("./Output/4C_joined_PINs_bills_and_exemptions.csv")

joined_pins <- left_join(joined_pins, pins_2021, by = c("pin"))
```


# Proportion of Residential Land in School Districts {.tabset .tabset-pills}



__Exemptions within each Elementary School District:__


Percent Residential is the Residential EAV outside of TIFs / School District EAV outside of TIFs.

```{r pct-exempt}
pct_exempt_MC <- joined_pins %>% 
  group_by(school_elementary_district_name, school_elementary_district_geoid) %>%
  mutate(total_eav = sum(eav), # eav from pin table. "original" eav before exemptions or tifs
         nonTIF_EAV_post_exemps = sum(final_tax_to_dist/(tax_code_rate/100)),
        # nonTIF_EAV_pre_exemps = sum(final_tax_to_dist +  )
         pin_count  = n(),
         township_exemptEAV = sum(all_exemptions)) %>%
  ungroup() %>%
  group_by(school_elementary_district_name, school_elementary_district_geoid, major_class_code, total_eav, pin_count, township_exemptEAV, nonTIF_EAV_post_exemps) %>%
  summarize(MC_eav =  sum(eav),
            mc_pc = n(),
            MC_exemptions = sum(all_exemptions),
            MC_nonTIF_EAV_post_exemps = sum(final_tax_to_dist / (tax_code_rate/100) ),
            
) %>%
  mutate(
         pct_eav_MC = MC_eav / total_eav,
       #  pct_pins_w_exe = mc_pc / pin_count,
        exemps_per_resPIN =  MC_exemptions / mc_pc)

pct_exempt_MC


pct_exempt_MC %>% ungroup() %>% anti_join(elem_schools_shp, by = c( "school_elementary_district_name" = "AGENCY_DESCRIPTION") ) %>% distinct(school_elementary_district_name)



pct_exempt_MC %>% 
  filter(major_class_code == 2) %>%
  full_join(elem_schools_shp, by = c( "school_elementary_district_name" = "AGENCY_DESCRIPTION") ) %>%
  ggplot(aes(fill = exemps_per_resPIN)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Average Exempt EAV for Residential PINs in Elementary School Dist", 
     #  caption = "The median township has 67.2% of its EAV from Class 2 Residential Properties"
     ) +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+# +#+
  scale_fill_steps2(
    high = "darkblue", low = "black",  #  guide = "legend",
    #midpoint = median(pct_residential$pct_eav_MC),
    show.limits=TRUE,
    na.value = NA,
    name = "Exempt EAV",
    labels = scales::dollar)
```

Chicago has 353 elementary school districts...? Woah. 

```{r}
pct_eav_MC <- joined_pins %>% 
  group_by(school_elementary_district_name, school_elementary_district_geoid) %>%
  mutate(total_eav =  sum(eav)) %>%
  ungroup() %>%
  group_by(school_elementary_district_name, school_elementary_district_geoid, major_class_code, total_eav) %>%
  summarize(MC_eav =  sum(eav)) %>%
  mutate(
         pct_eav_MC = MC_eav / total_eav )

pct_residential <- pct_eav_MC %>% filter(major_class_code == 2)
  
pct_eav_MC %>% 
  filter(major_class_code == 2) %>%
  full_join(elem_schools_shp, by = c( "school_elementary_district_name" = "AGENCY_DESCRIPTION") ) %>%
  ggplot(aes(fill = pct_eav_MC)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Residential EAV /  Total EAV in Elem. District", 
       #caption = "The median township has 67.2% of its EAV from Class 2 Residential Properties"
       ) +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+# +#+
  scale_fill_steps2(
    high = "darkblue", low = "black",  #  guide = "legend",
   # midpoint = median(pct_residential$pct_eav_MC),
    show.limits=TRUE,
    na.value = NA,
    name = "% Residential",
    labels = scales::percent)
```






