---
title: "Tax_bill() no exemptions simulation"
author: "Alea Wilbur"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)


library(tidyverse)
library(DBI)
library(data.table)
library(ggspatial)
library(gstat)
library(here)
library(httr)
library(jsonlite)
library(ptaxsim)
library(sf)
library(stars)
library(glue)



library(sf)
library(jsonlite)
library(httr)

# link to the API output as a JSON file
muni_shp <- read_sf("https://gis.cookcountyil.gov/traditional/rest/services/politicalBoundary/MapServer/2/query?outFields=*&where=1%3D1&f=geojson")




# Create the DB connection with the default name expected by PTAXSIM functions
ptaxsim_db_conn <- DBI::dbConnect(RSQLite::SQLite(), "./ptaxsim.db/ptaxsim-2021.0.4.db")

# has all potential property classes for pins
# downloaded from CCAO gitlab website
## I used this to merge additional information to the pins and class data later on.
class_dict <- read_csv("class_dict.csv")

options(digits=4, scipen = 999)

```

```{r agency-dt}
# has EAV values, extensions by agency_num
agency_dt <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT *
  FROM agency
  WHERE year = 2021
  "
)

# grabs all unique muni names. Would be needed if creating a loop for calculating all munis
# municipality names and their agency number
muni_agency_names <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT DISTINCT agency_num, agency_name, minor_type
  FROM agency_info
  WHERE minor_type = 'MUNI'
  OR agency_num = '020060000'  

  "
)

muni_agency_names <- muni_agency_names %>% 
  select(-minor_type)



#Makes a list of ALL taxing agencies, including TIFs, SSAs, etc.

# all agency names, numbers, and types
# includes TIF and non-TIF agencies
all_taxing_agencies <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT agency_num, agency_name, major_type, minor_type
  FROM agency_info
  "
  )


muni_agency_nums<- all_taxing_agencies %>% 
  filter(minor_type %in% c("MUNI")) %>%
   select(agency_num)

muni_tax_codes <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql("
  SELECT*
  FROM tax_code
  WHERE agency_num IN ({muni_agency_names$agency_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
)# %>% select(-agency_rate)

# Agency number and agency name for all TIFs
TIF_agencies <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT DISTINCT agency_num, agency_name, major_type, minor_type
  FROM agency_info
  WHERE minor_type = 'TIF'
  "
)

unique_tif_taxcodes <- DBI::dbGetQuery(
  ptaxsim_db_conn, 
  glue_sql("
  SELECT DISTINCT tax_code_num
  FROM tax_code
  WHERE agency_num IN ({TIF_agencies$agency_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
)



## Read in summarized tax code level data for exemptions and taxbills.

#taxbills_by_Class_per_TC <- read_csv("taxbills_inMunis_perTC.csv")  %>% 
#  mutate(tax_code = as.character(tax_code),
    #      tax_code_frozen_eav = replace_na(tax_code_frozen_eav, 0), 
    #      tax_code_tif_increment = ifelse((tax_code_frozen_eav > tax_code_eav), 0, (tax_code_eav-tax_code_frozen_eav)),
 #   )
       #  tax_code_tif_increment = tax_code_eav-tax_code_frozen_eav) %>%
  #select(-c(year:tax_code_rate, tax_code_eav, tax_code_revenue, tax_code_distribution_pct))
  


# get rid of rpm variables, tax_amt_exe,pre and post exemption variables,
#exemptions_by_class_per_TC <- read_csv("all_exemptions_by_TC.csv") %>%
#    mutate(tax_code_num = as.character(tax_code_num))

tif_distrib <- DBI::dbGetQuery(
  ptaxsim_db_conn, 
  glue_sql("
  SELECT *
  FROM tif_distribution
  WHERE tax_code_num IN ({muni_tax_codes$tax_code_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
) %>% mutate(tax_code_num = as.character(tax_code_num))



## Add agency names since those weren't included originally in the table:

# all_taxing_agencies <- all_taxing_agencies %>% 
#   left_join(muni_agency_names, by = c("first5", "first6")) %>% 
#   rename(muni_name =  agency_name.y,
#         muni_num = agency_num.y,
#         agency_name = agency_name.x,
#         agency_num = agency_num.x)
# 


# # combine taxing agency names and agency type to data table that has eav and extension values
# agency_data <- right_join(agency_dt, all_taxing_agencies) %>% 
#   # get rid of unneeded columns to make table outputs smaller
#   select(-c(cty_dupage_eav:cty_livingston_eav, lim_numerator, lim_denominator)) %>% # drop some of the unused variables
#   arrange(agency_num)
```


```{r eval=FALSE}
muni_pins <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql(
  "SELECT DISTINCT pin, class, tax_code_num
  FROM pin
  WHERE tax_code_num IN ({muni_tax_codes$tax_code_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
))

#all_muni_pins <-read_csv("all_muni_pins.csv")

# Normal output from lookup_pin() command. Includes all types of exemptions
exe_dt <- lookup_pin(2021, muni_pins$pin) %>%
  setDT(key = c("year", "pin"))




taxbills_current <- tax_bill(2021, 
                  muni_pins$pin, 
                  pin_dt = exe_dt, # default option
                  simplify = FALSE)


head(taxbills_current)

TC_bills_current <- taxbills_current %>%
  left_join(class_dict, by = c("class" = "class_code")) %>% # add major property types to pin data
  group_by(major_class_code, major_class_type, tax_code) %>%
  summarize(
            final_tax_to_dist = sum(final_tax_to_dist),
            final_tax_to_tif = sum(final_tax_to_tif),
            tax_amt_exe = sum(tax_amt_exe), # revenue lost due to exemptions
            tax_amt_pre_exe = sum(tax_amt_pre_exe), # total rev before all exemptions
            tax_amt_post_exe = sum(tax_amt_post_exe), # total rev after all exemptions
        rpm_tif_to_cps = sum(rpm_tif_to_cps), # not used
        rpm_tif_to_rpm = sum(rpm_tif_to_rpm), # not used
         rpm_tif_to_dist = sum(rpm_tif_to_dist), # not used
          tif_share = mean(tif_share), # not used
       ) %>%
  mutate(stage = "With exemptions")

TC_bills_current

write_csv(TC_bills_current, "TC_bills_current.csv")




# Changes all exemptions to 0 for all pins. 
# This table will then go INSIDE of the taxbill() comand
# for no exemptions simulation
no_exe_dt <- lookup_pin(2021, muni_pins$pin)# %>%
 # mutate(across(starts_with("exe_"), ~0)) %>%
 # setDT(key = c("year", "pin"))


no_exe_dt[, tax_code := lookup_tax_code(year, pin)]

exe_cols <- names(no_exe_dt)[startsWith(names(no_exe_dt), "exe_")]
taxcode_sum_no_exe <- no_exe_dt[,
  .(exe_total = sum(rowSums(.SD))),
  .SDcols = exe_cols,
  by = .(year, tax_code)
]

# recalculate the base
# change agency data table and their tax rate

t_agency_dt_no_exe <- lookup_agency(2021, no_exe_dt$tax_code)
t_agency_dt_no_exe[
  taxcode_sum_no_exe,
  on = .(year, tax_code),
  agency_total_eav := agency_total_eav + exe_total
]

no_exe_dt[, (exe_cols) := 0][, c("tax_code") := NULL]


taxbills_no_exemps <- tax_bill(2021, 
                  pin_vec = muni_pins$pin,
                  agency_dt = t_agency_dt_no_exe,
                  pin_dt = no_exe_dt,
                  simplify = FALSE
)[
  , stage := "No exemptions"
]

head(taxbills_no_exemps)

class_dict$class_code <- as.character(class_dict$class_code)



TC_bills_noexemps <- taxbills_no_exemps %>%
  left_join(class_dict, by = c("class" = "class_code")) %>% # add major property types to pin data
  group_by(major_class_code, major_class_type, tax_code) %>%
  summarize(
            final_tax_to_dist = sum(final_tax_to_dist),
            final_tax_to_tif = sum(final_tax_to_tif),
            tax_amt_exe = sum(tax_amt_exe), # revenue lost due to exemptions
            tax_amt_pre_exe = sum(tax_amt_pre_exe), # total rev before all exemptions
            tax_amt_post_exe = sum(tax_amt_post_exe), # total rev after all exemptions
        rpm_tif_to_cps = sum(rpm_tif_to_cps), # not used
        rpm_tif_to_rpm = sum(rpm_tif_to_rpm), # not used
         rpm_tif_to_dist = sum(rpm_tif_to_dist), # not used
          tif_share = mean(tif_share), # not used
       )

TC_bills_noexemps

write_csv(TC_bills_noexemps, "TC_bills_noexemps.csv")

rm(taxbills_no_exemps)

```

# Muni Land Use 
```{r}
TC_bills_current <- read_csv("TC_bills_current.csv") %>% mutate(tax_code = as.character(tax_code))
class_dict$class_code <- as.character(class_dict$class_code)
 

taxcodes_current <- full_join(TC_bills_current, muni_tax_codes, 
                      by = c("tax_code" = "tax_code_num")) 

Current_Taxrates <- taxcodes_current %>% 
  left_join(muni_agency_names) %>%
  group_by(agency_name) %>% 
  summarize(MuniLevy = sum(final_tax_to_dist, na.rm = TRUE), #amount billed by muni within cook county
            nonTIF_EAV_post_exemps = sum(final_tax_to_dist/(tax_code_rate/100), na.rm = TRUE), # within cook county
            TIF_increment_EAV = sum(final_tax_to_tif/(tax_code_rate/100), na.rm=TRUE),  # within cook county
            Exempt_EAV = sum(tax_amt_exe/(tax_code_rate/100), na.rm=TRUE), # within cook county
            Total_EAV = sum((tax_amt_exe+final_tax_to_dist+final_tax_to_tif)/(tax_code_rate/100), na.rm = TRUE))%>%

  mutate(tax_rate_current = MuniLevy/nonTIF_EAV_post_exemps,
         nonTIF_EAV_pre_exemps = nonTIF_EAV_post_exemps + Exempt_EAV,
         taxrate_new = MuniLevy/nonTIF_EAV_pre_exemps) %>% 
  select(agency_name, tax_rate_current, taxrate_new, everything()) %>% 
  arrange(desc(tax_rate_current))

land_use <- taxcodes_current %>% 
  left_join(muni_agency_names) %>%
  mutate( ResidentialProps = ifelse(major_class_code %in% c("2", "3", "9"), "Residential", "Commercial"),
         PropType = case_when(
           major_class_code %in% c("3","9") ~ "Multi-Family",
           major_class_code == "2" ~ "Single-Family",
           TRUE ~ "Commercial-Industrial")) %>%
  group_by(agency_name, ResidentialProps, agency_num) %>% 
  
  # All of the values calculated below are AFTER exemptions have been removed
  summarize(taxrev_from_proptype = sum(final_tax_to_dist, na.rm = TRUE),
            nonTIF_EAV = sum(final_tax_to_dist/(tax_code_rate/100), na.rm = TRUE),
            TIF_increment_EAV = sum(final_tax_to_tif/(tax_code_rate/100), na.rm=TRUE),
            Exempt_EAV = sum(tax_amt_exe/(tax_code_rate/100), na.rm=TRUE),
            Total_EAV = sum((tax_amt_exe+final_tax_to_dist+final_tax_to_tif)/(tax_code_rate/100), na.rm = TRUE) )

land_use

muni_eav <- land_use %>% group_by(agency_name, agency_num) %>% summarize(muni_EAV = sum(Total_EAV)) %>% ungroup()

perc_residential <- left_join(land_use, muni_eav) %>% 
  filter(ResidentialProps == "Residential") %>% 
  mutate(percent_residential = Total_EAV / muni_EAV) %>% 
  select(agency_name, percent_residential, agency_num) 

#lowest % of eav from residential property types
perc_residential %>%
  arrange(percent_residential) %>% head()

#highest % of eav from residential property types
perc_residential %>% 
  arrange(percent_residential) %>% tail()


perc_residential %>%
  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%
  ggplot(aes(fill = percent_residential)) + 
  geom_sf(aes(geometry = geometry), color = "black") + theme_void()+ 
  labs(title = "Residential EAV /  Total EAV") +
    theme_classic() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
   scale_fill_gradientn(colors = c("white", "darkblue"), 
                    #   limits = c(0,.6),
                       n = 5,
                        name = "% Residential", label = scales::percent)
```


# Tax Rates

```{r}
TC_bills_noexemps <- read_csv("TC_bills_noexemps.csv") %>% mutate(tax_code = as.character(tax_code))
# noexemps_rev<- taxcodes_noexemps  %>%   
#   left_join(muni_agency_names) %>%
#   select(tax_code, major_class_code, major_class_type, final_tax_to_dist, agency_name) %>% 
#   rename(final_tax_to_dist_noexemps = final_tax_to_dist)

taxcodes_noexemps <- full_join(TC_bills_noexemps, muni_tax_codes, 
                      by = c("tax_code" = "tax_code_num")) 


noexemps_munisummary <- taxcodes_noexemps %>% 
  left_join(muni_agency_names) %>%
  group_by(agency_name) %>% 
  summarize(MuniLevy = sum(final_tax_to_dist, na.rm = TRUE), #amount billed by muni within cook county
            nonTIF_EAV = sum(final_tax_to_dist/(tax_code_rate/100), na.rm = TRUE), # within cook county
           final_tax_to_dist = sum(final_tax_to_dist, na.rm=TRUE),
     TIF_increment_EAV = sum(final_tax_to_tif/(tax_code_rate/100), na.rm=TRUE),  # within cook county
            Exempt_EAV = sum(tax_amt_exe/(tax_code_rate/100), na.rm=TRUE), # within cook county
            Total_EAV = sum((tax_amt_exe+final_tax_to_dist+final_tax_to_tif)/(tax_code_rate/100), na.rm = TRUE)) %>%

  mutate(taxrate_noexemps = MuniLevy/nonTIF_EAV,
        # nonTIF_EAV= nonTIF_EAV + Exempt_EAV,
       ) 
 # select(agency_name, tax_rate_noexemps, everything()) %>% 
#  arrange(desc(tax_rate_noexemps))




```

6 Highest and 6 lowest tax rates. View the extremes on both sides

```{r}

Current_Taxrates %>%  filter(tax_rate_current > 0.25 )
                           
Current_Taxrates %>%  filter( tax_rate_current < 0.08)
```


```{r}
burden_with_exemptions <- taxcodes_current %>% 
 # left_join(muni_tax_codes) %>% 
  left_join(muni_agency_names) %>%   
  left_join(Current_Taxrates) %>% 
#  left_join(noexemps_rev) %>%
  select(-c(agency_rate)) %>%
  
  mutate(         ResidentialProps = ifelse(major_class_code %in% c("2", "3", "9"), "Residential", "Commercial"),
         
         PropType = case_when(
           major_class_code %in% c("3","9") ~ "Multi-Family",
           major_class_code == "2" ~ "Single-Family",
           TRUE~ "Commercial-Industrial")) %>% 
  arrange(tax_code) %>%

  group_by(agency_name, major_class_type, tax_rate_current, taxrate_new, PropType, ResidentialProps, agency_num, major_class_code) %>%
  summarize(final_tax_to_dist = sum(final_tax_to_dist, na.rm=TRUE),
           # final_tax_to_dist_noexemps = sum(final_tax_to_dist_noexemps, na.rm=TRUE),
            MuniLevy = first(MuniLevy)) %>%
  mutate(current_burden = final_tax_to_dist/MuniLevy) %>%
           # no_exemptions_burden = sum(no_exemptions_burden, na.rm = TRUE)) %>%
  select(agency_name, major_class_type, current_burden, tax_rate_current, taxrate_new, everything()) %>% ungroup()

# find outliers
burden_with_exemptions %>% arrange(desc(current_burden))

```

```{r}
burden_noexemps<- taxcodes_noexemps %>% 
 # left_join(muni_tax_codes) %>% 
  left_join(muni_agency_names) %>%   
  left_join(noexemps_munisummary) %>% 
#  left_join(noexemps_rev) %>%
  select(-c(agency_rate)) %>%
  
  mutate(ResidentialProps = ifelse(major_class_code %in% c("2", "3", "9"), "Residential", "Commercial"),
         
         PropType = case_when(
           major_class_code %in% c("3","9") ~ "Multi-Family",
           major_class_code == "2" ~ "Single-Family",
           TRUE~ "Commercial-Industrial")) %>% 
  arrange(tax_code) %>%

  group_by(agency_name, major_class_type, taxrate_noexemps, PropType, ResidentialProps, agency_num, major_class_code) %>%
  summarize(final_tax_to_dist_noexemps = sum(final_tax_to_dist, na.rm=TRUE),
           # final_tax_to_dist_noexemps = sum(final_tax_to_dist_noexemps, na.rm=TRUE),
           # MuniLevy = first(MuniLevy)
           ) %>%
#  mutate(burden_noexemps = final_tax_to_dist/MuniLevy) %>%
           # no_exemptions_burden = sum(no_exemptions_burden, na.rm = TRUE)) %>%
  select(agency_name, major_class_type,
      #   burden_noexemps,
         taxrate_noexemps, everything()) %>% ungroup()


#burden_noexemps %>% arrange(desc(burden_noexemps))

burden_shift <- full_join( burden_with_exemptions, burden_noexemps, 
                           by = c("agency_name", "major_class_type", "PropType", "ResidentialProps", "major_class_code", "agency_num")) %>%
  mutate(burden_noexemps = final_tax_to_dist_noexemps/MuniLevy,
       burden_change = current_burden - burden_noexemps,
                                                         taxrate_change = tax_rate_current - taxrate_new)


#burden_shift
```

Residential Burden Shift: 

```{r}
burden_shift %>% filter(major_class_code == 2)  %>% 
  select(agency_name, burden_change, taxrate_change, tax_rate_current, taxrate_new, taxrate_noexemps) %>% arrange(burden_change)




burden_shift_table <- burden_shift %>% 
  mutate(burden_change = round(current_burden - burden_noexemps, digits =5),
         taxrate_change = tax_rate_current - taxrate_new) %>% 
  select(agency_name, major_class_type, burden_change, taxrate_change, tax_rate_current, taxrate_new, PropType, ResidentialProps) %>% 
  arrange(burden_change)


#burden_shift_table %>% write.csv("burden_shift_July8.csv")

```

# Map of Tax Rate Change


```{r}


burden_shift %>% 
    filter(major_class_code == 2) %>% # all property types have same composite tax rate
  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%
  ggplot(aes(fill = tax_rate_current)) + 
      theme_classic() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
   scale_fill_gradientn(colors = c("white", "maroon"), 
                      limits = c(0,.32),
                       n = 6,
                        name = "Tax Rate", label = scales::percent)+
  geom_sf(aes(geometry = geometry), color = "black") +  
  labs(title = "Current composite tax rates with current exemptions in place" ,    
         caption = "Highest composite tax rate is in Park Forest (41.4%.)
       Lowest composite tax rate is in Oak Brook (6.3%) and Chicago (6.7%).")
```

```{r}
# as a dot graph ## 

order <- burden_shift %>% ungroup %>% as_tibble() %>%
# filter(YEAR == 2019 & NAME != "NA" & CIHISPEED == "10") %>% 
  select(agency_name, tax_rate_current)



# look at ones that changed the most
burden_shift %>%  
  filter(tax_rate_current > 0.23 |tax_rate_current < 0.085) %>% 
  filter(PropType == "Single-Family") %>%
  ungroup() %>% 
  select(agency_name, tax_rate_current, taxrate_new) %>% 
  pivot_longer(c("tax_rate_current", "taxrate_new"), 
               names_to = "type", values_to = "tax_rate") %>% 
  left_join(order) %>%
  ggplot(aes(x = tax_rate*100, y= reorder(agency_name, tax_rate_current)))+
  geom_line(aes(group = agency_name))+ 
   geom_point(aes(color=type), size=3 )+
  theme_minimal() + 
  theme( 
    legend.title = element_blank(),
               plot.title.position = "plot",
     #   panel.background = element_rect(fill='transparent'), #transparent panel bg
    plot.background = element_rect(fill='transparent', color=NA) #transparent plot bg
   )+
        scale_color_brewer(palette="Paired", labels = c("Exemptions", "No Exemptions"), direction = 1)+

  labs(title = "Change in Composite Tax Rate if there were No Exemptions", x = "Composite Tax Rate (%)", y = "" , caption = "For the highest and lowest composite tax rates")
```

Table of composite tax rate change:

```{r}



burden_shift %>% 
  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%
  ggplot(aes(fill = taxrate_change*100)) + 
  geom_sf(aes(geometry = geometry), color = "black") +  
      theme_classic() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
   scale_fill_gradientn(colors = c("white", "maroon"), 
                    #   limits = c(0,.6),
                       n = 7,
                        name = "Percentage Point Change")+
  labs(title = "Change in Composite Tax Rate if Exemptions are Removed")

```


# Single-Family (Class 2) Burden Shift


Change in Tax Burden for Class 2 Properties:

```{r}
burden_shift %>%
      filter(major_class_code == 2) %>%

  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%

  ggplot(aes(fill =current_burden)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  theme_classic() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
   scale_fill_gradientn(colors = c("white", "darkblue"), 
                    #   limits = c(0,.6),
                       n = 6,
                        name = "Burden with \nExemptions", label = scales::percent)+
  labs(title = "Current share of property tax burden", subtitle = "for Class = 2 Property Types")


burden_shift %>% 
      filter(major_class_code == 2) %>%

  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%

  ggplot(aes(fill = burden_noexemps)) + 
  geom_sf(aes(geometry = geometry), color = "black") +     
  theme_classic() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
   scale_fill_gradientn(colors = c("white", "darkblue"), 
                    #   limits = c(0,.6),
                       n = 6,
                        name = "Burden w/o \nExemptions", label = scales::percent) +
  labs(title = "New share of property tax burden", subtitle = "for Class = 2 Property Types")
```



```{r}
burden_shift %>% 
  filter(major_class_code == 2) %>%
  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%
  ggplot(aes(fill = burden_change)) + 
  geom_sf(aes(geometry = geometry), color = "black") + theme_void()+ 
  labs(title = "Change in Residential Share of Tax Burden") +
    theme_classic() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
   scale_fill_gradientn(colors = c("darkblue","white"), 
                    #   limits = c(0,.6),
                       n = 6,
                        name = "Change in Burden", label = scales::percent)
  
   # geom_sf(data = countyIL, fill=NA, color="dark gray")

```

```{r collapse=TRUE}
# as a dot graph ## 

order <- burden_shift %>% 
  ungroup %>% as_tibble() %>%
  #  filter(ResidentialProps == "Residential") %>%

  filter(PropType == "Single-Family") %>%
  select(agency_name, current_burden)



# burder_shift_ordered <-  burden_shift %>% 
#   ungroup() %>% 
#   select(agency_name, current_burden, no_exemptions_burden) %>%    
#   pivot_longer(c("current_burden", "no_exemptions_burden"), 
#                names_to = "type", values_to = "pct_burden") %>% 
#   left_join(order)

# look at ones that changed the most
burden_shift %>%  
  filter(current_burden > 0.9 |current_burden < .3 ) %>% 
  #filter(ResidentialProps == "Residential") %>%
    filter(PropType == "Single-Family") %>%

  ungroup() %>% 
  select(agency_name, current_burden, burden_noexemps) %>% 
  pivot_longer(c("current_burden", "burden_noexemps"), 
               names_to = "type", values_to = "pct_burden") %>% 
  inner_join(order) %>%
  ggplot(aes(x = pct_burden, y= reorder(agency_name, current_burden)))+
  geom_line(aes(group = agency_name))+ 
   geom_point(aes(color=type), size=3 )+
  theme_minimal() + 
  theme(#legend.position = "none", 
    legend.title = element_blank(),
               plot.title.position = "plot",
     #   panel.background = element_rect(fill='transparent'), #transparent panel bg
    plot.background = element_rect(fill='transparent', color=NA) #transparent plot bg
   )+
        scale_color_brewer(palette="Paired", labels = c("Exemptions", "No Exemptions"), direction = 1)+

  labs(title = "Change in Single-family Residential Tax Burden", x = "Share of Levy (%)", y = "" , caption = "Residential Tax Burden is the Share of the property tax collected that was paid for by 
     single-family home owners in property classes 2.") #+
# scale_x_continuous(label = scales::percent)

```

# Summary table of Burden Shift: 3 Property Categories

Same data that is used above but property types are combined into 3 categories:

Single-family (Class 2),   
Multi-Family (Class 3 & 9),   
and Commercial-Industrial (all other Classes: 1,4, 5, 6, 7, 8, 9)  

Current Burden:

```{r}
munis_3property_types <- burden_shift %>%
  ungroup() %>% 
  group_by(agency_name, PropType) %>%
  summarize(final_tax_to_dist = sum(final_tax_to_dist, na.rm = TRUE),
            current_burden = sum(current_burden, na.rm = TRUE),
            burden_noexemps = sum(burden_noexemps, na.rm = TRUE),
            burden_change = sum(burden_change, na.rm = TRUE)) 

# munis_3property_types

munis_3property_types %>% pivot_wider( id_cols = agency_name , names_from = "PropType", values_from = "current_burden") %>% arrange(-`Single-Family`)
```

__Tax Burden if there were no exemptions:__

```{r}

munis_3property_types %>% pivot_wider( id_cols = agency_name , names_from = "PropType", values_from = "burden_noexemps") %>% arrange(-`Single-Family`)
```

__Change in Share of Burden if there were no exemptions:__

```{r}
# burden_shift %>% ungroup() %>% 
#   group_by(agency_name, PropType) %>%
#   summarize(district_rev_collected = sum(district_rev_collected),
#             current_burden = sum(current_burden),
#             no_exemptions_burden = sum(no_exemptions_burden)) 

props_wide <- munis_3property_types %>%
  pivot_wider(id_cols = agency_name , 
               names_from = "PropType", 
               values_from = "burden_change") %>% 
  arrange(-`Single-Family`) 

props_wide[2:4] <- sapply(props_wide[2:4], function(x) scales::percent(x, accuracy=.001))
props_wide
```

