---
title: "Tax_bill() no exemptions simulation"
author: "Alea Wilbur"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)


library(tidyverse)
library(DBI)
library(data.table)
library(ggspatial)
library(gstat)
library(here)
library(httr)
library(jsonlite)
library(ptaxsim)
library(sf)
library(stars)
library(glue)
library(cmapplot)


library(sf)
library(jsonlite)
library(httr)

# link to the API output as a JSON file
muni_shp <- read_sf("https://gis.cookcountyil.gov/traditional/rest/services/politicalBoundary/MapServer/2/query?outFields=*&where=1%3D1&f=geojson")




# Create the DB connection with the default name expected by PTAXSIM functions
ptaxsim_db_conn <- DBI::dbConnect(RSQLite::SQLite(), "./ptaxsim.db/ptaxsim-2021.0.4.db")

# has all potential property classes for pins
# downloaded from CCAO gitlab website
## I used this to merge additional information to the pins and class data later on.
class_dict <- read_csv("class_dict.csv")

options(digits=4, scipen = 999)

```

```{r agency-dt}
# has EAV values, extensions by agency_num
agency_dt <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT *
  FROM agency
  WHERE year = 2021
  "
)

# grabs all unique muni names. Would be needed if creating a loop for calculating all munis
# municipality names and their agency number
muni_agency_names <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT DISTINCT agency_num, agency_name, minor_type
  FROM agency_info
  WHERE minor_type = 'MUNI'
  OR agency_num = '020060000'  

  "
)

muni_agency_names <- muni_agency_names %>% 
  select(-minor_type)



#Makes a list of ALL taxing agencies, including TIFs, SSAs, etc.

# all agency names, numbers, and types
# includes TIF and non-TIF agencies
all_taxing_agencies <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT agency_num, agency_name, major_type, minor_type
  FROM agency_info
  "
  )


muni_agency_nums<- all_taxing_agencies %>% 
  filter(minor_type %in% c("MUNI")) %>%
   select(agency_num)

muni_tax_codes <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql("
  SELECT*
  FROM tax_code
  WHERE agency_num IN ({muni_agency_names$agency_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
)# %>% select(-agency_rate)

# Agency number and agency name for all TIFs
TIF_agencies <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT DISTINCT agency_num, agency_name, major_type, minor_type
  FROM agency_info
  WHERE minor_type = 'TIF'
  "
)

unique_tif_taxcodes <- DBI::dbGetQuery(
  ptaxsim_db_conn, 
  glue_sql("
  SELECT DISTINCT tax_code_num
  FROM tax_code
  WHERE agency_num IN ({TIF_agencies$agency_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
)



## Read in summarized tax code level data for exemptions and taxbills.

#taxbills_by_Class_per_TC <- read_csv("taxbills_inMunis_perTC.csv")  %>% 
#  mutate(tax_code = as.character(tax_code),
    #      tax_code_frozen_eav = replace_na(tax_code_frozen_eav, 0), 
    #      tax_code_tif_increment = ifelse((tax_code_frozen_eav > tax_code_eav), 0, (tax_code_eav-tax_code_frozen_eav)),
 #   )
       #  tax_code_tif_increment = tax_code_eav-tax_code_frozen_eav) %>%
  #select(-c(year:tax_code_rate, tax_code_eav, tax_code_revenue, tax_code_distribution_pct))
  


# get rid of rpm variables, tax_amt_exe,pre and post exemption variables,
#exemptions_by_class_per_TC <- read_csv("all_exemptions_by_TC.csv") %>%
#    mutate(tax_code_num = as.character(tax_code_num))

tif_distrib <- DBI::dbGetQuery(
  ptaxsim_db_conn, 
  glue_sql("
  SELECT *
  FROM tif_distribution
  WHERE tax_code_num IN ({muni_tax_codes$tax_code_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
) %>% mutate(tax_code_num = as.character(tax_code_num))



## Add agency names since those weren't included originally in the table:

# all_taxing_agencies <- all_taxing_agencies %>% 
#   left_join(muni_agency_names, by = c("first5", "first6")) %>% 
#   rename(muni_name =  agency_name.y,
#         muni_num = agency_num.y,
#         agency_name = agency_name.x,
#         agency_num = agency_num.x)
# 


# # combine taxing agency names and agency type to data table that has eav and extension values
# agency_data <- right_join(agency_dt, all_taxing_agencies) %>% 
#   # get rid of unneeded columns to make table outputs smaller
#   select(-c(cty_dupage_eav:cty_livingston_eav, lim_numerator, lim_denominator)) %>% # drop some of the unused variables
#   arrange(agency_num)
```


```{r eval=FALSE}
muni_pins <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql(
  "SELECT DISTINCT pin, class, tax_code_num
  FROM pin
  WHERE tax_code_num IN ({muni_tax_codes$tax_code_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
))

#all_muni_pins <-read_csv("all_muni_pins.csv")

# Normal output from lookup_pin() command. Includes all types of exemptions
exe_dt <- lookup_pin(2021, muni_pins$pin) %>%
  setDT(key = c("year", "pin"))




taxbills_current <- tax_bill(2021, 
                  muni_pins$pin, 
                  pin_dt = exe_dt, # default option
                  simplify = FALSE)


head(taxbills_current)

TC_bills_current <- taxbills_current %>%
  left_join(class_dict, by = c("class" = "class_code")) %>% # add major property types to pin data
  group_by(major_class_code, major_class_type, tax_code) %>%
  summarize(
            final_tax_to_dist = sum(final_tax_to_dist),
            final_tax_to_tif = sum(final_tax_to_tif),
            tax_amt_exe = sum(tax_amt_exe), # revenue lost due to exemptions
            tax_amt_pre_exe = sum(tax_amt_pre_exe), # total rev before all exemptions
            tax_amt_post_exe = sum(tax_amt_post_exe), # total rev after all exemptions
        rpm_tif_to_cps = sum(rpm_tif_to_cps), # not used
        rpm_tif_to_rpm = sum(rpm_tif_to_rpm), # not used
         rpm_tif_to_dist = sum(rpm_tif_to_dist), # not used
          tif_share = mean(tif_share), # not used
       ) %>%
  mutate(stage = "With exemptions")

TC_bills_current

write_csv(TC_bills_current, "TC_bills_current.csv")




# Changes all exemptions to 0 for all pins. 
# This table will then go INSIDE of the taxbill() comand
# for no exemptions simulation
no_exe_dt <- lookup_pin(2021, muni_pins$pin)# %>%
 # mutate(across(starts_with("exe_"), ~0)) %>%
 # setDT(key = c("year", "pin"))


no_exe_dt[, tax_code := lookup_tax_code(year, pin)]

exe_cols <- names(no_exe_dt)[startsWith(names(no_exe_dt), "exe_")]
taxcode_sum_no_exe <- no_exe_dt[,
  .(exe_total = sum(rowSums(.SD))),
  .SDcols = exe_cols,
  by = .(year, tax_code)
]

# recalculate the base
# change agency data table and their tax rate

t_agency_dt_no_exe <- lookup_agency(2021, no_exe_dt$tax_code)
t_agency_dt_no_exe[
  taxcode_sum_no_exe,
  on = .(year, tax_code),
  agency_total_eav := agency_total_eav + exe_total
]

no_exe_dt[, (exe_cols) := 0][, c("tax_code") := NULL]


taxbills_no_exemps <- tax_bill(2021, 
                  pin_vec = muni_pins$pin,
                  agency_dt = t_agency_dt_no_exe,
                  pin_dt = no_exe_dt,
                  simplify = FALSE
)[
  , stage := "No exemptions"
]

head(taxbills_no_exemps)

class_dict$class_code <- as.character(class_dict$class_code)



TC_bills_noexemps <- taxbills_no_exemps %>%
  left_join(class_dict, by = c("class" = "class_code")) %>% # add major property types to pin data
  group_by(major_class_code, major_class_type, tax_code) %>%
  summarize(
            final_tax_to_dist = sum(final_tax_to_dist),
            final_tax_to_tif = sum(final_tax_to_tif),
            tax_amt_exe = sum(tax_amt_exe), # revenue lost due to exemptions
            tax_amt_pre_exe = sum(tax_amt_pre_exe), # total rev before all exemptions
            tax_amt_post_exe = sum(tax_amt_post_exe), # total rev after all exemptions
        rpm_tif_to_cps = sum(rpm_tif_to_cps), # not used
        rpm_tif_to_rpm = sum(rpm_tif_to_rpm), # not used
         rpm_tif_to_dist = sum(rpm_tif_to_dist), # not used
          tif_share = mean(tif_share), # not used
       )

TC_bills_noexemps

write_csv(TC_bills_noexemps, "TC_bills_noexemps.csv")

rm(taxbills_no_exemps)

```

# Muni Land Use 

## EAV 

using the pin data from ptaxsim, I calculate the eav of all properties, exempt EAV, the current taxbase, and taxbase if there were not exemptios


- exempt EAV (summed from all exemption types)

- current taxbase (using the `tif_distrib` table and percent of taxcode rev that goes to the tif). If a taxcode is a TIF taxcode, then do (EAV-exempt EAV) * (1-%rev that goes to TIF). If the tax code is not a TIF tax code, then use the EAV-exempt EAV.

- taxbase without exemptions: If a taxcode is a TIF taxcode, then do EAV * (1-%rev that goes to TIF). If the tax code is not a TIF tax code, then use the eav




```{r}
class_dict <- read_csv("class_dict.csv")

# use exemptions in tax codes to summarize EAV. 
# More accurate that calculating it from revenue collected.tax rate in tax bill data
exemptions_by_class_per_TC <- read_csv("all_exemptions_by_TC.csv") %>% 
  mutate(tax_code_num = as.character(tax_code_num)) %>%
  left_join(muni_tax_codes) %>%
  full_join(muni_agency_names) %>%
  # merge with TIF distrib table to calculate the percent of the EAV that goes to the TIF vs the district

  left_join(tif_distrib, by=c("tax_code_num", "year", "tax_code_rate")) %>%
  mutate(exempt_EAV = (exe_homeowner + exe_senior + exe_freeze + exe_longtime_homeowner + 
         exe_disabled + exe_vet_returning + exe_vet_dis + exe_abate) ,
         in_TIF = ifelse(tax_code_num %in% unique_tif_taxcodes$tax_code_num, 1, 0),
         tax_base_current = ifelse(in_TIF==1, (eav-exempt_EAV)*(1-tax_code_distribution_pct/100), eav-exempt_EAV),
         tax_base_noexemps = ifelse(in_TIF==1, (eav)*(1-tax_code_distribution_pct/100), eav),
         ResidentialProps = ifelse(major_class_code %in% c("2", "3", "9"), "Residential", "Commercial"),
         
         PropType = case_when(
           major_class_code %in% c("3","9") ~ "Multi-Family",
           major_class_code == "2" ~ "Single-Family",
           TRUE~ "Commercial-Industrial"))

grouped_exemptions <- exemptions_by_class_per_TC %>% 
   # group_by(agency_name, major_class_code, major_class_type, ResidentialProps, PropType) %>%
  group_by(agency_name, ResidentialProps) %>%
  summarize(eav = sum(eav),
           exempt_EAV = sum(exempt_EAV, exe_abate, na.rm=TRUE),
         tax_base_current = sum(tax_base_current, na.rm=TRUE),
         tax_base_noexemps = sum(tax_base_noexemps, na.rm=TRUE)) %>% ungroup()
  
grouped_exemptions

# calculate totals with ONLY EAV outside of TIFs
muni_eav <- grouped_exemptions %>%  
  group_by(agency_name) %>% 
  summarize(muni_EAV_includesTIF = sum(eav), # all EAV in the municipality that exists
            muni_tax_base_current=sum(tax_base_current), # taxable EAV based on current exemptions
            muni_tax_base_noexemps = sum(tax_base_noexemps)) %>%  # taxable EAV pre-exemptions
  ungroup() 

perc_residential <- full_join(grouped_exemptions, muni_eav) %>% 
  filter(ResidentialProps == "Residential") %>% 
  mutate(percent_residential = eav / muni_EAV_includesTIF)# %>% select()

perc_residential


```

Percent Residential is the Residential EAV outside of TIFs / Municipality EAV outside of TIFs.


```{r}
# #lowest % of eav from residential property types
# perc_residential %>% select(-ResidentialProps) %>%
#   arrange(percent_residential) %>% head()
# 
# #highest % of eav from residential property types
# perc_residential %>%  select(-ResidentialProps) %>%
#   arrange(percent_residential) %>% tail()
# 

perc_residential %>% group_by(ResidentialProps)%>%
  
  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%
  ggplot(aes(fill = percent_residential)) + 
  geom_sf(aes(geometry = geometry), color = "black") + theme_void()+ 
  labs(title = "Residential EAV /  Total EAV") +
    theme_classic() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +#+
   scale_fill_stepsn(colors = c("white","darkblue"),
                    #   limits = c(0,.6),
                       n.breaks = 5, show.limits=TRUE,
                        name = "% Residential", label = scales::percent)
```
 Homeowner's Exemptions Exempt EAV / Residential EAV
 
```{r}

exemptions_to_resEAV_ratios <- grouped_exemptions %>% 
  filter(ResidentialProps == "Residential") %>%
  mutate(exemptEAV_pctof_resEAV = exempt_EAV/eav,
         nontif_ratio = exempt_EAV / tax_base_noexemps) %>% 
  select(agency_name, exemptEAV_pctof_resEAV, nontif_ratio) %>% 
  arrange(nontif_ratio)

exemptions_to_resEAV_ratios %>%
  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%
  ggplot(aes(fill = exemptEAV_pctof_resEAV)) + 
  geom_sf(aes(geometry = geometry), color = "black") + theme_void()+ 
  labs(title = "Exemptions / Residential EAV (in and out of TIFs)") +
    theme_classic() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
   scale_fill_stepsn(colors = c("white", "darkblue"), 
                    #   limits = c(0,.6),
                       n.breaks = 5, show.limits=TRUE,
                        name = "% Residential EAV \nthat is exempt", label = scales::percent)

exemptions_to_resEAV_ratios %>%
  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%
  ggplot(aes(fill = nontif_ratio)) + 
  geom_sf(aes(geometry = geometry), color = "black") + theme_void()+ 
  labs(title = "Non-TIF EAV only: Homestead Exemptions / Residential EAV", caption = "Village of Phoenix skews graph. Dropped in map below") +
    theme_classic() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
   scale_fill_stepsn(colors = c("white", "darkblue"), 
                    #   limits = c(0,.6),
                       n.breaks = 5, show.limits=TRUE,
                        name = "% Residential EAV \nthat is exempt", label = scales::percent)

exemptions_to_resEAV_ratios %>% 
  filter(nontif_ratio<.5) %>%
  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%
  ggplot(aes(fill = nontif_ratio)) + 
  geom_sf(aes(geometry = geometry), color = "black") + theme_void()+ 
  labs(title = "Percent of Residential EAV that is Tax Exempt", 
  subtitle = "Non-TIF EAV only: Homestead Exemptions / Residential EAV", caption = "Drops Village of Phoenix because it skews map colors 
       (78% of their residential EAV is tax exempt)") +
    theme_classic() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
   scale_fill_stepsn(colors = c("white", "darkblue"), 
                    #   limits = c(0,.6),
                       n.breaks = 5, show.limits=TRUE,
                        name = "% Residential EAV \nthat is exempt", label = scales::percent)
```


## District Shares

Tax base in the tables above are the total EAV outside of TIF areas. 


```{r}
TC_bills_current <- read_csv("TC_bills_current.csv") %>% 
  mutate(tax_code = as.character(tax_code))

class_dict$class_code <- as.character(class_dict$class_code)
 

taxcodes_current <- full_join(TC_bills_current, muni_tax_codes, 
                      by = c("tax_code" = "tax_code_num")) 

Current_Taxrates <- taxcodes_current %>% 
  left_join(muni_agency_names) %>%
  group_by(agency_name) %>% 
  summarize(MuniLevy = sum(final_tax_to_dist, na.rm = TRUE), # amount billed by munis with current exemptions in place
            nonTIF_EAV_post_exemps = sum(final_tax_to_dist/(tax_code_rate/100), na.rm = TRUE), # within cook county
            TIF_increment_EAV = sum(final_tax_to_tif/(tax_code_rate/100), na.rm=TRUE),  # within cook county
            Exempt_EAV = sum(tax_amt_exe/(tax_code_rate/100), na.rm=TRUE), # within cook county
            Total_EAV = sum((tax_amt_exe+final_tax_to_dist+final_tax_to_tif)/(tax_code_rate/100), na.rm = TRUE))%>%

  mutate(tax_rate_current = MuniLevy/nonTIF_EAV_post_exemps,
         nonTIF_EAV_pre_exemps = nonTIF_EAV_post_exemps + Exempt_EAV,
         taxrate_new = MuniLevy/nonTIF_EAV_pre_exemps,
         taxrate_change = tax_rate_current-taxrate_new) %>% 
  select(agency_name, tax_rate_current, taxrate_new, everything()) %>% 
  arrange(desc(tax_rate_current))

land_use <- taxcodes_current %>% 
  left_join(muni_agency_names) %>% 
  left_join(Current_Taxrates) %>%
  mutate(ResidentialProps = ifelse(major_class_code %in% c("2", "3", "9"), "Residential", "Commercial"),
         PropType = case_when(
           major_class_code %in% c("3","9") ~ "Multi-Family",
           major_class_code == "2" ~ "Single-Family",
           TRUE ~ "Commercial-Industrial")) %>%
  group_by(agency_name, ResidentialProps, agency_num, tax_rate_current, taxrate_new, taxrate_change) %>% 
  
  # All of the values calculated below are AFTER exemptions have been removed
  summarize(taxrev_from_proptype = sum(final_tax_to_dist, na.rm = TRUE),
            nonTIF_EAV = sum(final_tax_to_dist/(tax_code_rate/100), na.rm = TRUE),
            TIF_increment_EAV = sum(final_tax_to_tif/(tax_code_rate/100), na.rm=TRUE),
            Exempt_EAV = sum(tax_amt_exe/(tax_code_rate/100), na.rm=TRUE),
            Total_EAV = sum((tax_amt_exe+final_tax_to_dist+final_tax_to_tif)/(tax_code_rate/100), na.rm = TRUE) ) %>% ungroup()

Current_Taxrates
```
Current composite tax rates for each municipality are above. Table also includes the new tax rate if there were no exemptions, the levy (aka amount collected by the municipality from final_tax_to_dist variable), EAV outside of TIFs, amount of exempt EAV, and additional variables. 


Below: Take EAV values within each property type, joine it with muni level EAV values, and calculate ...

```{r}

grouped_exemptions <- exemptions_by_class_per_TC %>% 
   # group_by(agency_name, major_class_code, major_class_type, ResidentialProps, PropType) %>%
  group_by(agency_name, major_class_type, ResidentialProps, PropType, major_class_code) %>%
  summarize(eav = sum(eav),
           exempt_EAV = sum(exempt_EAV, exe_abate, na.rm=TRUE),
         tax_base_current = sum(tax_base_current, na.rm=TRUE),
         tax_base_noexemps = sum(tax_base_noexemps, na.rm=TRUE)) %>% ungroup()
  
grouped_exemptions

# calculate totals with ONLY EAV outside of TIFs
muni_eav <- grouped_exemptions %>%  
  group_by(agency_name) %>% 
  summarize(muni_EAV_includesTIF = sum(eav), # all EAV in the municipality that exists
            muni_tax_base_current=sum(tax_base_current), # taxable EAV based on current exemptions
            muni_tax_base_noexemps = sum(tax_base_noexemps)) %>%  # taxable EAV pre-exemptions
  ungroup() 

pct_property_types <- left_join(grouped_exemptions, muni_eav) %>% 
  #filter(ResidentialProps == "Residential") %>% 
  mutate(pct_PropType = eav / muni_EAV_includesTIF)# %>% select()

pct_property_types


burden_table <- pct_property_types %>% 
  left_join(Current_Taxrates, by = c("agency_name")) %>%
  mutate(rev_collected_current = tax_base_current * tax_rate_current,
         rev_collected_new = tax_base_noexemps*taxrate_new,
         burden_current = rev_collected_current/MuniLevy,
         burden_noexemps = rev_collected_new/MuniLevy, 
         burden_change = burden_current - burden_noexemps,
         burden_noexemps = ifelse(burden_noexemps >1, 1, burden_noexemps)) %>%
  mutate(burden_current = ifelse(is.na(burden_current), 0, burden_current),
         burden_noexemps = ifelse(is.na(burden_noexemps), 0, burden_noexemps)) %>%
  mutate(burden_noexemps = ifelse(burden_noexemps>1, 1, burden_noexemps),
         burden_current = ifelse(burden_current>1, 1, burden_current))


burden_table %>% 
   
  arrange(desc(burden_noexemps))

burden_table

burden_shift <- burden_table # only to not change code below in graphs. 
```


Compare to 6 municipalities calculated by JD:

Only class 2 property types in output below. Looks good! Yay.

```{r}
burden_table %>%  filter(agency_name %in% c("CITY OF CHICAGO", "VILLAGE OF BRIDGEVIEW", "VILLAGE OF DOLTON", "VILLAGE OF HOFFMAN ESTATES","VILLAGE OF MIDLOTHIAN","VILLAGE OF OAK PARK") & major_class_code == 2) %>% select(-c(ResidentialProps, major_class_type, PropType)) %>%
  select(agency_name, muni_EAV_includesTIF, MuniLevy, exempt_EAV,burden_current, burden_noexemps)
```



```{r eval=FALSE, include=FALSE}
burden_with_exemptions <- taxcodes_current %>% 
  left_join(muni_agency_names) %>%   
  left_join(Current_Taxrates) %>% 
#  left_join(noexemps_rev) %>%
  select(-c(agency_rate)) %>%
  
  mutate(         ResidentialProps = ifelse(major_class_code %in% c("2", "3", "9"), "Residential", "Commercial"),
         
         PropType = case_when(
           major_class_code %in% c("3","9") ~ "Multi-Family",
           major_class_code == "2" ~ "Single-Family",
           TRUE~ "Commercial-Industrial")) %>% 
  arrange(tax_code) %>%

  group_by(agency_name, major_class_type, tax_rate_current, taxrate_new, PropType, ResidentialProps, agency_num, major_class_code) %>%
  summarize(final_tax_to_dist = sum(final_tax_to_dist, na.rm=TRUE),
           # final_tax_to_dist_noexemps = sum(final_tax_to_dist_noexemps, na.rm=TRUE),
            MuniLevy = first(MuniLevy)) %>%
  mutate(current_burden = final_tax_to_dist/MuniLevy) %>%
           # no_exemptions_burden = sum(no_exemptions_burden, na.rm = TRUE)) %>%
  select(agency_name, major_class_type, current_burden, tax_rate_current, taxrate_new, everything()) %>% ungroup()

# find outliers
burden_with_exemptions %>% arrange(desc(current_burden))


burden_noexemps<- taxcodes_noexemps %>% 
 # left_join(muni_tax_codes) %>% 
  left_join(muni_agency_names) %>%   
  left_join(noexemps_munisummary) %>% 
#  left_join(noexemps_rev) %>%
  select(-c(agency_rate)) %>%
  
  mutate(ResidentialProps = ifelse(major_class_code %in% c("2", "3", "9"), "Residential", "Commercial"),
         
         PropType = case_when(
           major_class_code %in% c("3","9") ~ "Multi-Family",
           major_class_code == "2" ~ "Single-Family",
           TRUE~ "Commercial-Industrial")) %>% 
  arrange(tax_code) %>%

  group_by(agency_name, major_class_type, taxrate_noexemps, PropType, ResidentialProps, agency_num, major_class_code) %>%
  summarize(final_tax_to_dist_noexemps = sum(final_tax_to_dist, na.rm=TRUE),
           # final_tax_to_dist_noexemps = sum(final_tax_to_dist_noexemps, na.rm=TRUE),
           # MuniLevy = first(MuniLevy)
           ) %>%
#  mutate(burden_noexemps = final_tax_to_dist/MuniLevy) %>%
           # no_exemptions_burden = sum(no_exemptions_burden, na.rm = TRUE)) %>%
  select(agency_name, major_class_type,
      #   burden_noexemps,
         taxrate_noexemps, everything()) %>% ungroup()


#burden_noexemps %>% arrange(desc(burden_noexemps))

burden_shift <- full_join( burden_with_exemptions, burden_noexemps, 
                           by = c("agency_name", "major_class_type", "PropType", "ResidentialProps", "major_class_code", "agency_num")) %>%
  mutate(burden_noexemps = final_tax_to_dist_noexemps/MuniLevy,
       burden_change = current_burden - burden_noexemps,
                                                         taxrate_change = tax_rate_current - taxrate_new)


#burden_shift
```

```{r eval=FALSE, include=FALSE}
burden_shift %>% filter(major_class_code == 2)  %>% 
  select(agency_name, burden_change, taxrate_change, tax_rate_current, taxrate_new, taxrate_noexemps) %>% arrange(burden_change)




burden_shift_table <- burden_shift %>% 
  mutate(burden_change = round(current_burden - burden_noexemps, digits =5),
         taxrate_change = tax_rate_current - taxrate_new) %>% 
  select(agency_name, major_class_type, burden_change, taxrate_change, tax_rate_current, taxrate_new, PropType, ResidentialProps) %>% 
  arrange(burden_change)


muni_eav <- land_use %>% group_by(agency_name, agency_num) %>% 
  summarize(muni_EAV = sum(Total_EAV)) %>% ungroup()

perc_residential <- left_join(land_use, muni_eav) %>% 
  filter(ResidentialProps == "Residential") %>% 
  mutate(percent_residential = Total_EAV / muni_EAV) 


table_ResidentialEAV <- land_use %>% 
  filter(ResidentialProps == "Residential" ) %>% 
  left_join(perc_residential) %>%
  select(agency_name,  Total_EAV, Exempt_EAV, percent_residential, tax_rate_current, taxrate_new, taxrate_change, 
         -c(agency_num, ResidentialProps, nonTIF_EAV, TIF_increment_EAV, taxrev_from_proptype)) %>% 
  rename("Residential EAV" = Total_EAV,
         "Exempt EAV" = Exempt_EAV,
         Municipality = agency_name) 

table_ResidentialEAV %>%
  mutate(percent_residential = scales::percent(percent_residential, accuracy = 0.01),
         tax_rate_current = scales::percent(tax_rate_current, accuracy = 0.01),
         taxrate_new = scales::percent(taxrate_new, accuracy = 0.01),
         taxrate_change = scales::percent(taxrate_change, accuracy = 0.01)) %>% arrange(desc(taxrate_change))

#lowest % of eav from residential property types
perc_residential %>%
  arrange(percent_residential) %>% head()

#highest % of eav from residential property types
perc_residential %>% 
  arrange(percent_residential) %>% tail()


perc_residential %>%
  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%
  ggplot(aes(fill = percent_residential)) + 
  geom_sf(aes(geometry = geometry), color = "black") + theme_void()+ 
  labs(title = "Residential EAV /  Total EAV") +
    theme_classic() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
   scale_fill_gradientn(colors = c("white", "darkblue"), 
                    #   limits = c(0,.6),
                       n = 5,
                        name = "% Residential", label = scales::percent)
```


# Tax Rates

```{r}
TC_bills_noexemps <- read_csv("TC_bills_noexemps.csv") %>% mutate(tax_code = as.character(tax_code))
# noexemps_rev<- taxcodes_noexemps  %>%   
#   left_join(muni_agency_names) %>%
#   select(tax_code, major_class_code, major_class_type, final_tax_to_dist, agency_name) %>% 
#   rename(final_tax_to_dist_noexemps = final_tax_to_dist)

taxcodes_noexemps <- left_join(TC_bills_noexemps, muni_tax_codes, 
                      by = c("tax_code" = "tax_code_num"))  %>%
    left_join(muni_agency_names)


noexemps_munisummary <- taxcodes_noexemps %>% 
  group_by(agency_name) %>% 
  summarize(MuniLevy = sum(final_tax_to_dist, na.rm = TRUE), #amount billed by muni within cook county
            nonTIF_EAV = sum(final_tax_to_dist/(tax_code_rate/100), na.rm = TRUE), # within cook county
           final_tax_to_dist = sum(final_tax_to_dist, na.rm=TRUE),
     TIF_increment_EAV = sum(final_tax_to_tif/(tax_code_rate/100), na.rm=TRUE),  # within cook county
            Exempt_EAV = sum(tax_amt_exe/(tax_code_rate/100), na.rm=TRUE), # within cook county
            Total_EAV = sum((tax_amt_exe+final_tax_to_dist+final_tax_to_tif)/(tax_code_rate/100), na.rm = TRUE)) %>%

  mutate(taxrate_noexemps = MuniLevy/nonTIF_EAV,
        # nonTIF_EAV= nonTIF_EAV + Exempt_EAV,
       ) 
 # select(agency_name, tax_rate_noexemps, everything()) %>% 
#  arrange(desc(tax_rate_noexemps))


noexemps_munisummary

```

6 Highest and 6 lowest tax rates. View the extremes on both sides

```{r}

Current_Taxrates %>%  filter(tax_rate_current > 0.25 )
                           
Current_Taxrates %>%  filter( tax_rate_current < 0.08)
```


# Map of Tax Rate Change


```{r}


burden_shift %>% 
    filter(major_class_code == 2) %>% # all property types have same composite tax rate
  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%
  ggplot(aes(fill = tax_rate_current)) + 
      theme_classic() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
  # scale_fill_gradientn(colors = c("white", "maroon"), 
  scale_fill_stepsn(colors = c("white", "#76C3D6", "#1A5E7D"),
                      limits = c(0,.35),
                    show.limits=TRUE,
                       n.breaks = 5,
                        name = "Tax Rate", label = scales::percent)+
  geom_sf(aes(geometry = geometry), color = "black") +  
  labs(title = "Current composite tax rates with current exemptions in place" ,    
         caption = "Highest composite tax rate is in Park Forest (41.4%.)
       Lowest composite tax rate is in Oak Brook (6.3%) and Chicago (6.7%).")



burden_shift %>% 
    filter(major_class_code == 2) %>% # all property types have same composite tax rate
  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%
  ggplot(aes(fill = taxrate_new)) + 
      theme_classic() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
     scale_fill_stepsn(colors = c("white", "#76C3D6", "#1A5E7D"),
       #colors = c("white", "maroon"), 

 #  scale_fill_gradientn(colors = c("white", "maroon"), 
  #scale_fill_stepsn(colors = c("","white", "#235D53"),
                      limits = c(0,.35),
                       n.breaks = 6, show.limits=TRUE,
                        name = "Tax Rate", label = scales::percent)+
  geom_sf(aes(geometry = geometry), color = "black") +  
  labs(title = "New composite tax rates if exemptions were eliminated" ,    
         caption = "Data from CCAO & PTAXSIM. Unincorporated areas are filled with gray.")
```

```{r}
# as a dot graph ## 

order <- burden_shift %>% ungroup %>% as_tibble() %>%
# filter(YEAR == 2019 & NAME != "NA" & CIHISPEED == "10") %>% 
  select(agency_name, tax_rate_current)



# look at ones that changed the most
burden_shift %>%  
  filter(tax_rate_current > 0.23 |tax_rate_current < 0.085) %>% 
  filter(PropType == "Single-Family") %>%
  ungroup() %>% 
  select(agency_name, tax_rate_current, taxrate_new) %>% 
  pivot_longer(c("tax_rate_current", "taxrate_new"), 
               names_to = "type", values_to = "tax_rate") %>% 
  left_join(order) %>%
  ggplot(aes(x = tax_rate*100, y= reorder(agency_name, tax_rate_current)))+
  geom_line(aes(group = agency_name))+ 
   geom_point(aes(color=type), size=3 )+
  theme_minimal() + 
  theme( 
    legend.title = element_blank(),
               plot.title.position = "plot",
     #   panel.background = element_rect(fill='transparent'), #transparent panel bg
    plot.background = element_rect(fill='transparent', color=NA) #transparent plot bg
   )+
        scale_color_brewer(palette="Paired", labels = c("Exemptions", "No Exemptions"), direction = 1)+

  labs(title = "Difference in Composite Tax Rate if there were No Exemptions", x = "Composite Tax Rate (%)", y = "" , caption = "For the highest and lowest composite tax rates")
```

```{r}



burden_shift %>% 
  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%
  ggplot(aes(fill = taxrate_change*100)) + 
  geom_sf(aes(geometry = geometry), color = "black") +  
      theme_classic() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
   scale_fill_gradientn(colors = c("white", "maroon"), 
                    #   limits = c(0,.6),
                       n = 7,
                        name = "Percentage Point Change")+
  labs(title = "Change in Composite Tax Rate if Exemptions are Removed")

```


# Single-Family (Class 2) Burden Shift


Change in Tax Burden for Class 2 Properties:

```{r}
burden_shift %>%
      mutate(burden_current = ifelse(burden_current>1, 1, burden_current)) %>%

      filter(major_class_code == 2) %>%

  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%

  ggplot(aes(fill = burden_current)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  theme_classic() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
  # scale_fill_gradientn(
   scale_fill_steps(low="white", high="darkblue", 
                        show.limits=TRUE, n.breaks = 6,
                        name = "Burden with \nExemptions"#, labels = scales::percent
                    )+
  labs(title = "Current share of property tax burden", subtitle = "for Class = 2 Property Types")


burden_shift %>% 
      filter(major_class_code == 2) %>%

  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%

  ggplot(aes(fill = burden_noexemps)) + 
  geom_sf(aes(geometry = geometry), color = "black") +     
  theme_classic() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
   #scale_fill_gradientn(
    # scale_fill_binned(
     #colors = c("#f0f9e8", "#0868ac"),
   #  colors = c('#a6611a','#018571'),
                       #limits = c(0,.6),
           #            n = 6,
   #   name = "Burden w/o \nExemptions", label = scales::percent)+
  
  scale_fill_stepsn(colors = c("#ffffcc","#a1dab4" ,"#41b6c4","#2c7fb8", "#253494"),
 # low = "#ffffcc", mid = "#41b6c4", high = "#253494",
#  space = "Lab",
  n.breaks = 6,
  na.value = "grey50",
  guide = #"legend", 
   "coloursteps",
  aesthetics = "fill",
  show.limit = TRUE,
  name = "Burden w/o \nExemptions"
)+
#viridis::scale_fill_viridis(option = "mako", )+
 labs(title = "New share of property tax burden", subtitle = "for Class = 2 Property Types")
```



```{r}
burden_shift %>% 
  filter(major_class_code == 2) %>%
  mutate(burden_change = ifelse(burden_change>0,0, burden_change))%>%
  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%
  ggplot(aes(fill = burden_change)) + 
  geom_sf(aes(geometry = geometry), color = "black") + theme_void()+ 
  labs(title = "Change in Residential Share of Tax Burden") +
    theme_classic() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
   scale_fill_steps( low="darkblue", high = "white",
     #colors = c("darkblue","white"), 
     show.limits = TRUE,
                       n.breaks = 6,
                        name = "Difference in Burden", label = scales::percent)
  
   # geom_sf(data = countyIL, fill=NA, color="dark gray")

```

```{r collapse=TRUE}
# as a dot graph ## 

order <- burden_shift %>% 
  ungroup %>% as_tibble() %>%
  #  filter(ResidentialProps == "Residential") %>%

  filter(PropType == "Single-Family") %>%
  select(agency_name, burden_current)



# burder_shift_ordered <-  burden_shift %>% 
#   ungroup() %>% 
#   select(agency_name, current_burden, no_exemptions_burden) %>%    
#   pivot_longer(c("current_burden", "no_exemptions_burden"), 
#                names_to = "type", values_to = "pct_burden") %>% 
#   left_join(order)

# look at ones that changed the most
burden_shift %>%  
  filter(burden_current > 0.85 |burden_current < .3) %>% 
  #filter(ResidentialProps == "Residential") %>%
    filter(PropType == "Single-Family") %>%

  ungroup() %>% 
  select(agency_name, burden_current, burden_noexemps) %>% 
  mutate( 
         burden_noexemps = ifelse(burden_noexemps > 1, 1, burden_noexemps)) %>%
  pivot_longer(c("burden_current", "burden_noexemps"), 
               names_to = "type", values_to = "pct_burden") %>% 
  inner_join(order) %>%
  ggplot(aes(x = pct_burden*100, y= reorder(agency_name, burden_current)))+
  geom_line(aes(group = agency_name))+ 
   geom_point(aes(color=type), size=3 )+
  theme_minimal() + 
  theme(#legend.position = "none", 
    legend.title = element_blank(),
               plot.title.position = "plot",
     #   panel.background = element_rect(fill='transparent'), #transparent panel bg
    plot.background = element_rect(fill='transparent', color=NA) #transparent plot bg
   )+
        scale_color_brewer(palette="Paired", labels = c("No Exemptions", "Exemptions"), direction = 1)+

  labs(title = "Change in Single-family Residential Tax Burden", 
  x = "Share of Levy (%)", y = "" , 
  caption = "Residential Tax Burden is the Share of the property tax collected that was paid for by 
     single-family home owners in property classes 2.") #+
# scale_x_continuous(label = scales::percent)

```

# Summary table of Burden Shift: 3 Property Categories

Same data that is used above but property types are combined into 3 categories:

Single-family (Class 2),   
Multi-Family (Class 3 & 9),   
and Commercial-Industrial (all other Classes: 1,4, 5, 6, 7, 8, 9)  

Current Burden:

```{r}
munis_3property_types <- burden_shift %>%# mutate(
     #       burden_noexemps = ifelse(is.na(burden_noexemps), 0, burden_noexemps)) %>%
  #    mutate(burden_current = ifelse(burden_current>1, 1, burden_current)) %>%
  ungroup() %>% 
  group_by(agency_name, PropType) %>%
  summarize(#final_tax_to_dist = sum(final_tax_to_dist, na.rm = TRUE),
            burden_current = sum(burden_current, na.rm = TRUE),
            burden_noexemps = sum(burden_noexemps, na.rm = TRUE),
            burden_change = sum(burden_change, na.rm = TRUE)) 

#write.csv(munis_3property_types, "1_usemetable.csv")

proptypes3_current <- munis_3property_types %>% 
    mutate(burden_current = ifelse(is.na(burden_current), 0, burden_current)) %>%

  pivot_wider( id_cols = agency_name , names_from = "PropType", values_from = "burden_current",names_prefix="current" ) %>% 
  arrange(-`currentSingle-Family`)
```

__Tax Burden if there were no exemptions:__

```{r}

proptypes3_noexemps <- munis_3property_types %>% 
  mutate(burden_noexemps = ifelse(is.na(burden_noexemps), 0, burden_noexemps)) %>%
  pivot_wider( id_cols = agency_name , names_from = "PropType", values_from = "burden_noexemps", names_prefix = "W/O Exemptions")  
```

__Change in Share of Burden if there were no exemptions:__

```{r}
# burden_shift %>% ungroup() %>% 
#   group_by(agency_name, PropType) %>%
#   summarize(district_rev_collected = sum(district_rev_collected),
#             current_burden = sum(current_burden),
#             no_exemptions_burden = sum(no_exemptions_burden)) 

props_wide <- munis_3property_types %>%
  pivot_wider(id_cols = agency_name , 
               names_from = "PropType", 
               values_from = "burden_change") %>% 
  arrange(-`Single-Family`) 

props_wide[2:4] <- sapply(props_wide[2:4], function(x) scales::percent(x, accuracy=.001))
props_wide


proptypes3_comparisontable<- left_join(proptypes3_current, proptypes3_noexemps, by = "agency_name")
proptypes3_comparisontable

proptypes3_comparisontable %>% write.csv("BurdenshareTable.csv")
```


```{r}
proptypes3_burdenchange <- munis_3property_types %>% 
  mutate(burden_noexemps = ifelse(is.na(burden_change), 0, burden_change)) %>%
  pivot_wider( id_cols = agency_name , names_from = "PropType", values_from = "burden_change") %>%
  arrange(desc(`Single-Family`))

proptypes3_burdenchange

write.csv(proptypes3_burdenchange, "proptypes3_burdenchange.csv")
```

# Exporting data to Excel file


```{r}
library(openxlsx)

dataset_names <- list(
  'Land Use' = perc_residential.
  'Burden Table' = burden_table,
 # 'EAVoutsideTIFs' = EAV_outside_TIFS_byClass,
  'TaxcodeData-NoExemptions'=taxcodes_noexemps,
  'TaxcodeData-Current' = taxcodes_current,
  'Composite Tax Rates' = Current_Taxrates,
                      'MuniEAV' = muni_eav,
 # 'ResidentialEAV' = table_ResidentialEAV, 
 # 'Burden with Exemps' = burden_with_exemptions,
 # 'Burden without Exemps' = burden_noexemps,
                      'Burden Shift-3PropTypes' = proptypes3_burdenchange,
                      '3 Property Types'= munis_3property_types,
                      'Tax Codes - Current'
 = taxcodes_current,
 'Tax Codes - No Exemptions' = taxcodes_noexemps)

write.xlsx(dataset_names, file = 'data_for_slides.xlsx')
```

