---
title: "Incentive PROJECT Models"
date: "August 10, 2024"
date-modified: last-modified
format: 
  html:
    embed-resources: true
    theme: lumen
    code-fold: true
    code-line-numbers: true
    code-overflow: wrap
    toc: true
    toc-location: left
    df-print: paged
knitr: 
  opts_chunk:
    warning: true
    message: false
---

## Data prep

```{r warning = FALSE}
library(tidyverse)
library(jsonlite)


ptax_pins <- read_csv("./Output/comm_ind_PINs_2011-2022_balanced.csv")

parcuniverse_keypins <- readxl::read_xlsx("./Inputs/parceluniverse_keypins_20240725.xlsx", 
                                          sheet = "keypins_20240725") %>%
  mutate(pin14 = str_pad(as.character(pin), width = 14, side = "left", pad = "0"),
         keypin = str_pad(as.character(proration_key_pin), width = 14, side = "left", pad = "0"),
         pin10 = str_sub(pin14,1,10),
         pin7 = str_sub(pin14,1,7), .before = "pin",) %>%
  select(-c(pin_7dig, pin, Column1)) %>%
  filter(class != "EX")

base_url <- "https://datacatalog.cookcountyil.gov/resource/tx2p-k2g9.json"

nbh_pins <- GET(
  base_url,
  query = list(
    tax_year = 2022,
    `$select` = paste0(c("pin", # "pin10", 
                         "nbhd_code"
    ),
    collapse = ","),
    `$limit` = 500000000L
  )
)

nbh_pins <- fromJSON(rawToChar(nbh_pins$content))
nbh_pins <- nbh_pins |> filter(pin %in% ptax_pins$pin)
```


```{r warning = FALSE}
puni_keypins <- parcuniverse_keypins %>% 
  select(pin14, 
         puni_keypin = keypin, 
         proration_rate2022 = proration_rate, 
         mailing_name2022 = mailing_name )

# Workaround for identifying more project IDs. 
# Used Appeal ID to create unique identifier to group PINs.
bor <- read_csv("Output/borappeals.csv") %>%
  mutate(project_appellant = paste(project_id, sep = "-", appellant))

bor_pins <- bor %>% 
  group_by(pin) %>% 
  arrange(desc(tax_year)) %>%
  summarize(pin = first(pin),
            class_bor = paste(list(unique(class)), collapse = ", "),
            taxyear_appealed = paste(list(tax_year), collapse = ", "),
            appellant = first(appellant),
            project_id = first(project_id), 
            timesappealed = n() ) %>%
  mutate(proj_appeallant = paste(project_id, "-", appellant))

ptax_proj <- ptax_pins %>% 
  left_join(bor_pins) %>%
  left_join(puni_keypins, by = c("pin" = "pin14"))  %>%
  left_join(nbh_pins)


proj_xwalk <- read_csv("Output/all_keypins.csv")               
ptax_proj <- ptax_proj %>% left_join(proj_xwalk, by = "pin", relationship = "many-to-many")


ptax_proj <- ptax_proj %>% 
  mutate(actual_keypin = ifelse(is.na(proj_appeallant), puni_keypin, proj_appeallant))%>%
  mutate(actual_keypin = ifelse(is.na(actual_keypin), keypin, actual_keypin)) %>%
  select(year, actual_keypin, pin, nbhd_code, tax_code_num, class, class_bor, tax_code_rate, fmv, av_clerk, tax_bill_total, everything())
```

```{r manual-pin-codings}

ptax_proj <- ptax_proj |> 
mutate(actual_keypin = case_when(
  pin %in% c(09253201421002, 09253201421007) ~ "ICUSS - 7720 W. Touhy", 
  pin %in% c(04333020971001:04333020971020) ~  "611 Milwaukee Ave",
  pin %in% c(18282000461001) ~ "ADCON EXPRESS INC", 
  pin %in% c(18282000461003, 20084040010000: 20084040060000 ) ~ "959 W 51st St.",

    pin %in% c(17221101220000, 17221101230000, 17221101240000 ) ~ "1231 S. Prairie Ave.",
  TRUE ~ as.character(actual_keypin)))
 
```

```{r}
modelsummary::datasummary_skim(ptax_proj)
```

**Which came first: The TIF or the incentive class??**

```{r}
#| label: tbl-whichcamefirst
#| tbl-cap: "PINS that were in a TIF before receiving an incentive classification. 254 PINs had incentive classes before they were in a TIF. 1373 PINs were in a TIF before they had an incentive classification"


change_table <- ptax_proj |>
  ungroup() |>
  arrange(pin, year) |>
  group_by(pin) |>
  summarize(first_tif_year = first(year[in_tif==1]),
                     first_incent_year = first(year[incent_prop=="Incentive"])) |>
  filter(!is.na(first_tif_year) & !is.na(first_incent_year)) |>
  select(pin, first_tif_year, first_incent_year)|> 
  distinct() 

#254 PINs had incentive classes before they were in a TIF
# change_table |> filter(first_tif_year > first_incent_year)

#1373 PINs were in a TIF before they had an incentive classification
change_table |> filter(first_tif_year < first_incent_year)
```


```{r }
#| label: tbl-keypin-tiforincent
#| tbl-cap: "Projects that were in a TIF before they also got an incentive classification."

change_table <- ptax_proj |>
  ungroup() |>
  arrange(pin, year) |>
  group_by(actual_keypin) |>
  summarize(first_tif_year = first(year[in_tif==1]),
                     first_incent_year = first(year[incent_prop=="Incentive"])) |>
  filter(!is.na(first_tif_year) & !is.na(first_incent_year)) |>
  select(actual_keypin, first_tif_year, first_incent_year)|> 
  distinct() 

# 152 projects were in a tif
# change_table |> filter(first_tif_year > first_incent_year) 
# 329 projects were in a TIF before they received incentive classes
change_table |> filter(first_tif_year < first_incent_year) 
```


**How many Commercial PINs become Industrial PINs? If commercial pins are the comparison group, we should know the sample size of PINs that change to other land uses.**

```{r}
#| label: tbl-commercial-toOther-pins
#| tbl-cap: "Commercial PINs that became Industrial or other land use types"


change_table <- ptax_proj |>
  ungroup() |>
  arrange(pin, year) |>
  group_by(pin) |>
  summarize(
    first_commerc_year = first(year[land_use == "Commercial"]),
    first_indust_year = first(year[land_use == "Industrial"]),
    first_other_year = first(year[!land_use %in% c("Commercial", "Industrial")]))  |>
  select(pin, first_commerc_year, first_indust_year, first_other_year)|> 
  distinct() 

change_table |> filter(first_commerc_year < first_indust_year)
  
change_table |> filter(first_commerc_year < first_other_year)
```


```{r }
#| label: tbl-keypinchange
#| tbl-cap: "Projects where the first year as a commercial land use is before becoming commercial."

change_table <- ptax_proj |>
  ungroup() |>
  arrange(pin, year) |>
  group_by(actual_keypin) |>
  summarize(
    first_commerc_year = first(year[land_use == "Commercial"]),
    first_indust_year = first(year[land_use == "Industrial"]),
    first_other_year = first(year[!land_use %in% c("Commercial", "Industrial")])) |>
  select(actual_keypin, first_commerc_year, first_indust_year, first_other_year)|> 
  distinct() 

change_table |> filter(first_commerc_year < first_other_year)
  
change_table |> filter(first_commerc_year < first_indust_year)


```
```{r}
table(ptax_proj$incent_prop)
ptax_proj %>% 
  arrange(pin, year) %>%
 # select(-c(landuse_change, land_use, incent_change, tif_change, pin_classes_2011, pin_classes_2022) ) %>%
  filter(incent_change2 %in% c("Incentive")) %>%
           write_csv("Output/incentive_project_summaries.csv")
```

```{r}
ptax_proj_summs <- ptax_proj %>% 
  
  group_by(year, actual_keypin) %>%
  summarize(
    incent_pins = sum(ifelse(class >= 600 & class < 900, 1, 0)),
    project_pins = n(),
    fmv = sum(fmv, na.rm=TRUE),
    tax_code = mean(tax_code_num, na.rm=TRUE),
    in_tif = mean(in_tif, na.rm=TRUE),
    pin_classes = paste(sort(unique(class)), collapse = ", "),
    land_use = paste(sort(unique(land_use)), collapse = ", "),
    incent_prop = paste(sort(unique(incent_prop)), collapse = ", "),
    triad = first(Triad),
    exe_abate = sum(exe_abate, na.rm = TRUE),
    clean_name = first(clean_name),
    landuse_change= paste(sort(unique(landuse_change)), collapse = ", "),
    incent_change = paste(sort(unique(incent_change)), collapse = ", "),
    tif_change = paste(sort(unique(tif_change)), collapse = ", "),
    ) %>%
  ungroup() |>
  group_by(actual_keypin) |>
  mutate(
    pin_classes_2011 = pin_classes[year == 2011],
    pin_classes_2022 = pin_classes[year == 2022],
    base_year_fmv_2011 = fmv[year == 2011],
    end_year_fmv_2022 = fmv[year == 2022],
    fmv_growth_2011 = fmv/base_year_fmv_2011,
    fmv_2022_2011_diff = end_year_fmv_2022 - base_year_fmv_2011) %>% 
  ungroup()

ptax_proj_summs <- ptax_proj_summs %>%
  mutate(landuse_change2 = ifelse(landuse_change!= "Always Commercial" & landuse_change != "Always Industrial", "Changes", landuse_change),
         land_use2 = ifelse(!land_use %in% c("Commercial", "Industrial"), "Multi Land Uses", land_use),
         tif_change2 = ifelse(!tif_change %in% c("Always TIF", "Never TIF"), "Changes", tif_change),
         incent_change2 = ifelse(!incent_change %in% c("Always Incentive", "Never Incentive"), "Changes", incent_change),
         in_tif2 = ifelse(in_tif != 0 & in_tif != 1, "Partly in TIF", in_tif)) %>%
  mutate(across(c(fmv, tax_code), round, digits = 0))
```


```{r}
ptax_proj_summs %>% 
  select(-c(landuse_change, land_use, incent_change, tif_change, pin_classes_2011, pin_classes_2022) ) %>%
           write_csv("Output/project_summaries.csv")

ptax_proj_summs %>% 
  select(-c(landuse_change, land_use, incent_change, tif_change, pin_classes_2011, pin_classes_2022) ) %>%
  filter(incent_change2 %in% c("Always Incentive", "Changes")) %>%
           write_csv("Output/incentive_project_summaries.csv")
```


```{r eval=FALSE}
# ptax_proj_summs %>% is.finite() %>% modelsummary::datasummary_skim()

ptax_proj_summs %>% 
  select(-c(land_use, incent_prop, landuse_change, tif_change, incent_change)) %>% 
  modelsummary::datasummary_skim("categorical")

```


```{r}
#| label: tbl-exemptprojects
#| tbl-cap: Projects that were exempt or became exempt. 

were_exempt <- ptax_proj_summs %>%
  group_by(actual_keypin) |>
  filter(fmv == 0 & year == 2011)
were_exempt

became_exempt <- ptax_proj_summs %>%
  group_by(actual_keypin) |>
  filter(fmv == 0 & year == 2022)  |> arrange(desc(project_pins))
became_exempt

#write_csv(were_exempt, "./Output/properties_were_exempt.csv")

#write_csv(became_exempt, "./Output/properties_became_exempt.csv")

ptax_proj_summs <- ptax_proj_summs %>% 
  filter(!actual_keypin %in% were_exempt$actual_keypin) %>% ## creates Infinite growth but these are the observations we do want to examine more for vacant to non-vacant properties.
  filter(!actual_keypin %in% became_exempt$actual_keypin)


```

852 projects have a combined FMV of \$0 in 2011 but have value in 2022. 2,294 PINs have FMV > $0 in 2011 but become tax exempt by 2022. 


**Problem example 1:** 3714512 - DAVITA INC is a project where all PINs were exempt in 2011 and had FMV values of 0 but became an incentive project in 2015 with 800 level PINs. Using FMV growth from 2011 results in a growth of INF and would need to be dropped from the model.

**Problem example 2:** 7308877 - LYCEE FRANCAIS DE CHIC	 became tax exempt. Filtering out PINs


> making binary variable for if a PIN went from vacant to not-vacant or not might work in a logit model though.


**Large FMV Growth PINS**:

- 17221101230000 for ACRON Navy Pier Hotel LLC was land class 100 in 2011, became 590 and gained value after it opened? 

- 17221101230000 had large growth but it should be added to a project. (PINs in project also experienced subdivision and changed PIN numbers?)


```{r eval=FALSE}
modelsummary::datasummary_skim(ptax_proj_summs)

ptax_proj_summs %>%  
  select(-c(land_use, incent_prop, landuse_change, tif_change, incent_change)) %>% 
  modelsummary::datasummary_skim("categorical")
```

## Fixed Effect Model using Project Level

```{r eval=FALSE}
library(plm)

proj_panel <- ptax_proj_summs %>% 
  select(actual_keypin, year, incent_change2, tif_change2, incent_prop, 
         landuse_change2, land_use2, in_tif2, triad, fmv_growth_2011, 
         base_year_fmv_2011, project_pins, incent_pins, clean_name)  %>% 
  mutate( fmv_growth_2011 = as.numeric(fmv_growth_2011)) %>% 
  arrange(actual_keypin, year) %>% 
  filter(!is.na(actual_keypin ) & !is.na(fmv_growth_2011) & !is.na(base_year_fmv_2011) )  %>%
   %>%
  mutate(across(year:triad, as.factor))


proj_panel <- proj_panel %>%
  pdata.frame(index = c("actual_keypin", "year") )
is.pbalanced(proj_panel)

plm(fmv_growth_2011 ~ incent_pins*in_tif2, 
    data = proj_panel,
    model = "within",
    effect = "twoways") %>% summary()


plm(fmv_growth_2011 ~ lag(incent_prop) + lag(incent_prop, 2), 
    data = proj_panel,
    model = "within",
    effect = "twoways") %>% summary()

plm(fmv_growth_2011 ~ land_use2*incent_prop, 
    data = proj_panel,
    model = "within",
    effect = "twoways") %>% summary()


plm(fmv_growth_2011 ~ land_use2, 
    data = proj_panel,
    model = "within",
    effect = "twoways") %>% summary()
```


```{r}
library(fixest)

proj_panel <- ptax_proj_summs %>% 
  select(actual_keypin, year, incent_change2, tif_change2, incent_prop,
         landuse_change2, land_use2, in_tif2, triad, fmv_growth_2011, 
         base_year_fmv_2011, project_pins, incent_pins, clean_name)  %>% 
  mutate( fmv_growth_2011 = as.numeric(fmv_growth_2011)) %>% 
  arrange(actual_keypin, year) %>% 
  filter(!is.na(actual_keypin ) & !is.na(fmv_growth_2011) & !is.na(base_year_fmv_2011) )  %>%
  mutate(across(year:triad, as.factor),
         land_use2 = as.character(land_use2),
         incent_prop2 = factor(incent_prop, levels= c("Non-Incentive", "Incentive"), labels = c("NonIncentive", "Incentive"), ordered = TRUE))

table(proj_panel$incent_prop2)

write_csv(proj_panel, "Output/project_panel_summaries.csv")

```

# Models 


```{r}
projectlevel_2FE = feols(fmv_growth_2011 ~ land_use2 + incent_prop2 + land_use2*incent_prop2  + incent_prop2 | actual_keypin + year, panel.id = c("actual_keypin", "year"), vcov = "twoway", data = proj_panel)

projectlevel_3FE = feols(fmv_growth_2011 ~ land_use2 + incent_prop2 + land_use2*incent_prop2  + incent_prop2 | actual_keypin + clean_name + year, panel.id = c("actual_keypin", "year"), vcov = "twoway", data = proj_panel)

munilevel = feols(fmv_growth_2011 ~ land_use2*incent_prop2 + incent_prop2 + land_use2 | clean_name + year, panel.id = c("actual_keypin", "year"), vcov = "twoway", data = proj_panel)

etable(projectlevel_2FE, projectlevel_3FE, munilevel)
```

**TIF and and incentive property interactions**
```{r}
projectlevel = feols(fmv_growth_2011 ~  land_use2*incent_prop2 + in_tif2*incent_prop2 | actual_keypin + year, panel.id = c("actual_keypin", "year"), vcov = "twoway", data = proj_panel)

munilevel = feols(fmv_growth_2011 ~ land_use2*incent_prop2 + land_use2*in_tif2| clean_name + year, panel.id = c("actual_keypin", "year"), vcov = "twoway", data = proj_panel)

etable(projectlevel, munilevel)
```


** some other variable combination **

```{r}
projectlevel = feols(fmv_growth_2011 ~ land_use2*incent_prop2 + in_tif2  | actual_keypin + year, panel.id = c("actual_keypin", "year"), vcov = "twoway", data = proj_panel)

projectlevel2 = feols(fmv_growth_2011 ~ land_use2*incent_prop2 + in_tif2  | actual_keypin + clean_name+ year, panel.id = c("actual_keypin", "year"), vcov = "twoway", data = proj_panel)

munilevel = feols(fmv_growth_2011 ~ land_use2*incent_prop2 + incent_prop2 + in_tif2 | clean_name + year, panel.id = c("actual_keypin", "year"), vcov = "twoway", data = proj_panel)

etable(projectlevel, projectlevel2, munilevel)

```

