---
title: "Incentive PROJECT Models"
date: "August 10, 2024"
date-modified: last-modified
format: 
  html:
    embed-resources: true
    theme: lumen
    code-fold: true
    code-line-numbers: true
    code-overflow: wrap
    toc: true
    toc-location: left
    df-print: paged
knitr: 
  opts_chunk:
    warning: true
    message: false
---

## Data prep

> Switched to timeseries CSV on September 1st isntead of balanced panel csv.

```{r warning = FALSE}
library(tidyverse)
library(jsonlite)
library(httr)
library(fixest)
```


```{r warning = FALSE}
project_pins <- readxl::read_xlsx("./Output/projects_checked_MAINFILE.xlsx")

ptax_pins <- read_csv("./Output/comm_ind_PINs_2011to2022_timeseries.csv") |> 
  filter(year >= 2011)


ptax_proj <- ptax_pins %>% 
  left_join(project_pins, by = c("pin"), relationship = "many-to-many") |>
  rename(Triad = Triad.x,
         clean_name = clean_name.x,
         incent_prop = incent_prop.x) |>
  select(-c(Triad.y, clean_name.y, incent_prop.y,))
```


```{r eval=FALSE}
# ptax_pins <- read_csv("./Output/comm_ind_PINs_2011-2022_balanced.csv")
ptax_pins <- read_csv("./Output/comm_ind_PINs_2011to2022_timeseries.csv") |> 
  filter(year >= 2011)

parcuniverse_keypins <- readxl::read_xlsx("./Inputs/parceluniverse_keypins_20240725.xlsx", 
                                          sheet = "keypins_20240725") %>%
  mutate(pin14 = str_pad(as.character(pin), width = 14, side = "left", pad = "0"),
         keypin = str_pad(as.character(proration_key_pin), width = 14, side = "left", pad = "0"),
         pin10 = str_sub(pin14,1,10),
         pin7 = str_sub(pin14,1,7), .before = "pin",) %>%
  select(-c(pin_7dig, pin, Column1)) %>%
  filter(class != "EX")

base_url <- "https://datacatalog.cookcountyil.gov/resource/tx2p-k2g9.json"

nbh_pins <- GET(
  base_url,
  query = list(
    tax_year = 2022,
    `$select` = paste0(c("pin",
                         "nbhd_code", 
                         "lat", "lon"
    ),
    collapse = ","),
    `$limit` = 500000000L
  )
)

nbh_pins <- fromJSON(rawToChar(nbh_pins$content))
nbh_pins <- nbh_pins |> filter(pin %in% ptax_pins$pin)
```


```{r warning = FALSE, eval=FALSE}
puni_keypins <- parcuniverse_keypins %>% 
  select(pin14, 
         puni_keypin = keypin, 
         proration_rate2022 = proration_rate, 
         mailing_name2022 = mailing_name )

# Workaround for identifying more project IDs. 
# Used Appeal ID to create unique identifier to group PINs.
bor <- read_csv("Output/borappeals.csv") %>%
  mutate(project_appellant = paste(project_id, sep = "-", appellant))

bor_pins <- bor %>% 
  group_by(pin) %>% 
  arrange(desc(tax_year)) %>%
  summarize(pin = first(pin),
            class_bor = paste(list(unique(class)), collapse = ", "),
            taxyear_appealed = paste(list(tax_year), collapse = ", "),
            appellant = first(appellant),
            project_id = first(project_id), 
            timesappealed = n() ) %>%
  mutate(proj_appeallant = paste(project_id, "-", appellant))
```


```{r warning = FALSE, eval=FALSE}
ptax_proj <- ptax_proj %>% 
  mutate(main_keypin = ifelse(is.na(proj_appeallant), puni_keypin, proj_appeallant))%>%
  mutate(main_keypin = ifelse(is.na(main_keypin), keypin, main_keypin)) %>%
  select(year, main_keypin, pin, nbhd_code, tax_code_num, class, class_bor, tax_code_rate, fmv, av_clerk, tax_bill_total, everything())
```


```{r}
modelsummary::datasummary_skim(ptax_proj)
```


```{r}
# modelsummary::datasummary_skim(project_pins)

```

**Which came first: The TIF or the incentive class??**

```{r}
#| label: tbl-whichcamefirst
#| tbl-cap: "PINS that were in a TIF before receiving an incentive classification. 254 PINs had incentive classes before they were in a TIF. 1373 PINs were in a TIF before they had an incentive classification"

change_table <- ptax_proj |>
  ungroup() |>
  arrange(pin, year) |>
  group_by(pin) |>
  summarize(first_tif_year = first(year[in_tif==1]),
            first_incent_year = first(year[incent_prop=="Incentive"])) |>
  filter(!is.na(first_tif_year) & !is.na(first_incent_year)) |>
  select(pin, first_tif_year, first_incent_year)|> 
  distinct() 

# 284 PINs had incentive classes before they were in a TIF
# change_table |> filter(first_tif_year > first_incent_year)

# 1534 PINs were in a TIF before they had an incentive classification
change_table |> filter(first_tif_year < first_incent_year)
```


```{r}
#| label: tbl-keypin-tiforincent
#| tbl-cap: "Projects that were in a TIF before they also got an incentive classification."


change_table <- ptax_proj |>
  ungroup() |> 
  arrange(pin, year) |> 
  group_by(main_keypin) |>
  summarize(first_tif_year = first(year[in_tif==1]),
            first_incent_year = first(year[incent_prop=="Incentive"])) |>
  filter(!is.na(first_tif_year) & !is.na(first_incent_year)) |>
  select(main_keypin, first_tif_year, first_incent_year)|> 
  distinct() 

# 156 projects were in a tif
change_table |> filter(first_tif_year > first_incent_year) 
# 397 projects were in a TIF before they received incentive classes
change_table |> filter(first_tif_year < first_incent_year) 
```


**How many Commercial PINs become Industrial PINs? If commercial pins are the comparison group, we should know the sample size of PINs that change to other land uses.**

```{r}
#| label: tbl-commercial-toOther-pins
#| tbl-cap: "Commercial PINs that became Industrial or other land use types"
#| eval: false
#| include: false

change_table <- ptax_proj |>
  ungroup() |>
  arrange(pin, year) |>
  group_by(pin) |>
  summarize(
    first_commerc_year = first(year[land_use == "Commercial"]),
    first_indust_year = first(year[land_use == "Industrial"]),
    first_other_year = first(year[!land_use %in% c("Commercial", "Industrial")]))  |>
  select(pin, first_commerc_year, first_indust_year, first_other_year)|> 
  distinct() 

# 864
change_table |> filter(first_commerc_year < first_indust_year)
  
# 7154
change_table |> filter(first_commerc_year < first_other_year)
```


```{r }
#| label: tbl-keypinchange
#| tbl-cap: "Projects where the first year as a commercial land use is before becoming commercial."


change_table <- ptax_proj |>
  ungroup() |>
  arrange(pin, year) |>
  group_by(main_keypin) |>
  summarize(
    first_commerc_year = first(year[land_use == "Commercial"]),
    first_indust_year = first(year[land_use == "Industrial"]),
    first_other_year = first(year[!land_use %in% c("Commercial", "Industrial")])) |>
  select(main_keypin, first_commerc_year, first_indust_year, first_other_year)|> 
  distinct() 

#713 
change_table |> filter(first_commerc_year < first_other_year)
  
# 318
change_table |> filter(first_commerc_year < first_indust_year)


```

```{r}
# ptax_proj %>% 
#   arrange(pin, year) %>%
#  # select(-c(landuse_change, land_use, incent_change, tif_change, pin_classes_2011, pin_classes_2022) ) %>%
#   filter(incent_change %in% c("Always Incentive", "Changes Sometime")) %>%
#            write_csv("Output/incentive_project_summaries.csv")
```


```{r blocksums, eval=FALSE}

block_sums <- ptax_proj %>% 
  mutate(block7 = str_sub(pin, 1,7)) %>%
  group_by(year, block7 ) %>%
  summarize(
    incent_pins = sum(ifelse(class >= 600 & class < 900, 1, 0)),
    project_pins = n_distinct(main_keypin),
    fmv = sum(fmv, na.rm=TRUE),
    tax_code = mean(tax_code_num, na.rm=TRUE),
    in_tif = mean(in_tif, na.rm=TRUE),
    # pin_classes = paste(sort(unique(class)), collapse = ", "),
    land_use = paste(sort(unique(land_use)), collapse = ", "),
    incent_prop = paste(sort(unique(incent_prop)), collapse = ", "),
    triad = first(Triad),
    clean_name = first(clean_name),
    reassessed_year = first(reassessed_year),
    exe_abate = sum(exe_abate, na.rm = TRUE),
   
    landuse_change= paste(sort(unique(landuse_change)), collapse = ", "),
    incent_change = paste(sort(unique(incent_change)), collapse = ", "),
    tif_change = paste(sort(unique(tif_change)), collapse = ", "),
    av_mailed = sum(av_mailed, na.rm=TRUE),
    av_clerk = sum(av_clerk, na.rm=TRUE) 
    )  %>%
   ungroup()

timespan = 12 

block_sums <- block_sums %>%
  
  group_by(block7) |>
  mutate(
    years_existed = n(),
    base_year_fmv_2011 = ifelse(years_existed == timespan, 
                                fmv[year == 2011], NA),
    end_year_fmv_2022 = ifelse(years_existed == timespan, fmv[year == 2022], NA),
    fmv_2022_2011_diff = ifelse(!is.na(end_year_fmv_2022) & !is.na(base_year_fmv_2011), end_year_fmv_2022 - base_year_fmv_2011, NA)
  ) %>% 
  ungroup() %>%
  mutate(fmv_growth_2011 = fmv/base_year_fmv_2011) |>
  group_by(year) %>%
  mutate(fmv_w = DescTools::Winsorize(fmv, quantile(fmv, probs = c(0.01, 0.99), na.rm=TRUE)),
         fmv_growth_2011_w = DescTools::Winsorize(fmv_growth_2011, quantile(fmv_growth_2011, probs = c(0.01, 0.99), na.rm=TRUE)),
         fmv_2022_2011_diff_w = DescTools::Winsorize(fmv_2022_2011_diff, quantile(fmv_2022_2011_diff, probs = c(0.01, 0.99), na.rm=TRUE))) |>
  ungroup()

block_sums <- block_sums %>%  mutate(
    # commented out to make it run faster 
    # landuse_change2 = ifelse(landuse_change!= "Always Commercial" & landuse_change != "Always Industrial", "Changes", landuse_change),
         land_use2 = ifelse(!land_use %in% c("Commercial", "Industrial", "Land", "Exempt"), "Multi Land Uses", land_use),
         # tif_change2 = ifelse(!tif_change %in% c("Always TIF", "Never TIF"), "Changes", tif_change),
        #  incent_change2 = ifelse(!incent_change %in% c("Always Incentive", "Never Incentive"), "Changes", incent_change),
         in_tif2 = ifelse(in_tif != 0 & in_tif != 1, "Partly in TIF", in_tif)) %>%
  mutate(across(c(fmv, tax_code), round, digits = 0))


block_panel <- block_sums %>% 
    filter(year >= 2012) %>% # drop first year of data when fmv growth is 1
  select(block7, year, fmv, fmv_w, incent_prop,
         land_use2, in_tif2, triad, fmv_growth_2011, years_existed,
        base_year_fmv_2011, reassessed_year, incent_pins,
         project_pins, incent_pins, clean_name, fmv_growth_2011_w)  %>% 
  arrange(block7, year) %>% 

  mutate(#across(year:triad, as.factor),
         land_use2 = as.character(land_use2),
         incent_prop2 = factor(incent_prop, levels= c("Non-Incentive", "Incentive"), labels = c("NonIncentive", "Incentive"), ordered = TRUE))  %>%
  ungroup() |>
  group_by(year) %>%
  mutate(fmv_w = DescTools::Winsorize(fmv, quantile(fmv, probs = c(0.005, 0.995), na.rm=TRUE)),
         fmv_growth_2011_w = DescTools::Winsorize(fmv_growth_2011, quantile(fmv_growth_2011, probs = c(0.01, 0.99), na.rm=TRUE)))

# table(proj_panel$incent_prop2)
table(block_panel$years_existed)

summary(block_panel$fmv_growth_2011)
summary(block_panel$fmv_growth_2011_w)
#write_csv(block_panel, "Output/block_panel_summaries.csv")


block_panel %>% filter(year == 2022) %>%
  summarize(pin_count = sum(project_pins))
```

```{r}

ptax_proj_summs <- ptax_proj %>% 
 # left_join(project_pins, by = "pin") |>
  filter(!is.na(main_keypin)) |>
 # mutate(main_keypin = ifelse(is.na(main_keypin), pin, main_keypin) ) |>
  
  group_by(year, main_keypin) %>%
  summarize(
    incent_pins = sum(ifelse(class_ptaxsim >= 600 & class_ptaxsim < 900, 1, 0)),
    project_pins = n(),
    fmv = sum(fmv, na.rm=TRUE),
    tax_code = mean(tax_code_num, na.rm=TRUE),
    in_tif = mean(in_tif, na.rm=TRUE),
    # pin_classes = paste(sort(unique(class)), collapse = ", "),
    land_use = paste(sort(unique(land_use)), collapse = ", "),
    incent_prop = paste(sort(unique(incent_prop)), collapse = ", "),
    triad = first(Triad),
    clean_name = first(clean_name),
    reassessed_year = first(reassessed_year),
    exe_abate = sum(exe_abate, na.rm = TRUE),
   
    landuse_change= paste(sort(unique(landuse_change)), collapse = ", "),
    incent_change = paste(sort(unique(incent_change)), collapse = ", "),
    #tif_change = paste(sort(unique(tif_change)), collapse = ", "),
    av_mailed = sum(av_mailed, na.rm=TRUE),
    av_clerk = sum(av_clerk, na.rm=TRUE) 
    )  %>%
   ungroup()

timespan = 12 

ptax_proj_summs <- ptax_proj_summs %>%
  
  group_by(main_keypin) |>
  mutate(
    years_existed = n(),
    # commenting out just to make it run faster
    # but maybe interesting for looking at types of PINs included in a project
   # pin_classes_2011 = ifelse(years_existed == timespan, pin_classes[year == 2011], NA),
    #pin_classes_2022 = ifelse(years_existed == timespan, pin_classes[year == 2022], NA),
    base_year_fmv_2011 = ifelse(years_existed == timespan, 
                                fmv[year == 2011], NA),
    end_year_fmv_2022 = ifelse(years_existed == timespan, fmv[year == 2022], NA),
    fmv_2022_2011_diff = ifelse(!is.na(end_year_fmv_2022) & !is.na(base_year_fmv_2011), end_year_fmv_2022 - base_year_fmv_2011, NA)
  ) %>% 
  ungroup() %>%
  mutate(fmv_growth_2011 = fmv/base_year_fmv_2011) |>
  group_by(year) %>%
  mutate(fmv_w = DescTools::Winsorize(fmv, quantile(fmv, probs = c(0.01, 0.99), na.rm=TRUE)),
         # fmv_growth_2011_w = DescTools::Winsorize(fmv_growth_2011, quantile(fmv_growth_2011, probs = c(0.01, 0.99), na.rm=TRUE)),
         # fmv_2022_2011_diff_w = DescTools::Winsorize(fmv_2022_2011_diff, quantile(fmv_2022_2011_diff, probs = c(0.01, 0.99), na.rm=TRUE)),
         # 
         
       #  base_year_fmv_2006_w =DescTools::Winsorize(base_year_fmv_2006,quantile(base_year_fmv_2006, probs = c(0.01,0.99), na.rm=TRUE)),
  #  fmv_growth_2006_w =DescTools::Winsorize(fmv_growth_2006, quantile(fmv_growth_2006, probs = c(0.01,0.99), na.rm=TRUE)),
    base_year_fmv_2011_w =
      DescTools::Winsorize(base_year_fmv_2011,
                           quantile(base_year_fmv_2011, probs = c(0.01,0.99), na.rm=TRUE)),
    fmv_growth_2011_w =
      DescTools::Winsorize(fmv_growth_2011,
                           quantile(fmv_growth_2011, probs = c(0.01,0.99), na.rm=TRUE))) |>

  ungroup() 

ptax_proj_summs <- ptax_proj_summs %>%
  mutate(
    # commented out to make it run faster 
    # landuse_change2 = ifelse(landuse_change!= "Always Commercial" & landuse_change != "Always Industrial", "Changes", landuse_change),
         land_use2 = ifelse(!land_use %in% c("Commercial", "Industrial", "Land", "Exempt"), "Multi Land Uses", land_use),
         # tif_change2 = ifelse(!tif_change %in% c("Always TIF", "Never TIF"), "Changes", tif_change),
        #  incent_change2 = ifelse(!incent_change %in% c("Always Incentive", "Never Incentive"), "Changes", incent_change),
     #    in_tif2 = ifelse(in_tif != 0 & in_tif != 1, "Partly in TIF", in_tif)
        ) %>%
  mutate(across(c(fmv, tax_code), round, digits = 0))   %>%
  filter(years_existed == 12) 


```


```{r eval=FALSE, include = FALSE}
ptax_proj_summs %>% 
 # select(-c(landuse_change, land_use, incent_change) ) %>%
           write_csv("Output/Pin to Project Files/project_summaries_2025_Nov10.csv")

```


```{r eval=FALSE}
# ptax_proj_summs %>% is.finite() %>% modelsummary::datasummary_skim()

ptax_proj_summs %>% 
#  select(-c(land_use, incent_prop, landuse_change, tif_change, incent_change)) %>% 
  modelsummary::datasummary_skim()

```


```{r eval=FALSE}
#| label: tbl-exemptprojects
#| tbl-cap: Projects that were exempt or became exempt.

summary(ptax_proj_summs$fmv_growth_2011)
summary(ptax_proj_summs$fmv_growth_2011_w)

ptax_proj_summs %>% filter(is.na(fmv_growth_2011))

exempt_anyyear <- ptax_proj_summs %>%
  filter(fmv == 0) # |> distinct(main_keypin)
# 4138 unique projects, 31,076 observations


lowFMV <- ptax_proj_summs %>%
  filter(fmv < 100 & fmv != 0)  # |> distinct(main_keypin)
# 206 unique projects, 1645 observations


# projects exempt in 2011. Cause infinite growth in DV
were_exempt <- ptax_proj_summs %>%
  group_by(main_keypin) |>
  filter(fmv == 0 & year == 2011) |> 
  arrange(desc(fmv), main_keypin)
were_exempt 
# 1822 unique projects exempt in 2011

# projects exempt in 2022. Cause DV to be 0. (not the worst thing)
became_exempt <- ptax_proj_summs %>%
  group_by(main_keypin) |>
  filter(fmv == 0 & year == 2022)  |> 
  arrange(desc(fmv), main_keypin)
# 3289 projects
became_exempt


#write_csv(were_exempt, "./Output/properties_were_exempt.csv")

#write_csv(became_exempt, "./Output/properties_became_exempt.csv")

ptax_proj_summs <- ptax_proj_summs %>% 
  filter(!main_keypin %in% were_exempt$main_keypin) %>% ## creates Infinite growth but these are the observations we do want to examine more for vacant to non-vacant properties.
  filter(!main_keypin %in% became_exempt$main_keypin) %>%
  #filter(!main_keypin %in% exempt_anyyear$main_keypin) %>%
 # filter(!main_keypin %in% lowFMV$main_keypin) %>%
  arrange(desc(fmv), main_keypin)

summary(ptax_proj_summs$fmv_growth_2011)
summary(ptax_proj_summs$fmv_growth_2011_w)

ptax_proj_summs %>% filter(year == 2022) %>%
  summarize(project_count = n_distinct(main_keypin),
            pin_count = sum(project_pins))

```

OLD NUMBERS: 

- 852 projects have a combined FMV of \$0 in 2011 but have value in 2022. (1,824 when filtering from the timeseries data instead of the "balanced panel" data.)  

- 2,294 PINs have FMV > $0 in 2011 but become tax exempt by 2022.  (3,289 when filtering from the timeseries data instead of "balanced panel" data)   


**Problem example 1:** 3714512 - DAVITA INC is a project where all PINs were exempt in 2011 and had FMV values of 0 but became an incentive project in 2015 with 800 level PINs. Using FMV growth from 2011 results in a growth of INF and would need to be dropped from the model.

**Problem example 2:** 7308877 - LYCEE FRANCAIS DE CHIC	 became tax exempt. Filtering out PINs


**Large FMV Growth PINS**:

- 17221101230000 for ACRON Navy Pier Hotel LLC was land class 100 in 2011, became 590 and gained value after it opened? 

- 17221101230000 had large growth but it should be added to a project. (PINs in project also experienced subdivision and changed PIN numbers?)


```{r eval=FALSE}
modelsummary::datasummary_skim(ptax_proj_summs)

ptax_proj_summs %>%  
  select(-c(land_use, incent_prop, landuse_change, tif_change, incent_change)) %>% 
  modelsummary::datasummary_skim("categorical")
```

## Fixed Effect Model using Projects

Create winsorized FMV growth at this stage, after removing properties the properties with infinite growth. Too many properties had Inf or 0's as values for the winsorizing to do anything. 



```{r make-proj-panel, eval = FALSE}
library(dplyr)
proj_panel <- ptax_proj_summs %>% 
    filter(year >= 2012) %>% # drop first year of data when fmv growth is 1
  select(main_keypin, year, fmv, fmv_w, incent_prop,
         land_use2, #in_tif2, 
         triad, fmv_growth_2011, years_existed,
        base_year_fmv_2011, reassessed_year, incent_pins,
         project_pins, incent_pins, clean_name, fmv_growth_2011_w)  %>% 
  arrange(main_keypin, year) %>% 

  mutate(
         land_use2 = as.character(land_use2),
         incent_prop2 = factor(incent_prop, levels= c("Non-Incentive", "Incentive"), labels = c("NonIncentive", "Incentive"), ordered = TRUE)
       )  %>%
  ungroup() |>
  group_by(year) %>%
  mutate(fmv_w = DescTools::Winsorize(fmv, quantile(fmv, probs = c(0.005, 0.995), na.rm=TRUE)),
         fmv_growth_2011_w = DescTools::Winsorize(fmv_growth_2011, quantile(fmv_growth_2011, probs = c(0.005, 0.995), na.rm=TRUE)))

write_rds(proj_panel, file = "Output/proj_panel2.RDS")
```


```{r read-proj-panel}

proj_panel <- read_rds("Output/proj_panel2.RDS")

# table(proj_panel$incent_prop2)
table(proj_panel$years_existed)

summary(proj_panel$fmv_growth_2011)
summary(proj_panel$fmv_growth_2011_w)
# write_csv(proj_panel, "Output/project_panel_summaries_2025Sept8.csv")


proj_panel %>% filter(year == 2022) %>%
  summarize(project_count = n_distinct(main_keypin),
            pin_count = sum(project_pins))
```



```{r}
#| label: tbl-growthcomparison-nonwinsor
#| tbl-cap: "Non-winsorized growth for projects."

models <- list(
  projectlevel_2FE = feols(fmv_growth_2011 ~  land_use2*incent_prop2 | 
                             main_keypin + year, 
                           panel.id = c("main_keypin", "year"),                                     
                        vcov = ~clean_name,
                           data = proj_panel)
                
)


modelsummary::modelsummary(model = models, output = "flextable",
                           stars = TRUE,
                           gof_omit = "AIC|BIC")
```


### With Reassessment interactions or Triad FE

```{r }
#| label: tbl-winsorized-reassessed
#| tbl-cap: "With reassessment years interaction with Winsorized FMV Growth"

models <- list(
  reassess_year = feols(fmv_growth_2011_w ~ land_use2*incent_prop2  + reassessed_year | 
                          main_keypin + clean_name + year, 
                        panel.id = c("main_keypin", "year"),                                     
                        vcov = ~clean_name,
data = proj_panel),
  
  reassess_interact = feols(fmv_growth_2011_w ~  land_use2*incent_prop2 + reassessed_year*incent_prop2 |
                              main_keypin  + year, 
                            panel.id = c("main_keypin", "year"),                                     
                        vcov = ~clean_name,
  data = proj_panel),
  
   triadFE1 = feols(fmv_growth_2011_w ~ land_use2*incent_prop2  + reassessed_year | 
                      main_keypin + year^triad, 
                    panel.id = c("main_keypin", "year"),                                     
                        vcov = ~clean_name,
 data = proj_panel),
  
  triadFE2 = feols(fmv_growth_2011_w ~  land_use2*incent_prop2 + reassessed_year*incent_prop2 |
                     main_keypin + year^triad, 
                   panel.id = c("main_keypin", "year"),                                    
                        vcov = ~clean_name,
  data = proj_panel)
)



modelsummary::modelsummary(model = models, 
                           output = "flextable", 
                          stars = TRUE,
                           gof_omit = "AIC|BIC")
```


```{r }
#| label: tbl-nonwinsorized-reassessed
#| tbl-cap: "With reassessment years interaction with FMV Growth"

models <- list(
  reassess_year = feols(fmv_growth_2011 ~ land_use2*incent_prop2  + reassessed_year | 
                          main_keypin  + year, 
                        panel.id = c("main_keypin", "year"),                                    
                        vcov = ~clean_name,
                        data = proj_panel),
  
  reassess_interact = feols(fmv_growth_2011 ~  land_use2*incent_prop2 + reassessed_year*incent_prop2 | 
                              main_keypin + year, 
                            panel.id = c("main_keypin", "year"),                                     
                        vcov = ~clean_name,
                            data = proj_panel),
  
   triadFE1 = feols(fmv_growth_2011 ~ land_use2*incent_prop2  + reassessed_year | main_keypin + year, panel.id = c("main_keypin", "year"),                                     
                        vcov = ~clean_name,
                    data = proj_panel),
  
  triadFE2 = feols(fmv_growth_2011 ~  land_use2*incent_prop2 + reassessed_year*incent_prop2 | 
                     main_keypin + year^triad, 
                   panel.id = c("main_keypin", "year"),                                     
                        vcov = ~clean_name,
                   data = proj_panel)
)



modelsummary::modelsummary(model = models, 
                           output = "flextable", 
                          stars = TRUE,
                           gof_omit = "AIC|BIC")
```


