---
title: "Incentive PROJECT Models"
format: html
---

## Fixed Effect Model using Project Level

```{r}
library(tidyverse)

ptax_pins <- read_csv("./Output/comm_ind_PINs_2011-2022_balanced.csv")

parcuniverse_keypins <- readxl::read_xlsx("./Inputs/parceluniverse_keypins_20240725.xlsx", 
                                          sheet = "keypins_20240725") %>%
  mutate(pin14 = str_pad(as.character(pin), width = 14, side = "left", pad = "0"),
         keypin = str_pad(as.character(proration_key_pin), width = 14, side = "left", pad = "0"),
         pin10 = str_sub(pin14,1,10),
         pin7 = str_sub(pin14,1,7), .before = "pin",) %>%
  select(-c(pin_7dig, pin, Column1)) %>%
  filter(class != "EX")

puni_keypins <- parcuniverse_keypins %>% 
  select(pin14, 
         puni_keypin = keypin, 
         proration_rate2022 = proration_rate, 
         mailing_name2022 = mailing_name)

# Workaround for identifying more project IDs. 
# Used Appeal ID to create unique identifier to group PINs.
bor <- read_csv("Output/borappeals.csv") %>%
  mutate(project_appellant = paste(project_id, sep = "-", appellant))

bor_pins <- bor %>% 
  group_by(pin) %>% 
  arrange(desc(tax_year)) %>%
  summarize(pin = first(pin),
            class_bor = list(class),
            taxyear_appealed = list(tax_year),
            appellant = first(appellant),
            project_id = first(project_id), 
            timesappealed = n() ) %>%
  mutate(proj_appeallant = paste(project_id, "-", appellant))

ptax_proj <- ptax_pins %>% 
  left_join(bor_pins) %>%
  left_join(puni_keypins, by = c("pin" = "pin14")) 

proj_xwalk <- read_csv("Output/all_keypins.csv")               
ptax_proj <- ptax_proj %>% left_join(proj_xwalk, by = "pin")


ptax_proj <- ptax_proj %>% 
  mutate(actual_keypin = ifelse(is.na(proj_appeallant), puni_keypin, proj_appeallant))%>%
  mutate(actual_keypin = ifelse(is.na(actual_keypin), keypin, actual_keypin)) %>%
  select(year, actual_keypin, pin, tax_code_num, class, class_bor, tax_code_rate, fmv, av_clerk, tax_bill_total, everything())
```


```{r}
ptax_proj_summs <- ptax_proj %>% 
  group_by(year, actual_keypin) %>%
  summarize(
    incent_pins = sum(ifelse(class >= 600 & class < 900, 1, 0)),
    project_pins = n(),
    fmv = sum(fmv, na.rm=TRUE),
    tax_code = mean(tax_code_num, na.rm=TRUE),
    in_tif = mean(in_tif, na.rm=TRUE),
    pin_classes = ifelse(project_pins < 10, list(class), list(unique(class))),
    land_use = paste(unique(land_use), collapse = ","),
    triad = first(Triad),
    exe_abate = sum(exe_abate, na.rm = TRUE),
    clean_name = first(clean_name),
    landuse_change= paste(unique(landuse_change), collapse = ","),
    incent_change = paste(unique(incent_change), collapse = ","),
    tif_change = paste(unique(tif_change),collapse = ","),
    ) %>%
  ungroup() |>
  group_by(actual_keypin) |>
  mutate(
    pin_classes_2011 = pin_classes[year == 2011],
    pin_classes_2022 = pin_classes[year == 2022],
    base_year_fmv_2011 = fmv[year == 2011],
    end_year_fmv_2022 = fmv[year == 2022],
    fmv_growth_2011 = fmv/base_year_fmv_2011,
    fmv_2022_2011_diff = end_year_fmv_2022 - base_year_fmv_2011,
   # land_use = paste(unlist(land_use), collapse = ","),
    #incent_change = paste(unlist(incent_change), collapse = ","),
   # landuse_change = paste(distinct(unlist(landuse_change)), collapse = ",")
    ) %>% 
  ungroup()

were_exempt <- ptax_proj_summs %>%
  group_by(actual_keypin) |>
  filter(fmv == 0 & year == 2011)
were_exempt

became_exempt <- ptax_proj_summs %>%
  group_by(actual_keypin) |>
  filter(fmv == 0 & year == 2022)  |> arrange(desc(project_pins))
became_exempt


ptax_proj_summs <- ptax_proj_summs %>% 
  filter(!actual_keypin %in% were_exempt$actual_keypin) %>% ## creates Infinite growth but these are the observations we do want to examine more for vacant to non-vacant properties.
  filter(!actual_keypin %in% became_exempt$actual_keypin)

```

852 projects have a combined FMV of \$0 in 2011 but have value in 2022. 2,294 PINs have FMV > $0 in 2011 but become tax exempt by 2022. 


> Problem example 1: 3714512 - DAVITA INC is a project where all PINs were exempt in 2011 and had FMV values of 0 but became an incentive project in 2015 with 800 level PINs. Using FMV growth from 2011 results in a growth of INF and would need to be dropped from the model.

Problem example 2: 7308877 - LYCEE FRANCAIS DE CHIC	 became tax exempt. Filtering out PINs

> making binary variable for if a PIN went from vacant to not-vacant or not might work in a logit model though.

```{r}
#| eval: false 

ptax_proj <- ptax_proj %>%
  mutate(#incent_change = as.factor(incent_change),
        # landuse_change = as.factor(landuse_change),
       #  triad = as.factor(Triad),
       #  in_tif = as.factor(in_tif),
         land_use = as.factor(land_use),
         incent_prop = as.factor(incent_prop),
         clean_name = as.factor(clean_name),
         fmv_growth_2011 = round(fmv_growth_2011)) |>
  
  mutate(incent_change = relevel(incent_change, ref = "Never Incentive"),
         landuse_change = relevel(landuse_change, ref = "Always Commercial"),
         incent_prop = relevel(incent_prop, ref = "Non-Incentive"),
         
         #triad = relevel(triad, ref = "North"),
         land_use = relevel(land_use, ref = "Commercial")
  ) |>
  # Create binary variables
  mutate(incent_change_bin = as.factor(ifelse(incent_change == 
                                                "Changes Sometime", 1, 0)),
         prop_use_change_bin = as.factor(ifelse(landuse_change == 
                                                  "Changes Land Use", 1, 0)),
         is_chicago = as.factor(ifelse(clean_name == "Chicago", 1, 0))
  ) |>
  rename(fmv_2011 = base_year_fmv_2011) |>
  arrange(pin, year)
```

```{r eval=FALSE}
library(plm)

proj_panel <- ptax_proj_summs %>% 
  select(actual_keypin, year, in_tif, incent_pins, project_pins, 
         fmv_growth_2011, base_year_fmv_2011)  %>% 
  (is.numeric(fmv_growth_2011)) %>% 
  arrange(actual_keypin, year) %>% 
  filter(!is.na(actual_keypin ) ) 

proj_panel <- proj_panel %>%
  pdata.frame(index = c("actual_keypin", "year"))

plm(fmv_growth_2011 ~   incent_pins + project_pins + in_tif, 
    data = proj_panel,
    model = "within",
    effect = "twoways") %>% summary()
```

