---
title: "All Muni Calculations"
author: "Alea Wilbur"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    df_print: paged
---

```{r setup, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)


library(tidyverse)
library(DBI)
library(data.table)
library(ggspatial)
library(gstat)
library(here)
library(httr)
library(jsonlite)
library(ptaxsim)
library(sf)
library(stars)
library(glue)

# Create the DB connection with the default name expected by PTAXSIM functions
ptaxsim_db_conn <- DBI::dbConnect(RSQLite::SQLite(), "./ptaxsim.db/ptaxsim-2021.0.4.db")

# has all potential property classes for pins
# downloaded from CCAO gitlab website
## I used this to merge additional information to the pins and class data later on.
class_dict <- read_csv("class_dict.csv")

options(digits=4, scipen = 999)

```

[Example from gitlab site](https://ccao-data-science---modeling.gitlab.io/packages/ptaxsim/articles/reassessment.html#chicago-ward)

Cook County Assessors Data description page and other [datasets](https://datacatalog.cookcountyil.gov/stories/s/i22y-9sd2).

# Data Prep

Pull all agency names that exist, then use agency numbers associated with MUNI types to pull only the `muni_agency_names` object.

There are 946 unique taxing agencies that existed in 2021.

```{r agency-dt}
# has EAV values, extensions by agency_num
agency_dt <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT *
  FROM agency
  WHERE year = 2021
  "
)

# show variables that exist in the data table
agency_dt
```

This table is used for the municipality levy in later calculations.

`agency_dt` has all taxing agencies (BUT NOT TIFS!) that existed in 2021 and includes their total taxable base (cty_cook_eav), their levy, taxing rate, binary variables for if a municipality is home rule or not, as well as many other variables. Also currently doesn't have agency_names to go with the agency numbers. That will be merged in shortly.

The table has waaaay too many columns to display all at once.Click the arrow on the right-hand side of the table to see additional columns that are hidden initially in the display.

```{r all-muni-names}
# grabs all unique muni names. Would be needed if creating a loop for calculating all munis
# municipality names and their agency number
muni_agency_names <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT DISTINCT agency_num, agency_name, minor_type
  FROM agency_info
  WHERE minor_type = 'MUNI'
  OR agency_num = '020060000'  

  "
)

muni_agency_names <- muni_agency_names %>% 
  mutate(first6 = str_sub(agency_num,1,6),
         first5 = str_sub(agency_num,1,5)) %>% 
  select(-minor_type)

```

Makes a list of ALL taxing agencies, including TIFs, SSAs, etc.

First 5 and 6 digit codes variables are created to nest TIFs and other taxing agencies within the municipality. Still experimental, variables are not used in the code below.

There are 1,878 taxing agencies. When grouped by minor_type, there are 133 muni agencies, 639 tif agencies, 30 townships, etc.

```{r all-taxingagency-names}
# all agency names, numbers, and types
# includes TIF and non-TIF agencies
all_taxing_agencies <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT agency_num, agency_name, major_type, minor_type
  FROM agency_info
  "
)

# create same first 5 and 6 digit codes 
all_taxing_agencies <- all_taxing_agencies %>%
  mutate(first6 = str_sub(agency_num,1,6),
         first5 = str_sub(agency_num,1,5))

all_taxing_agencies

# 133 muni agencies, 639 tif agencies, 30 townships, etc.
# we use the 133 muni numbers and Cicero's agency number to get the 134 municipalities. 
table(all_taxing_agencies$minor_type)



table(all_taxing_agencies$major_type)
```

All agencies have an agency name, but there isn't an indicator for which municipality each taxing agency is within. Merging the `muni_agency_names` to `all_taxing_agencies` hopefully adds a variable that allows for grouping and summing at the municipality level that includes SSAs, (if we want to do that). Does not work for School districts or other taxing districts which begin with completely different values (e.g. school districts begin with 4)

> When people are saying "the municipality" are they mentally picturing the services provided with property taxes like parks, libraries, schools, etc. Or just the singular taxing agency associated with each municipality that only contains a few goods/services provided by the area? Is this even a good comparison between cities? More thoughts below:

-   Are all libraries paid for library taxing agencies or do some cities just include the cost of a public library as part of their municipal levy? Same for Parks

```{r all-taxingagency-names2}
# none of these are included in their geographic municipality areas automatically. have different agency number code beginnings too. 
all_taxing_agencies %>% 
  filter(major_type == "MISCELLANEOUS") %>% 
  group_by(minor_type) %>%
  summarize(count = n()) 
#all_taxing_agencies %>% filter(minor_type == "FIRE") 
```


Add agency names since those weren't included originally in the table:
```{r all-taxingagency-names3}
all_taxing_agencies <- all_taxing_agencies %>% 
  left_join(muni_agency_names, by = c("first5", "first6")) %>% 
  rename(muni_name =  agency_name.y,
        muni_num = agency_num.y,
        agency_name = agency_name.x,
        agency_num = agency_num.x)



# combine taxing agency names and agency type to data table that has eav and extension values
agency_data <- right_join(agency_dt, all_taxing_agencies) %>% 
  # get rid of unneeded columns to make table outputs smaller
  select(-c(cty_dupage_eav:cty_livingston_eav, lim_numerator, lim_denominator)) %>% # drop some of the unused variables
  arrange(first5, agency_num)

```

Using the agency numbers for each municipality, I pull all tax codes that have an agency_num included in the muni_agency_names object. By narrowing the agencies down to just Municipality types, this prevents duplicate tax_codes from being pulled.

There are 3774 tax codes within Cook County that are taxed by Municipalities.

```{r municiple-taxcodes}
# muni_agency_nums<- all_taxing_agencies %>% 
#   filter(minor_type %in% c("MUNI")) %>%
#   select(agency_num)

muni_tax_codes <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql("
  SELECT*
  FROM tax_code
  WHERE agency_num IN ({muni_agency_names$agency_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
)
```

## Levies


```{r}
muni_levies <- agency_data %>% 
  filter(minor_type == "MUNI" | 
           agency_num == "020060000"  # Cicero is an oddball and considered Township

  )

# variables that exist for the 134 municipalities for levies and extensions values 
# originally from agency dt
muni_levies
```

## TIF Dummy variable for tax codes

I find all TIF taxing agencies and tax codes that are taxed by TIF agencies.

TFI_distrib table total EAV within each tax code, the frozen EAV amount, tax code rate, and tax code revenue variables. Super helpful table, however, I don't directly use the table.

```{r}
# Agency number and agency name for all TIFs
TIF_agencies <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT DISTINCT agency_num, agency_name, major_type, minor_type
  FROM agency_info
  WHERE minor_type = 'TIF'
  "
)

unique_tif_taxcodes <- DBI::dbGetQuery(
  ptaxsim_db_conn, 
  glue_sql("
  SELECT DISTINCT tax_code_num
  FROM tax_code
  WHERE agency_num IN ({TIF_agencies$agency_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
)



tif_distrib <- DBI::dbGetQuery(
  ptaxsim_db_conn, 
  glue_sql("
  SELECT *
  FROM tif_distribution
  WHERE tax_code_num IN ({muni_tax_codes$tax_code_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
  )
)

tif_distrib

```

Using the TIF distribution tables and tax_code tables included in ptaxsim, I calculate the total revenue collected by each tax code (and the amount that whent to TIFs or not.)
```{r}
tif_distrib <- tif_distrib %>% 
  mutate(#first6= str_sub(agency_num,1,6),
         first5 = str_sub(agency_num,1,5),
         #tif_increment = tax_code_eav - tax_code_frozen_eav,
         total_taxcode_revenue = tax_code_eav*tax_code_rate/100,
         district_revenue_withinTIF = tax_code_frozen_eav*tax_code_rate/100,
         # checking to see if equal to tax_code_revenue
         # same but has the negative revenue values instead of zeros.
         tif_revenue_withinTIF = total_taxcode_revenue - district_revenue_withinTIF)

# add 
#tif_distrib <- tif_distrib %>% left_join(muni_tax_codes) 
tif_distrib

```


## Pins 

There are 1,825,816 pins in 2021. 

```{r eval=FALSE}
muni_pins <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  glue_sql(
  "SELECT DISTINCT pin, class, tax_code_num
  FROM pin
  WHERE tax_code_num IN ({muni_tax_codes$tax_code_num*})
  AND year = 2021
  ",
  .con = ptaxsim_db_conn
))

# change variable type to character so the join works.
class_dict$class_code <- as.character(class_dict$class_code)

# finds all pins within a municipality
pin_data <- lookup_pin(2021, muni_pins$pin) %>%
  left_join(muni_pins, by = c("pin", "class"))


# use the property class to make the major property types
# joins the class_dict file to the pin_data classes
pin_data <- class_dict %>%
  select(-c(assessment_level:reporting_group, class_desc:max_size)) %>%
  right_join(pin_data, by = c("class_code" = "class"))

write_csv(pin_data, "all_pins_2021.csv")

exemptions_by_class_per_TC <- pin_data %>%
  group_by(tax_code_num, major_class_code, major_class_type) %>%
  summarize(eav=sum(eav, na.rm=TRUE),
            exe_homeowner = sum(exe_homeowner, na.rm=TRUE),
            exe_senior = sum(exe_senior, na.rm=TRUE),
            exe_freeze = sum(exe_freeze, na.rm=TRUE),
            exe_longtime_homeowner = sum(exe_longtime_homeowner, na.rm=TRUE),
            exe_disabled = sum(exe_disabled, na.rm=TRUE),
            exe_vet_returning = sum(exe_vet_returning, na.rm=TRUE),
            exe_vet_dis = sum(exe_vet_dis_lt50 + exe_vet_dis_50_69 + exe_vet_dis_ge70, na.rm=TRUE),
            exe_abate = sum(exe_abate, na.rm=TRUE),
            pin_count = n() # number of pins within each tax code and property combo

)
write_csv(exemptions_by_class_per_TC, "all_exemptions_by_TC.csv")
```

## Tax Bills

The tax_bill command returns informatin for erach taxing agency that taxed a pin. EACH pin can be taxed by 15+ taxing agencies and each taxing agency has its own taxing rate. 

Each tax code has a unique combination of taxing agencies taxing properties within in. So, each tax code has a summed EAV value from each pin within it and a "composite tax rate" made from the taxing agencies and their taxable base. In the code below, its called the weighted tax rate but that isn't the best name for it... It is more like the amount of money that each taxing agency gets from each property.

`revenue collected`(previously named weighted_tax_rate in earlier code versions) was calculated for every tax bill observation for every pin and is equal to the eav of a property X agency_tax_rate. Then all pins within tax codes were summed together to get a tax code level EAV and tax code level tax rate. _Note: There is a tax_code_rate variable included in the `muni_tax_codes` object made from the `tax_code` dt included in ptaxsim_ 


```{r pull-taxbills, eval =FALSE}
# creates the 4.5 GB file of all taxbills for 2021. 
bills <- tax_bill(2021, muni_pins$pin, simplify = FALSE)

bills_byClass_pertaxcode <- bills %>%
  left_join(class_dict, by = c("class" = "class_code")) %>% # add major property types to pin data
  mutate(revenue_collected = agency_tax_rate*eav) %>% # $ that each taxing agency gets from each pin
  group_by(major_class_code, major_class_type, tax_code) %>%
  summarize(
            final_tax_to_dist = sum(final_tax_to_dist, na.rm=TRUE),
            final_tax_to_tif = sum(final_tax_to_tif, na.rm=TRUE),
            tax_amt_exe = sum(tax_amt_exe, na.rm=TRUE),
            tax_amt_pre_exe = sum(tax_amt_pre_exe, na.rm=TRUE),
            tax_amt_post_exe = sum(tax_amt_post_exe, na.rm=TRUE),
        agency_tax_rate = mean(agency_tax_rate, na.rm=TRUE),
        rpm_tif_to_cps =sum(rpm_tif_to_cps, na.rm=TRUE),
        rpm_tif_to_rpm = sum(rpm_tif_to_rpm, na.rm=TRUE),
        rpm_tif_to_dist = sum(rpm_tif_to_dist, na.rm=TRUE),
           tif_share = mean(tif_share, na.rm=TRUE),
        rev_perTC = sum(revenue_collected, na.rm = TRUE))

# na.rm = TRUE option is probably only needed for claculating the mean values of variables. I just didn't delete it yet. Avoids getting an error if there is a zero when calculating mean.

write_csv(bills_byClass_pertaxcode, "taxbills_byPropClass_perTC.csv")

write_csv(bills, "all_taxbills.csv")

```

# Combining Data

## Tax Bills and Levies for Taxcodes and Municipalities

Tax bill data was pulled from ptaxsim using the tax_bill() function for all pins within Cook County in 2021. This was a 4.5 GB file. In order to make the file manageable and combine it with exemption data for all municipalities, I summarized values to the Municipality-TaxCode-PropertyClass level.

Tax bills by property class has 9716 observations. (3774 tax codes X multiple property types).



```{r}
taxbills_by_Class_per_TC <- read_csv("taxbills_byPropClass_perTC.csv") %>%
  mutate(tax_code = as.character(tax_code),
         rev_perTC = round(rev_perTC)) %>%
  left_join(muni_tax_codes, by = c("tax_code" = "tax_code_num"))


taxbills_by_Class_per_TC <- left_join(taxbills_by_Class_per_TC , muni_agency_names)

taxbills_by_Class_per_TC 


taxbills_by_Class_per_TC %>% mutate(
                            "Current Tax Revenue(District+TIF)" = scales::dollar(tax_amt_post_exe),
                            "District Revenue" = scales::dollar(final_tax_to_dist),
                            "TIF Revenue" = scales::dollar(final_tax_to_tif),
                            "Lost Revenue from Exempt." = scales::dollar(tax_amt_exe), # same as tax_amt_pre_exe-tax_amt_post_exe
                            "Revenue Collected" = rev_perTC # within each Taxcode
                          
                            ) %>% 
 # mutate(`EAV in and out of TIFs` = `EAV in TIF (Increment)`+`EAV outside TIF`,
 #            EAVbeforeExemptions = `EAV in and out of TIFs`+Exemptions) %>%
  select(-c(tax_amt_post_exe, final_tax_to_dist:tax_amt_pre_exe, first5, first6, year, rev_perTC, rpm_tif_to_cps:rpm_tif_to_dist, agency_tax_rate, tif_share, major_class_code)) 


#class_exemptions <- full_join(exemptions_by_class,  class_eav, by = "major_class_code") %>%
#  select(major_class_code, major_class_type, eav, final_tax_to_dist, tax_share)
#class_exemptions
```


```{r}
# all munis + Cicero
MuniLevy <- taxbills_by_Class_per_TC %>% 
  group_by(agency_name)%>%
  summarize(MuniLevy = round(sum(final_tax_to_dist)),
            Levy_plus_TIFrev =round(sum(final_tax_to_dist + final_tax_to_tif)))

MuniLevy
```

Comparison of ptaxsim/Alea calculations and the numbers for the Levy value in Josh Drucker's Excel files. Pretty close but not perfect matches (which is fine, I'm not surprised by that).

Chicago: 6,536,952,735 vs. 6,490,629,279 JD
Bridgeview: 65,534,363 vs. 65,133,190 JD
Dolton: 54,957,926 vs. 56,755,459 JD
Hoffman Estates: 177,316,159	vs. 182,363,626 JD
Midlothian: 30,988,554 vs 30,877,423 JD
Oak Park: 234,665,831 vs 233,132,867 JD.




## Exemptions summarized by Tax Code

Uses exemption data from the lookup_pin() command for all pins in Cook County in 2021. Values are then summarized by Tax codes and property types (so each Municipality has a lot of rows at this point BUT this stage of the data is the most flexible for calculations and future research questions.)


```{r}
exemptions_by_class_per_TC <- read_csv("all_exemptions_by_TC.csv") %>% 
  mutate(tax_code_num = as.character(tax_code_num))


muni_tax_codes <- full_join(muni_tax_codes, muni_agency_names)

# make tax code variables character formatted for the join
muni_tax_codes$tax_code_num <- as.character(muni_tax_codes$tax_code_num)
exemptions_by_class_per_TC$tax_code_num <- as.character(exemptions_by_class_per_TC$tax_code_num)


# make dummy variable for if a tax code is a tif tax code or not
exemptions_by_class_per_TC <- left_join(exemptions_by_class_per_TC, muni_tax_codes, by =c("tax_code_num")) %>%
  mutate(in_TIF = ifelse(tax_code_num %in% unique_tif_taxcodes$tax_code_num, 1,0) )

exemptions_by_class_per_TC
```


```{r}
exemptions_by_class_per_TC <- left_join(exemptions_by_class_per_TC, muni_levies, by = c("agency_name", "agency_num", "year")) %>%
  mutate(in_TIF = ifelse(tax_code_num %in% unique_tif_taxcodes$tax_code_num, 1,0) )

exemps_inTIF_byClass_perMuni <- exemptions_by_class_per_TC %>% 
  mutate(taxamount = tax_code_rate/100*eav) %>%
  group_by(agency_name, agency_num, major_class_code, in_TIF,
           major_class_type, agency_rate) %>%
  summarize(taxamount_perClass = round(sum(taxamount)),
            eav_pre_exe_perClass  = sum(eav, na.rm=TRUE),
            exe_homeowner = sum(exe_homeowner, na.rm=TRUE),
            exe_senior = sum(exe_senior, na.rm=TRUE),
            exe_freeze = sum(exe_freeze, na.rm=TRUE),
            exe_longtime_homeowner = sum(exe_longtime_homeowner, na.rm=TRUE),
            exe_disabled = sum(exe_disabled, na.rm=TRUE),
            exe_vet_returning = sum(exe_vet_returning, na.rm=TRUE),
            exe_vet_dis = sum(exe_vet_dis, na.rm=TRUE),
            exe_abate = sum(exe_abate, na.rm=TRUE)
           ) %>%
  
  mutate(Exemp_in_PropClass = exe_homeowner + exe_senior+ exe_freeze +
           exe_longtime_homeowner + exe_disabled+ exe_vet_returning +exe_vet_dis + exe_abate,
          
            EAV_post_exe_PerClass = eav_pre_exe_perClass-Exemp_in_PropClass, 
         tax_share_post_exemp = taxamount_perClass/ EAV_post_exe_PerClass #current situation
         )

exemps_inTIF_byClass_perMuni
```


Table below has the EAV (pre and post exemptions), total exemption amounts, agency name, property class types,j and current tax burden for each Property Class within Municipality. 

```{r}
summtable <- exemps_inTIF_byClass_perMuni %>% 
  select(agency_name, major_class_code, in_TIF , tax_share_post_exemp, taxamount_perClass, 
         eav_pre_exe_perClass, Exemp_in_PropClass, EAV_post_exe_PerClass, agency_num, agency_rate, major_class_type)
summtable
```


```{r}
exemptions_perMuni <- exemps_inTIF_byClass_perMuni %>%
  group_by(agency_name, agency_num, agency_rate) %>% 
  summarize(taxamount_perMuni = sum(taxamount_perClass),
            eav_pre_exe_perMuni = sum(eav_pre_exe_perClass),
            TotalExemptions=sum(Exemp_in_PropClass),
            HomeownerExemps = sum(exe_homeowner),
         ResidentialExemp = sum(exe_homeowner +exe_senior+exe_freeze + exe_longtime_homeowner
                                + exe_disabled+ exe_vet_returning +exe_vet_dis, na.rm=TRUE),
         CommercialExemp = sum(exe_abate, na.rm=TRUE)) %>%
  mutate(
         eav_post_exe_perMuni = eav_pre_exe_perMuni-TotalExemptions) %>% 
  arrange(agency_num)

exemptions_perMuni

EAV_tifs_and_exemps_byClass_andMuni <- summtable %>% 
  left_join(exemptions_perMuni , by = c("agency_name", "agency_num", "agency_rate"))
```

# Scenarios

## Current Tax Burden and if there were no exemptions 

Current burden is the % of the levy paid by each property class.

Assumption: Levy does not change - so the money the municipality needs divided by the pre_exe_eav would be the new tax rate if exemptions didn't exist. 


```{r}
# 1602 rows: 134 munis with around 5 or more property types split between tif and nontif eav values. 
burden_shift_byMuni_PropClass <- EAV_tifs_and_exemps_byClass_andMuni %>% #ungroup() %>% 
  select(agency_name, major_class_type, in_TIF, agency_rate, eav_pre_exe_perClass, EAV_post_exe_PerClass, agency_num, eav_pre_exe_perMuni, eav_post_exe_perMuni) %>% 
  left_join(MuniLevy) %>%
  mutate(noexemps_agency_rate = MuniLevy / eav_pre_exe_perMuni,
         exemps_agency_rate = MuniLevy / eav_post_exe_perMuni)
burden_shift_byMuni_PropClass

burden_shift_byMuni_PropClass <- burden_shift_byMuni_PropClass%>%
  group_by(agency_name, agency_num, major_class_code, major_class_type, in_TIF#, noexemps_agency_rate, exemps_agency_rate
           ) %>%
   summarize(agency_rate_no_exemps = noexemps_agency_rate,
             agency_rate_w_exemps= exemps_agency_rate,
             MuniLevy = MuniLevy, 
             eav_pre_exe_perClass = eav_pre_exe_perClass,
             EAV_post_exe_PerClass = EAV_post_exe_PerClass,
             eav_post_exe_perMuni = eav_post_exe_perMuni, 
             eav_pre_exe_perMuni = eav_pre_exe_perMuni) %>% 
  mutate(
         # the current revenue collected with current exemptions
     rev_w_exemps = round(agency_rate_w_exemps* EAV_post_exe_PerClass), 
     
       # revenue collected if rate didn't change but the eav did
     inbetween_step = round(agency_rate_w_exemps * eav_pre_exe_perClass),
     
    # revenue collected if there were no exemptions. Rate changes because levy didn't change
     rev_w_no_exemps = round(agency_rate_no_exemps * eav_pre_exe_perClass)
    ) %>%
     

     

  ungroup() %>%
  ungroup() %>%
  mutate(
       # percent of levy paid by each property class based on current exemptions
     tax_burden_current = round(EAV_post_exe_PerClass * agency_rate_w_exemps/ MuniLevy, digits = 5),
     
     # for indirect effect
      tax_burden_inbetween = round(eav_pre_exe_perClass* agency_rate_w_exemps/MuniLevy, digits = 5),
     
     # percent of levy paid by each property class if there were no exemptions
     tax_burden_noexemps = round(eav_pre_exe_perClass *agency_rate_no_exemps / MuniLevy, digits = 5),
     

     # difference in percent of the levy being paid by each property class

    overall_burden_shift = tax_burden_noexemps - tax_burden_current,
    direct_effect = tax_burden_current - tax_burden_inbetween,
    indirect_effect = tax_burden_inbetween - tax_burden_noexemps ) 


burden_shift_byMuni_PropClass

write.csv(burden_shift_byMuni_PropClass, "burden_shift_allPropertyClasses_allMunis.csv")
```


Note: Burden shift is a percentage point change!

Because of current exemptions (all types), class 2 properties are currently shifting 4% of the levy on to other property groups in Berwyn.

Comparing the origial 6 municipalities in the JD excel files:


```{r}
burden_shift_byMuni_PropClass %>% filter(agency_num %in% c("030120000", "030210000", "030800000", "030920000","030580000","030310000") &
                                           major_class_code == "2"& in_TIF==0) %>% 
  select(-c(in_TIF, agency_num, major_class_code))


## Only includes EAV OUTSIDE of TIFs!
burden_shift_byMuni_PropClass %>% filter(major_class_code == "2"& in_TIF==0) %>% 
  select(-c(in_TIF, agency_num, major_class_code)) %>% write.csv("residential_burden_shift_all_munis.csv")


```

__Interpretation: __

2.5% of the total levy for Chicago that would be paid by class 2 properties is being shifted to other property types due to current exemptions.

Nearly 9% of the total levy in Dolton is currently shifted from class 2 residential properties to other types of properties due to current exemptions in place. 


_Probably ignore table below: More complicated version that includes the amoutn inside and outside of TIFs. Probably just stick with the table above. (Same data, but I filter the data so only property outside of TIFs is included)_
```{r}
burden_shift_byMuni_PropClass %>% filter(agency_num %in% c("030120000", "030210000", "030800000", "030920000","030580000","030310000")) 

```



## If there were no Homeowners Exemptions


Main difference for this chunk is that all exemptions except homeowners exemptions are subtracted from the EAV value to calculate the post exemption value. 

```{r}

exemptions_by_class_per_TC <- read_csv("all_exemptions_by_TC.csv") %>% 
  mutate(tax_code_num = as.character(tax_code_num))


muni_tax_codes <- full_join(muni_tax_codes, muni_agency_names)

# make tax code variables character formatted for the join
muni_tax_codes$tax_code_num <- as.character(muni_tax_codes$tax_code_num)
exemptions_by_class_per_TC$tax_code_num <- as.character(exemptions_by_class_per_TC$tax_code_num)


# make dummy variable for if a tax code is a tif tax code or not
exemptions_by_class_per_TC <- left_join(exemptions_by_class_per_TC, muni_tax_codes, by =c("tax_code_num")) %>%
  mutate(in_TIF = ifelse(tax_code_num %in% unique_tif_taxcodes$tax_code_num, 1,0) )

exemptions_by_class_per_TC <- left_join(exemptions_by_class_per_TC, muni_levies, by = c("agency_name", "agency_num", "year")) %>%
  mutate(in_TIF = ifelse(tax_code_num %in% unique_tif_taxcodes$tax_code_num, 1,0) )

exemps_inTIF_byClass_perMuni <- exemptions_by_class_per_TC %>% 
  mutate(taxamount = tax_code_rate/100*eav) %>%
  group_by(agency_name, agency_num, major_class_code, in_TIF,
           major_class_type, agency_rate) %>%
  summarize(taxamount_perClass = round(sum(taxamount)),
            eav_pre_exe_perClass  = sum(eav, na.rm=TRUE),
            exe_homeowner = sum(exe_homeowner, na.rm=TRUE),
            exe_senior = sum(exe_senior, na.rm=TRUE),
            exe_freeze = sum(exe_freeze, na.rm=TRUE),
            exe_longtime_homeowner = sum(exe_longtime_homeowner, na.rm=TRUE),
            exe_disabled = sum(exe_disabled, na.rm=TRUE),
            exe_vet_returning = sum(exe_vet_returning, na.rm=TRUE),
            exe_vet_dis = sum(exe_vet_dis, na.rm=TRUE),
            exe_abate = sum(exe_abate, na.rm=TRUE)
           ) %>%
  
  mutate(Exemp_in_PropClass = exe_homeowner + exe_senior+ exe_freeze + exe_longtime_homeowner + exe_disabled+ exe_vet_returning +exe_vet_dis + exe_abate,
          
         # if only homeowners exemptions were removed
         # adds
        EAV_post_exe_PerClass = eav_pre_exe_perClass - Exemp_in_PropClass + exe_homeowner, 

         )

# Now EAV post exemptions is only after homeowner exemptions are  
exemps_inTIF_byClass_perMuni
```


```{r}
summtable <- exemps_inTIF_byClass_perMuni %>% 
  select(agency_name, major_class_code, in_TIF , taxamount_perClass, 
         eav_pre_exe_perClass, Exemp_in_PropClass, EAV_post_exe_PerClass, agency_num, agency_rate, major_class_type)

```


```{r}
exemptions_perMuni <- exemps_inTIF_byClass_perMuni %>%
  group_by(agency_name, agency_num, agency_rate) %>% 
  summarize(taxamount_perMuni = sum(taxamount_perClass),
            eav_pre_exe_perMuni = sum(eav_pre_exe_perClass),
            TotalExemptions=sum(Exemp_in_PropClass),
            HomeownerExemps = sum(exe_homeowner),
         ResidentialExemp = sum(exe_homeowner +exe_senior+exe_freeze + exe_longtime_homeowner
                                + exe_disabled+ exe_vet_returning +exe_vet_dis, na.rm=TRUE),
         CommercialExemp = sum(exe_abate, na.rm=TRUE)) %>%
  mutate(
         eav_post_exe_perMuni = eav_pre_exe_perMuni-TotalExemptions+HomeownerExemps) %>% 
  arrange(agency_num)

exemptions_perMuni %>%
  select(agency_name, eav_pre_exe_perMuni,HomeownerExemps, eav_post_exe_perMuni, agency_rate )

EAV_tifs_and_exemps_byClass_andMuni <- summtable %>% 
  left_join(exemptions_perMuni , by = c("agency_name", "agency_num", "agency_rate")) 
```


```{r}
burden_shift_byMuni_PropClass <- EAV_tifs_and_exemps_byClass_andMuni %>% #ungroup() %>% 
  select(agency_name, major_class_type, in_TIF, agency_rate, HomeownerExemps, eav_pre_exe_perClass, EAV_post_exe_PerClass, agency_num, eav_pre_exe_perMuni, eav_post_exe_perMuni) %>% 
 
   left_join(MuniLevy) %>%
  
  mutate(noexemps_agency_rate = MuniLevy / eav_pre_exe_perMuni, 
         exemps_agency_rate = MuniLevy / eav_post_exe_perMuni # if all exemptions existed except hoemowners exemption
         ) %>%
  
  group_by(agency_name, agency_num, major_class_code, major_class_type, in_TIF#, noexemps_agency_rate, exemps_agency_rate
           ) %>%
   summarize(agency_rate_no_exemps = noexemps_agency_rate,
             agency_rate_w_exemps= exemps_agency_rate,
             MuniLevy = MuniLevy, 
             eav_pre_exe_perClass = eav_pre_exe_perClass,
             EAV_post_exe_PerClass = EAV_post_exe_PerClass,
             eav_post_exe_perMuni = eav_post_exe_perMuni, 
             eav_pre_exe_perMuni = eav_pre_exe_perMuni) %>% 
  mutate(
    # if there were no exemptions
     rev_w_no_exemps = round(agency_rate_no_exemps * eav_pre_exe_perClass),
     
     # the current situation with current exemptions
     rev_w_exemps = round(agency_rate_w_exemps* EAV_post_exe_PerClass)) %>%
  ungroup() %>%
  ungroup() %>%
  mutate(
     
    # percent of levy paid by each property class if there were no exemptions
     tax_burden_noexemps = round(eav_pre_exe_perClass *agency_rate_no_exemps / MuniLevy, digits = 5),
     
     # percent of levey paid by each property class based on current exemptions
     tax_burden_current = round(EAV_post_exe_PerClass * agency_rate_w_exemps/ MuniLevy, digits = 5),
     
     # difference in percent of the levy being paid by each property class
     burden_shift_from_current_exemps = tax_burden_noexemps - tax_burden_current ) 

burden_shift_byMuni_PropClass

burden_shift_byMuni_PropClass %>% 
  filter(major_class_code == 2) %>% 
  arrange(-burden_shift_from_current_exemps) %>%
  select(agency_name, burden_shift_from_current_exemps, tax_burden_current, tax_burden_noexemps,everything())
```

If there were no homeowner exemptions, Calumet Park, Broadview, Maywood, Thornton, etc. would experience the largest increase in revenue collected from the elimination of homeowners exemptions. This also means that  homeowners in these municipalities are being subsidized the most (as a percentage of the levy) by other types of properties.


# Read section above first!


> If you liked the numbers and process above then stop reading here.

# Combined Data while ignoring the existance of TIFs


Values in this section do not consider TIFs or the EAV within TIFs. 

I consider the section above more accurate but included this just in case. 

```{r}
exemps_byClass_perMuni <- exemptions_by_class_per_TC %>% 
  mutate(taxamount = tax_code_rate/100*eav) %>%
  group_by(agency_name, agency_num, major_class_code, major_class_type, agency_rate) %>%
  summarize(taxamount_perClass = round(sum(taxamount)),
            eav_pre_exe_perClass  = sum(eav, na.rm=TRUE),
            exe_homeowner = sum(exe_homeowner, na.rm=TRUE),
            exe_senior = sum(exe_senior, na.rm=TRUE),
            exe_freeze = sum(exe_freeze, na.rm=TRUE),
            exe_longtime_homeowner = sum(exe_longtime_homeowner, na.rm=TRUE),
            exe_disabled = sum(exe_disabled, na.rm=TRUE),
            exe_vet_returning = sum(exe_vet_returning, na.rm=TRUE),
            exe_vet_dis = sum(exe_vet_dis, na.rm=TRUE),
            exe_abate = sum(exe_abate, na.rm=TRUE)
           ) %>%
  
  mutate(Exemp_in_PropClass = exe_homeowner + exe_senior+ exe_freeze +
           exe_longtime_homeowner + exe_disabled+ exe_vet_returning +exe_vet_dis + exe_abate,
          
            EAV_post_exe_PerClass = eav_pre_exe_perClass-Exemp_in_PropClass, 
     #    tax_share_pre_exemp = taxamount / eav_pre_exe_perClass,
          
         tax_share_post_exemp = taxamount_perClass/ EAV_post_exe_PerClass #curreny situation
         )

exemps_byClass_perMuni


summtable <- exemps_byClass_perMuni %>% 
  select(agency_name, major_class_code, tax_share_post_exemp, taxamount_perClass, 
         eav_pre_exe_perClass, Exemp_in_PropClass, EAV_post_exe_PerClass, agency_num, agency_rate, major_class_type)
summtable

```




```{r}
exemptions_perMuni <- exemps_byClass_perMuni %>%
  group_by(agency_name, agency_num, agency_rate) %>% 
  summarize(taxamount_perMuni = sum(taxamount_perClass),
            eav_pre_exe_perMuni = sum(eav_pre_exe_perClass),
            TotalExemptions=sum(Exemp_in_PropClass),
         ResidentialExemp = sum(exe_homeowner +exe_senior+exe_freeze + exe_longtime_homeowner
                                + exe_disabled+ exe_vet_returning +exe_vet_dis, na.rm=TRUE),
         CommercialExemp = sum(exe_abate, na.rm=TRUE)) %>%
  mutate(
         eav_post_exe_perMuni = eav_pre_exe_perMuni-TotalExemptions) %>% 
  arrange(agency_num)

exemptions_perMuni

EAV_and_exemps_byClass_andMuni <- summtable %>% 
  left_join(exemptions_perMuni , by = c("agency_name", "agency_num", "agency_rate"))



burden_shift_byMuni_PropClass <- EAV_and_exemps_byClass_andMuni %>% #ungroup() %>% 
  select(agency_name, major_class_type, agency_rate, eav_pre_exe_perClass, EAV_post_exe_PerClass, agency_num, eav_pre_exe_perMuni, eav_post_exe_perMuni) %>% 
  left_join(MuniLevy) %>%
  mutate(noexemps_agency_rate = MuniLevy / eav_pre_exe_perMuni,
         exemps_agency_rate = MuniLevy / eav_post_exe_perMuni) %>%
  group_by(agency_name, agency_num, major_class_code, major_class_type#, noexemps_agency_rate, exemps_agency_rate
           ) %>%
   summarize(agency_rate_no_exemps = noexemps_agency_rate,
             agency_rate_w_exemps= exemps_agency_rate,
             MuniLevy = MuniLevy, 
             eav_pre_exe_perClass = eav_pre_exe_perClass,
             EAV_post_exe_PerClass = EAV_post_exe_PerClass,
             eav_post_exe_perMuni = eav_post_exe_perMuni, 
             eav_pre_exe_perMuni = eav_pre_exe_perMuni) %>% 
  mutate(
    # if there were no exemptions
     rev_w_no_exemps = round(agency_rate_no_exemps * eav_pre_exe_perClass),
     
     # the current situation with current exemptions
     rev_w_exemps = round(agency_rate_w_exemps* EAV_post_exe_PerClass)) %>%
  ungroup() %>%
  ungroup() %>%
  mutate(
     
     # percent of levy paid by each property class if there were no exemptions
     tax_burden_noexemps = round(eav_pre_exe_perClass / MuniLevy, digits = 5),
     
     # percent of levey paid by each property class based on current exemptions
     tax_burden_current = round(EAV_post_exe_PerClass / MuniLevy, digits = 5),
     
     # difference in percent of the levy being paid by each property class
     burden_shift_from_current_exemps = tax_burden_noexemps - tax_burden_current ) 

# write.csv(EAV_and_exemps_byClass_andMuni, "EAV_and_exemps_byClass_andMuni.csv")

burden_shift_byMuni_PropClass
```

Six community Comparison:

```{r}
burden_shift_byMuni_PropClass %>% filter(agency_num %in% c("030120000", "030210000", "030800000", "030920000","030580000","030310000"))

burden_shift_byMuni_PropClass %>% filter(agency_num %in% c("030120000", "030210000", "030800000", "030920000","030580000","030310000") &
                                           major_class_code == "2")
```

Overall burden shift is change in percentage points.



> To do: Create residential & commercial categories and EAV in each. Then calculate the share of exemptions/residential eav, exemptions/total_eav, and residential eav/total_eav.







