---
format: 
  html:
    df-print: paged
    embed-resources: true
    theme: lumen
    code-fold: true
    code-line-numbers: true
    code-overflow: wrap
    toc: true
    toc-location: left
knitr: 
  opts_chunk:
    warning: true
    message: false
---

# 2011-2022 Panel Data Cross Tables

```{r setup}
#| output: FALSE

# Load packages

library(tidyverse)
library(corrr)
library(glue)
library(DT)
library(flextable)
library(kableExtra)
library(crosstable)
library(scales)

# Set table formatting defaults

set_flextable_defaults(theme_fun = theme_vanilla, 
                       padding = 2,
                       #line_spacing = 1,
                       big.mark = ","
                       )

options(DT.options = list())

FitFlextableToPage <- function(ft, pgwidth = 6){
  ft_out <- ft %>% autofit()
  ft_out <- width(ft_out, width = dim(ft_out)$widths*pgwidth /(flextable_dim(ft_out)$widths))
  return(ft_out)
}

comm_ind <- read_csv("./Output/comm_ind_2011-2022_balanced.csv") 

## set variable types 
comm_ind <- comm_ind %>%
  mutate(across(c(class, improvement_ind, has_AB_exemp,fmv_NA_indicator, in_tif), as.character) ) 

```


## Descriptive Stats for All Years Together

```{r}
library(modelsummary)
datasummary_skim(comm_ind)
datasummary_skim(comm_ind, type  = "categorical")
```

## Descriptive Stats for 2011 Strata of data

```{r}
commind2011 <- comm_ind %>% filter(year == 2011)

datasummary_skim(commind2011)
```

## Descriptive Stats for 2022 Strata of data

```{r}

commind2022 <- comm_ind %>% filter(year == 2022)

datasummary_skim(commind2022)
datasummary_skim(commind2022, type = "categorical")
```

```{r}
datasummary_balance(~incent_change, data = commind2022, title = "PINs from 2022", output = "flextable")
```


# Correlations

```{r}
datasummary_correlation(commind2022, output = "flextable")
```

# Muni Check

Park Forest is there again! 118 Munis have commercial or industrial PINs during tax year 2022.

```{r}
comm_ind |> filter(year==2022) |> reframe(pincount= n(), fmv = sum(fmv), .by = clean_name) |> arrange(clean_name)
```

# Model Things

- Drop unnecessary variables from dataframe. Helps with speed of `pdata.frame()` command.    
- Set index of panel data forthe PIN and year (unit of observation and time) 
- Fixed effects will drop anything that doesn't change: Municipalty, Triad, distance to Indiana, etc.   
   
- DV: FMV Growth since 2011? 

- 5760 PINs were exempt at least 1 year in the dataframe...

```{r eval=FALSE}
library(dplyr)


# # 5760 PINs were exempt at least 1 year in the dataframe...
# comm_ind %>% 
#   group_by(pin) %>% 
#   mutate(exempt_sum = sum(exempt_indicator)) %>%  # was a PIN tax exempt ever?
#   ungroup() %>%
#   filter(exempt_sum > 0) %>% distinct(pin)

data <- comm_ind %>% 
  group_by(pin) %>% 
  mutate(exempt_sum = sum(exempt_indicator),
         years_existed2 = n()) %>%  # was a PIN tax exempt ever?
  ungroup() %>%
  filter(exempt_sum == 0 & years_existed2 == 12) %>%
  filter(year >= 2012) %>%        ## Drop first year of panel data

  select(pin, year, av_clerk, has_AB_exemp, land_use, incent_prop) 
# %>% mutate_at(c(pin, year, has_AB_exemp), as.character)

library(plm)

df_panel <- pdata.frame(data, index = c("pin", "year"))   
is.pbalanced(df_panel)
plm::is.pseries(df_panel)
plm::pdim(df_panel)
plm::make.pbalanced(df_panel)


plm::pdata.frame(data = comm_ind, )
```

