# Federal COVID Money Received {#sec-covid-federal-funds}

```{r setup}
#| warning: false
#| message: false
#| echo: false

knitr::opts_chunk$set(warning=FALSE, message = FALSE)
library(tidyverse)
library(haven)
#install.packages("ggsankey")
#install.packages("ggalluvial")

library(lubridate)
library(smooth)
library(forecast)
library(scales)

library(ggplot2)
library(readxl)
library(tidyverse)
library(data.table)
library(quantmod)
library(geofacet)
library(fredr)
library(sf) 
library(usmap)
library(janitor)

library(RColorBrewer)
#library(ggsankey)
library(ggalluvial)
library(readxl)

comma <- function(x) format(x, digits = 2, big.mark = ",")
options(scipen = 999)
```


##  Data



::: callout-note
*Variable names in the sankeyattempt2022 file do not have the most useful names due to the information within morphing over time as I figured out the format to necessary for making the graphs. Any variable names used while making these graphs might not actually contain the data one would expect given the variable names. I do intend on renaming the file and changing the variable names and code to something more intuitive but have not had time to complete that task.*
:::


```{r}
#| layout-ncol: 2


sankey <- read_csv("./Necessary_Files/sankey2023.csv") %>% 
  #filter(Total == "Total") %>% 
  mutate(value = as.numeric(value))



sankey %>% 
  filter( level == "IncentiveTotal") %>%
  ggplot( 
       aes(y = value, 
          #  axis3=Total, 
           axis2=Total, 
           axis1 = Incentive, label = "stratum")) +
  geom_flow(aes(fill = Incentive), color = "black", reverse=FALSE) +
  geom_stratum(reverse=FALSE)+
coord_flip()+
   scale_fill_brewer(palette = "YlOrRd", direction = -1)+
  theme_void() +
  theme(legend.position = "bottom") + 
      geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 2, reverse=FALSE)+
  labs(title = "665 billion FMV in Cook County. ")

sankey %>% 
  filter(level == "Type Total") %>% 
  ggplot( 
       aes(y = value, 
          #  axis3=Total, 
           axis2=Total, 
           axis1 = Commercial, label = "stratum")) +
  geom_flow(aes(fill = Commercial), color = "black", reverse=FALSE) +
  geom_stratum(reverse=FALSE)+
coord_flip()+
   scale_fill_brewer(palette = "YlOrRd", direction = -1)+
  theme_void() +
  theme(legend.position = "bottom") + 
      geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 2, reverse=FALSE)+
  labs(title = "665 billion FMV in Cook County. ")
```

```{r}
sankey <- read_csv("./Necessary_Files/sankey2023.csv") %>% 
  mutate(value = as.numeric(value))





```


```{r}
sankey %>% 
  filter( level == "Class Total") %>%
  ggplot( 
       aes(y = value, 
          axis4 =Total, 
          axis3 = Incentive,
           axis2=Commercial, 
          # axis1 = CountyType,
          label = "stratum")) +
  geom_flow(aes(fill = MajorClass), color = "black", reverse=FALSE) +
  geom_stratum(reverse=FALSE)+
coord_flip()+
   scale_fill_brewer(palette = "RdGy", direction = 1)+
  theme_void() +
  theme(legend.position = "bottom") + 
      geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 2, reverse=FALSE)+
  labs(title = "665 billion FMV in Cook County. ")

sankey %>% 
  filter( level == "Class Total") %>%
  ggplot( 
       aes(y = value, 
          axis4 =Total, 
         # axis3 = MajorClass,
           axis2=Commercial, 
           axis1 = MajorClass,
          label = "stratum")) +
  geom_flow(aes(fill = Incentive), color = "black", reverse=FALSE) +
  geom_stratum(reverse=FALSE)+
coord_flip()+
   scale_fill_brewer(palette = "RdGy", direction = 1)+
  theme_void() +
  theme(legend.position = "bottom") + 
      geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 2, reverse=FALSE)+
  labs(title = "665 billion FMV in Cook County. ")
```


```{r}
sankey %>% 
  filter(Total == "Total" & Commercial %in% c("Commercial", "Industrial")) %>%
  ggplot( 
       aes(y = value, 
          axis4 =Total, 
         # axis3 = MajorClass,
           axis2=Commercial, 
           axis1 = MajorClass,
          label = "stratum")) +
  geom_flow(aes(fill = Incentive), color = "black", reverse=FALSE) +
  geom_stratum(reverse=FALSE)+
coord_flip()+
   scale_fill_brewer(palette = "BlGn", direction = 1)+
  theme_void() +
  theme(legend.position = "bottom") + 
      geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 2, reverse=FALSE)+
  labs(title = "$112 billion Commercial and Industrial FMV in Cook County. ")

sankey %>% 
  filter(Total == "Total" & Commercial %in% c("Industrial")) %>%
  ggplot( 
       aes(y = value, 
          axis4 =Total, 
         # axis3 = MajorClass,
           axis2=Commercial, 
           axis1 = MajorClass,
          label = "stratum")) +
  geom_flow(aes(fill = Incentive), color = "black", reverse=FALSE) +
  geom_stratum(reverse=FALSE)+
coord_flip()+
   scale_fill_brewer(palette = "BlGn", direction = 1)+
  theme_void() +
  theme(legend.position = "bottom") + 
      geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 2, reverse=FALSE)+
  labs(title = "$24.7 billion Industrial FMV in Cook County. ")



sankey %>% 
  filter(Total == "Total" & Commercial %in% c("Commercial")) %>%
  ggplot( 
       aes(y = value, 
          axis4 =Total, 
         # axis3 = MajorClass,
           axis2=Commercial, 
           axis1 = MajorClass,
          label = "stratum")) +
  geom_flow(aes(fill = Incentive), color = "black", reverse=FALSE) +
  geom_stratum(reverse=FALSE)+
coord_flip()+
   scale_fill_brewer(palette = "BlGn", direction = 1)+
  theme_void() +
  theme(legend.position = "bottom") + 
      geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 2, reverse=FALSE)+
  labs(title = "$88 billion Commercial FMV in Cook County. ")


```



```{r}
sankey %>% 
  filter( level == "Class Total") %>%
  ggplot( 
       aes(y = value, 
           axis4 = Total,
          axis3=Commercial, 
           axis2=Incentive, 
           axis1 = MajorClass, label = "stratum")) +
  geom_flow(aes(fill = Incentive), color = "black", reverse=FALSE) +
  geom_stratum(reverse=FALSE)+
coord_flip()+
   scale_fill_brewer(palette = "YlOrRd", direction = -1)+
  theme_void() +
  theme(legend.position = "bottom") + 
      geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 2, reverse=FALSE)+
  labs(title = "665 billion FMV in Cook County. ")
```


```{r}
sankey %>% 
  filter(Total == "FMV") %>%
  ggplot( 
       aes(y = value, 
           axis4 = Total,
          axis3=Commercial, 
           axis2=Incentive, 
        #   axis1 = MajorClass, 
          label = "stratum")) +
  geom_flow(aes(fill = Commercial), color = "black", reverse=FALSE) +
  geom_stratum(reverse=FALSE)+
coord_flip()+
   scale_fill_brewer(palette = "YlOrRd", direction = -1)+
  theme_void() +
  theme(legend.position = "bottom") + 
      geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 2, reverse=FALSE)+
  labs(title = "665 billion FMV in Cook County. ")

sankey %>% 
  filter(Total == "FMV") %>%
  ggplot( 
       aes(y = value, 
           axis4 = Total,
          axis3=Commercial, 
           axis2=residential, 
           axis1 = Incentive, 
          label = "stratum")) +
  geom_flow(aes(fill = Commercial), color = "black", reverse=FALSE) +
  geom_stratum(reverse=FALSE)+
coord_flip()+
   scale_fill_brewer(palette = "YlOrRd", direction = -1)+
  theme_void() +
  theme(legend.position = "bottom") + 
      geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 2, reverse=FALSE)+
  labs(title = "665 billion FMV in Cook County. ")

sankey %>% 
  filter(Total == "FMV") %>%
  ggplot( 
       aes(y = value, 
           axis4 = Total,
           axis2=residential, 
           axis3 = Incentive, 
                     axis1=Commercial, 

          label = "stratum")) +
  geom_flow(aes(fill = Commercial), color = "black", reverse=FALSE) +
  geom_stratum(reverse=FALSE)+
coord_flip()+
   scale_fill_brewer(palette = "YlOrRd", direction = -1)+
  theme_void() +
  theme(legend.position = "bottom") + 
      geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 2, reverse=FALSE)+
  labs(title = "665 billion FMV in Cook County. ")


sankey %>% 
  filter(Total == "FMV") %>%
  ggplot( 
       aes(y = value, 
           axis5 = Total,
           axis4=residential, 
           axis3 = Incentive,
           axis2 = CountyType,
          axis1=Commercial, 

          label = "stratum")) +
  geom_flow(aes(fill = Commercial), color = "black", reverse=FALSE) +
  geom_stratum(reverse=FALSE)+
coord_flip()+
   scale_fill_brewer(palette = "YlOrRd", direction = -1)+
  theme_void() +
  theme(legend.position = "bottom") + 
      geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 2, reverse=FALSE)+
  labs(title = "665 billion FMV in Cook County. ")


```


```{r}
sankey %>% 
  filter(Total == "Total") %>%
  ggplot( 
       aes(y = final_tax_to_dist, 
           axis5 = Total,
           axis4=residential, 
           axis3 = Incentive,
          axis1=Commercial, 

          label = "stratum")) +
  geom_flow(aes(fill = Commercial), color = "black", reverse=FALSE) +
  geom_stratum(reverse=FALSE)+
coord_flip()+
   scale_fill_brewer(palette = "RdGy", direction = -1)+
  theme_void() +
  theme(legend.position = "bottom") + 
      geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 2, reverse=FALSE)+
  labs(title = "$16 Billion collected by Local Governments (excludes TIFs)")

sankey %>% 
  filter(Total == "Total") %>%
  ggplot( 
       aes(y = value, 
           axis5 = Total,
           axis4=residential, 
           axis3 = Incentive,
          axis1=Commercial, 

          label = "stratum")) +
  geom_flow(aes(fill = Commercial), color = "black", reverse=FALSE) +
  geom_stratum(reverse=FALSE)+
coord_flip()+
   scale_fill_brewer(palette = "RdGy", direction = -1)+
  theme_void() +
  theme(legend.position = "bottom") + 
      geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 2, reverse=FALSE)+
  labs(title = "FMV in Cook County")
```

