---
title: "Senior Freeze Proposed Policy Change"
format: 
  html:
    toc: true
    code-tools: true
    code-fold: true
    
---

"Low-income Senior Citizens Assessment Freeze Homestead Exemption (SCAFHE)
A person qualifies for this exemption if the person

- is at least 65 years old;  
- has a total household income of $65,000 or less; and  
- meets certain other qualifications.  

This exemption "freezes" the senior citizen's property's equalized assessed value the year that the senior citizen qualifies for the exemption. The property's equalized assessed value does not increase as long as qualification for the exemption continues. The tax bill may still increase if any tax rates are increased or if improvements are added that increase the value of the property.

This exemption allows senior citizens who meet the qualifications to elect to maintain the equalized assessed value (EAV) of their homes at the base year EAV and prevent any increase in that value due to inflation. The amount of the exemption benefit is determined each year based on (1) the property's current EAV minus the frozen base year value (the property's prior year's EAV for which the applicant first qualifies for the exemption), and (2) the applicant's total household maximum income limitation.

Each year applicants must file a Form PTAX-340, Low-income Senior Citizens Assessment Freeze Homestead Exemption Application and Affidavit, with the Chief County Assessment Office.

[(35 ILCS 200/15-170)](http://www.ilga.gov/legislation/ilcs/fulltext.asp?DocName=003502000K15-170) 
Additional Info - [Illinois.gov](https://tax.illinois.gov/localgovernments/property/taxrelief.html#:~:text=This%20program%20allows%20persons%2065%247%2C500)



```{r warning=FALSE, message=FALSE, results='hide'}
library(tidyverse)
library(ipumsr)
library(tigris)
library(srvyr)
library(naniar)
library(survey)

knitr::opts_chunk$set(message = FALSE, warning = FALSE)

# PUMA shapefiles
pumasIL2020 <- pumas("IL", cb=T, year=2020) 

pumasIL2020 <- pumasIL2020 |>
  mutate(County = ifelse(
    str_sub(GEOID20, 1, 5) == "17031", "Cook", NA), 
    YEAR="2023") |>
  select(-c(ALAND20, AWATER20, STATEFP20, AFFGEOID20, LSAD20, STUSPS20, ST_NAME20)) |>
  rename(GEOID=GEOID20, 
         puma = PUMACE20,
         puma_area = NAMELSAD20)



pumasIL2018 <- pumas("IL", cb=T, year=2018) 

pumasIL2018 <- pumasIL2018 |>
  mutate(County = ifelse(
    (str_sub(GEOID10, 1, 5) == "17034" |
       str_sub(GEOID10, 1, 5) == "17035") == TRUE, "Cook", NA) ,
    YEAR="2018") |>
  select(-c(ALAND10, AWATER10, STATEFP10, AFFGEOID10, LSAD10 ) ) |>
  rename(GEOID = GEOID10,
         puma = PUMACE10,
         puma_area = NAME10
  )


pumas <- rbind(pumasIL2018, pumasIL2020)  |>
  mutate(uniqueid = paste0(puma, "_", YEAR),
           puma_area = str_remove_all(puma_area, " PUMA")) |>
  filter(County == "Cook")

ddi <- read_ipums_ddi("inputs/usa_00026.xml")
data2018 <- read_ipums_micro(ddi) |>     
  mutate(PUMA = str_pad(PUMA, 5,side = "left", pad="0"),
         YEAR = as.character(YEAR)
         ) |>
  select(YEAR, PUMA, PERWT, HHWT, AGE, SEX, PERNUM, HHINCOME, INCEARN, VALUEH, OWNERSHP, STRATA, 
         CLUSTER, HHTYPE, SERIAL, SAMPLE) |>
  left_join(pumasIL2018, by = c("PUMA" = "puma", "YEAR"))


ddi <- read_ipums_ddi("inputs/usa_00025.xml")
data2023 <- read_ipums_micro(ddi) |> 
  select(YEAR, PUMA, PERWT, HHWT, AGE, SEX, PERNUM, HHINCOME, INCEARN, VALUEH, OWNERSHP, STRATA, 
         CLUSTER, HHTYPE, SERIAL, SAMPLE) |>
  mutate(PUMA = str_pad(PUMA, 5, side = "left", pad="0"),
                  YEAR = as.character(YEAR)) |>
  left_join(pumasIL2020, by = c("PUMA" = "puma", "YEAR"))


data <- rbind(data2018, data2023)

data <- data |> 
  mutate(
    VALUEH = ifelse(VALUEH == 9999999, NA, VALUEH),
    HHINCOME = ifelse(HHINCOME == 9999999 | HHINCOME == 9999998, NA, HHINCOME)
  )

data <- data |> 
  mutate(age_cat = 
           
           case_when(AGE < 24 ~ "16to24",
                     AGE > 24 & AGE < 35 ~ "25to34",
                     AGE > 34 & AGE < 45 ~ "35to44",
                     AGE > 44 & AGE < 55 ~ "45to54",
                     AGE > 54 & AGE < 65 ~ "55to64",
                     AGE > 64 ~ "65+"),
         sex_cat = case_when(SEX == 1 ~ "Male",
                             SEX == 2 ~ "Female"))

# #check coding make sure it is the same for both censuses
# data <-  data |> mutate(white = if_else(RACE ==1, 1, 0),
#                          black = if_else(RACE ==2, 1, 0), 
#                          asian = if_else(RACE %in% c(4,5,6), 1, 0),
#                          otherrace = if_else(RACE %in% c(3,7,8,9),1,0)) 

```

```{r}
#| include: false

head(data$HHTYPE)
summarytools::freq(data$HHTYPE)
head(data$OWNERSHP)
```


```{r include=FALSE}
data |> 
  filter(YEAR == "2023") |>
  group_by(SERIAL) |>  ## Household ID number
  summarize(multi_peep = n(),
            HHWT = mean(HHWT),
            owned = max(OWNERSHP),
            hhinc = max(HHINCOME),
            house_value = max(VALUEH),
            oldest = max(AGE),
            youngest = min(AGE)) |> 
 # filter(multi_peep > 1) |> 
  arrange(desc(multi_peep) )

data |> 
    filter(YEAR == "2023") |>
  #group_by(SERIAL) |>  ## Household ID number
  summarize(obs = n(),
            HHWT = sum(HHWT),
            PERWT = sum(PERWT))

```

38,112 obs in data for Cook County downloaded from IPUMS. summed HHWT representing 5,029,986 households and 5,086,576 individuals. 

10,461 unweighted household observations with more than 1 person in them. 17,119 unweighted household observations in IPUMs data for Cook County. 

```{r}
#| include: false


obs_perPUMA <- data |> 
   group_by(YEAR, PUMA, age_cat) |> 
   dplyr::summarize(weightedcount=sum(HHWT), #weighted 
                    unweightedcount = n()) |>
  arrange(PUMA)
obs_perPUMA

obs_perPUMA |> 
  select(-unweightedcount) |> 
  pivot_wider(names_from = "age_cat", values_from = "weightedcount")


# 1 = OWNED, 2 = RENTED, 3 = N/A
obs_perPUMA<- data |> 
  filter(PERNUM == 1) |> # only use one person from each household, the "householder"?
   group_by(PUMA, age_cat, OWNERSHP ) |> 
   dplyr::summarize(
     hhs_wghted = sum(HHWT),
     unweightedcount = n()) |>
  arrange(PUMA)

obs_perPUMA
```


```{r}
#| layout-ncol: 2

# INCEARN
# AGE
# VETSTAT
# HHINCOME

data |> filter(YEAR == 2018) |>
  ggplot() + geom_histogram(aes(x=HHINCOME, weight = HHWT))


data |> filter(YEAR == 2023) |>
  ggplot() + geom_histogram(aes(x=HHINCOME, weight = HHWT))

data |> filter(YEAR == 2018) |>
  ggplot() + geom_histogram(aes(x=INCEARN, weight = PERWT))

data |> filter(YEAR == 2023) |>
  ggplot() + geom_histogram(aes(x=INCEARN, weight = PERWT))
```


```{r}

# inc_quantiles <- survey::svyquantile(~HHINCOME, design=HHdesign, 
#                     quantiles = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1) ,
#                     na.rm=TRUE, ci = FALSE  )
# inc_quantiles
# 
#          0   0.1   0.2   0.3   0.4   0.5    0.6    0.7    0.8    0.9       1
# [1,] -7900 21600 40000 58700 77000 97000 120000 148400 185000 261000 1744000
#          
## numbers used for income breaks are calculated in Income Deciles section. 
# created now so that the variable exists in the joined dataset before creating the survey design object
data <- data |> 
  mutate(
    below_freeze = ifelse(HHINCOME >= 65000, "Not Eligible", "Eligible"),

    ## current senior freeze threshold!
    new_freeze = ifelse(HHINCOME >= 85000, "Not Eligible", "Eligible"),
    )


# filter to household head only so households are counted 
# multiple times for each person in the household
HHdesign <- survey::svydesign(id = ~SERIAL, nest=TRUE, strata = ~STRATA, weights = ~HHWT, data = data|> filter(PERNUM==1)) 
```

## Current Policy - $65K

```{r}
incometable <- svytable(~ YEAR + PUMA + OWNERSHP  + age_cat + below_freeze, design = HHdesign)

mapPUMAboth <- incometable |> 
  as_tibble() |>
  mutate(PUMA = str_pad(PUMA, 5, side= "left", pad = "0"),
  ) |>
  group_by(YEAR, PUMA) |>
  mutate(
    Prop = n/sum(n),
    Total = sum(n)
  ) 

mapPUMAboth <- mapPUMAboth |> 
  filter(below_freeze == "Eligible" & age_cat == "65+" & OWNERSHP == 1) |>
  inner_join(pumas, by = c("PUMA" = "puma", "YEAR"))

mapPUMAboth |> 
  ungroup() |> 
  group_by(puma_area) |> 
  pivot_wider( id_cols = c(puma_area), names_from = "YEAR", values_from = "n") |> 
  mutate(change = `2023`-`2018`) |>
  select(change, puma_area# Prop, puma_area
  ) |> 
  arrange(desc(change))
```


```{r}
#| label: fig-current-income-threshold
#| fig-cap: "Share of all Own. Occ. Households 65+ w/ HH incomes < $65K. Shows number of households eligible for the Low-Income Senior Freeze Exemption in 2018, when the income threshold was last updated, and the number of households that are eligible in tax year 2023."
#| fig-height: 10
#| fig-width: 16
#| column: screen-inset


current_policy <- ggplot(mapPUMAboth, 
                         aes(fill = Prop, 
                             geometry = geometry,
                             label = paste0(
                               scales::percent(Prop, accuracy = 0.01), "\n",  
                              #  scales::comma(n), " Eligible", "\n",
                               scales::comma(Total), " Total HH")
                               #" ", "frac(", n, ",", Total, ")"
                         ))  +
  geom_sf(color = "black")+ 
  geom_sf_text(size = 3) +
  #labs(title = "Share of all Own. Occ. Households 65+ w/ HH incomes < $65K",
  #     subtitle = "Currently Eligible for Senior Freeze"#,
       # caption = "n represents the number of households eligible") + 
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
 # theme(legend.position = "bottom")+
  scale_fill_binned(high = "darkblue", low = "white", 
                    show.limits=TRUE,
                    nice.breaks=FALSE,
                    labels=scales::percent,
                    name = "% with HH Inc < 65K\n& Age 65+") +
  facet_wrap(~YEAR)

current_policy
```

## New Policy - $85K

```{r include=FALSE}
HHdesign <- survey::svydesign(id = ~SERIAL, nest=TRUE, strata = ~STRATA, weights = ~HHWT, data = data |> filter(PERNUM==1)) 

agetable <- svytable(~YEAR + PUMA + age_cat, design = HHdesign)


mapPUMAboth <- agetable |> 
  as_tibble() |>
  mutate(PUMA = str_pad(PUMA, 5, side= "left", pad = "0")) |>
  group_by(YEAR, PUMA) |> 
     mutate(
       Prop = n/sum(n)) |>
     filter(age_cat =="65+") |>
  ungroup() |>
  left_join(pumas, by = c("PUMA" = "puma", "YEAR" = "YEAR"))
  
 
 
#mapPUMAboth |> select(YEAR, countyFIP, PUMA, Prop, n, NAME10)



figure <- ggplot(mapPUMAboth, aes(fill = Prop)) +
  geom_sf(aes(geometry = geometry), color = "black")+ 
  labs(title = "Senior Households are up to 30% of all households in some areas", 
       subtitle = "Percent of PUMA Households in 2023 that are 65+") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+

     scale_fill_binned(
                      high = "darkblue", low = "white", 
                       show.limits=TRUE,
                       nice.breaks=FALSE,
                       labels=scales::percent,
                        name = "% HH 65+") +
  facet_wrap(~YEAR)
figure
```


```{r include = FALSE}
incometable <- svytable(~ YEAR + PUMA + new_freeze, design = HHdesign)

mapPUMAboth <- incometable |> 
  as_tibble() |>
  mutate(PUMA = str_pad(PUMA, 5, side= "left", pad = "0")) |>
  group_by(YEAR, PUMA) |>
     mutate(
       Prop = n/sum(n)) |>
     filter(new_freeze == "Eligible")|>
  left_join(pumas, by = c("PUMA" = "puma", "YEAR" = "YEAR"))
 
#mapPUMAboth |> select(YEAR, countyFIP, PUMA, Prop, n, NAME10)



figure5c <- ggplot(mapPUMAboth, aes(fill = Prop)) +
  geom_sf(aes(geometry = geometry), color = "black")+ 
  labs(title = "Share of ALL households below 85K") + 
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+

     scale_fill_binned(high = "darkblue", low = "white", 
                       show.limits=TRUE,
                       nice.breaks=FALSE,
                       labels=scales::percent,
                        name = "% with HH Inc < 85K ") +
  facet_wrap(~YEAR)
figure5c
```

```{r include=FALSE}
incometable <- svytable(~ YEAR + PUMA + age_cat + new_freeze, design = HHdesign)

mapPUMAboth <- incometable |> 
  as_tibble() |>
  mutate(PUMA = str_pad(PUMA, 5, side= "left", pad = "0"),
        # new_freeze = ifelse(hhincdecile_w >= 5, "Not Eligible", "Eligible")
         ) |>
  group_by(YEAR, PUMA) |>
     mutate(
       Prop = n/sum(n)) |>
     filter(new_freeze == "Eligible" & age_cat == "65+") |>
    inner_join(pumas, by = c("PUMA" = "puma", "YEAR"))
 
#mapPUMAboth |> select(YEAR, PUMA, Prop, n, age_cat)



figure5c <- ggplot(mapPUMAboth, aes(fill = Prop)) +
  geom_sf(aes(geometry = geometry), color = "black")+ 
  labs(title = "Share of all households 65+ w/ HH incomes < $85K") + 
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+

     scale_fill_binned(high = "darkblue", low = "white", 
                       show.limits=TRUE,
                       nice.breaks=FALSE,
                       labels=scales::percent,
                        name = "% with HH Inc < 85K ") + facet_wrap(~YEAR)
figure5c
```


```{r}
#| label: fig-proposed-income-threshold
#| fig-cap: "Share of all Own. Occ. Households 65+ w/ HH incomes < $85K."
#| fig-height: 10
#| fig-width: 16
#| column: screen-inset

incometable <- svytable(~ YEAR + PUMA + OWNERSHP  + age_cat + new_freeze, design = HHdesign)

mapPUMAboth <- incometable |> 
  as_tibble() |>
  mutate(PUMA = str_pad(PUMA, 5, side= "left", pad = "0"),
         ) |>
  group_by(YEAR, PUMA) |>
     mutate(
       Prop = n/sum(n),
       Total = sum(n)) |>
     filter(new_freeze == "Eligible" & age_cat == "65+" & OWNERSHP == 1) |>
    inner_join(pumas, by = c("PUMA" = "puma", "YEAR"))


new_policy <- ggplot(mapPUMAboth, 
                         aes(fill = Prop, 
                             geometry = geometry,
                             label = paste0(
                               scales::percent(Prop, accuracy = 0.01), "\n",  
                                scales::comma(n), " Eligible", "\n",
                               scales::comma(Total), " Total HH")
                         ))  +
  
  geom_sf(aes(geometry = geometry), color = "black")+ 
  geom_sf_text(size = 3) +
#  labs(title = "Share of all Own. Occ. Households 65+ w/ HH incomes < $85K",
 #      subtitle = "Eligible for the Senior Freeze Exemption if Household Income increased to $85K") + 
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+

     scale_fill_binned(high = "darkblue", low = "white", 
                       show.limits=TRUE,
                       nice.breaks=FALSE,
                       labels=scales::percent,
                        name = "% with HH Inc < 85K\n& Age 65+") +
  facet_wrap(~YEAR)
new_policy
```








