---
title: "Agency Funds"
format: 
  html:
    code-fold: true
    toc: true
    toc-location: left
    tbl-cap-location: margin
    fig-cap-location: margin
    df-print: paged
---


```{r warning = FALSE, message = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
options(scipen=999)

library(tidyverse)
library(ptaxsim)
library(DBI)
library(glue)
library(data.table)
```

# Data Collection

## PTAXSIM Part

> REMEMBER: **senior freeze exemption is called `exe_freeze`** in ptaxsim. The senior exemption is called `exe_senior`


```{r}
# #| eval: false

ptaxsim_db_conn <- DBI::dbConnect(RSQLite::SQLite(),
  "./ptaxsim.db/ptaxsim-2023.0.0.db")

pins <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT *
  FROM pin
  WHERE exe_freeze > 0 
  "
  )

sf_pins <- pins |> distinct(pin) |> select(pin)

pins <- DBI::dbGetQuery(
  ptaxsim_db_conn,
    glue_sql(
  "SELECT *
  FROM pin
  WHERE pin IN ({sf_pins$pin*})"
  ,     
  .con = ptaxsim_db_conn
 ))

n_distinct(pins$year)

pins <- pins |> 
  group_by(pin) |> 
  arrange(pin, year) |>
  mutate(years_with_sfexe = sum(exe_freeze > 0),
         gain_lose = case_when(
           years_with_sfexe == 18 ~ "Always had SF Exe",
           exe_freeze > 0 & lag(exe_freeze) == 0 ~ "Gained SF Exe",
         exe_freeze == 0 & lag(exe_freeze) > 0 ~ "Lost SF Exe",
         TRUE~"No Change"
         )) 

table(pins$year, pins$gain_lose)
```

10 million+ observations for pin-years for senior exemption.  
5,879,046 pin-year combos for **senior freeze exemption**.   

583,163 distinct pins had senior exemption at least 1 year.   
328,836 had the senior freeze exemption at least 1 year.

50,241 properties had the senior exemption every year from 2006 to 2023.
7,075 PINs had the senior freeze exemption every year from 2006 to 2023. 


```{r}
pins |> filter(year ==2023) |> group_by(gain_lose) |> summarize(n = n())
```
```{r}
pins |> 
  group_by(year) |> 
  summarize(had_SF_exe = sum(exe_freeze>0)) |> 
  ggplot() +
  geom_line(aes(x=year, y = had_SF_exe)) +
  geom_vline(aes(xintercept=2017), color= "red", lwd=1) +
  labs(y= "# PINs", x="", title = "PIN Count: Senior Freeze Exemption", caption = "Tax Year 2017 was the first year for senior freeze exemption increase")+
  theme_bw()
```

Senior freeze exemption was last increased and reflected in Tax Year 2017. [Civic Federation](https://www.civicfed.org/iifs/blog/new-state-law-increases-cook-county-property-tax-homestead-exemptions)


```{r}
pins2 <- pins |> ungroup() |> filter(years_with_sfexe != 18)

table(pins2$years_with_sfexe[pins2$year==2023])

pins2 |> ungroup() |> filter(gain_lose=="Lost SF Exe" & years_with_sfexe > 5) |> arrange(pin)

pins2 |> 
  group_by(pin)|>
  summarize(bill_wo_exemp = mean(tax_bill_total[gain_lose== "Lost SF Exe"]),
            bill_w_exemp = mean(tax_bill_total[gain_lose == "Gained SF Exe"])) |>
  mutate(avg_diff = bill_wo_exemp - bill_w_exemp)
```

**Looking at a couple random pins:**

For PIN 01011000430000, their bill went up \$2500 from 2010 to 2011 and  increased $1400 from 2013 to 2014. 

```{r}
pins |> filter(pin == "01011000430000") # lost exemption in 2011 and 2014

pins  |> filter(pin == "01011240370000") # lost exemption in 2019 and 2023 
# bill went up $1400 in 2023

pins |> filter(pin == "02112010020000")
```

```{r}
pins2 |> ungroup() |> filter(gain_lose=="Gained SF Exe") |> arrange(pin)

```

