---
title: "Elasticity Regression w/ Lags - AWM Edits"
author: "MVH & AWM"
format:
  html: 
    code-fold: true
    toc: true
    toc-location: left
---

# Initial Stuff

Stuff like loading and cleaning data sets and then running the old regression.

## Loading Packages/ptaxsim

```{r setup, message=FALSE}

options(scipen = 999)

library(tidyverse)
library(DBI)
library(data.table)
library(ggspatial)
library(gstat)
library(here)
library(httr)
library(jsonlite)
library(ptaxsim)
library(sf)
library(stars)
library(glue)
library(lmtest)
library(huxtable)
library(jtools)
library(plm)
library(modelsummary)

#ptaxsim_db_conn <- DBI::dbConnect(RSQLite::SQLite(), "./ptaxsim.db")

# base_url <- "https://datacatalog.cookcountyil.gov/resource/tx2p-k2g9.json"

ptaxsim_db_conn <- DBI::dbConnect(RSQLite::SQLite(), "./ptaxsim.db/ptaxsim-2021.0.4.db")

knitr::opts_chunk$set(warning = FALSE, message = FALSE)


```

## Generating Data

```{r}

agency_dt <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT *
  FROM agency
  "
) %>%
  mutate(first6 = str_sub(agency_num,1,6),
         first5 = str_sub(agency_num,1,5)
         )

muni_agency_names <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT DISTINCT agency_num, agency_name, minor_type
  FROM agency_info
  WHERE minor_type = 'MUNI'
  OR agency_num = '020060000'
  "
  ) %>% 
  mutate(first6 = str_sub(agency_num,1,6),
         first5 = str_sub(agency_num,1,5)
         ) %>%
  select(-minor_type)

# bring in Triads (North, South, Chicago)
triads <-  read_csv("muni_agency_triads.csv")

muni_agency_names <- muni_agency_names %>% 
  left_join(triads)

table(muni_agency_names$triad)



all_taxing_agencies <- DBI::dbGetQuery(
  ptaxsim_db_conn,
  "SELECT agency_num, agency_name, major_type, minor_type
  FROM agency_info
  "
)

all_taxing_agencies <- all_taxing_agencies %>%
  mutate(first6 = str_sub(agency_num,1,6),
        first5 = str_sub(agency_num,1,5)
         )

all_taxing_agencies <- all_taxing_agencies %>%
left_join(muni_agency_names, by = c("first6", "first5")) %>%
 rename(muni_name =  agency_name.y,
        muni_num = agency_num.y,
        agency_name = agency_name.x,
        agency_num = agency_num.x)

agency_dt <- left_join(agency_dt, all_taxing_agencies, by = c("agency_num", "first5", "first6"))



# grouped_munis <- agency_dt %>% 
#   mutate(cty_cook_eav = as.numeric(cty_cook_eav),
#          total_final_levy = as.numeric(total_final_levy),
#          cty_cook_eav = ifelse(cty_cook_eav < 1, NA, cty_cook_eav), total_final_levy = ifelse(total_final_levy <1, NA, total_final_levy)) %>%
#   group_by(muni_name, year) %>%
#   summarize(levy_sum = sum(total_final_levy, na.rm = TRUE), 
#             eav = first(cty_cook_eav), 
#             log_levy = log(levy_sum),
#             log_eav = log(eav),
#             year=first(year)
# )
# grouped_munis
# table(grouped_munis$year)
# 
# grouped_munis_rectangle <- grouped_munis %>% filter(!is.na(muni_name) & !is.na(year))
# 
# grouped_munis_rectangle
# 
# table(grouped_munis_rectangle$year)

agency_data <- agency_dt %>%
  mutate(first2 = str_sub(agency_num, 1,2),
         last2 = str_sub(agency_num,8,9),
         in_SSA = ifelse(minor_type == "SSA", 1,0),
         in_chicago = ifelse(str_detect(agency_name, "CHICAGO"),1,0)) %>%
  select(-c(cty_dupage_eav:cty_livingston_eav)) 

agency_data <- agency_data %>% 
  mutate(cty_cook_eav = as.numeric(cty_cook_eav),
         total_final_levy = as.numeric(total_final_levy),
         cty_cook_eav = ifelse(cty_cook_eav < 1, NA, cty_cook_eav), total_final_levy = ifelse(total_final_levy <1, NA, total_final_levy)) %>%
  mutate(log_eav = log(cty_cook_eav), 
         log_levy = log(total_final_levy))

panel_data <- agency_data %>% 
  filter(minor_type!="TIF") %>%
  select(year, agency_name, agency_num, major_type, minor_type, 
         cty_cook_eav, total_final_levy, log_eav, log_levy, total_final_rate, 
         in_chicago, in_SSA, triad
  ) 
```

Let's write a csv.

```{r}

write.csv(panel_data, "panel_data_mvh_7-13.csv")
```

# Prior Models

## All agencies, agency and year FEs

```{r}

all_agencies<-pdata.frame(panel_data, index = c("agency_name", "year"))

all_m1<- plm(log_levy ~ log_eav, 
             index = c("agency_name", "year"), 
             model = "within", effect= "twoways",                    
             data = all_agencies)

model_names <- c(all_m1 = "All Agencies-Agency & Year FE")

export_summs(all_m1, robust = "HC3",  
             model.names = model_names, 
             statistics = c(N = "nobs",
                            adjR2 = "adj.r.squared"))

```

## Just minor type = munis

```{r}

minortype_munis <- panel_data %>% 
  filter(agency_num %in% muni_agency_names$agency_num  
 ) %>%
  select(year, agency_name, agency_num, 
         major_type, minor_type, cty_cook_eav, 
         total_final_levy,  total_final_rate,
         log_eav, log_levy
         ) 

minortype_munis %>% ggplot()+
  geom_point(aes(x=log(cty_cook_eav), y=log(total_final_levy), col=agency_name, alpha = .1)) +
  theme(legend.position = "none")+
  ggtitle("Minor Type: Logged EAV vs Logged Levy \n Homerule & non-Homerule Together")


minortype_munis %>% 
  ggplot(aes(x=year, 
             y= log(total_final_levy)/log(cty_cook_eav), 
             group = agency_name
               )  )+
  geom_line(alpha = .1)+
  ggtitle("134 Municipalities based on Minor Type: \nLogged EAVLogg/ed Levy Over Time \n Homerule & non-Homerule Together")
```

## Home Rule vs. Non-Home Rule

### Minor type munis

Home Rule

```{r}

homerule_minortype_munis <- agency_data %>% 
  filter(agency_num %in% muni_agency_names$agency_num & 
           home_rule_ind==1
 ) %>% 
  select(year, agency_name, agency_num, 
         major_type, minor_type, cty_cook_eav, 
         total_final_levy,  total_final_rate,
         log_eav, log_levy
         ) 

hr_muni_minor <- plm(log_levy ~ log_eav, 
             index = c("agency_name", "year"), 
             model = "within", effect= "twoways",           
             data = homerule_minortype_munis)

plot(hr_muni_minor)

homerule_minortype_munis %>% 
  ggplot(aes(x=year, 
             y= log(total_final_levy)/log(cty_cook_eav), 
             group = agency_name, color= minor_type
               )  )+
  geom_line(alpha = .3)+
  ggtitle("Homerule Muni: Logged EAV/Logged Levy over Time")
```

Non-Home Rule

```{r}

non_homerule_munis <- agency_data %>% 
  filter(agency_num %in% muni_agency_names$agency_num & 
           home_rule_ind==0 ) %>% 
  select(year, agency_name, agency_num, 
         major_type, minor_type, cty_cook_eav, 
         total_final_levy,  total_final_rate,
         log_eav, log_levy
         ) 


nonhr_muni_minor <- plm(log_levy ~ log_eav, 
             index = c("agency_name", "year"), 
             model = "within", effect= "twoways", 
             data = non_homerule_munis)

summary(nonhr_muni_minor)

non_homerule_munis %>% 
  ggplot(aes(x=year, 
             y= log(total_final_levy)/log(cty_cook_eav), 
             group = agency_name, color= minor_type
               )  )+
  geom_line(alpha = .5)+theme(legend.position = "none")+
  ggtitle("Non Homerule Municipalities: Logged EAV/Logged Levy over Time")
```

### Major type munis

Home Rule

```{r}

homerule_majortype_munis <- agency_data %>% 
  filter(major_type == "MUNICIPALITY/TOWNSHIP" & 
           home_rule_ind==1) %>% 
  select(year, agency_name, agency_num, 
         major_type, minor_type, cty_cook_eav, 
         total_final_levy,  total_final_rate,
         log_eav, log_levy
         ) 

hr_muni1 <- plm(log_levy ~ log_eav, 
             index = c("agency_name", "year"), 
             model = "within", effect= "twoways",                    
             data = homerule_majortype_munis)

summary(hr_muni1)
```

Non-Home Rule

```{r}

nonhr_muni_major <- agency_data %>% 
  filter(major_type == "MUNICIPALITY/TOWNSHIP" & 
           home_rule_ind==0
 ) %>% 
  select(year, agency_name, agency_num, 
         major_type, minor_type, cty_cook_eav, 
         total_final_levy,  total_final_rate,
         log_eav, log_levy
         ) 

nonhr_muni_maj <- plm(log_levy ~ log_eav, 
             index = c("agency_name", "year"), 
             model = "within", effect= "twoways",
             data = nonhr_muni_major)

summary(nonhr_muni_maj)
```

## Schools

Making school data set.

```{r}

schools <- agency_data %>% 
  filter(major_type=="SCHOOL")  %>% 
  select(agency_name, year, agency_num, log_levy, log_eav, minor_type)

table(schools$minor_type)

schools %>% 
  ggplot(aes(x =year, 
             y = log_levy/log_eav, 
             col = minor_type, label = agency_name))+
  geom_point(alpha=.5)
```

Write "schools" csv.

```{r}

write.csv(schools, "schools_7-13.csv")
```

Models

```{r}

school_mod_major <- plm(log_levy ~ log_eav, index = c("agency_name", "year"), 
                  model = "within",  effect= "twoways", data = schools)

school_mod_secondary <- plm(log_levy ~ log_eav, 
                        subset = minor_type == "SECONDARY", 
                        index = c("agency_name", "year"), model = "within",  
                        effect= "twoways", data = schools)

school_mod_unified <- plm(log_levy ~ log_eav, 
                        subset = minor_type == "UNIFIED", 
                        index = c("agency_name", "year"), model = "within",  
                        effect= "twoways", data = schools)

school_mod_elementary <- plm(log_levy ~ log_eav, subset = minor_type == "ELEMENTARY", 
                             index = c("agency_name", "year"), model = "within",  
                             effect= "twoways", data = schools)

school_mod_comcol <- plm(log_levy ~ log_eav, index = c("agency_name", "year"), 
                         subset = minor_type == "COMM COLL", model = "within",  
                         effect= "twoways", data = schools)




model_names = c("All Schools", "Secondary", "Elementary", "Com. Colleges", "Unified" )

export_summs(school_mod_major, school_mod_secondary, school_mod_elementary, school_mod_comcol, school_mod_unified, model.names = model_names, robust=TRUE, statistics = c(N = "nobs", adjR2 = "adj.r.squared"))

plot(school_mod_elementary)
```

# New models!

## New Variables

We want to add the following:

-   Lag EAV (2 years)

-   Reassessment Year

-   Total change in EAV

### Lag EAV (3x)

```{r}

schools <- schools %>%
  arrange(year) %>%
  group_by(agency_num) %>%
  mutate(eav_lag1 = lag(log_eav, n = 1, order_by = year, default = NA)) %>%
  ungroup()

schools <- schools %>%
  arrange(year) %>%
  group_by(agency_num) %>%
  mutate(eav_lag2 = lag(log_eav, n = 2, order_by = year, default = NA)) %>%
  ungroup()

schools <- schools %>%
  arrange(year) %>%
  group_by(agency_num) %>%
  mutate(eav_lag3 = lag(log_eav, n = 3, order_by = year, default = NA)) %>%
  ungroup()
```

### Total change in EAV

When using fixed effects, independent variables must have within-group variability. Since we are using agency fixed effects, the total change in EAV for a given agency will not work in the regression model.

### Reassessment

We still haven't entirely figured this one out.

## Models: All schools!

### One lag!

```{r}

school_all_1yr <- plm(log_levy ~ log_eav + eav_lag1, index = c("agency_name", "year"),
                        model = "within",  effect= "twoways", data = schools)

summary(school_all_1yr)
```

### Two lags!

```{r}

school_all_2yr <- plm(log_levy ~ log_eav + eav_lag1 + eav_lag2, index = c("agency_name", "year"),
                      model = "within",  effect= "twoways", data = schools)

summary(school_all_2yr)
```

## Model comparisons for all schools

#### One lag, fixed effects vs. fixed effects

```{r}

f_test_1 <- pFtest(school_all_1yr, school_mod_major)

f_test_1
```

#### Two lags, fixed effects vs fixed effects

```{r}

f_test_2 <- pFtest(school_all_2yr, school_mod_major)

f_test_2
```

#### Two lags, fixed effects vs one lag, fixed effects

```{r}

f_test_1v2 <- pFtest(school_all_2yr, school_all_1yr)

f_test_1v2
```

### Conclusion:

For all schools, two lags provides the best fit.

## Models: Elementary schools

### One lagged variable.

```{r}

plm_elementary_lag1 <- plm(log_levy ~ log_eav + eav_lag1, subset = minor_type == "ELEMENTARY", 
                             index = c("agency_name", "year"), model = "within",  
                             effect= "twoways", data = schools)

summary(plm_elementary_lag1)
```

### Two lagged variables

```{r}

plm_elementary_lag2 <- plm(log_levy ~ log_eav + eav_lag1 + eav_lag2, subset = minor_type == "ELEMENTARY", 
                             index = c("agency_name", "year"), model = "within",  
                             effect= "twoways", data = schools)

summary(plm_elementary_lag2)
```

### Model Comparisons

One lag

```{r}

f_test_ele_1 <- pFtest(plm_elementary_lag1, school_mod_elementary)

f_test_ele_1
```

Two lags

```{r}

f_test_ele_2 <- pFtest(plm_elementary_lag2, school_mod_elementary)

f_test_ele_2
```

Comparing lagged models

```{r}

f_test_ele_lagged <- pFtest(plm_elementary_lag2, plm_elementary_lag1)

f_test_ele_lagged
```

## Table

```{r}

model_names = c("All Schools (No lag)", "All Schools (Lag)", "Elementary (No lag)", "Elementary (Lag)") 

lag_table <- export_summs(school_mod_major, school_all_2yr, school_mod_elementary, plm_elementary_lag2, model.names = model_names, robust=TRUE, statistics = c(N = "nobs", adjR2 = "adj.r.squared"))

lag_table
```

```{r}
#quick_docx(lag_table)
```
