---
title: "Yearly Trends - Incentive Classifications"
format: 
  html:
    code-fold: true
    toc: true
    toc-location: left
    tbl-cap-location: margin
    fig-cap-location: margin
    df-print: paged
---

```{r setup}
#| output: false

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

options(scipen = 999)

library(tidyverse)
library(glue)
library(sf)
library(DT)
library(flextable)

nicknames <- readxl::read_excel("../Necessary_Files/muni_shortnames.xlsx")

cook_sums <- read_csv("../Output/ptaxsim_cook_level_2006to2023_new.csv") 

muni_shp <- read_sf("../Necessary_Files/muni_shp.gpkg")


knitr::opts_chunk$set(warning = FALSE, message = FALSE)

set_flextable_defaults(theme_fun = theme_vanilla, 
                       padding = 2,
                       line_spacing = 1,
                       big.mark = ",",
                       )

options(DT.options = list())

FitFlextableToPage <- function(ft, pgwidth = 6){

  ft_out <- ft %>% autofit()

  ft_out <- width(ft_out, width = dim(ft_out)$widths*pgwidth /(flextable_dim(ft_out)$widths))
  return(ft_out)
}

# Create an empty data frame with a column named "year"
params <- data.frame(year = numeric(0))

# Add the value 2021 to the "year" column
params <- rbind(params, data.frame(year = 2023))

min_year = min(cook_sums$year)
max_year = max(cook_sums$year)
```



## Time Trends - FMV Growth 2011-2023 

### Pin Counts

```{r}
year_count <- 13

comm_ind_pins <- read_csv("../Output/comm_ind_PINs_2006to2023_timeseries.csv")

```

```{r}

commind_together <- comm_ind_pins  %>%
#  left_join(class_dict |> select(class, land_use)) |>
 # filter(land_use %in% c("Industrial", "Commercial")) |>
  group_by(year) %>%
  summarize(
    proptype_pincount = n(),
            n_incent = sum(ifelse(incent_prop == "Incentive", 1, 0)),
            n_nonincent = sum(ifelse(incent_prop == "Non-Incentive", 1, 0)))  %>% mutate(pct_incent = n_incent / proptype_pincount)

  

table <- comm_ind_pins %>% 
    filter(land_use %in% c("Industrial", "Commercial")) |>

  group_by(year, land_use) %>%
  summarize(
    proptype_pincount = n(),
            n_incent = sum(ifelse(incent_prop == "Incentive", 1, 0)),
            n_nonincent = sum(ifelse(incent_prop == "Non-Incentive", 1, 0))) %>%
  mutate(pct_incent = n_incent / proptype_pincount) %>%
  arrange(year) %>%
  mutate(year = as.character(year)) %>%
filter(!land_use %in% c("Exempt", "Other Residential", "Owner Occupied", "Rental") ) %>% 
  select(year, land_use, n_incent, pct_incent) |>
  pivot_wider(id_cols = year, names_from = land_use, 
              values_from = c(n_incent,pct_incent)) %>% ungroup() |>
  select(-year)

cbind(commind_together, table) %>% 
  select(Year = year, 
         "C&I Incentive Parcels" = n_incent, 
         "Share of C&I" = pct_incent,
         "Commercial Pin Count" = n_incent_Commercial, 
         "Share of Commercial" = pct_incent_Commercial,
         "Industrial Pin Count" = n_incent_Industrial, 
         "Share of Industrial" = pct_incent_Industrial,
         "C&I PIN Count" = proptype_pincount) %>% 
  mutate(Year = as.character(Year)) |>
  mutate(across(contains("Share"), scales::percent, 
                #digits = 2 ## doesn't work!! 
                accuracy = 0.01)) |>
  flextable() %>%
  add_header_row(values = c(" ", "Incent C&I", "Commercial", "Industrial","All C&I"), colwidths = c(1,2,2,2, 1))

```
```{r}
comm_ind_pins %>% 
  filter(land_use %in% c("Commercial", "Industrial")) %>%
 group_by(year, land_use) %>%
  summarize(
  PC = sum(incent_prop=="Incentive") )%>%
  ggplot() + 
  geom_col(aes(x=year, y = PC, fill = land_use)) + 
  scale_x_continuous(
    limits = c(
      #min(comm_ind_pins$year), params$year
min_year-1, max_year + 1),                                      
    breaks = c(min_year,
               2010, 2015, 2020,
              max_year
               )) + 
  theme_classic() + 
  theme(legend.position =  "bottom", legend.title = element_blank(), 
                          axis.title.x =   element_blank())
```


```{r}
commind_together_fmv <- comm_ind_pins %>%
  filter(land_use %in% c("Industrial", "Commercial")) |>
  group_by(year) %>%
  summarize(
    fmv_CI = sum(fmv, na.rm=TRUE),
    fmv_incent = sum(ifelse(incent_prop == "Incentive", fmv, 0))
          )  %>% 
  mutate(pct_incent = (fmv_incent / fmv_CI))

  

table_fmv <- comm_ind_pins %>% 
    filter(land_use %in% c("Industrial", "Commercial")) |>

  group_by(year, land_use) %>%
  summarize(
    fmv_incent = sum(ifelse(incent_prop == "Incentive", fmv, 0), na.rm=TRUE),
        fmv = sum(fmv, na.rm=TRUE),

          )  %>% 
  mutate(pct_incent = (fmv_incent / fmv)) %>%  
  
  arrange(year) %>%
  select(year, land_use, fmv_incent, pct_incent) |>
  pivot_wider(id_cols = year, names_from = land_use, 
             values_from = c(fmv_incent, pct_incent)) %>% ungroup() |>
  select(-year)


cbind(commind_together_fmv, table_fmv) %>% 
  select(Year = year, "C&I Incentive FMV" = fmv_incent, "Share of C&I" = pct_incent,
                  "Commercial Incent. FMV" = fmv_incent_Commercial, "Share of Commercial" = pct_incent_Commercial,

         "Industrial Incent. FMV" = fmv_incent_Industrial, "Share of Industrial" = pct_incent_Industrial, fmv_CI) %>% 
    mutate(Year = as.character(Year)) %>%
  mutate(across(contains("Share"), scales::percent, accuracy = 0.01)) |>

  flextable() %>%
  add_header_row(values = c(" ", "Commercial & Industrial", "Commercial", "Industrial","All C&I"), colwidths = c(1,2,2,2, 1))
```


```{r}
comm_ind_pins %>% 
  filter(land_use %in% c("Commercial", "Industrial")) %>%
 group_by(year, land_use) %>%
  summarize(
  fmv = sum(ifelse(incent_prop=="Incentive", fmv/1000000000, 0) ))%>%
  ggplot() + 
  geom_col(aes(x=year, y = fmv, fill = land_use)) + 
  scale_x_continuous(
    limits = c(min(comm_ind_pins$year), params$year), 
    breaks = c(min(comm_ind_pins$year), 2009, 2012, 2015, 2018, params$year)) + 
  theme_classic() +
  labs(y = "FMV ($ Billions)")+
  theme(legend.position =  "bottom", legend.title = element_blank(), 
                          axis.title.x =   element_blank())
```


The file ~~`comm_ind_inmunis_timeseries_2006to2023.csv`~~ `comm_ind_PINs_2011to2023_timeseries.csv` contains all PINs that had an incentive property class for at least 1 year. It includes all observations for a property during the years that it existed, even if it is not an incentive class property in that year.

Only includes PINs in incorporated areas.  



```{r}
comm_ind_pins %>%
  filter(incent_prop == "Incentive") %>%
  # group_by(year, land_use) %>%
  reframe(PC=n(), .by= c(year, land_use)) %>%
  distinct() %>%
   reframe(land_use, PC,
     both_PC = sum(PC), .by=year) %>% distinct() %>% 
  arrange(year) %>%
  group_by(land_use) %>%
  mutate(ratio_2011 = PC/ PC[year==2011],
         ratio_both_2011 = both_PC/both_PC[year==2011]) %>% 
  ggplot() +
  geom_line(aes(x=year, y = ratio_2011, group = land_use, color = land_use)) + 
  geom_line(aes(x=year, y =ratio_both_2011), lty = 1) + 
  theme_bw() + 
    scale_x_continuous(
    limits = c(min(comm_ind_pins$year), params$year), 
    breaks = c(min(comm_ind_pins$year), 2009,  2015, 2018, 2021, params$year))
```


```{r}
comm_ind_pins %>%
  filter(incent_prop == "Incentive") %>%
   reframe(fmv,
           type_fmv = sum(fmv, na.rm=TRUE), .by = c(year,land_use)
          ) %>%
   reframe(land_use, type_fmv,
          fmv_incent = sum(fmv, na.rm=TRUE), .by=year) %>% distinct() %>% 
  arrange(year) %>%
  group_by(land_use) %>%
  mutate(ratio_2011 = type_fmv/ type_fmv[year==2011],
         ratio_both_2011 = fmv_incent/fmv_incent[year==2011]) %>% 
  ggplot() +
  geom_line(aes(x=year, y = ratio_2011, group = land_use, color = land_use)) + 
  geom_line(aes(x=year, y =ratio_both_2011), lty = 1) + 
    scale_x_continuous(
    limits = c(min(comm_ind_pins$year), params$year), 
    breaks = c( 2009, 2012, 2015, 2018, 2021, params$year)) + 
  theme_classic()
```



```{r}
#| label: tbl-categorycountsforcomparison-2023
#| tbl-cap: "Number of PINs that gain or lose incentive classification each year."


pins_per_year <- comm_ind_pins %>% 
  group_by(year) %>%
  summarize("Gains Incent" = sum(gain_incent, na.rm=TRUE),
            "Loses Incent" = sum(lose_incent, na.rm=TRUE),
            "PINs with Incent. Class." = sum(incent_prop=="Incentive"),
            "PINs in Sample" = n(),
            "Became Exempt" = sum(became_exempt, na.rm=TRUE))

pins_per_year
```



```{r}
df_2011_class8change <- comm_ind_pins|>
  group_by(pin) |>
      mutate(class8years = sum(class_1dig == 8)) |>

  mutate(class8_change = case_when(
      class8years == year_count ~ "Always Class 8",
      class8years == 0 ~ "Never Class 8",
      TRUE ~ "Changes to C8")
  ) |> 
  ungroup() |>
group_by(year, class8_change) |>
  summarize(
    group_pin_count = n(),
    group_year_fmv = sum(fmv, na.rm=TRUE),
  ) |> ungroup() |>
  group_by( class8_change) %>%
  mutate(
    base_year_fmv = group_year_fmv[year == min(year)],
    fmv_group_growth = (group_year_fmv / base_year_fmv) -1)  



# df_2011_class8change %>% 
#   filter(year == 2022) %>%
#   select(class8_change, fmv_group_growth, base_year_fmv, group_year_fmv, everything(), -year)

df_2011_class8change %>% 
  filter(year == 2023) %>%
  select(class8_change, fmv_group_growth, base_year_fmv, group_year_fmv, everything(), -year)
```


```{r}
#| label: tbl-class8eachyear-2023
#| tbl-cap: "Number of PINs that became each major class type each year."

comm_ind_pins <- comm_ind_pins %>% 
  group_by(pin) |>
  mutate(class8years = sum(class_1dig == 8),
         class8_change = case_when(
           class8years == year_count ~ "Always Class 8",
           class8years == 0 ~ "Never Class 8",
           sum(lag(class_1dig) != 8 & class_1dig == 8) > 0 ~ "Becomes Class 8",
           sum(class_1dig != 8 & lag(class_1dig) == 8 ) > 0 ~ "Leaves Class 8",
           TRUE ~ "Other"),
         becomes8 = ifelse(lag(class_1dig) != 8 & class_1dig == 8, 1, 0),
         becomes7 = ifelse(lag(class_1dig) != 7 & class_1dig == 7, 1, 0),
         becomes6 = ifelse(lag(class_1dig) != 6 & class_1dig == 6, 1, 0),
         switches_to8 = ifelse(sum(incent_prop == "Incentive")> 1 & becomes8 ==1, 1, 0),
         # is_exempt = ifelse(class %in% c("0", "EX"), 1, 0),
         #  was_exempt = ifelse(lag(class) %in% c("0","EX"), 1, 0)
  ) %>%
        ungroup()

 classpins_per_year <- comm_ind_pins |>
   group_by(year) %>%
  summarize("Becomes 8" = sum(becomes8==1, na.rm=TRUE),
            "Total 8" = sum(class_1dig == 8),

            "6 or 7 to 8" = sum(switches_to8 == 1, na.rm=TRUE),

            "Becomes 7" = sum(becomes7 == 1, na.rm=TRUE),
            "Total 7" = sum(class_1dig == 7),

            "Becomes 6" = sum(becomes6 == 1, na.rm=TRUE),
            "Total 6" = sum(class_1dig == 6),
            "PINs with Incent. Class." = sum(incent_prop=="Incentive"),
          #  "Tax Exempt" = sum(is_exempt),
            "PINs in Sample" = n(),
            #"Became Exempt" = sum(became_exempt, na.rm=TRUE)
          )
classpins_per_year %>%  flextable::flextable()
```

### Explore PINs that became Class 8 or Other Incentive in South Triad

```{r}
#| label: tbl-class8eachyearsouthtriad-2023
#| tbl-cap: "Number of PINs that became each major class 8 in the South triad."
#| 
# comm_ind_pins %>% 
#   filter(class8_change == "Becomes Class 8") %>% 
#   left_join(bor, by = c("pin", "year" = "tax_year")) %>%
#   filter(!is.na(appellant)) %>%
#   
#   arrange(pin, desc(year)) %>%
#   select(year, appellant, project_id, class.x, everything())

# example of PIN that was class 8 and then became un-incentivized in 2017
# comm_ind_pins %>% filter(pin == "16271000330000")


comm_ind_pins %>% 
  filter(class8_change == "Becomes Class 8" & Triad == "South") %>% 
  arrange(pin) %>%
  select(year, class, everything())
```


```{r}
#| label: tbl-southtriadpins2023
#| tbl-cap: "South Triad Number of PINs that became each major class type each year."
#| layout-ncol: 3

comm_ind_pins %>% 
  filter(becomes8 == 1 & Triad == "South") %>% 
  arrange(pin) %>%
  select(year, class, everything())

comm_ind_pins %>% 
  filter(becomes7== 1& Triad == "South") %>% 
  arrange(pin) %>%
  select(year, class, everything())

comm_ind_pins %>% 
  filter(becomes6 == 1 & Triad == "South") %>% 
  arrange(pin) %>%
  select(year, class, everything())
```

> note to self: status is coded incorrectly. There are 'Gains Incentive' PINs that are also Controls if they existed less than 12 years. Became_taxed is wrong too

```{r}
#| label: tbl-incentchange-2023indexed
#| tbl-cap: "Aggregate FMV Growth by Incentive Classification. `Changes Sometime` includes properties that gained or lost an incentive classification. `incent_status` in tables below breaks up Changes Sometime into more detailed categories. "

df_2011_incentchange <- comm_ind_pins|>
group_by(year, incent_change) |>
  summarize(
    group_pin_count = n(),
    group_year_fmv = sum(fmv, na.rm=TRUE),
  ) |> ungroup() |>
  group_by( incent_change) %>%
  mutate(
    base_year_fmv = group_year_fmv[year == min(year)],
    fmv_group_growth = (group_year_fmv / base_year_fmv) -1)  



df_2011_incentchange %>% 
  #filter(year == params$year) %>%
  filter(year == 2023) |>
  select(incent_change, fmv_group_growth, base_year_fmv, group_year_fmv, everything(), -year)
```




```{r}
#| label: tbl-growthcalculations-byincentstatus-2023
#| tbl-cap: "Aggregate FMV Growth by `incent_status`. Calculations for PINs counts in header of table in report."

comm_ind_pins %>%
  ungroup() %>% 
  group_by(year, incent_status) |>
  summarize(
    group_pin_count = n(),
    group_year_fmv = sum(fmv, na.rm=TRUE)) |> 
  ungroup() |>
group_by(incent_status) %>%
  mutate(
    base_year_fmv = group_year_fmv[year == min(year)],
    fmv_group_growth = (group_year_fmv / base_year_fmv) -1) |> 
  filter(year == params$year)  %>%
  select(incent_status, fmv_group_growth, group_pin_count, everything(), -year)

```


```{r}
#| label: tbl-growthcalculations-bylandusechange-2023
#| tbl-cap: !expr paste0("Aggregate FMV Growth from 2011 to ",  params$year, " by `landuse_change`." )

comm_ind_pins %>%
  group_by(year, landuse_change) |>
  summarize(
    group_pin_count = n(),
    group_year_fmv = sum(fmv, na.rm=TRUE)) |> 
  ungroup() |>
  group_by(landuse_change) %>%
  mutate(
    base_year_fmv = group_year_fmv[year == min(year)],
    base_year_n = group_pin_count[year == min(year)],
    n_group_growth = group_pin_count/base_year_n -1,
    fmv_group_growth = (group_year_fmv / base_year_fmv) -1) %>%
    filter(year == params$year) %>%   
  select(landuse_change, fmv_group_growth, group_pin_count)
```

```{r}
#| label: tbl-growthcalculations-landusechange-incentstatus-balanced-2023
#| tbl-cap: !expr paste0("Growth from 2011 to ",  params$year, " - Change in Land Use by Incentive Class Status. Non-winsorized version of table used in Table X of report.")

df_2011_bal <- comm_ind_pins %>%
  group_by(year, landuse_change, incent_status) |>
  summarize(
    group_pin_count = n(),
    group_year_fmv = sum(fmv, na.rm=TRUE)) |> 
  ungroup() |>
  group_by(landuse_change, incent_status) %>%
  mutate(
    base_year_fmv = group_year_fmv[year == min(year)],
    base_year_n = group_pin_count[year == min(year)],
    n_group_growth = group_pin_count/base_year_n -1,
    fmv_group_growth = (group_year_fmv / base_year_fmv) -1) 

df_2011_bal %>% 
  select(year, landuse_change, incent_status, fmv_group_growth, group_pin_count) %>% 
  filter(year == params$year) %>%   
  select(-year)


```



```{r}
#| label: fig-landusebyincentstatus-2023
#| fig-cap: "Includes the 'Excluded' Category"

df_2011_bal %>% 
  ggplot() + 
  geom_line(aes(x=year, fmv_group_growth, group = incent_status, color = incent_status)) + 
  labs( title = "FMV Growth Since 2011 by Land Use",) + 
  theme_bw()  +  
  facet_wrap(~landuse_change)
```


```{r}
#| label: fig-landusebyincentstatus-clean-2023
#| fig-cap: "Excludes the 'Excluded' Category and 'Exempt Sometime' properties."

df_2011_bal %>% 
  filter(incent_status != "Excluded" & landuse_change != "Excluded" & landuse_change != "Exempt Sometime") %>%
  ggplot() + 
  geom_line(aes(x=year, fmv_group_growth, group = incent_status, color = incent_status)) + 
  labs( title = "FMV Growth Since 2011 by Land Use", 
        caption = "PINs that did not exist during all years of the sample frame were excluded from the image (n=10,809)."
        ) + 
  scale_color_brewer(direction = -1) + 
  theme_bw()  +  
    scale_x_continuous(
    limits = c(min(comm_ind_pins$year), params$year), 
    breaks = c( 2009, 2012, 2015, 2018, 2021, params$year)) + 
  facet_wrap(~landuse_change)

```


```{r}
#| label: fig-growth-facetby-incentstatus-2023
#| fig-cap: !expr paste0("Growth from 2011 to ", params$year,  ". Faceted by if a PIN changed landuse during the sample period. Indexed to fair market value during 2011")


df_2011_bal %>%  
  mutate(year = as.factor(year)) %>%
  ggplot() + 
  geom_line(aes(x=year, y=fmv_group_growth, group = landuse_change, color = landuse_change)) +
  theme_bw()  +  
  facet_wrap(~incent_status, nrow = 1) +
  scale_x_discrete(breaks = c(2012, params$year)) + 
  scale_y_continuous(breaks = c(-1, 0, 1, 2, 3, 4, 10), labels = scales::percent,) +
  scale_color_brewer(direction = -1) + 

  labs(title= paste0("Growth from 2011 to ", params$year),
       subtitle = "Incentive Classification Status by Land Use Change",
       y = "FMV Growth since 2011", x = NULL,
       caption =  "Values are indexed to 2011 FMV") +
  theme(legend.title = element_blank(), legend.position = "bottom")
```


```{r}
#| label: fig-growth-facetby-incentstatus-clean-2023
#| fig-cap: !expr paste0("Aggregate FMV Growth from 2011 to ", params$year, ". Faceted by PIN  incentive status during the sample period. Indexed to fair market value during 2011.")

df_2011_bal %>%  
  filter(landuse_change != "Excluded" & incent_status != "Excluded" & landuse_change != "Exempt Sometime") %>%
  mutate(year = as.factor(year)) %>%
  ggplot() + 
  geom_line(aes(x=year, y=fmv_group_growth, group = landuse_change, color = landuse_change)) +
  theme_bw()  +  
  facet_wrap(~incent_status, nrow = 1) +
  scale_x_discrete(breaks = c(2012, params$year), ) + 
  scale_y_continuous(breaks = c(-1, 0, 1, 2, 3, 4, 10), labels = scales::percent,) +
  labs(title= "Growth from 2011", 
       subtitle = "Incentive Classification Status by Land Use Change",
       y = "FMV Growth since 2011", x = NULL,
       caption =  paste0("Values are indexed to 2011 FMV. Excludes PINs that were tax exempt some years or did not exist for all years between 2011 and ", params$year)) +
  theme(legend.title = element_blank(), legend.position = "bottom")

```



## Trends for Incentive Report - FMV Growth 2011-2022
### Pin Counts

```{r}
# Create an empty data frame with a column named "year"
params <- data.frame(year = numeric(0))

# Add the value 2021 to the "year" column
params <- rbind(params, data.frame(year = 2022))


year_count <- 12

comm_ind_pins <- read_csv("../Output/comm_ind_PINs_2011to2022_timeseries.csv")

```

```{r}

commind_together <- comm_ind_pins  %>%
#  left_join(class_dict |> select(class, land_use)) |>
 # filter(land_use %in% c("Industrial", "Commercial")) |>
  group_by(year) %>%
  summarize(
    proptype_pincount = n(),
            n_incent = sum(ifelse(incent_prop == "Incentive", 1, 0)),
            n_nonincent = sum(ifelse(incent_prop == "Non-Incentive", 1, 0)))  %>% mutate(pct_incent = n_incent / proptype_pincount)

  

table <- comm_ind_pins %>% 
    filter(land_use %in% c("Industrial", "Commercial")) |>

  group_by(year, land_use) %>%
  summarize(
    proptype_pincount = n(),
            n_incent = sum(ifelse(incent_prop == "Incentive", 1, 0)),
            n_nonincent = sum(ifelse(incent_prop == "Non-Incentive", 1, 0))) %>%
  mutate(pct_incent = n_incent / proptype_pincount) %>%
  arrange(year) %>%
  mutate(year = as.character(year)) %>%
filter(!land_use %in% c("Exempt", "Other Residential", "Owner Occupied", "Rental") ) %>% 
  select(year, land_use, n_incent, pct_incent) |>
  pivot_wider(id_cols = year, names_from = land_use, 
              values_from = c(n_incent,pct_incent)) %>% ungroup() |>
  select(-year)

cbind(commind_together, table) %>% 
  select(Year = year, 
         "C&I Incentive Parcels" = n_incent, 
         "Share of C&I" = pct_incent,
         "Commercial Pin Count" = n_incent_Commercial, 
         "Share of Commercial" = pct_incent_Commercial,
         "Industrial Pin Count" = n_incent_Industrial, 
         "Share of Industrial" = pct_incent_Commercial,
         "C&I PIN Count" = proptype_pincount) %>% 
  mutate(Year = as.character(Year)) |>
  mutate(across(contains("Share"), scales::percent, digits = 2)) |>
  flextable() %>%
  add_header_row(values = c(" ", "Incent C&I", "Commercial", "Industrial","All C&I"), colwidths = c(1,2,2,2, 1))

```
```{r}
comm_ind_pins %>% 
  filter(land_use %in% c("Commercial", "Industrial")) %>%
 group_by(year, land_use) %>%
  summarize(
  PC = sum(incent_prop=="Incentive") )%>%
  ggplot() + 
  geom_col(aes(x=year, y = PC, fill = land_use)) + 
  scale_x_continuous(
    limits = c(min(comm_ind_pins$year), params$year),                                      
    breaks = c(min(comm_ind_pins$year), 2010, 2015, 2020, params$year)) + 
  theme_classic() + 
  theme(legend.position =  "bottom", legend.title = element_blank(), 
                          axis.title.x =   element_blank())
```


```{r}
commind_together_fmv <- comm_ind_pins %>%
  filter(land_use %in% c("Industrial", "Commercial")) |>
  group_by(year) %>%
  summarize(
    fmv_CI = sum(fmv, na.rm=TRUE),
    fmv_incent = sum(ifelse(incent_prop == "Incentive", fmv, 0))
          )  %>% 
  mutate(pct_incent = (fmv_incent / fmv_CI))

  

table_fmv <- comm_ind_pins %>% 
    filter(land_use %in% c("Industrial", "Commercial")) |>

  group_by(year, land_use) %>%
  summarize(
    fmv = sum(fmv, na.rm=TRUE),
    fmv_incent = sum(ifelse(incent_prop == "Incentive", fmv, 0), na.rm=TRUE)
          )  %>% 
  mutate(pct_incent = (fmv_incent / fmv) / 10000) %>%  ## should standardize percentage formatting....
  arrange(year) %>%
  select(year, land_use, fmv_incent, pct_incent) |>
  pivot_wider(id_cols = year, names_from = land_use, 
             values_from = c(fmv_incent, pct_incent)) %>% ungroup() |>
  select(-year)

cbind(commind_together_fmv, table_fmv) %>% 
  select(Year = year, "C&I Incentive FMV" = fmv_incent, "Share of C&I" = pct_incent,
                  "Commercial Incent. FMV" = fmv_incent_Commercial, "Share of Commercial" = pct_incent_Commercial,

         "Industrial Incent. FMV" = fmv_incent_Industrial, "Share of Industrial" = pct_incent_Industrial, fmv_CI) %>% 
    mutate(Year = as.character(Year)) %>%
  mutate(across(contains("Share"), scales::percent, digits = 2)) |>

  flextable() %>%
  add_header_row(values = c(" ", "Commercial & Industrial", "Commercial", "Industrial","All C&I"), colwidths = c(1,2,2,2, 1))
```


```{r}
comm_ind_pins %>% 
  filter(land_use %in% c("Commercial", "Industrial")) %>%
 group_by(year, land_use) %>%
  summarize(
  fmv = sum(ifelse(incent_prop=="Incentive", fmv/1000000000, 0) ))%>%
  ggplot() + 
  geom_col(aes(x=year, y = fmv, fill = land_use)) + 
  scale_x_continuous(
    limits = c(min(comm_ind_pins$year), params$year), 
    breaks = c(min(comm_ind_pins$year), 2009, 2012, 2015, 2018, params$year)) + 
  theme_classic() +
  labs(y = "FMV ($ Billions)")+
  theme(legend.position =  "bottom", legend.title = element_blank(), 
                          axis.title.x =   element_blank())
```


The file ~~`comm_ind_inmunis_timeseries_2006to2022.csv`~~ `comm_ind_PINs_2011to2022_timeseries.csv` contains all PINs that had an incentive property class for at least 1 year. It includes all observations for a property during the years that it existed, even if it is not an incentive class property in that year.

Only includes PINs in incorporated areas.  



```{r}
comm_ind_pins %>%
  filter(incent_prop == "Incentive") %>%
  # group_by(year, land_use) %>%
  reframe(PC=n(), .by= c(year, land_use)) %>%
  distinct() %>%
   reframe(land_use, PC,
     both_PC = sum(PC), .by=year) %>% distinct() %>% 
  arrange(year) %>%
  group_by(land_use) %>%
  mutate(ratio_2011 = PC/ PC[year==2011],
         ratio_both_2011 = both_PC/both_PC[year==2011]) %>% 
  ggplot() +
  geom_line(aes(x=year, y = ratio_2011, group = land_use, color = land_use)) + 
  geom_line(aes(x=year, y =ratio_both_2011), lty = 1) + 
  theme_bw() + 
    scale_x_continuous(
    limits = c(min(comm_ind_pins$year), params$year), 
    breaks = c(min(comm_ind_pins$year), 2009,  2015, 2018, 2021, params$year))
```


```{r}
comm_ind_pins %>%
  filter(incent_prop == "Incentive") %>%
   reframe(fmv,
           type_fmv = sum(fmv, na.rm=TRUE), .by = c(year,land_use)
          ) %>%
   reframe(land_use, type_fmv,
          fmv_incent = sum(fmv, na.rm=TRUE), .by=year) %>% distinct() %>% 
  arrange(year) %>%
  group_by(land_use) %>%
  mutate(ratio_2011 = type_fmv/ type_fmv[year==2011],
         ratio_both_2011 = fmv_incent/fmv_incent[year==2011]) %>% 
  ggplot() +
  geom_line(aes(x=year, y = ratio_2011, group = land_use, color = land_use)) + 
  geom_line(aes(x=year, y =ratio_both_2011), lty = 1) + 
    scale_x_continuous(
    limits = c(min(comm_ind_pins$year), params$year), 
    breaks = c( 2009, 2012, 2015, 2018, 2021, params$year)) + 
  theme_classic()
```



```{r}
#| label: tbl-categorycountsforcomparison
#| tbl-cap: "Number of PINs that gain or lose incentive classification each year."


pins_per_year <- comm_ind_pins %>% 
  group_by(year) %>%
  summarize("Gains Incent" = sum(gain_incent, na.rm=TRUE),
            "Loses Incent" = sum(lose_incent, na.rm=TRUE),
            "PINs with Incent. Class." = sum(incent_prop=="Incentive"),
            "PINs in Sample" = n(),
            "Became Exempt" = sum(became_exempt, na.rm=TRUE))

pins_per_year
```



```{r}
df_2011_class8change <- comm_ind_pins|>
  group_by(pin) |>
      mutate(class8years = sum(class_1dig == 8)) |>

  mutate(class8_change = case_when(
      class8years == year_count ~ "Always Class 8",
      class8years == 0 ~ "Never Class 8",
      TRUE ~ "Changes to C8")
  ) |> 
  ungroup() |>
group_by(year, class8_change) |>
  summarize(
    group_pin_count = n(),
    group_year_fmv = sum(fmv, na.rm=TRUE),
  ) |> ungroup() |>
  group_by( class8_change) %>%
  mutate(
    base_year_fmv = group_year_fmv[year == min(year)],
    fmv_group_growth = (group_year_fmv / base_year_fmv) -1)  



# df_2011_class8change %>% 
#   filter(year == 2022) %>%
#   select(class8_change, fmv_group_growth, base_year_fmv, group_year_fmv, everything(), -year)

df_2011_class8change %>% 
  filter(year == 2023) %>%
  select(class8_change, fmv_group_growth, base_year_fmv, group_year_fmv, everything(), -year)
```


```{r}
#| label: tbl-class8eachyear
#| tbl-cap: "Number of PINs that became each major class type each year."

comm_ind_pins <- comm_ind_pins %>% 
  group_by(pin) |>
  mutate(class8years = sum(class_1dig == 8),
         class8_change = case_when(
           class8years == year_count ~ "Always Class 8",
           class8years == 0 ~ "Never Class 8",
           sum(lag(class_1dig) != 8 & class_1dig == 8) > 0 ~ "Becomes Class 8",
           sum(class_1dig != 8 & lag(class_1dig) == 8 ) > 0 ~ "Leaves Class 8",
           TRUE ~ "Other"),
         becomes8 = ifelse(lag(class_1dig) != 8 & class_1dig == 8, 1, 0),
         becomes7 = ifelse(lag(class_1dig) != 7 & class_1dig == 7, 1, 0),
         becomes6 = ifelse(lag(class_1dig) != 6 & class_1dig == 6, 1, 0),
         switches_to8 = ifelse(sum(incent_prop == "Incentive")> 1 & becomes8 ==1, 1, 0),
         # is_exempt = ifelse(class %in% c("0", "EX"), 1, 0),
         #  was_exempt = ifelse(lag(class) %in% c("0","EX"), 1, 0)
  ) %>%
        ungroup()

 classpins_per_year <- comm_ind_pins |>
   group_by(year) %>%
  summarize("Becomes 8" = sum(becomes8==1, na.rm=TRUE),
            "Total 8" = sum(class_1dig == 8),

            "6 or 7 to 8" = sum(switches_to8 == 1, na.rm=TRUE),

            "Becomes 7" = sum(becomes7 == 1, na.rm=TRUE),
            "Total 7" = sum(class_1dig == 7),

            "Becomes 6" = sum(becomes6 == 1, na.rm=TRUE),
            "Total 6" = sum(class_1dig == 6),
            "PINs with Incent. Class." = sum(incent_prop=="Incentive"),
          #  "Tax Exempt" = sum(is_exempt),
            "PINs in Sample" = n(),
            #"Became Exempt" = sum(became_exempt, na.rm=TRUE)
          )
classpins_per_year %>%  flextable::flextable()
```

### Explore PINs that became Class 8 or Other Incentive in South Triad

```{r}
#| label: tbl-class8eachyearsouthtriad
#| tbl-cap: "Number of PINs that became each major class 8 in the South triad."
#| 
# comm_ind_pins %>% 
#   filter(class8_change == "Becomes Class 8") %>% 
#   left_join(bor, by = c("pin", "year" = "tax_year")) %>%
#   filter(!is.na(appellant)) %>%
#   
#   arrange(pin, desc(year)) %>%
#   select(year, appellant, project_id, class.x, everything())

# example of PIN that was class 8 and then became un-incentivized in 2017
# comm_ind_pins %>% filter(pin == "16271000330000")


comm_ind_pins %>% 
  filter(class8_change == "Becomes Class 8" & Triad == "South") %>% 
  arrange(pin) %>%
  select(year, class, everything())
```


```{r}
#| label: tbl-southtriadpins
#| tbl-cap: "South Triad Number of PINs that became each major class type each year."
#| layout-ncol: 3

comm_ind_pins %>% 
  filter(becomes8 == 1 & Triad == "South") %>% 
  arrange(pin) %>%
  select(year, class, everything())

comm_ind_pins %>% 
  filter(becomes7== 1& Triad == "South") %>% 
  arrange(pin) %>%
  select(year, class, everything())

comm_ind_pins %>% 
  filter(becomes6 == 1 & Triad == "South") %>% 
  arrange(pin) %>%
  select(year, class, everything())
```

> note to self: status is coded incorrectly. There are 'Gains Incentive' PINs that are also Controls if they existed less than 12 years. Became_taxed is wrong too

```{r}
#| label: tbl-incentchange-2022indexed
#| tbl-cap: "Aggregate FMV Growth by Incentive Classification. `Changes Sometime` includes properties that gained or lost an incentive classification. `incent_status` in tables below breaks up Changes Sometime into more detailed categories. "

df_2011_incentchange <- comm_ind_pins|>
group_by(year, incent_change) |>
  summarize(
    group_pin_count = n(),
    group_year_fmv = sum(fmv, na.rm=TRUE),
  ) |> ungroup() |>
  group_by( incent_change) %>%
  mutate(
    base_year_fmv = group_year_fmv[year == min(year)],
    fmv_group_growth = (group_year_fmv / base_year_fmv) -1)  



df_2011_incentchange %>% 
  #filter(year == params$year) %>%
  filter(year == 2023) |>
  select(incent_change, fmv_group_growth, base_year_fmv, group_year_fmv, everything(), -year)
```




```{r}
#| label: tbl-growthcalculations-byincentstatus
#| tbl-cap: "Aggregate FMV Growth by `incent_status`. Calculations for PINs counts in header of table in report."

comm_ind_pins %>%
  ungroup() %>% 
  group_by(year, incent_status) |>
  summarize(
    group_pin_count = n(),
    group_year_fmv = sum(fmv, na.rm=TRUE)) |> 
  ungroup() |>
group_by(incent_status) %>%
  mutate(
    base_year_fmv = group_year_fmv[year == min(year)],
    fmv_group_growth = (group_year_fmv / base_year_fmv) -1) |> 
  filter(year == params$year)  %>%
  select(incent_status, fmv_group_growth, group_pin_count, everything(), -year)

```


```{r}
#| label: tbl-growthcalculations-bylandusechange
#| tbl-cap: !expr paste0("Aggregate FMV Growth from 2011 to ",  params$year, " by `landuse_change`." )

comm_ind_pins %>%
  group_by(year, landuse_change) |>
  summarize(
    group_pin_count = n(),
    group_year_fmv = sum(fmv, na.rm=TRUE)) |> 
  ungroup() |>
  group_by(landuse_change) %>%
  mutate(
    base_year_fmv = group_year_fmv[year == min(year)],
    base_year_n = group_pin_count[year == min(year)],
    n_group_growth = group_pin_count/base_year_n -1,
    fmv_group_growth = (group_year_fmv / base_year_fmv) -1) %>%
    filter(year == params$year) %>%   
  select(landuse_change, fmv_group_growth, group_pin_count)
```

```{r}
#| label: tbl-growthcalculations-landusechange-incentstatus-balanced
#| tbl-cap: !expr paste0("Growth from 2011 to ",  params$year, " - Change in Land Use by Incentive Class Status. Non-winsorized version of table used in Table X of report.")

df_2011_bal <- comm_ind_pins %>%
  group_by(year, landuse_change, incent_status) |>
  summarize(
    group_pin_count = n(),
    group_year_fmv = sum(fmv, na.rm=TRUE)) |> 
  ungroup() |>
  group_by(landuse_change, incent_status) %>%
  mutate(
    base_year_fmv = group_year_fmv[year == min(year)],
    base_year_n = group_pin_count[year == min(year)],
    n_group_growth = group_pin_count/base_year_n -1,
    fmv_group_growth = (group_year_fmv / base_year_fmv) -1) 

df_2011_bal %>% 
  select(year, landuse_change, incent_status, fmv_group_growth, group_pin_count) %>% 
  filter(year == params$year) %>%   
  select(-year)


```



```{r}
#| label: fig-landusebyincentstatus
#| fig-cap: "Includes the 'Excluded' Category"

df_2011_bal %>% 
  ggplot() + 
  geom_line(aes(x=year, fmv_group_growth, group = incent_status, color = incent_status)) + 
  labs( title = "FMV Growth Since 2011 by Land Use",) + 
  theme_bw()  +  
  facet_wrap(~landuse_change)
```


```{r}
#| label: fig-landusebyincentstatus-clean
#| fig-cap: "Excludes the 'Excluded' Category and 'Exempt Sometime' properties."

df_2011_bal %>% 
  filter(incent_status != "Excluded" & landuse_change != "Excluded" & landuse_change != "Exempt Sometime") %>%
  ggplot() + 
  geom_line(aes(x=year, fmv_group_growth, group = incent_status, color = incent_status)) + 
  labs( title = "FMV Growth Since 2011 by Land Use", 
        caption = "PINs that did not exist during all years of the sample frame were excluded from the image (n=10,809)."
        ) + 
  scale_color_brewer(direction = -1) + 
  theme_bw()  +  
    scale_x_continuous(
    limits = c(min(comm_ind_pins$year), params$year), 
    breaks = c( 2009, 2012, 2015, 2018, 2021, params$year)) + 
  facet_wrap(~landuse_change)

```


```{r}
#| label: fig-growth-facetby-incentstatus
#| fig-cap: !expr paste0("Growth from 2011 to ", params$year,  ". Faceted by if a PIN changed landuse during the sample period. Indexed to fair market value during 2011")


df_2011_bal %>%  
  mutate(year = as.factor(year)) %>%
  ggplot() + 
  geom_line(aes(x=year, y=fmv_group_growth, group = landuse_change, color = landuse_change)) +
  theme_bw()  +  
  facet_wrap(~incent_status, nrow = 1) +
  scale_x_discrete(breaks = c(2012, params$year)) + 
  scale_y_continuous(breaks = c(-1, 0, 1, 2, 3, 4, 10), labels = scales::percent,) +
  scale_color_brewer(direction = -1) + 

  labs(title= paste0("Growth from 2011 to ", params$year),
       subtitle = "Incentive Classification Status by Land Use Change",
       y = "FMV Growth since 2011", x = NULL,
       caption =  "Values are indexed to 2011 FMV") +
  theme(legend.title = element_blank(), legend.position = "bottom")
```


```{r}
#| label: fig-growth-facetby-incentstatus-clean
#| fig-cap: !expr paste0("Aggregate FMV Growth from 2011 to ", params$year, ". Faceted by PIN  incentive status during the sample period. Indexed to fair market value during 2011.")

df_2011_bal %>%  
  filter(landuse_change != "Excluded" & incent_status != "Excluded" & landuse_change != "Exempt Sometime") %>%
  mutate(year = as.factor(year)) %>%
  ggplot() + 
  geom_line(aes(x=year, y=fmv_group_growth, group = landuse_change, color = landuse_change)) +
  theme_bw()  +  
  facet_wrap(~incent_status, nrow = 1) +
  scale_x_discrete(breaks = c(2012, params$year), ) + 
  scale_y_continuous(breaks = c(-1, 0, 1, 2, 3, 4, 10), labels = scales::percent,) +
  labs(title= "Growth from 2011", 
       subtitle = "Incentive Classification Status by Land Use Change",
       y = "FMV Growth since 2011", x = NULL,
       caption =  paste0("Values are indexed to 2011 FMV. Excludes PINs that were tax exempt some years or did not exist for all years between 2011 and ", params$year)) +
  theme(legend.title = element_blank(), legend.position = "bottom")

```

## Export Tables

```{r}
#| eval: false

library(readxl)

tablelist <- list(
  "Muni Stats" = munilevel,
  "Muni Rate Change" = muni_ratechange,
  
  "Cook Sums" = table_cook,
  "Cook Class Sums" = cty_MC_table,
  "Cook Rate Change" = cook_ratechange
)




#writexl::write_xlsx(tablelist, "Output/calculations_report_content_20240927.xlsx")

readme <- c(
  "Pin Change: Count of PINs gaining or losing an incentive class each year.",
  "Major Class Change: Counts of PINs becoming major class types each year.",
  "Top 10 Commercial Incents is based on summed FMV from properties that have commercial incentive classes.",
  "Top 10 Industrial Incents is based on summed FMV from properties that have industrial incentive classes.",
  
  "Muni Shares shows the top 5, middle 5, and bottom 5 munis based on the share of FMV that has an incentive property out of all FMV in the municipality. Note: Municipalities without commercial or industrial property are excluded from this list.",
  "Commerc_Share shows the top 5, middle 5, and bottom 5 municipalities based on share of commercial incentive FMV out of the commercial FMV within the municipality.Note: There are at least 50 municipalities that do not have commercial incentive properties. The bottom 5 displayed are alphabetically last, but there are 50 tied at 0%.",
            "Indust_Share shows the top 5, middle 5, and bottom 5 municipalities based on the share of industrial incentive FMV out of the industrial FMV within the municipality. Note: There are at least 38 municipalities that do not have industrial incentive properties. The bottom 5 displayed are alphabetically last, but there are 38 tied at 0%.",
  
  "Muni Share_all includes all municipalities and their share of FMV with incentives out of all FMV in the municipality.",
  "IndustShare_all is same as Indust_share but includes all munis.",
  "CommercShare_all is same as Commerc_share but includes all munis.") %>% as.data.frame()

paper_tables <- list(
  "README"= readme,
  
  "Pin Change per Year" = pins_per_year,
  "Top 10 Commercial Incents" = commerc_top10,
  "Top 10 Indust Incents" = indust_top10,
  
  "Muni Shares" = muni_incentshare_slice,
  "Indust_Share" = indust_share,
  "Commerc_Share" = commerc_share,

  "Muni Share_all" = muni_incent_share,
  "IndustShare_all" = indust_share_full,
  "Commerc_Share_all" = commerc_share_full,
  
  "Rate Change" = muni_ratechange_sliced
  
)

# writexl::write_xlsx(paper_tables, "../Output/incentive_report_tables.xlsx")

```

