---
title: "Additional Findings"
---

```{r setup, include=FALSE}
library(tidyverse)
library(ptaxsim)
library(DBI)
library(httr)
library(jsonlite)
library(glue)
library(sf)

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

```

## Residential EAV

```{r}
muni_shp <- read_sf("https://gis.cookcountyil.gov/traditional/rest/services/politicalBoundary/MapServer/2/query?outFields=*&where=1%3D1&f=geojson")


nicknames <- readxl::read_excel("../Necessary_Files/muni_shortnames.xlsx")

```

```{r}
muni_sums <- read_csv("../Output/ptaxsim_muni_level_2006-2022.csv") %>% filter(year == 2021) %>% left_join(nicknames)

MuniLevel_CompRates <- muni_sums %>% 
  select(clean_name, muni_current_rate_avg, muni_zero_bill, muni_levy )
  #select(clean_name, cur_comp_muni_rate, new_comp_muni_rate, tif_share, zero_bills, total_bill_current) %>%
  mutate(cur_comp_muni_rate = cur_comp_muni_rate/100,
new_comp_muni_rate = new_comp_muni_rate / 100 )

#install.packages("DT")
library(DT)

datatable(MuniLevel_CompRates, rownames = FALSE,
          colnames = c('Municipality' = 'clean_name', 'Current Comp. Rate' = 'cur_comp_muni_rate', 'Hypothetical Rate' = 'new_comp_muni_rate', '%Rev to TIF' = 'tif_share', 'Count of $0 Tax Bills' = 'zero_bills', 'Composite Levy' = 'total_bill_current'),
          caption = "Table 1: Current and Hypothetical Composite Tax Rates if GHE $0") %>%
  formatPercentage(c('Current Comp. Rate', 'Hypothetical Rate', '%Rev to TIF'), digits = 2) %>%  
  formatCurrency('Composite Levy', digits = 0)
          
```

```{r}
muni_MC_sums <- read_csv("../Output/ptaxsim_muni_MC_2006-2021.csv") %>% filter(year == 2021) %>% left_join(nicknames)

fig_data <- muni_MC_sums %>% 
  group_by(clean_name, agency_name) %>%
  mutate(total_eav = sum(equalized_av)) %>%
  ungroup() %>%
  mutate(pct_eav = equalized_av / total_eav) %>%
  filter(major_class_code == 2) 

fig_data %>%
    mutate(agency_name = ifelse(agency_name == "TOWN CICERO", "CITY OF CICERO", agency_name)) %>%
  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%
  ggplot(aes(fill = pct_eav)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Residential EAV / Total EAV", 
  caption = "Residential Property includes 200 level property classes.
       The median municipality has 68.9% of its EAV from Class 2 Properties.") +
    theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+# +#+
    scale_fill_steps2(
    high = "darkblue", low = "orange", 
    mid = "gray50", #  guide = "legend",
     midpoint = median(fig_data$pct_eav),
           na.value = NA,   
                       n.breaks = 6,
  show.limits=TRUE,
                        name = "% Residential",
  labels = scales::percent)
```

### Composite Tax Rates

```{r}
Current_Taxrates <- taxcodes_current %>% 
  filter(!agency_num %in% cross_county_lines) %>%
  left_join(muni_agency_names) %>%
  left_join(nicknames, by = c("agency_name" = "agency_name")) %>%
  #filter(!agency_num %in% cross_county_lines) %>%
  group_by(clean_name, agency_name) %>%
  summarize(MuniLevy = sum(final_tax_to_dist, na.rm = TRUE), # amount billed by munis with current exemptions in place
            nonTIF_EAV_post_exemps = sum(final_tax_to_dist/(tax_code_rate/100), na.rm = TRUE),
            TIF_increment_EAV = sum(final_tax_to_tif/(tax_code_rate/100), na.rm=TRUE),  
            Exempt_EAV = sum(tax_amt_exe/(tax_code_rate/100), na.rm=TRUE), 
            Total_EAV = sum((tax_amt_exe+final_tax_to_dist+final_tax_to_tif)/(tax_code_rate/100), na.rm = TRUE)) %>%

  mutate(tax_rate_current = MuniLevy/nonTIF_EAV_post_exemps,
         nonTIF_EAV_pre_exemps = nonTIF_EAV_post_exemps + Exempt_EAV,
         taxrate_new = MuniLevy/nonTIF_EAV_pre_exemps,
         taxrate_change = tax_rate_current-taxrate_new) %>% 
  select(clean_name, taxrate_change, tax_rate_current, taxrate_new, everything()) %>% 
  arrange(desc(tax_rate_current))


# 
# land_use <- taxcodes_current %>% 
#   left_join(muni_agency_names) %>% 
#   left_join(Current_Taxrates) %>%
#   left_join(class_dict, by = c("class" = "class_code")) %>%
#   
# #  mutate(class_1dig = str_sub(class, 1, 1),
# #    ResidentialProps = ifelse(class_1dig %in% c("2", "3", "9"), "Residential", "Non-Residential")) %>%
#   #        PropType = case_when(
#   #          major_class_code %in% c("3","9") ~ "Multi-Family",
#   #          major_class_code == "2" ~ "Single-Family",
#   #          TRUE ~ "Commercial-Industrial")) %>%
#   group_by(clean_name, major_class_code, agency_num, tax_rate_current, taxrate_new, taxrate_change, agency_name) %>% 
#   
#   # All of the values calculated below are AFTER exemptions have been removed
#   summarize(taxrev_from_proptype = sum(final_tax_to_dist, na.rm = TRUE),
#             nonTIF_EAV = sum(final_tax_to_dist/(tax_code_rate/100), na.rm = TRUE),
#             TIF_increment_EAV = sum(final_tax_to_tif/(tax_code_rate/100), na.rm=TRUE),
#             Exempt_EAV = sum(tax_amt_exe/(tax_code_rate/100), na.rm=TRUE),
#             Total_EAV = sum((tax_amt_exe+final_tax_to_dist+final_tax_to_tif)/(tax_code_rate/100), na.rm = TRUE) ) %>% ungroup()
# 
# land_use %>% filter(clean_name %in% c( "Dolton", "Chicago", "Winnetka")) %>% select(-agency_name)


```

```{r Figure-AWM1}
# made in Excel originally

Current_Taxrates %>%  filter(clean_name %in% c("Park Forest","Dolton", "Hillside", "Riverside", "Chicago", "Winnetka", "Rosemont")) %>% select(clean_name, tax_rate_current, taxrate_new) %>% pivot_longer(cols = c(tax_rate_current,taxrate_new), names_to = "names", values_to = "values")  %>%
  ggplot(aes(x = clean_name, y = values, fill = names)) +
  geom_col(position = "dodge" ) + theme_classic() + labs(x="", y = "Composite Tax Rate")


Current_Taxrates %>%  filter(clean_name %in% c("Markham", "Chicago", "Westchester", "Winnetka")) %>% select(clean_name, tax_rate_current, taxrate_new) %>% pivot_longer(cols = c(tax_rate_current,taxrate_new), names_to = "names", values_to = "values")  %>%
  ggplot(aes(x = clean_name, y = values, fill = names)) +
  geom_col(position = "dodge" ) + theme_classic() + labs(x="", y = "Composite Tax Rate")

```

```{r results='hold', out.width= "50%", echo=FALSE}
# 
# burden_shift %>% 
#   filter(major_class_code == 2) %>%
#   summarize(median_currentburden = median(burden_current),
#             median_new_burden = median(burden_noexemps),
#             median_burdenchange = median(burden_change),
#             median_currenttaxrate = median(tax_rate_current),
#             median_taxratechange = median(taxrate_change),
#             median_taxratenew = median(taxrate_new)) %>% 
#   pivot_longer(cols = everything(), names_to = "Variable", values_to = "Median Value")


Current_Taxrates %>% 
    mutate(agency_name = ifelse(agency_name == "TOWN CICERO", "CITY OF CICERO", agency_name) 
    ) %>%
  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%
  
  ggplot(aes(fill = tax_rate_current)) +
  geom_sf(aes(geometry = geometry), color = "black") + 
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
  scale_fill_steps2(#colors = colors,
    high = "plum4", low = "#008080",# mid = "lavenderblush2",
    midpoint = median(Current_Taxrates$tax_rate_current), #  midpoint = 0.123,
    #breaks = breaks_sd,
    limits = c(0,.45),
    show.limits=TRUE,
    na.value = NA,
    nice.breaks=FALSE,
    n =6,
    name = "Tax Rate",
    label = scales::percent )+
  
  labs(title = "Current composite tax rates" ,    
       caption = "The current median composite tax rate is 12.1%. 
         Highest composite tax rate is in Park Forest (41.4%.)
       Lowest composite tax rate is in Chicago (6.7%).")

Current_Taxrates %>% 
      mutate(agency_name = ifelse(agency_name == "TOWN CICERO", "CITY OF CICERO", agency_name) 
    ) %>%
  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%
  ggplot(aes(fill = taxrate_new)) + 
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
  scale_fill_steps2(
    high = "plum4", low = "#008080",
    limits = c(0,.45),
    midpoint = median(Current_Taxrates$taxrate_new),  # midpoint = .1052,
    nice.breaks=FALSE,
    show.limits=TRUE,
    na.value=NA,
    n=6,
    name = "Tax Rate", label = scales::percent)+
  geom_sf(aes(geometry = geometry), color = "black") +  
  labs(title = "New composite tax rates if exemptions were eliminated" ,    
       caption = "The new median composite tax rate would be approximately 10.5% 
       if exemptions were removed.")
```

```{r Figure-AWM2, echo=FALSE}
Current_Taxrates %>% 
  mutate(agency_name = ifelse(agency_name == "TOWN CICERO", "CITY OF CICERO", agency_name) ) %>%
  # filter(clean_name != "Park Forest") %>%
  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%
  ggplot(aes(fill = taxrate_change)) + 
  geom_sf(aes(geometry = geometry), color = "black") +  
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
  
  scale_fill_steps2(high = "blue", low = "black",
    n=6, #midpoint = 0.01388, 
    show.limits=TRUE,
    nice.breaks=FALSE,
    na.value = NA,
    labels = scales::percent,
    name = "Percentage Point \nDifference")+
  labs(title = "Change in Composite Tax Rate if Exemptions are Removed",
       caption = "The median change in composite tax rate is 1.43 percentage points")
```

## Percent of Residential EAV that is tax exempt

For this document, residential EAV includes Class 2 properties only. Residential property can be in and outside of a TIF. Any EAV greater than the frozen EAV within a TIF becomes TIF revenue when it is taxed. We try to only include the frozen EAV and non-TIF EAV in our residential calculations.

```{r}
exemptions_to_class2EAV_ratios <- perc_residential %>% 
  mutate(exemptEAV_pctof_resEAV = exempt_EAV/eav,
         exemptEAV_pctof_totalEAV = exempt_EAV / muni_EAV_includesTIF,
         exemptEAV_pctof_taxbase_current = exempt_EAV / tax_base_current,
         exemptEAV_pctof_taxbase_noexemps = exempt_EAV / tax_base_noexemps,
         nontif_ratio = exempt_EAV / tax_base_noexemps) %>% 
  select(agency_name, clean_name, exemptEAV_pctof_resEAV, nontif_ratio, everything()) %>% 
  arrange(nontif_ratio)

exemptions_to_class2EAV_ratios %>%
    mutate(agency_name = ifelse(agency_name == "TOWN CICERO", "CITY OF CICERO", agency_name) ) %>%
 full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%
  ggplot(aes(fill = exemptEAV_pctof_resEAV)) + 
  geom_sf(aes(geometry = geometry), color = "black") + theme_void()+ 
  labs(title = "Exemptions / Residential EAV (in and out of TIFs)") +
    theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
   scale_fill_steps2(high = "darkblue", low = "black",  mid = "beige",
                      # limits = c(0,.8),
                       n.breaks = 7, show.limits=TRUE,
                       na.value = NA,
                    nice.breaks = FALSE,
                    midpoint = median(exemptions_to_class2EAV_ratios$exemptEAV_pctof_resEAV),
                        name = "% Residential EAV \nthat is exempt", label = scales::percent)
```

```{r, fig.show='hold', out.width="50%"}
exemptions_to_class2EAV_ratios %>%
  mutate(agency_name = ifelse(agency_name == "TOWN CICERO", "CITY OF CICERO", agency_name) ) %>%
  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%
  ggplot(aes(fill = nontif_ratio)) + 
  geom_sf(aes(geometry = geometry), color = "black") + theme_void()+ 
  labs(title = "Non-TIF EAV only: Homestead Exemptions / Residential EAV", caption = "Village of Phoenix skews graph. Dropped in map below") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
  scale_fill_steps2(high = "darkblue", low = "black", mid="beige",
                    #colors = c("white", "darkblue"), 
                    # limits = c(0),
                    midpoint = median(exemptions_to_class2EAV_ratios$nontif_ratio),
                    n.breaks = 5, 
                    na.value = NA,
                    show.limits=TRUE,
                    name = "% Residential EAV \nthat is exempt", label = scales::percent) + 
  labs( caption = "Median value is 20.0% ")


exemptions_to_class2EAV_ratios %>%
  mutate(agency_name = ifelse(agency_name == "TOWN CICERO", "CITY OF CICERO", agency_name) ) %>%
  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%
  ggplot(aes(fill = nontif_ratio)) + 
  geom_sf(aes(geometry = geometry), color = "black") + theme_void()+ 
  labs(title = "Non-TIF EAV only: Homestead Exemptions / Residential EAV", caption = "Village of Phoenix skews graph. Dropped in map below") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
  scale_fill_steps2(high = "darkblue", low = "black",mid ="beige",
                    #colors = c("white", "darkblue"), 
                    # limits = c(0),
                    midpoint = median(exemptions_to_class2EAV_ratios$nontif_ratio),
                    n.breaks = 7, 
                    na.value = NA,
                    show.limits=TRUE,
                    name = "% Residential EAV \nthat is exempt", label = scales::percent) + 
  labs( caption = "Median value is 20.0% ")
```

```{r , fig.show='hold', out.width="50%"}
exemptions_to_class2EAV_ratios %>%  filter(nontif_ratio<.6) %>%
    mutate(agency_name = ifelse(agency_name == "TOWN CICERO", "CITY OF CICERO", agency_name) ) %>%
  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%
  ggplot(aes(fill = nontif_ratio)) + 
  geom_sf(aes(geometry = geometry), color = "black") + theme_void()+ 
  labs(title = "Percent of Residential EAV that is Tax Exempt", 
       subtitle = "% of Non-TIF Residential EAV only: \nHomestead Exemptions / Residential EAV", 
       caption = "") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
  scale_fill_steps2(high = "darkblue", low = "black",mid ="beige",
                    #  colors = c("white", "darkblue"), 
                    limits = c(0,.4),
                    na.value = NA,
                    midpoint = median(exemptions_to_class2EAV_ratios$nontif_ratio),
                    n.breaks = 5, show.limits=TRUE,
                    name = "% Residential EAV \nthat is tax exempt", label = scales::percent)


exemptions_to_class2EAV_ratios %>%  filter(nontif_ratio<.6) %>%
  mutate(agency_name = ifelse(agency_name == "TOWN CICERO", "CITY OF CICERO", agency_name) ) %>%
  full_join(muni_shp, by = c("agency_name" = "AGENCY_DESC")) %>%
  ggplot(aes(fill = nontif_ratio)) + 
  geom_sf(aes(geometry = geometry), color = "black") + theme_void()+ 
  labs(title = "Percent of Residential EAV that is Tax Exempt", 
       subtitle = "% of Non-TIF Residential EAV only: \nHomestead Exemptions / Residential EAV", 
       caption = "") +
  theme_void() + 
  theme(axis.ticks = element_blank(), axis.text = element_blank())+
  scale_fill_steps2(high = "darkblue", low = "black",mid ="beige",
                    #  colors = c("white", "darkblue"), 
                    limits = c(0,.4),
                    na.value = NA,
                    midpoint = median(exemptions_to_class2EAV_ratios$nontif_ratio),
                    n.breaks = 7, show.limits=TRUE,
                    name = "% Residential EAV \nthat is tax exempt", label = scales::percent)
```

Drops Village of Phoenix because it skews map colors (78% of their residential EAV is tax exempt). Median municipality can not get tax revenue from 1/5th of their residential EAV due to exemptions. 20.85% of Single-family home EAV is not taxed and transferred to other tax payers.

In Cook County there is `r scales::dollar(sum(perc_residential$tax_base_current))` in Class 2 Residential Property EAV.

There is `r scales::dollar(sum(perc_residential$tax_base_noexemps))` in Class 2 Residential Property EAV before all exemptions.

The Total EAV for Cook County `r scales::dollar(sum(perc_residential$muni_EAV_includesTIF))`. Total EAV includes TIF increment and exempt EAV.

The current taxable EAV for Cook County `r scales::dollar(sum(perc_residential$muni_tax_base_current))`.

The hypothetical taxable EAV for Cook County `r scales::dollar(sum(perc_residential$muni_tax_base_noexemps))`.

## Average exempt EAV per PIN in tax year 2021

```{r}
NBHs <- read_sf("https://datacatalog.cookcountyil.gov/resource/pcdw-pxtg.geojson")

nbh_sum <- read_csv("../Output/nbh_summary.csv") %>%
    mutate(nbhd_code = as.character(nbhd_code),
           township_code = str_sub(nbhd_code, 1,2))

nbh_MC_sum <- read_csv("../Output/nbh_MC_summary.csv") %>%
    mutate(nbhd_code = as.character(nbhd_code))

nbh_sum %>% 
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>% 
  ggplot(aes(fill = nbh_res_exe/nbh_PC)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Average Exemptions/PIN in Neighborhoods", 
  caption = "Excludes Abatements") +
  theme_void() + 
   theme(axis.ticks = element_blank(), axis.text = element_blank()) +
scale_fill_steps2(
  high = "darkblue", low = "orange",
  midpoint = median(nbh_sum$nbh_res_exe/nbh_sum$nbh_PC, na.rm = TRUE),
 nice.breaks = FALSE,
  show.limits=TRUE,
na.value = "gray",
  name = "Dollars",
  labels = scales::dollar
) 

nbh_sum %>% 
  filter(nbh_exe_abate > 0) %>%
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>% 
  ggplot(aes(fill = nbh_exe_abate) ) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Exempt EAV from Abatements in Assessor Neighborhoods", 
  caption = "Abatements  comes from exe_abate in PTAXSIM data.") +
  theme_void() + 
   theme(axis.ticks = element_blank(), axis.text = element_blank()) +
scale_fill_steps2(
  high = "darkblue", low = "orange",
 # midpoint = mean(nbh_sum$nbh_exe_abate),
 nice.breaks = FALSE,
  show.limits=TRUE,
na.value = NA,
  name = "Dollars",
  labels = scales::dollar
)

nbh_sum %>% 
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>% 
  ggplot(aes(fill = nbh_current_rate)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Composite Tax Rate in Assessor Neighborhoods") +
  theme_void() + 
   theme(axis.ticks = element_blank(), axis.text = element_blank()) +
scale_fill_steps2(
  high = "darkblue", low = "orange",
  midpoint = median(nbh_sum$nbh_current_rate, na.rm = TRUE),
 nice.breaks = FALSE,
  show.limits=TRUE,
na.value = "gray",
  name = "Tax Rate",
  labels = scales::percent
) 
```

```{r}
nbh_MC_sum %>%
    group_by(nbhd_code) %>%
    summarize(pct_taxable_EAV = nbh_nonTIF_EAV_post_exemps /Total_EAV
  ) %>%
 arrange(pct_taxable_EAV) %>%
  full_join(NBHs, by = c("nbhd_code" = "town_nbhd")) %>% 
  ggplot(aes(fill = pct_taxable_EAV)) + 
  geom_sf(aes(geometry = geometry), color = "black") + 
  labs(title = "Percent of EAV that is taxable by taxing districts",
       caption = "Non-taxable EAV includes TIF EAV above the frozen EAV 
       and tax exempt EAV due to various exemptions") +
  theme_void() + 
   theme(axis.ticks = element_blank(), axis.text = element_blank()) +
scale_fill_steps2(
  high = "darkblue", low = "white",
# midpoint = median(pct_taxable_EAV, na.rm = TRUE),
 nice.breaks = FALSE,
  show.limits=TRUE,
na.value = "gray",
  name = "Percent EAV taxable",
  labels = scales::percent
) 
```
